### 2015-09-09 exampleOutline.tex
add example outline tex file

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers

\newcommand{\mbh}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments

\title{Historical demography of maize during diffusion from its center of origin}

\author[1]{Li Wang}%author information
\author[2]{Timothy M. Beissinger}
\author[2,3]{Jeffrey Ross-Ibarra}
\author[1,4]{Matthew B. Hufford}
\affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
\affil[2]{Department of Plant Sciences, University of California Davis, Davis, CA, USA}
\affil[3]{Center for Population Biology and Genome Center, University of California Davis, Davis, CA, USA}
\affil[4]{Corresponding Author: Matthew B. Hufford; 339A Bessey Hall, Iowa State University, Ames, IA, USA; phone: 1-515-294-8511; email: mhufford@iastate.edu}
\date{}

\begin{document}

\maketitle

\section*{Introduction}
\subsection*{Hypothesis on demography of maize diffusion}

Maize as a main crop worldwide, has been mainly studied done on domestication, little on diffusion. Cite van Etten, Vigouroux, van Heerwaarden, Fonseca, Shohei. 
Maize, compared to its wild progenitor, has undergone a domestication bottleneck, during which, its diversity has been largely reduced (Hufford et al. 2012).
Post domestication, maize colonized multiple highland habitats in Americas during its migration. 
When adapted to the highland habitat, does maize population undertake a second bottleneck? 
Or, are they experiencing the serial founder effects with migrating away from its origin center?
Describing the concept of serial founder effects and listed examples of serial founder effects found in Human population history.

We know little about demography except some introgression (and hints from Ross-Ibarra 2009).
However, the gene exchange between maize and its wild progenitor is not well-known to us. 
Ross-Ibarra et al. (2009) pointed out the evidence of introgression between multiple teosinte species and maize solely based on a dozen of nuclear genes.
Can we identify a signature of introgression when revisiting the question with whole genome sequencing data?
Hufford et al. (2013) discovered the adaptive introgression from \emph{mexicana} to maize in Mexican Highland, but question remained unanswered. 
For example, does this kind of introgression is widespread to other highland populations, even those populations without direct contact with \emph{mexicana} populations ?

\subsection*{}

little about how post-domestication evolution has affected genomic diversity (except in landraces in highlands).

\section*{Methods}

\jri{you can start writing these up already}

\section*{Results}

\subsection*{Diversity and Population Structure}

\subsubsection*{SEEDS data}
\begin{itemize}
\item PCA (Fig1A)
\item diversity moving away from center of origin (Fig 1B)
\end{itemize}

\subsubsection*{whole genome sequence data}

To investigate the genomic impact of diffusion post domesticaiton, we sampled 30 genomes from across the range, including from Andes which are undersampled.

\begin{itemize}
\item Sampling
\item PCA and ngsadmix (Figure 2A)
\item genetic diversity and tajima's D (Figure2B)
\item ngsF
\jri{not sure this figure 2 is useful, thoughts?}
\end{itemize}

\subsection*{Impact of introgression}
\begin{itemize}
\item D stats with mexicana (Figure 3A is either Dstats with mex or Gmin with mex, but something mex-introgression related)
\item Gmin with mexicana (inv4m, other regions)
\item D stats with parviglumis, luxurians and diploperrennis (compare to R-I 2009, Doebley, etc.) (FIGURE 3B)
\end{itemize}

\subsection*{Demographic change}
\begin{itemize}
\item MSMC -- consistent with SeeDs?, check consistency with sims (Figure 4 of MSMC)
\item dadi?
\item show introgression not messing this up (supplement)
\end{itemize}

\subsection*{Impacts of a recent strong bottleneck in S. Am.}
\begin{itemize}
\item ROH, GERP homozygosity (Figure 5)
\item burden test?
\end{itemize}

\section*{Discussion}

blah blah lots more to learn.

%
%\begin{enumerate}
%\item \textbf{General characterization of populations}
%\begin{itemize}
%\item Population structure
%
%PCA analyses (ngsTools, ngsCovar):  
%the first PC demonstrates the geographic north-south differentiation; Southwest US and Andes populations are at the two extremes.
%\item Genetic diversity ($\hat\theta_\pi$ from ANGSD)
%\begin{itemize}
%\item Andes exhibited significantly lowest genetic diversity
%\item Southwest US, Guatemala High, SA Low, Mexican Low form intermediate, indistinguishable group
%\item Mexican Highland population has the highest diversity
%
%\item Inbreeding coeffecients (ngsF)
%
%No differences among populations observed, F very low for all individuals, which provides little evidence for recent inbreeding.
%
%
%
%Do introgressed regions contribute to the higher diversity in the Mexican Highlands?  
%\jri{make introgression a separate section, include results from luxurians. we have a lane of PE 100 from diploperennis you could use too. iplant/rossibarra/diplo. just ask and i can give you access.  this lets you test some of the arguments made in Ross-Ibarra 2009}
%
%Is there evidence of \emph{mexicana} introgression in other highland populations (\textit{i.e.}, did \textit{mexicana} haplotypes spread to nearby, allopatric highland regions in Guatemala and the Southwest US or are they limited to the Mexican Highlands?)
%
%Detection of introgression (ABBA-BABA and Gmin scan) further validate the introgression between the Mexican Highland population and \textit{Zea mays} ssp. \textit{mexicana}.
%
%\textcolor{blue}
%{To Do: ABBA-BABA statistics and Gmin scan for highland regions outside of Mexico vs. \textit{mexicana}.}
%
%\jri{need to test D against all lowland mex. also include lowland mex. in test.}
%
%
%\textcolor{blue}
%{Plot $\hat\theta_\pi$ in introgressed regions relative to the genome-wide distribution for each population.}
%\end{itemize}
%
%
%
%\item Demography suggested by Tajima's D (ANGSD)
%
%Andes is the only population with median Tajima's D \textgreater{} 0 (significantly more positive than other pops), indicating an absence of rare alleles, and possibly stronger bottleneck.
%\end{itemize}
%
%\item \textbf{What is the pattern of demography in maize populations?}
%\begin{itemize}
%\item MSMC analyses
%
%Our results suggest serial founder effects, but do we trust these given the runaway behavior and smoothing over demographic changes that have previously been observed in PSMC and Dical?
%
%\textcolor{blue}{To Do: Simulation to confirm applicability of method in maize: simulate a continuous population size reduction and two bottlenecks} \textcolor{red}{Tim: also run in dadi to check the likelihoods of serial founder effect model compared to others?}
%
%\item We observe a significant negative correlation between proportion of polymorphic sites and geographic distance from the Balsas using SEEDS data
%\end{itemize}
%
%\item \textbf{The outcomes of an extreme bottleneck in the Andes}
%\begin{itemize}
%
%
%\item Runs of Homozygosity (PLINK)
%
%In general, the Andes shows more runs of homozygosity (ROH), particularly small ROHs ($<1$MB), consistent with a stronger bottleneck in this population. 
%When looking only at larger ROHs ($>1$MB), which might reflect more recent inbreeding events, there is no difference among populations (confirms ngsF result).
%\mbh{Li, we should summarize this result based on genetic distances}
%
%\item Mutation load
%
%The Andes does not show higher mutation load when considering both the heterozygous and homozygous non-reference sites, but does exhibit higher load when considering \emph{ONLY} the non-reference homozygous sites, indicating a potential role of dominance in the mutation load. 
%With a higher GERP cutoff (moderately and strongly deleterious sites), the difference in mutation load between the Andes and other populations is even stronger. 
%Additionally, the Andes has a higher percentage of derived homozygotes (polarized vs. \textit{Tripsacum}), consistent with less efficient selection due to a stronger bottleneck.
%\jri{consider adding the burden ratio test here}
%\end{itemize}
%
%
%\item \textbf{Potential additional analyses}
%\begin{itemize}
%\item \textcolor{blue}{Running MSMC without introgressed regions or with exclusively introgression regions.}
%\end{itemize}
%\end{enumerate}

\end{document}

### 2015-09-15 crop_adaptation_gene_flow_gj_outline_temp.tex
Create crop_adaptation_gene_flow_gj_outline_temp.tex
I will delete this file when I successfully link my local directory to this repository.

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers

\newcommand{\mbh}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments

\title{Crop adaptation through gene flow (outline)}

\begin{document}

\maketitle

\section*{Outline}
\subsection*{Outline parts}
Introgression in Crops (General Information)
Crop Examples
    Maize
    Wheat/Rye
    Barley
    Sunflower
    Tomato?
Non-Crop System Examples?

1\subsection*{Introgression in Crops (General)}

From {label22}
A very good review.
Hybrid zones make for a great model system of study.
Introgression from wild populations into agronomic populations is potentially beneficial for breeding programmes.
It is generally easier to introgress into crops than introgressing into wild populations (stronger selection, overcoming LD).

From {label32}
Another very good review, with more of a molecular emphasis

From {label34}
The difficulties of utilizing wild relatives in breeding programmes effectively.

From {label6} {label7} {label26} {Label27}
Lots of papers focus on gene flow from crop to wild relative, particularly about transgene escape

\subsection*{Maize}

From {label8}
"All things considered, there appears to be little or no clear-cut evidence to support the idea that the teosintes have been greatly altered by maize introgression."

From {label7}
Lists papers citing the sources of barriers to gene flow between teosinte and maize

From {label1}, {label2}, {label20}
Nobogame Valley, Mexico, where teosinte genes flow into local maize.  This gene flow was measured morphologically and imparted positive agronomic results.

From {label10}
"Several races of maize that contain isozyme alleles that have apparently been transferred to maize from subsp. mexicana by introgression." (quote from {label9}

From {label4}
Table 4, STRUCTURE analysis of admixture in maize, source and destination included (gene flow between teosinte and maize, both directions)

From {label11}
Gene flow is probably greater from maize to teosinte.
Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte.
Gene flow from maize to teosinte occurs most easily when teosinte pollinates maize

From {label12}
Gene flow from Parv and "Chalco" teosinte into maize for the past 200 years, may still be ongoing
These teosintes should be preserved for their potential to fuel new breeding programs.

From {label13} {label14}
Maize-maize introgression
Research on introgression between highland and lowland maize shows adaptive traits being transferred

From {label16}
Gene flow between mexicana and maize is low enough that bad alleles are unlikely to be passed, only neutral/beneficial alleles. (from Slatkin 1987)

From {label17}
mexicana introgression into highland maize landraces
introgression into maize favored over introgression into teosinte

From {label18}
higher mexican gene flow into landraces Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno, but also into mexican maize (non-landrace) at the same elevation

From {label19}
Similar to label18 and label17, mexicana and parviglumis into maize (simulation)

From {label32}
most common with mexicana (introgression is)

\subsection*{wheat}

From {label24}
review/overview.  See table 1

From {label25}
Wheat was given leaf rust resistance genes from wild populations (summarized in Table 1)

From {label26}
A few examples (wild genes into wheat) given.  Originally, only with cross-compatible species, now with cross-incompatible species.
Cites and tests several QTLs

From {label31}
"suggestions are made concerning techniques for exploitation of the wild diploid species in wheat breeding programs."

\subsection*{barley}

From {label24}
review/overview.  See table 1

From {label30}
"wild barley does harbour valuable alleles, which can enrich the genetic basis of cultivated barley and improve quantitative agronomic traits."

\subsection*{sunflower}

From {label33}
yeah, introgression with wild relatives may be common, unsure about direction
intermediates found in the field

\subsection*{non-crop systems}
From {label3}
Bacteria, malaria mosquitoes, blackflies, Darwin's finches, butterflies (each example has a cited source, not included in the bibliography of this outline).
Also, non-crop plants, like trees (oak, larch).

From {label15}
Helianthus, Iris
Mouse, Salamander

From {label21}
Iris, Helianthus, Cowania/Purshia
Dacus (now Bactrocera), Anopheles
Haemophilus influenza, Trypanosoma cruzi

From {label23}
165 proposed cases of introgression, 65 "deemed to be sufficiently documented" (plants)

From {label28}
Milkweeds

From {label29}
White clover

\begin{thebibliography}{1}
\bibitem{label1} H. G. Wilkes (1977) 
Hybridization of Maize and Teosinte, in Mexico and Guatemala and the Improvement of Maize. 
{\em Economic Botany\} 31:254-293

\bibitem{label2} C. Lumholtz (1902)
Unknown Mexico. Charles Scribner's Sons, New York. Vol I. 530 pp.
%This citation was problematic

\bibitem{label3} J. Mallet (2005)
Hybridization as an invasion of the genome.
{\em TRENDS in Ecology and Evolutin\} 20:229-237

\bibitem{label4} J. Ross-Ibarra, M. Tenaillon, B. S. Gaut (2009)
Historical Divergence and Gene Flow in the Genus Zea.
{\em Genetics\} 181:1399-1413

\bibitem{label5} N. C. Ellstrand, H. C. Prentice, J. F. Hancock (1990)
Gene Flow and Introgression from Domesticated Plants into their Wild Relatives.
{\em Annual Review of Ecology and Systematics\} 30:539-563

\bibitem{label6} D. Quist, I. H. Chapela (2001)
Transgenic DNA introgressed into traditional maize landraces in Oaxaca, Mexico.
{\em Nature\} 414:541-543
%RETRACTED?

\bibitem{label7} J. Doebley (1990)
Molecular Evidence for Gene Flow among {\em Zea\} Species.
{\em BioScience\} 40:443-448

\bibitem{label8} J. F. Doebley (1984)
Maize Introgression into Teosinte - A Reappraisal.
{\em Annals of the Missouri Botanical Garden\} 71:1100-1113

\bibitem{label9} J. Doebley (1990)
Molecular Evidence and the Evolution of Maize.
{\em Economic Botany\} 44:6-27
%a review paper

\bibitem{label10} J. Doebley, M. M. Goodman, C. W. Stuber (1987)
Patterns of isozyme variation between maize and Mexican annual teosinte.
{\em Economic Botany\} 41:234-246

\bibitem{label11} B. M. Baltazar, J. de J. Sanchez-Gonzalez (2005)
Pollination between maize and teosinte: an important determinant of gene flow in Mexico
{\em Theoretical Applied Genetics\} 110:519-526
%special character for the Sanchez name, plus how to abbreviate "de"

\bibitem{label12} M. L. Warburton, G. Wilkes, S. Taba, A. Charcosset, C. Mir, F. Dumas, D. Madur, S. Dreisigacker, C. Bedoya, B. M. Prasanna, C. X. Xie, S. Hearne, J. Franco (2011)
Gene flow among different teosinte taxa and into the domesticated maize gene pool.
{\em Genetic Resources and Crop Evolution\} 58:1243-1261

\bibitem{label13} C. Jiang, G. O. Edmeades, I. Armstead, H. R. Lafitte, M. D. Hayward, D. Hoisington (1999)
Genetic analysis of adaptation of adaptation differences between highland and lowland tropical maize using molecular markers.
{\em Theoretical Applied Genetics\} 99:1106-1119

\bibitem{label14} E. Bitocchi, L. Nanni, M. Rossi, D. Rau, E. Bellucci, A. Giardini, A. Buonamici, G. G. Vendramin, R. Papa (2009)
Introgression from modern hybrid varieties into landrace populations of maize ({\em Zea mays\} ssp. {\em mays\} L.) in central Italy.
{\em Molecular Ecology\} 18:603-621

\bibitem{label15} M. L. Arnold, N. H. Martin (2009)
Adaptation by introgression.
{\em Journal of Biology\} 8:82
%mini review of introgression

\bibitem{label16} N. C. Ellstrand, C. Garner, S. Hegde, R. Guadagnuolo, L. Blancas (2007)
Spontaneous hybridization between Maize and Teosinte.
{\em Journal of Heredity\} 98:183-187

\bibitem{label17} M. B. Hufford, P. Lubinsky, T. Pyhajarvi, M. T. Devengenzo, N. C. Ellstrand, J. Ross-Ibarra (2013)
The genomic signature of crop-wild introgression in maize.
{\em PLOS Genetics\} 9:
%special characters, page numbers?

\bibitem{label18} Y. Matsuoka, Y. Vigouroux, M. M. Goodman, J. Sanchez G., E. Buckler, J. Doebley (2002)
A single domestication for maize shown by multilocus microsatellite genotyping.
{\em PNAS\} 99:6080-6084

\bibitem{label19} J. van Heerwaarden, J. Doebley, W. H. Briggs, J. C. Glaubitz, M. M. Goodman, J. de Jesus Sanchez Gonzalez, J. Ross-Ibarra (2010)
Genetic signals of origin, spread, and introgression in a large sample of maize landraces.
{\em PNAS\} 108:1088-1092
%Gonzalez, "van"

\bibitem{label20} H. G. Wilkes (1970)
Teosinte introgression in the maize of the Nobogame Valley
{\em Botanical Museum Leaflets\} 22:297-311

\bibitem{label21} M. L. Arnold (2004)
Transfer and origin of adaptations through natural hybridization: Were Anderson and Stebbins Right?
{\em The Plant Cell\} 16:562-570
%introgression review

\bibitem{label22} C. N. Stewart Jr., M. D. Halfhill, S. Warwick (2003)
Transgene introgression from genetically modified crops to their wild relatives.
{\em Nature\} 4:806-817
%is nature reviews the same thing as nature?
%good review of introgression with crops
%Table 1 shows crops and their progenitor species and main wild relatives

\bibitem{label23} L. H. Rieseberg, J. F. Wendel (1993)
Hybrid zones and the evolutionary process.
{\em Journal of Evolutionary Biology\} 7:631-634

\bibitem{label24} E. Nevo, G. Chen (2010)
Drought and salt tolerances in wild relatives for wheat and barley improvement
{\em Plant, Cell & Environment\} 33:670-685

\bibitem{label25} E. Autrique, R. P. Singh, S. D. Tanksley, M. E. Sorrells (1995)
Molecular markers for four lef rust resistance genes introgressed into wheat from wild relatives
{\em Genome\} 38:75-83

\bibitem{label26} S. G. Hegde, J. G. Waines (2004)
Hybridization and introgression between bread wheat and wild and weedy relatives in North America.
{\em Crop Science\} 44:1145-1155

\bibitem{label27} M. A. Chapman, J. M. Burke (2006)
Letting the gene out of the bottle: The population genetics of genetically modified crops.
{\em New Phytologist\} 170:429-443

\bibitem{label28} S. B. Broyles (2002)
Hybrid bridges to gene flow: A case study in milkweeds ({\em Asclepias\}).
{\em Evolution\} 56:1943-1953

\bibitem{label29} S. W. Hussain, W. M. Williams (1997)
Development of a fertile genetic bridge between {\em Trifolum ambiguum\} M. Bieb. and {\em T. repens\} L.
{\em Theoretical Applied Genetics\} 95:678-690

\bibitem{label30} M. von Korff, H. Wang, J. Leon, K. Pillen (2006)
AB-QTL analysis in spring barley: II. Detection of favourable exotic alleles for agronomic traits introgressed from wild barley ({\em H. vulgare\} ssp. {\em spontaneum\}).
{\em Theoretical Applied Genetics\} 112:1221-1231
%special characters

\bibitem{label31} D. Zohary, J. R. Harlan, A. Vardi (1969)
The wild diploid progenitors of wheat and their breeding value.
{\em Euphytica\} 18:58-65

\bibitem{label32} [[A huge fricken list of names]] (2013)
Hybridization and speciation.
{\em Journal of Evolutionary Biology\} 26:229-246

\bibitem{label32} K. Fukunaga, J. Hill, Y. Virouroux, Y. Matsuoka, J. Sanchez G., K. Liu, E. S. Buckler, J. Doebley (2005)
Genetic diversity and Population Structure of Teosinte.
{\em Genetics\} 169:2241-2254

\bibitem{label33} M. Reagon, A. A. Snow (2006)
Cultivated {\em Helianthus annus\} (Asteraceae) volunteers as a genetic "bridge" to weedy sunflower populations in North America.
{\em Ammerican Journal of Botany\} 93:127-133

\bibitem{label34} J. A. Able, P. Langridge, A. S. Milligan (2006)
Capturing diversity in the cereals: many options but little promiscuity.
{\em TRENDS in Plant Science\} 12:1360-1385

\end{document}

### 2015-09-22 Li_outline.tex
Li added outline tex

\documentclass[11pt]{article}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
\usepackage[normalem]{ulem}
\usepackage{multibib}

\title{Crop adaptation with gene flow from wild relatives}

\author[1]{author 1}%author information
\author[1]{author 2}
\author[1, 4]{author 3}
\affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
\affil[4]{Corresponding Author: Matthew B. Hufford; 339A Bessey Hall, Iowa State University, Ames, IA, USA; phone: 1-515-294-8511; email: mhufford@iastate.edu}
\date{}

\begin{document}

\maketitle

\section{Perspectives and key issues}
\begin{itemize}
%\item ways to detect gene flow 
%compromise between sample size of the source population and the SNP density across the genome
%\item ways to detect selection
%The methods to detect both gene flow and selection are summarized in Racimo et al. (2015) (archaic adaptive introgression in humans)
\item What is the proportion of immigrants showing the signatures of adaptation across the genome
\item The extent to which adapted immigrants are physically linked on chromosomes and reside within structural features restricting recombination, such as inversions
\item the importance of genome structures on local adaptation
An antagonism exists between selection for combination of locally adapted genes, and migration and recombination tending to break them down.
Therefore, genome structures, such as inversion and centromeres, which contrains the recombination, can boost the effectiveness of selection by extending the linkage disequilibrium.
\end{itemize}

\section{theory of adaptative introgression}

\section{Empirical Data}
\subsection{maize}
Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it is sympatric with a lot of its wild relatives with a range of different divergence time \cite{hufford2013}.
\emph{Zea mays} ssp. \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize is a well-studied pair.
Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
However, \citep{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAMPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled. 
Among the nine regions, three spans the centromeres of chromosomes 5, 6 and 10 and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
It was also revealed that in the growth chamber experiment maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
To put it together, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
%Finally, the contribution of \emph{mexicana} genetic resource to a broader maize panel was evaluated by analyzing patterns of identity by state (IBS) between maize and \emph{mexicana} as well as between maize and \emph{Zea mays} ssp. \emph {pulviglumis} (referred to \emph {pulviglumis} hereafter).

Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported \cite{hufford2013}, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations, which are allopatric to \emph{mexicana}. 
\citep{Takuno2015} found little empirical and theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly owing to the difference in the potential for adaptive introgression from wild relatives.
Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.
Furthermore, the adaptive SNPs in Mexican highland population were more likely also showing signatures of local adaptation in \emph{mexicana} and \emph {pulviglumis} populations than those South American highland population.
To sum up, the adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.
A recent study \cite{Wang2015manuscript} revealed that gene flow exist from \emph{mexicana} not only to Mexican highland maize, but also to Guatamalan highland maize, as well as to some individuals in Southwestern US Highland and even Andes, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations depending on the distance away from the distribution of \emph{mexicana}. 
However, whether the introgression from \emph{mexicana} to those highland populations contributes to the highland adaptation is awaiting for future research.

\subsection{barley}
\subsection{wheat}
\subsection{sunflower}

\section{Concluding remarks and future directions}

\bibliography{Reference}
%\bibliographystyle{...}
%\onecolumn
%\input{supportingInfo}

% taxa in selection/migration equilibrium
%transition from individual gene to whole-genome understanding
% extent of local adaptation to environmental factors, as well as patterns of gene flow
% migrant alleles or migrant haplotypes

\end{document}

### 2015-09-22 Li_outline.tex
Li added outline

@@
 \usepackage{color}
-\usepackage{pgf}
-\usepackage[normalem]{ulem}
-\usepackage{multibib}
+\usepackage{authblk}%allows footnote format for authors
+\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
 
@@
 
-\author[1]{author 1}%author information
-\author[1]{author 2}
-\author[1, 4]{author 3}
+\author[1] {author1}
+\author[1] {author2}
+\author[1,2] {author3}
 \affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
-\affil[4]{Corresponding Author: Matthew B. Hufford; 339A Bessey Hall, Iowa State University, Ames, IA, USA; phone: 1-515-294-8511; email: mhufford@iastate.edu}
+\affil[2]{Corresponding Author: Matthew B. Hufford; 339A Bessey Hall, Iowa State University, Ames, IA, USA; phone: 1-515-294-8511; email: mhufford@iastate.edu}
 \date{}
@@
 \begin{itemize}
-%\item ways to detect gene flow 
-%compromise between sample size of the source population and the SNP density across the genome
-%\item ways to detect selection
-%The methods to detect both gene flow and selection are summarized in Racimo et al. (2015) (archaic adaptive introgression in humans)
+\item ways to detect gene flow
+
+compromise between sample size of the source population and the SNP density across the genome
+\item ways to detect selection
+
+The methods to detect both gene flow and selection are summarized in Racimo et al. (2015) (archaic adaptive introgression in humans), probably we should not write the first two items.
 \item What is the proportion of immigrants showing the signatures of adaptation across the genome
+\item Most studies targeted quite long introgression region (in Mb); it is difficult to pinpoint the loci which are introgressed and caused the linkage.
 \item The extent to which adapted immigrants are physically linked on chromosomes and reside within structural features restricting recombination, such as inversions
 \item the importance of genome structures on local adaptation
+
 An antagonism exists between selection for combination of locally adapted genes, and migration and recombination tending to break them down.
-Therefore, genome structures, such as inversion and centromeres, which contrains the recombination, can boost the effectiveness of selection by extending the linkage disequilibrium.
+Therefore, genome structures, such as inversion and centromeres, which constrains the recombination, can boost the effectiveness of selection by extending the linkage disequilibrium.
 \end{itemize}
 
-\section{theory of adaptative introgression}
+\section{Theory of adaptative introgression}
 
@@
 Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported \cite{hufford2013}, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations, which are allopatric to \emph{mexicana}.
-\citep{Takuno2015} found little empirical and theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly owing to the difference in the potential for adaptive introgression from wild relatives.
+A recent study \cite{Takuno2015} found little empirical and theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly owing to the difference in the potential for adaptive introgression from wild relatives.
 Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.

### 2015-09-29 crop_adaptation_gene_flow_gj_outline.tex
updated crop_adaptation_gene_flow_gj_outline.tex

--- 2015-09-15 crop_adaptation_gene_flow_gj_outline_temp.tex
@@
 
+Single Center of Origin:
+Wheat - With wild relatives
+Maize - With wild relatives
+Potato - With wild relatives
+(Cassava)
+
+Other:
+Barley - With wild relatives
+Tomato (unknown Center of Origin) - With wild relatives
+Sunflower (unknown Center of Origin)
+Rice - With wild relatives (both domesticated monophyletic varieties)
+Common Bean - Two regions of domestication, one or more domestication events within each
+(Cotton)
+(Millet)
+(Squash)
+
+\subsection*{Introduction}
+ "Whenever sympatric populations of crops and their wild relatives show a similar trait, there will be several possible explanations: (1) retention of an ancestral trait, (2) convergent or parallel evolution, and (3) introgression.
+Perhaps even more difficult is establishing the direction of introgression; does gene flow go from crop into wild species, the reverse, or in both directions?" and "For this reason, introgression may be best studied between crops and their more distant wild relatives where the probability of finding taxon-specific genetic markers is increased." \cite{doebley1990molecular}
+
 1\subsection*{Introgression in Crops (General)}
 
-From {label22}
+From {stewart2003transgene}
 A very good review.
@@
 
+Order:
+the crop and the wild ancestor
+    where
+    how long ago
+history of the study of that relationship
+traits that are conferred
+current application (realized or hypothetical)
+
 \subsection*{Maize}
 
-From {label8}
-"All things considered, there appears to be little or no clear-cut evidence to support the idea that the teosintes have been greatly altered by maize introgression."
+Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it still grows in sympatry with many of its wild relatives \cite{hufford2013}.
 
-From {label7}
-Lists papers citing the sources of barriers to gene flow between teosinte and maize
+Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite {matsuoka2002single} from teosinte.
 
-From {label1}, {label2}, {label20}
-Nobogame Valley, Mexico, where teosinte genes flow into local maize.  This gene flow was measured morphologically and imparted positive agronomic results.
+Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
 
-From {label10}
-"Several races of maize that contain isozyme alleles that have apparently been transferred to maize from subsp. mexicana by introgression." (quote from {label9}
+In particular, \emph{Zea mays} ssp.
+\emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
+\emp{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
 
-From {label4}
-Table 4, STRUCTURE analysis of admixture in maize, source and destination included (gene flow between teosinte and maize, both directions)
+Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
+[[Li's and my bib tags might be in conflict, e.g.
+wilkes1977 is likely my wilkes1977hybridization.]]
+.
+ Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
 
-From {label11}
-Gene flow is probably greater from maize to teosinte.
-Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte.
+Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
+
+However, \cite{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
+The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled.
+Among the nine regions, three spans the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
+By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
+Growth chamber experiments with maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
+
+Gene flow between \emph{mexicana} and maize is low enough that only fitness-neutral or advantageous alleles are likely to be passed \cite{slatkin1987gene}.
+
+From \cite{baltazar2005pollination}
+Gene flow is probably greater from teosinte to maize.
+Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte and slow rates of introgression.
 Gene flow from maize to teosinte occurs most easily when teosinte pollinates maize
+[[what paper can i cite to mention the gene tb1 and its role in hybridization, viz a viz restricting directionality of teosinte pollen into maize
 
-From {label12}
-Gene flow from Parv and "Chalco" teosinte into maize for the past 200 years, may still be ongoing
-These teosintes should be preserved for their potential to fuel new breeding programs.
-
-From {label13} {label14}
+From \cite{jiang1999genetic}, {bitocchi2009introgression}
 Maize-maize introgression
@@
 
-From {label16}
-Gene flow between mexicana and maize is low enough that bad alleles are unlikely to be passed, only neutral/beneficial alleles. (from Slatkin 1987)
-
-From {label17}
-mexicana introgression into highland maize landraces
-introgression into maize favored over introgression into teosinte
-
-From {label18}
-higher mexican gene flow into landraces Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno, but also into mexican maize (non-landrace) at the same elevation
-
-From {label19}
-Similar to label18 and label17, mexicana and parviglumis into maize (simulation)
-
-From {label32}
-most common with mexicana (introgression is)
+\cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
 
 \subsection*{wheat}
+
+Domesticated einkorn wheat (Triticum monococcum monococcum), wild relative einkorn wheat {T.
+m.
+boeoticum), and the progenitor T.
+m.
+boeoticum from the Karacadag [[special character]] mountains of southeast Turkey \cite{heun1997site}.
+Wild emmer wheat (Triticum turgidum subsp. dicoccoides), the progenitor of modern (tetraploid and hexaploid) wheat, is from the Fertile Crescent \cite{lev2000cradle}.
+Two distinct taxa of Triticum dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern maize was domesticated \cite{ozkan2005reconsideration}.
+Localizing domestication centers for other forms of wheat at a resolution greater than the Fertile Crescent is not yet possible \cite{nesbitt1998wheat}, but then \cite{lev2000cradle} point to "a small core area within the Fertile Crescent—near the upper reaches of the Tigris and Euphrates rivers [HN6] in present-day southeastern Turkey/northern Syria".
+So, wheat has one restricted center of domestication, with wild relatives nearby.
+From \cite{ozkan2005reconsideration}, "The geographical distribution of T.
+dicoccoides reported by Zohary and Hopf (2000) includes the western Fertile Crescent, the central part of southeastern Turkey and areas in eastern Iran and Iraq.
+Johnson (1975) reports that from southeastern Turkey to Iran and Iraq the species is progressively substituted by the wild tetraploid wheat T.
+araraticum.
+T.
+dicoccoides also grows on the basaltic rocky slopes of the Karacadag mountains in southeastern Turkey (Johnson 1975; Harlan and Zohary 1966).
+The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}.
+
+The following are just notes.
 
@@
 
+Polyphyletic \cite{azhaguvel2007phylogenetic}.
+\cite{von1995ecogeographical} states that barley and its progenitor are the same species, as they are not sufficiently genetically distinct and geneflow continues unhindered.
+The progenitor species is Hordeum spontaneum \cite{oka2012origin}.
+
+Notes:
+
 From {label24}
@@
 intermediates found in the field
+
+\subsection*{rice}
+
+The center of rice domestication is not known with confidence, with evidence pointing towards both China and India \cite{oka2012origin} [[see the full review on page 16]].
+There are main cultigens of rice; common rice, Oryza sativa, and African rice, O.
+glaberrima, which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and because O.
+glaberrima is annual \cite{oka2012origin}.
+Oryza sativa was domesticated from wild O. rufipogon "or the Asian form of O. perennis complex" \cite{oka2012origin}, whereas O. glaberrima was domesticated from O. breviligulata \cite{oka2012origin}.
+Both O. sativa/O. rufipogon and O. glaberrima/O.breviligulata produce weedy hybrids.
+The greater genetic diversity within O.
+sativa is likely due to introgression with wild relatives both during domestication and upon the dispersal of O.
+sativa into new environments and sympatry with new relatives \cite{second1982origin}.
+Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into O.
+sativa L.
+by researchers \cite{brar1997alien, khush1974inheritance}, but adaptive introgression has also happened without human intervention \cite{second1982origin}.
+
+\subsection*{potato}
+
+Modern Solanum tuberosum was domesticated in southern Peru from northern members of the polyphyletic S. brevicaule complex \cite{spooner2005single}.
+These northern members of the S. breviaule complex are not clearly defined, and may in fact be one singular species (in which case the species name would be S. bukasovii) \cite{spooner2005single}.
+Resistance genes have been experimentally introgressed from wild poptato relatives (S.
+tuberosum subsp.
+andigena, \cite{van1999tight}, S.
+bulbocastanum, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
+
+\subsection*{tomato}
+
+The domestication center for tomato is not known with certainty, but \cite{sims1979history} quotes Dr.
+Charles Rick (University of California, August 1978) stating that tomato was domesticated from Lycopersicon esculentum v.
+cerasiformae in the New World (most likely Mexico).
+Most of the genetic diversity resides within the wild relatives of the tomato \cite{rick1988tomato} \cite{miller1990rflp}, so breeders have long turned to these plants for their germplasm \cite{rick1982potential}.
+\cite{hanson2000mapping} mapped the tomato yellow leaf curl virus resistance gene in L.
+hirsutum f.
+glabratum accession 'B 6013' (line H24) (from \cite{banerjee1990transfer}) to an introgression on chromosome 11.
+
+\subsection*{common bean}
+
+\subsection*{Other Notes}
+
+\cite{zohary2012domestication} has Map 1, showing the domestication center of einkorn wheat, emmer wheat, barley, chickpea, flax, lentil, pea, and bitter vetch in the Fertile Crescent.
+Most of the wild progenitors of founder crops (excluding barley and flax) have a distribution restricted to the Fertile Crescent \cite{zohary2012domestication}.
+However, such exceptions as barley, flax, foxtail millet, and oat have wild progenitors which inhabit a wide range of habitats beyond the Fertile Crescent.
+
+\cite{jarvis1999wild} Table 2 is an incredible resource, cited examples of farmer selection and/or use of introgressed types, and confirmation of introgression
+
+https://books.google.com/books?hl=en&lr=&id=CvcbUopfa54C&oi=fnd&pg=PA7&dq=common+bean+domestication&ots=ggAF-EoWG8&sig=dI9wOyATpzAelxLf_cg-gs0J2gI#v=onepage&q=common%20bean%20domestication&f=false
 
@@
 \bibitem{label13} C. Jiang, G. O. Edmeades, I. Armstead, H. R. Lafitte, M. D. Hayward, D. Hoisington (1999)
-Genetic analysis of adaptation of adaptation differences between highland and lowland tropical maize using molecular markers.
+Genetic analysis of adaptation of differences between highland and lowland tropical maize using molecular markers.
 {\em Theoretical Applied Genetics\} 99:1106-1119
@@
 \bibitem{label24} E. Nevo, G. Chen (2010)
-Drought and salt tolerances in wild relatives for wheat and barley improvement
-{\em Plant, Cell & Environment\} 33:670-685
+Drought and salt tolerances in wild relatives for wheat and barley improvement.
+{\em Plant, Cell Environment\} 33:670-685
 
@@
 
+\end{thebibliography}{1}
+
 \end{document}

### 2015-09-29 Li_outline.tex

@@
 To sum up, the adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.
-A recent study \cite{Wang2015manuscript} revealed that gene flow exist from \emph{mexicana} not only to Mexican highland maize, but also to Guatamalan highland maize, as well as to some individuals in Southwestern US Highland and even Andes, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations depending on the distance away from the distribution of \emph{mexicana}.
-However, whether the introgression from \emph{mexicana} to those highland populations contributes to the highland adaptation is awaiting for future research.
+
+A recent study \cite{Wang2015manuscript} revealed that gene flow exist from \emph{mexicana} not only to Mexican highland maize, but also to Guatamalan highland maize, as well as to some individuals in Southwestern US Highland, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations depending on the distance away from the distribution of \emph{mexicana}.
+The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
+The stronger genetic drift introduced by the smaller historical population size improved the homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
+The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
+Even though it is clear that the absence of introgression from wild relatives provided less genetic resource for the highland adaptation and thus made the highland adaptation unique in the Andes, it is yet unknown that out of reach of wild relatives is a reason for reduced fitness in the Andes population.
+And the question whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
 

### 2015-10-19 Li_outline.tex
Li made modifications

@@
 
-A recent study \cite{Wang2015manuscript} revealed that gene flow exist from \emph{mexicana} not only to Mexican highland maize, but also to Guatamalan highland maize, as well as to some individuals in Southwestern US Highland, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations depending on the distance away from the distribution of \emph{mexicana}.
+However, has the introgression from \emph{mexicana} spread to multiple highland populations? 
+How does the adaptation differ among populations with and without introgression from \emph{mexicana}?
+A recent study \cite{Wang2015manuscript} first utilized the ABBA-BABA statistics to evaluate the existence or absence of introgression from  \emph{mexicana} to multiple highland populations, and then $\hat{f_{d}}$ statistic proposed by Martin et al.
+(2015) \citep{martin2015} were calculated to locate the introgressed loci.
+Figure \ref{fig:fd} showed the loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across two chromosomes (chr4 and chr5) in multiple comparisons. 
+On chromosome 4 (Figure \ref{fig:fd}) , Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
+(2013) \citep{hufford2013}.
+The signal of introgression is absent for the other three populations.
+On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
+More details on the other chromosomes could be referred to \cite{Wang2015manuscript}.
+Overall, both analyses revealed that gene flow exist from \emph{mexicana} not only to MexHigh maize, but also to GuaHigh maize, as well as to some individuals in SW\_US, but not to the Andes and South American lowland (SA\_Low) populations, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations.
 The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
-The stronger genetic drift introduced by the smaller historical population size improved the homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
+%The stronger genetic drift introduced by the smaller historical population size improved the homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
 The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
@@
 And the question whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+
+\begin{figure*}[b]
+\centering
+\includegraphics[width=.45\textwidth]{FigsAndTables/fdTrackChr4AllPop}
+\hspace{0.05\textwidth} 
+\includegraphics[width=.45\textwidth]{FigsAndTables/fdTrackChr5AllPop}
+\caption{Loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across the genome in   five independent comparisons: MexLow\-MexHigh\-mexicana\-outgroup, MexLow\-GuaHigh\-mexicana\-outgroup, MexLow\-SW\_US\-mexicana\-outgroup, MexLow\-Andes\-mexicana\-outgroup, MexLow\-SA_Low\-mexicana\-outgroup.
+A: Chromosome 4; B: Chromosome 5.
+\label{fig:fd}}
+\end{figure*}
 

### 2015-11-30 crop_adaptation_gene_flow_gj_outline.tex
updating crop_adaptation_gene_flow_gj_outline.tex

@@
 
-\title{Crop adaptation through gene flow (outline)}
+\title{Crop adaptation through gene flow}
 
@@
 \subsection*{Outline parts}
-Introgression in Crops (General Information)
-Crop Examples
-    Maize
-    Wheat/Rye
-    Barley
-    Sunflower
-    Tomato?
-Non-Crop System Examples?
-
-Single Center of Origin:
-Wheat - With wild relatives
-Maize - With wild relatives
-Potato - With wild relatives
-(Cassava)
-
-Other:
-Barley - With wild relatives
-Tomato (unknown Center of Origin) - With wild relatives
-Sunflower (unknown Center of Origin)
-Rice - With wild relatives (both domesticated monophyletic varieties)
-Common Bean - Two regions of domestication, one or more domestication events within each
-(Cotton)
-(Millet)
-(Squash)
 
@@
 
+\cite{harlan1966distribution}\ The premise; domestication takes place at the edges of a progenitor's range, where it barely grows.
+there, farmers work on the crop to make it more adapted.
+if this is the case, this could maintain progenitor populations long after domestication, and may increase gene flow between the original progenitor population and the expanding domesticate.
+This is probably unlikely to be advantageous for the domesticate, however.
+
 From {label32}
@@
 Lots of papers focus on gene flow from crop to wild relative, particularly about transgene escape
-
-Order:
-the crop and the wild ancestor
-    where
-    how long ago
-history of the study of that relationship
-traits that are conferred
-current application (realized or hypothetical)
 
@@
 
+\cite{hammer1987collection}\ cites an example in Italy of one domesticated maize adapting to coastal environment by introgressions from another domesticated maize.
+
 \subsection*{wheat}
 
-Domesticated einkorn wheat (Triticum monococcum monococcum), wild relative einkorn wheat {T.
+The first domesticated wheat was einkorn wheat (Triticum monococcum), domesticated from wild einkorn (T.
+m.
+boeoticum) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
+Subsequently, emmer wheat (Triticum dicoccum) was domesticated from hybrids of diverse populations of wild emmer wheat (T.
+dioccoides) in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure}.
+Emmer wheat is tetraploid, and hybridization with wild diploid Tausch's goatgrass (Aegilops tauschii) lead to the production of hexaploid wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
+Emmer wheat has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
+
+Durum wheat (T. turgidum ssp. durum) was developed from emmer wheat as well.
+
+Common bread wheat, T.
+aestivum, is an allohexaploid with chromosome sets from red wild einkorn wheat (T.
+urartu), Aegilops speltoides, and Aegilops tauschii \cite{mcfadden1946origin, petersen2006phylogenetic}\, and furthermore, each of these species is the result of ancient hybridizations \cite{marcussen2014ancient}\.
+Hexaploidization probably occured twice, in southeast Turkey/northern Syria and, more recently, Iran \cite{giles2006gludy}\.
+
+Domesticated einkorn wheat (Triticum monococcum monococcum), wild einkorn wheat (T.
 m.
@@
 m.
-boeoticum from the Karacadag [[special character]] mountains of southeast Turkey \cite{heun1997site}.
-Wild emmer wheat (Triticum turgidum subsp. dicoccoides), the progenitor of modern (tetraploid and hexaploid) wheat, is from the Fertile Crescent \cite{lev2000cradle}.
-Two distinct taxa of Triticum dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern maize was domesticated \cite{ozkan2005reconsideration}.
+boeoticum are from the Karacadag [[special character]] mountains of southeast Turkey \cite{heun1997site}.
+Wild emmer wheat (Triticum turgidum subsp. dicoccoides), the progenitor of modern (tetraploid and hexaploid) emmer wheat, is from the Fertile Crescent \cite{lev2000cradle, ozkan2002aflp}.
+Two distinct taxa of Triticum dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}.
 Localizing domestication centers for other forms of wheat at a resolution greater than the Fertile Crescent is not yet possible \cite{nesbitt1998wheat}, but then \cite{lev2000cradle} point to "a small core area within the Fertile Crescent—near the upper reaches of the Tigris and Euphrates rivers [HN6] in present-day southeastern Turkey/northern Syria".
 So, wheat has one restricted center of domestication, with wild relatives nearby.
-From \cite{ozkan2005reconsideration}, "The geographical distribution of T.
+From \cite{ozkan2005reconsideration}\, "The geographical distribution of T.
 dicoccoides reported by Zohary and Hopf (2000) includes the western Fertile Crescent, the central part of southeastern Turkey and areas in eastern Iran and Iraq.
@@
 dicoccoides also grows on the basaltic rocky slopes of the Karacadag mountains in southeastern Turkey (Johnson 1975; Harlan and Zohary 1966).
+
+The progressive domestication of wheat has involved many introgressive acts between geographically separated lineages, evinced by identification of genomes' various sources in wheats with elevated ploidy levels [[see the two papers i included for bread wheat background]]
+
 The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}.
-
-The following are just notes.
-
-From {label24}
-review/overview.  See table 1
-
-From {label25}
-Wheat was given leaf rust resistance genes from wild populations (summarized in Table 1)
-
-From {label26}
-A few examples (wild genes into wheat) given.  Originally, only with cross-compatible species, now with cross-incompatible species.
-Cites and tests several QTLs
-
-From {label31}
-"suggestions are made concerning techniques for exploitation of the wild diploid species in wheat breeding programs."
+Stem rust resistance bred into cultivated T. durum (tetraploid) from wild einkorn wheat \cite{gerechter1971transfer}\.
+Pm36, a novel podery mildew resistance gene, was also experimentally introgressed from T. turgidum var. dicoccoides into durum wheat \cite{blanco2008molecular}\.
 
@@
 
-Polyphyletic \cite{azhaguvel2007phylogenetic}.
-\cite{von1995ecogeographical} states that barley and its progenitor are the same species, as they are not sufficiently genetically distinct and geneflow continues unhindered.
-The progenitor species is Hordeum spontaneum \cite{oka2012origin}.
+Domesticated and wild barleys belong to the same species, Hordeum vulgare, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
+Barley (Hordeum vulgare subsp.
+vulgare) is believed to have been domesticated from wild subsp.
+spontaneum in the Fertile Crescent and by subsp.
+spontaneum var.
+agriocrithon in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
 
-Notes:
+Presently, the distribution of spontaneum consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living}\.
+These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}\.
 
-From {label24}
-review/overview.  See table 1
+The barley domestication process has reduced the number of alleles in the domesticate to only 40% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
 
-From {label30}
-"wild barley does harbour valuable alleles, which can enrich the genetic basis of cultivated barley and improve quantitative agronomic traits."
+Variety agriocrithon (H. vulgare ssp. vulgare f. agriocrithon) is genetically diverse, and is found throughout much of the range of Barley.
+Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic} dispute these theories, suggesting instead that it arose from ssp.
+spontaneum, perhaps more than once (in Israel as well as in Tibet).
+Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
+See art%0223-4 for list of papers that discuss the beneficial effect of exotic genes on biotic, abiotic, and quality traits.
+
+Oweing to the lack of evidence of adaptive introgression between wild and domesticated barley is the low rate of outcrossing of spontaneum, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
+
+Barley/H. spontaneum hybrids are fertile, but hybrids of other wild relatives generally are not.
+There has been little genetic investigation into spontaneous barley/spontaneum hybrids \cite{ellstrand2003dangerous}
+
+Look, there's no barrier to hybridization, there was a broad suite of wild barleys already in place, they're all locally adapted, this is a prime example of POSSIBLE introgression, but where's the evidence?
+
+Introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}\.
 
@@
 
-From {label33}
-yeah, introgression with wild relatives may be common, unsure about direction
-intermediates found in the field
+Helianthus annuus shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
+Helianthus is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}\.
+[[This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.]] This sentence may not help much, actually.
+We're betting on genetic adaptability via introg hyb
+
+H.
+annus adapted to the environment in Texas by hybridizing with H.
+debilis ssp.
+cucumerifolius, gaining advantageous alleles (these hybrids are now called H.
+annus ssp.
+texanus, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
+Helianthus has several genes for downy mildew resistance.
+Each imparts resistance to one or more races of P. halstedii, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+Some of these downy mildew resistence genes were found in wild relatives (including H.
+argophyllus, H.
+tuberosus, and H.
+praecox) and have been successfully bred into modern H.
+annus \cite{miller1991inheritance}\.
+PlArg, an allele found in wild silverleaf sunflowers (Helianthus argophyllus, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
+H. annus lives in sympatry with wild relatives like H. petiolaris and H. bolanderi and forms  stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
+H.
+annus shows signs of persistent introgressive hybridization with H.
+petiolaris with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
+\cite{kane2008genetics} finds that most gene flow in weedy Helianthus is within geographic regions, rather than between separate popultions.
+Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
+Weediness is a trait that has likely evolved many times \cite{kane2008genetics}\.
+
+Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and Phomopsis resistance breeding efforts \cite{besnard1997specifying}\.
 
@@
 
-The center of rice domestication is not known with confidence, with evidence pointing towards both China and India \cite{oka2012origin} [[see the full review on page 16]].
-There are main cultigens of rice; common rice, Oryza sativa, and African rice, O.
+There are two main cultigens of rice; Asian rice, Oryza sativa, and African rice, O.
 glaberrima, which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and because O.
 glaberrima is annual \cite{oka2012origin}.
-Oryza sativa was domesticated from wild O. rufipogon "or the Asian form of O. perennis complex" \cite{oka2012origin}, whereas O. glaberrima was domesticated from O. breviligulata \cite{oka2012origin}.
+
+The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for O.
+sativa, 8,200-13,500 years ago, from wild O.
+rufipogon  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ and the Upper Niger River delta in Mali, Africa for O.
+glaberrima, 2-3,000 years ago, from ancestor O.
+barthii \cite{ricepedia}\.
+
+The two main subspecies of Asian rice, indica and japonica, were domesticated from wild O.
+rufipogon "or the Asian form of O.
+perennis complex" \cite{oka2012origin}\ or O.
+nivara, which is closely related to O.
+rufipogon \cite{sharma2000origin, oka1988origin}\, whereas O.
+glaberrima was domesticated from O.
+breviligulata \cite{oka2012origin}\.
+
 Both O. sativa/O. rufipogon and O. glaberrima/O.breviligulata produce weedy hybrids.
@@
 sativa into new environments and sympatry with new relatives \cite{second1982origin}.
+
+O.
+sativa has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
+These subspecies' display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}\.
+
+There are about 20 other wild species within the genus Oryza \cite{ricepedia}\.
+
+There exists evidence of gene flow between rice subpopulations and into wild rice \cite{zhao2010genomic}\.
+
 Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into O.
@@
 
+\subsection*{rye}
+
+Rye (Secale cereale) was most likely domesticated from S. vavilovii \cite{stutz1972origin}\.
+The species Secale cereale contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
+Rye domestication has recieved comparatively little academic interest \cite
+N.
+I.
+Vavilov (the namesake of the species name of this plant) hypothesized that rye originated as a secondary crop during the domestications of wheat, barley, and/or peas \cite{vavilov1928geographical}\.
+
 \subsection*{potato}
 
-Modern Solanum tuberosum was domesticated in southern Peru from northern members of the polyphyletic S. brevicaule complex \cite{spooner2005single}.
-These northern members of the S. breviaule complex are not clearly defined, and may in fact be one singular species (in which case the species name would be S. bukasovii) \cite{spooner2005single}.
+Modern potato Solanum tuberosum is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
+The northern members of the polyphyletic S.
+brevicaule wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because s.
+tuberosum has a polyphyletic origin.
+Then again, these northern members of the S.
+breviaule complex are not clearly defined, and may in fact be one singular species (in which case the species name would be S.
+bukasovii) \cite{spooner2005single}.
+
+Although potatoes are usually propogated clonally, farmers also promote sexyal hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
+Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
+
+The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
+Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
+
+Contrary to expectations \cite{hamrick1983distribution}\, landrace tetraploid potatoes show more genetic diversity than diploids \cite{zimmerer1991geographical}\.
+
+Widespread introgression and hybridization in potatoes \cite{grun1990evolution}\.
+
+, potato landraces show a weaker correlation, due perhaps to confounding influences of farmer interactions (including environmental manipulation and artificial selection by farmers \cite{zimmerer1991geographical, [[cite other sources from page 45 in this paper}
+
+These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
+However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
+
+Natural selection may be subverted by selection and preference of farmers \cite{brush1981dynamics}\.
+
+Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the Central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
+
+Cultivated diploid potatoes have been shown introgressed with genes from S. sparsipilum \cite{rabinowitz1990high}\ and S. megistacrolobum \cite{johns1987relationships, huaman1980solanum}.
+
 Resistance genes have been experimentally introgressed from wild poptato relatives (S.
@@
 Charles Rick (University of California, August 1978) stating that tomato was domesticated from Lycopersicon esculentum v.
-cerasiformae in the New World (most likely Mexico).
-Most of the genetic diversity resides within the wild relatives of the tomato \cite{rick1988tomato} \cite{miller1990rflp}, so breeders have long turned to these plants for their germplasm \cite{rick1982potential}.
+cerasiformae [[solanum lycopersicum cerasiforme]] in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
+\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agrees.
+\cite{nesbitt2002comparative}, however, conclude that v.
+cerasiformae is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes L.
+esculentum and L.
+pimpinellifolium.
+Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
+Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato} \cite{miller1990rflp, rick1982potential}.
+Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
 \cite{hanson2000mapping} mapped the tomato yellow leaf curl virus resistance gene in L.
 hirsutum f.
-glabratum accession 'B 6013' (line H24) (from \cite{banerjee1990transfer}) to an introgression on chromosome 11.
+glabratum accession 'B 6013' (line H24) (from \cite{banerjee1990transfer}) to an introgression on chromosome 11, which originated from the wild relative L.
+hirsutum.
+Similar resistance traits has also been shown to have been imparted from L. chilense, L. pimpinellifolium, and L. cheesmanii [[various papers cited page 19 \cite{hanson2000mapping}\).
 
 \subsection*{common bean}
+
+Common bean was domesticated from a wild common bean at least twice (once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes) \cite{kwak2009putative}\.
+The domesticated common bean progenitor species is P. vulgaris var. aborigineus, which is found across the breadth of this region \cite{debouck1988implications]\.
+Wild common bean now occurs across the Americas, from Mexico to Argentina, inhabiting a broad range of environments.
+Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}/), rates of gene flow between it and wild relatives are expected to be lower.
+\cite{papa2003asymmetry}\ shows little gene flow from wild common bean to domesticated.
+Though some hybrids between wild and domesticated common bean show heterosis, others (especially F1 hybrids) do not, producing smaller plants and smaller seeds \cite{gutierrez1985heterosis}\.
+It may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
+In the meantime, farmers may generally be selecting against hybrids.
+Most common bean cultivation occurs in South America on small- and mediumholder farms \cite{paredes1995extensive}.
+Because F1 hybrids are intermediate in seed size and color, farmers may effectively identify and avoid hybrid stock.
+\cite{papa2003asymmetry}\ found that the spread of wild common bean populations may be facilitated by adaptive genes from domesticated common bean landraces through stable weedy hybrid populations, but little field evidence exists to support adaptive introgression in the direction of wild bean to domesticated.
+\cite{papa2003asymmetry}\ was unable to determine how much of the genetic variability of domesticated common bean landraces originated from weedy or wild types, but point to the allele for 'L' phaseolin as a possible example of an allele that has been introgressed from wild common beans into domesticated.
+For example, genetic material from cultivated common bean has been introgressed into wild populations, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
+This may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}\).
+These weedy intermediates are occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
+Common leaf blight and white mold resistances have experimentally bred into P.
+vulgaris L.
+from scarlet runner bean P.
+coccineus \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils Zabrotes subfasciatus and Acanthoscelides obtectus from Mexican wild beans \cite{kornegay1991inheritance}\.
+Between the Mesoamerican and Andean domesticated gene pools, there is little opportunity for admixture due to geographical isolation, though their respective ranges do meet in some places, such as Columbia \cite{gepts1986phaseolin}\.
+\cite{paredes1995extensive}\ suggests that common bean germplasm in Chile may be the result of introgression between the Mesoamerican and Andean gene pools, and breathes hope into the prospect of adaptive introgression in common bean breeding.
+(Actually, the above suggestion was confirmed by \cite{kwak2009structure]\, citing that as many as 70% of Chilean landraces may be hybrid in origin [[(and by hybrid, I'm pretty sure they mean hybrids between the two domesticated forms of common bean, so i'm not actually sure it "breathes hope" into natural adaptive introgressive hybridization.)]]
+\cite{kwak2009structure}\ also found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
+Furthermore, \cite{papa2003asymmetry}\ asks the reader to consider the view that the domestication of common bean was not a single event in the past, but is rather an ongoing "dynamic process resulting from selection, hybridization, and reselection over many years".
+phaseolin?
+
+\subsection*{Sorghum}
+Anthropological evidence points to a domestication event of sorghum ((Sorghum bicolor subsp. bicolor) some 5,000 or 6,000 years ago Ethiopian Sudan \cite{smith2000sorghum}\.
+However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern savanna.
+The progenitor was S. arundinaceum \cite{doggett1988sorghum, harlan1971toward}\.
+The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
+
+Members of Sorghum bicolor are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}\.
+
+fertility rates can be low due to incompatability and ploidy
+
+Rates of outcrossing within sorghum range from 0-30% \cite{doggett1988sorghum}\.
+
+Within Eu-Sorghum, there are currently five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum)
+
+There are currently five races of S. bicolor subsp. bicolor currently cultivated: bicolor, guinea, kafir, caudatum, and durra \cite{smith2000sorghum}\.
+Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\.
+Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
+These hybrids express diminished fertility
+
+Members of Sorghum bicolor have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum}\.
+
+Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera, \cite{de1976cytogenetics}\) for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}\.
+Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}\.
+\cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution
 
@@
 However, such exceptions as barley, flax, foxtail millet, and oat have wild progenitors which inhabit a wide range of habitats beyond the Fertile Crescent.
-
+\cite{papa2003asymmetry}\ In order to persist, cultivated-weedy-wild breeding complexes require a conducive environment in which all three biological components can coexist in close proximity.
+Relevant environmental factors include biotic, climatic, and anthropogenic pressures.
 \cite{jarvis1999wild} Table 2 is an incredible resource, cited examples of farmer selection and/or use of introgressed types, and confirmation of introgression
-
 https://books.google.com/books?hl=en&lr=&id=CvcbUopfa54C&oi=fnd&pg=PA7&dq=common+bean+domestication&ots=ggAF-EoWG8&sig=dI9wOyATpzAelxLf_cg-gs0J2gI#v=onepage&q=common%20bean%20domestication&f=false
@@
 Also, non-crop plants, like trees (oak, larch).
-
 From {label15}
@@
 Mouse, Salamander
-
 From {label21}
@@
 Haemophilus influenza, Trypanosoma cruzi
-
 From {label23}
 165 proposed cases of introgression, 65 "deemed to be sufficiently documented" (plants)
-
 From {label28}
 Milkweeds
-
 From {label29}
@@
 
-\begin{thebibliography}{1}
-\bibitem{label1} H. G. Wilkes (1977) 
-Hybridization of Maize and Teosinte, in Mexico and Guatemala and the Improvement of Maize. 
-{\em Economic Botany\} 31:254-293
-
-\bibitem{label2} C. Lumholtz (1902)
-Unknown Mexico. Charles Scribner's Sons, New York. Vol I. 530 pp.
-%This citation was problematic
-
-\bibitem{label3} J. Mallet (2005)
-Hybridization as an invasion of the genome.
-{\em TRENDS in Ecology and Evolutin\} 20:229-237
-
-\bibitem{label4} J. Ross-Ibarra, M. Tenaillon, B. S. Gaut (2009)
-Historical Divergence and Gene Flow in the Genus Zea.
-{\em Genetics\} 181:1399-1413
-
-\bibitem{label5} N. C. Ellstrand, H. C. Prentice, J. F. Hancock (1990)
-Gene Flow and Introgression from Domesticated Plants into their Wild Relatives.
-{\em Annual Review of Ecology and Systematics\} 30:539-563
-
-\bibitem{label6} D. Quist, I. H. Chapela (2001)
-Transgenic DNA introgressed into traditional maize landraces in Oaxaca, Mexico.
-{\em Nature\} 414:541-543
-%RETRACTED?
-
-\bibitem{label7} J. Doebley (1990)
-Molecular Evidence for Gene Flow among {\em Zea\} Species.
-{\em BioScience\} 40:443-448
-
-\bibitem{label8} J. F. Doebley (1984)
-Maize Introgression into Teosinte - A Reappraisal.
-{\em Annals of the Missouri Botanical Garden\} 71:1100-1113
-
-\bibitem{label9} J. Doebley (1990)
-Molecular Evidence and the Evolution of Maize.
-{\em Economic Botany\} 44:6-27
-%a review paper
-
-\bibitem{label10} J. Doebley, M. M. Goodman, C. W. Stuber (1987)
-Patterns of isozyme variation between maize and Mexican annual teosinte.
-{\em Economic Botany\} 41:234-246
-
-\bibitem{label11} B. M. Baltazar, J. de J. Sanchez-Gonzalez (2005)
-Pollination between maize and teosinte: an important determinant of gene flow in Mexico
-{\em Theoretical Applied Genetics\} 110:519-526
-%special character for the Sanchez name, plus how to abbreviate "de"
-
-\bibitem{label12} M. L. Warburton, G. Wilkes, S. Taba, A. Charcosset, C. Mir, F. Dumas, D. Madur, S. Dreisigacker, C. Bedoya, B. M. Prasanna, C. X. Xie, S. Hearne, J. Franco (2011)
-Gene flow among different teosinte taxa and into the domesticated maize gene pool.
-{\em Genetic Resources and Crop Evolution\} 58:1243-1261
-
-\bibitem{label13} C. Jiang, G. O. Edmeades, I. Armstead, H. R. Lafitte, M. D. Hayward, D. Hoisington (1999)
-Genetic analysis of adaptation of differences between highland and lowland tropical maize using molecular markers.
-{\em Theoretical Applied Genetics\} 99:1106-1119
-
-\bibitem{label14} E. Bitocchi, L. Nanni, M. Rossi, D. Rau, E. Bellucci, A. Giardini, A. Buonamici, G. G. Vendramin, R. Papa (2009)
-Introgression from modern hybrid varieties into landrace populations of maize ({\em Zea mays\} ssp. {\em mays\} L.) in central Italy.
-{\em Molecular Ecology\} 18:603-621
-
-\bibitem{label15} M. L. Arnold, N. H. Martin (2009)
-Adaptation by introgression.
-{\em Journal of Biology\} 8:82
-%mini review of introgression
-
-\bibitem{label16} N. C. Ellstrand, C. Garner, S. Hegde, R. Guadagnuolo, L. Blancas (2007)
-Spontaneous hybridization between Maize and Teosinte.
-{\em Journal of Heredity\} 98:183-187
-
-\bibitem{label17} M. B. Hufford, P. Lubinsky, T. Pyhajarvi, M. T. Devengenzo, N. C. Ellstrand, J. Ross-Ibarra (2013)
-The genomic signature of crop-wild introgression in maize.
-{\em PLOS Genetics\} 9:
-%special characters, page numbers?
-
-\bibitem{label18} Y. Matsuoka, Y. Vigouroux, M. M. Goodman, J. Sanchez G., E. Buckler, J. Doebley (2002)
-A single domestication for maize shown by multilocus microsatellite genotyping.
-{\em PNAS\} 99:6080-6084
-
-\bibitem{label19} J. van Heerwaarden, J. Doebley, W. H. Briggs, J. C. Glaubitz, M. M. Goodman, J. de Jesus Sanchez Gonzalez, J. Ross-Ibarra (2010)
-Genetic signals of origin, spread, and introgression in a large sample of maize landraces.
-{\em PNAS\} 108:1088-1092
-%Gonzalez, "van"
-
-\bibitem{label20} H. G. Wilkes (1970)
-Teosinte introgression in the maize of the Nobogame Valley
-{\em Botanical Museum Leaflets\} 22:297-311
-
-\bibitem{label21} M. L. Arnold (2004)
-Transfer and origin of adaptations through natural hybridization: Were Anderson and Stebbins Right?
-{\em The Plant Cell\} 16:562-570
-%introgression review
-
-\bibitem{label22} C. N. Stewart Jr., M. D. Halfhill, S. Warwick (2003)
-Transgene introgression from genetically modified crops to their wild relatives.
-{\em Nature\} 4:806-817
-%is nature reviews the same thing as nature?
-%good review of introgression with crops
-%Table 1 shows crops and their progenitor species and main wild relatives
-
-\bibitem{label23} L. H. Rieseberg, J. F. Wendel (1993)
-Hybrid zones and the evolutionary process.
-{\em Journal of Evolutionary Biology\} 7:631-634
-
-\bibitem{label24} E. Nevo, G. Chen (2010)
-Drought and salt tolerances in wild relatives for wheat and barley improvement.
-{\em Plant, Cell Environment\} 33:670-685
-
-\bibitem{label25} E. Autrique, R. P. Singh, S. D. Tanksley, M. E. Sorrells (1995)
-Molecular markers for four lef rust resistance genes introgressed into wheat from wild relatives
-{\em Genome\} 38:75-83
-
-\bibitem{label26} S. G. Hegde, J. G. Waines (2004)
-Hybridization and introgression between bread wheat and wild and weedy relatives in North America.
-{\em Crop Science\} 44:1145-1155
-
-\bibitem{label27} M. A. Chapman, J. M. Burke (2006)
-Letting the gene out of the bottle: The population genetics of genetically modified crops.
-{\em New Phytologist\} 170:429-443
-
-\bibitem{label28} S. B. Broyles (2002)
-Hybrid bridges to gene flow: A case study in milkweeds ({\em Asclepias\}).
-{\em Evolution\} 56:1943-1953
-
-\bibitem{label29} S. W. Hussain, W. M. Williams (1997)
-Development of a fertile genetic bridge between {\em Trifolum ambiguum\} M. Bieb. and {\em T. repens\} L.
-{\em Theoretical Applied Genetics\} 95:678-690
-
-\bibitem{label30} M. von Korff, H. Wang, J. Leon, K. Pillen (2006)
-AB-QTL analysis in spring barley: II. Detection of favourable exotic alleles for agronomic traits introgressed from wild barley ({\em H. vulgare\} ssp. {\em spontaneum\}).
-{\em Theoretical Applied Genetics\} 112:1221-1231
-%special characters
-
-\bibitem{label31} D. Zohary, J. R. Harlan, A. Vardi (1969)
-The wild diploid progenitors of wheat and their breeding value.
-{\em Euphytica\} 18:58-65
-
-\bibitem{label32} [[A huge fricken list of names]] (2013)
-Hybridization and speciation.
-{\em Journal of Evolutionary Biology\} 26:229-246
-
-\bibitem{label32} K. Fukunaga, J. Hill, Y. Virouroux, Y. Matsuoka, J. Sanchez G., K. Liu, E. S. Buckler, J. Doebley (2005)
-Genetic diversity and Population Structure of Teosinte.
-{\em Genetics\} 169:2241-2254
-
-\bibitem{label33} M. Reagon, A. A. Snow (2006)
-Cultivated {\em Helianthus annus\} (Asteraceae) volunteers as a genetic "bridge" to weedy sunflower populations in North America.
-{\em Ammerican Journal of Botany\} 93:127-133
-
-\bibitem{label34} J. A. Able, P. Langridge, A. S. Milligan (2006)
-Capturing diversity in the cereals: many options but little promiscuity.
-{\em TRENDS in Plant Science\} 12:1360-1385
-
-\end{thebibliography}{1}
-
 \end{document}

### 2015-12-09 crop_adaptation_gene_flow_gj_outline.tex
updating crop_adaptation_gene_flow_gj_outline.tex, not quite done yet

@@
 \cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
-
 \cite{hammer1987collection}\ cites an example in Italy of one domesticated maize adapting to coastal environment by introgressions from another domesticated maize.
@@
 boeoticum) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
-Subsequently, emmer wheat (Triticum dicoccum) was domesticated from hybrids of diverse populations of wild emmer wheat (T.
-dioccoides) in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure}.
-Emmer wheat is tetraploid, and hybridization with wild diploid Tausch's goatgrass (Aegilops tauschii) lead to the production of hexaploid wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
-Emmer wheat has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
+Subsequently, emmer wheat (Triticum dicoccum) was domesticated from hybrids of diverse populations of wild emmer wheat (Triticum dioccoides), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}.
+Two distinct taxa of T.
+dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
+The tetraploid durum wheat (Triticum turgidum) is the result of hybridization between red wild einkorn wheat (T.
+urartu) and wild goatgrass (Aegilops speltoides) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (Aegilops tauschii) produced the allohexaploid common bread wheat (T.
+aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
+This hexaploidization event probably occured twice, once in southeast Turkey/northern Syria and, more recently, in Iran \cite{giles2006gludy}\.
+Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}\.
 
-Durum wheat (T. turgidum ssp. durum) was developed from emmer wheat as well.
+The progressive domestication of wheat has involved many introgressive acts between geographically dispersed lineages, evinced by identification of genomes' various sources in wheats with elevated ploidy levels.
 
-Common bread wheat, T.
-aestivum, is an allohexaploid with chromosome sets from red wild einkorn wheat (T.
-urartu), Aegilops speltoides, and Aegilops tauschii \cite{mcfadden1946origin, petersen2006phylogenetic}\, and furthermore, each of these species is the result of ancient hybridizations \cite{marcussen2014ancient}\.
-Hexaploidization probably occured twice, in southeast Turkey/northern Syria and, more recently, Iran \cite{giles2006gludy}\.
-
-Domesticated einkorn wheat (Triticum monococcum monococcum), wild einkorn wheat (T.
-m.
-boeoticum), and the progenitor T.
-m.
-boeoticum are from the Karacadag [[special character]] mountains of southeast Turkey \cite{heun1997site}.
-Wild emmer wheat (Triticum turgidum subsp. dicoccoides), the progenitor of modern (tetraploid and hexaploid) emmer wheat, is from the Fertile Crescent \cite{lev2000cradle, ozkan2002aflp}.
-Two distinct taxa of Triticum dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}.
-Localizing domestication centers for other forms of wheat at a resolution greater than the Fertile Crescent is not yet possible \cite{nesbitt1998wheat}, but then \cite{lev2000cradle} point to "a small core area within the Fertile Crescent—near the upper reaches of the Tigris and Euphrates rivers [HN6] in present-day southeastern Turkey/northern Syria".
-So, wheat has one restricted center of domestication, with wild relatives nearby.
-From \cite{ozkan2005reconsideration}\, "The geographical distribution of T.
-dicoccoides reported by Zohary and Hopf (2000) includes the western Fertile Crescent, the central part of southeastern Turkey and areas in eastern Iran and Iraq.
-Johnson (1975) reports that from southeastern Turkey to Iran and Iraq the species is progressively substituted by the wild tetraploid wheat T.
-araraticum.
-T.
-dicoccoides also grows on the basaltic rocky slopes of the Karacadag mountains in southeastern Turkey (Johnson 1975; Harlan and Zohary 1966).
-
-The progressive domestication of wheat has involved many introgressive acts between geographically separated lineages, evinced by identification of genomes' various sources in wheats with elevated ploidy levels [[see the two papers i included for bread wheat background]]
-
+The convoluted domestication of wheat is beset by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
+Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
+Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
 The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}.
-Stem rust resistance bred into cultivated T. durum (tetraploid) from wild einkorn wheat \cite{gerechter1971transfer}\.
-Pm36, a novel podery mildew resistance gene, was also experimentally introgressed from T. turgidum var. dicoccoides into durum wheat \cite{blanco2008molecular}\.
+Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
+Modern breeding efforts between wild and domesticated wheat have shown some success.
+For example, some researchers have experimentally bred resistance traits into durum wheat (stem rust resistance from wild einkorn, \cite{gerechter1971transfer}\, and powdery mildew resistance from T.
+turgidum var.
+dicoccoides, \cite{blanco2008molecular}\.
 
@@
 Barley (Hordeum vulgare subsp.
-vulgare) is believed to have been domesticated from wild subsp.
-spontaneum in the Fertile Crescent and by subsp.
+vulgare) is believed to have been domesticated at least twice, once from wild subsp.
+spontaneum in the Fertile Crescent and once by subsp.
 spontaneum var.
@@
 
-Presently, the distribution of spontaneum consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living}\.
+Variety agriocrithon (H. vulgare ssp. vulgare f. agriocrithon) is genetically diverse, and is found throughout much of the range of Barley.
+Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp.
+spontaneum, perhaps more than once (in Israel as well as in Tibet).
+Presently, the distribution of spontaneum consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
 These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}\.
@@
 The barley domestication process has reduced the number of alleles in the domesticate to only 40% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
-
-Variety agriocrithon (H. vulgare ssp. vulgare f. agriocrithon) is genetically diverse, and is found throughout much of the range of Barley.
-Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic} dispute these theories, suggesting instead that it arose from ssp.
-spontaneum, perhaps more than once (in Israel as well as in Tibet).
 Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
-See art%0223-4 for list of papers that discuss the beneficial effect of exotic genes on biotic, abiotic, and quality traits.
-
-Oweing to the lack of evidence of adaptive introgression between wild and domesticated barley is the low rate of outcrossing of spontaneum, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
-
-Barley/H. spontaneum hybrids are fertile, but hybrids of other wild relatives generally are not.
-There has been little genetic investigation into spontaneous barley/spontaneum hybrids \cite{ellstrand2003dangerous}
-
-Look, there's no barrier to hybridization, there was a broad suite of wild barleys already in place, they're all locally adapted, this is a prime example of POSSIBLE introgression, but where's the evidence?
-
-Introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}\.
+Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
+Owing to this sparcity of evidence is the low rate of outcrossing of spontaneum, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
+There has been little genetic investigation into spontaneous barley/spontaneum hybrids \cite{ellstrand2003dangerous}.
+Barley/H.
+spontaneum hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}\.
 
@@
 
-Helianthus annuus shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
+The common sunflower (Helianthus annuus) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
 Helianthus is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}\.
@@
 We're betting on genetic adaptability via introg hyb
-
 H.
-annus adapted to the environment in Texas by hybridizing with H.
+annuus adapted to the environment in Texas by hybridizing with H.
 debilis ssp.
 cucumerifolius, gaining advantageous alleles (these hybrids are now called H.
-annus ssp.
+annuus ssp.
 texanus, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
+H. annuus lives in sympatry with wild relatives like H. petiolaris and H. bolanderi and forms  stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
+H.
+annuus shows signs of persistent introgressive hybridization with H.
+petiolaris with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
+Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
+Weediness is a trait that has likely evolved many times \cite{kane2008genetics}\.
+
 Helianthus has several genes for downy mildew resistance.
@@
 praecox) and have been successfully bred into modern H.
-annus \cite{miller1991inheritance}\.
+annuus \cite{miller1991inheritance}\.
 PlArg, an allele found in wild silverleaf sunflowers (Helianthus argophyllus, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
-H. annus lives in sympatry with wild relatives like H. petiolaris and H. bolanderi and forms  stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
-H.
-annus shows signs of persistent introgressive hybridization with H.
-petiolaris with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
-\cite{kane2008genetics} finds that most gene flow in weedy Helianthus is within geographic regions, rather than between separate popultions.
-Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
-Weediness is a trait that has likely evolved many times \cite{kane2008genetics}\.
-
 Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and Phomopsis resistance breeding efforts \cite{besnard1997specifying}\.
@@
 
-There are two main cultigens of rice; Asian rice, Oryza sativa, and African rice, O.
-glaberrima, which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and because O.
-glaberrima is annual \cite{oka2012origin}.
-
+There are two main cultigens of rice; Asian rice (Oryza sativa) and African rice (O.
+glaberrima), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
 The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for O.
 sativa, 8,200-13,500 years ago, from wild O.
-rufipogon  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ and the Upper Niger River delta in Mali, Africa for O.
-glaberrima, 2-3,000 years ago, from ancestor O.
-barthii \cite{ricepedia}\.
+rufipogon  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of O.
+perennis complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for O.
+glaberrima, 2-3,000 years ago, from a wild ancestor, perhaps O.
+barthii \cite{ricepedia}\ or O.
+breviligulata \cite{oka2012origin}\.
+Asian rice has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
+Indica and Japonica are likely the result of independent domestications from separate O. rufipogon populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
+These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}\.
 
-The two main subspecies of Asian rice, indica and japonica, were domesticated from wild O.
-rufipogon "or the Asian form of O.
-perennis complex" \cite{oka2012origin}\ or O.
-nivara, which is closely related to O.
-rufipogon \cite{sharma2000origin, oka1988origin}\, whereas O.
-glaberrima was domesticated from O.
-breviligulata \cite{oka2012origin}\.
-
-Both O. sativa/O. rufipogon and O. glaberrima/O.breviligulata produce weedy hybrids.
+Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty, \cite{ricepedia}\), and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
 The greater genetic diversity within O.
@@
 sativa into new environments and sympatry with new relatives \cite{second1982origin}.
-
-O.
-sativa has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
-These subspecies' display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}\.
-
-There are about 20 other wild species within the genus Oryza \cite{ricepedia}\.
-
-There exists evidence of gene flow between rice subpopulations and into wild rice \cite{zhao2010genomic}\.
-
 Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into O.
 sativa L.
-by researchers \cite{brar1997alien, khush1974inheritance}, but adaptive introgression has also happened without human intervention \cite{second1982origin}.
+by researchers \cite{brar1997alien, khush1974inheritance}.
+Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
+Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
+Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
 
@@
 
-Rye (Secale cereale) was most likely domesticated from S. vavilovii \cite{stutz1972origin}\.
-The species Secale cereale contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
-Rye domestication has recieved comparatively little academic interest \cite
-N.
-I.
-Vavilov (the namesake of the species name of this plant) hypothesized that rye originated as a secondary crop during the domestications of wheat, barley, and/or peas \cite{vavilov1928geographical}\.
+Rye domestication has recieved comparatively little academic interest.
+Although there is little evidence for or against the role of natural introgression in rye domestication, rye (Secale cereale) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
 
@@
 
-Modern potato Solanum tuberosum is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
+Modern potato (Solanum tuberosum) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
 The northern members of the polyphyletic S.
@@
 tuberosum has a polyphyletic origin.
-Then again, these northern members of the S.
+Also, these northern members of the S.
 breviaule complex are not clearly defined, and may in fact be one singular species (in which case the species name would be S.
-bukasovii) \cite{spooner2005single}.
+bukasovii, \cite{spooner2005single}\).
 

### 2015-12-11 crop_adaptation_gene_flow_gj_outline.tex
updating crop_adaptation_gene_flow_gj_outline.tex, all crops pretty well complete except Maize, to be reconciled with Li's work

@@
 \section*{Outline}
-\subsection*{Outline parts}
 
 \subsection*{Introduction}
+
  "Whenever sympatric populations of crops and their wild relatives show a similar trait, there will be several possible explanations: (1) retention of an ancestral trait, (2) convergent or parallel evolution, and (3) introgression.
@@
 
-1\subsection*{Introgression in Crops (General)}
+\subsection*{Maize}
 
-From {stewart2003transgene}
-A very good review.
-Hybrid zones make for a great model system of study.
-Introgression from wild populations into agronomic populations is potentially beneficial for breeding programmes.
-It is generally easier to introgress into crops than introgressing into wild populations (stronger selection, overcoming LD).
-
-\cite{harlan1966distribution}\ The premise; domestication takes place at the edges of a progenitor's range, where it barely grows.
-there, farmers work on the crop to make it more adapted.
-if this is the case, this could maintain progenitor populations long after domestication, and may increase gene flow between the original progenitor population and the expanding domesticate.
-This is probably unlikely to be advantageous for the domesticate, however.
-
-From {label32}
-Another very good review, with more of a molecular emphasis
-
-From {label34}
-The difficulties of utilizing wild relatives in breeding programmes effectively.
-
-From {label6} {label7} {label26} {Label27}
-Lots of papers focus on gene flow from crop to wild relative, particularly about transgene escape
-
-\subsection*{Maize}
+Supplement Li's work with the following ideas:
 
@@
 
-Although potatoes are usually propogated clonally, farmers also promote sexyal hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
+Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
 Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
+These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
+However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
+Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}\.
 
+Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}\.
 The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
-
-Contrary to expectations \cite{hamrick1983distribution}\, landrace tetraploid potatoes show more genetic diversity than diploids \cite{zimmerer1991geographical}\.
-
-Widespread introgression and hybridization in potatoes \cite{grun1990evolution}\.
-
-, potato landraces show a weaker correlation, due perhaps to confounding influences of farmer interactions (including environmental manipulation and artificial selection by farmers \cite{zimmerer1991geographical, [[cite other sources from page 45 in this paper}
-
-These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
-However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
-
-Natural selection may be subverted by selection and preference of farmers \cite{brush1981dynamics}\.
-
+Cultivated diploid potatoes have been shown introgressed with genes from S. sparsipilum \cite{rabinowitz1990high}\ and S. megistacrolobum \cite{johns1987relationships, huaman1980solanum}.
 Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the Central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
-
-Cultivated diploid potatoes have been shown introgressed with genes from S. sparsipilum \cite{rabinowitz1990high}\ and S. megistacrolobum \cite{johns1987relationships, huaman1980solanum}.
-
 Resistance genes have been experimentally introgressed from wild poptato relatives (S.
@@
 Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
-Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato} \cite{miller1990rflp, rick1982potential}.
+
+Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{hanson2000mapping} mapped the tomato yellow leaf curl virus resistance gene in L.
+\cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of L.
 hirsutum f.
-glabratum accession 'B 6013' (line H24) (from \cite{banerjee1990transfer}) to an introgression on chromosome 11, which originated from the wild relative L.
+glabratum \cite{banerjee1990transfer}\ to an introgression on chromosome 11, which originated from the wild relative L.
 hirsutum.
-Similar resistance traits has also been shown to have been imparted from L. chilense, L. pimpinellifolium, and L. cheesmanii [[various papers cited page 19 \cite{hanson2000mapping}\).
+Similar resistance traits has also been shown to have been imparted from L. chilense, L. pimpinellifolium, and L. cheesmanii \cite{hanson2000mapping}\.
 
-\subsection*{common bean}
+\subsection*{Common Bean}
 
-Common bean was domesticated from a wild common bean at least twice (once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes) \cite{kwak2009putative}\.
-The domesticated common bean progenitor species is P. vulgaris var. aborigineus, which is found across the breadth of this region \cite{debouck1988implications]\.
-Wild common bean now occurs across the Americas, from Mexico to Argentina, inhabiting a broad range of environments.
-Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}/), rates of gene flow between it and wild relatives are expected to be lower.
-\cite{papa2003asymmetry}\ shows little gene flow from wild common bean to domesticated.
-Though some hybrids between wild and domesticated common bean show heterosis, others (especially F1 hybrids) do not, producing smaller plants and smaller seeds \cite{gutierrez1985heterosis}\.
-It may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
-In the meantime, farmers may generally be selecting against hybrids.
-Most common bean cultivation occurs in South America on small- and mediumholder farms \cite{paredes1995extensive}.
+Common bean (Phaseolus vulgaris) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
+The domesticated common bean progenitor species (P.
+vulgaris var.
+aborigineus) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
+Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}\), rates of gene flow between it and wild relatives are expected to be lower.
+These expectations are supported by evidence provided by \cite{papa2003asymmetry}\.
+However, hybrid populations can still be found in sympatry with wild and domesticated common bean, forming domesticate-hybrid-weedy complexes.
+
+Some evidence for introgression exists in the direction of domesitcated to wild common bean, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
+Wild common beans with this introgression are also occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
+\cite{kwak2009structure}\ also found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
+The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}\).
+Farmer practices may also be diminishing introgressive gene flow.
+Though some hybrids show heterosis, others (especially F1 hybrids) do not, expressing smaller plant size, smaller seed size, and increased mortality \cite{gutierrez1985heterosis, paredes1995extensive}\.
 Because F1 hybrids are intermediate in seed size and color, farmers may effectively identify and avoid hybrid stock.
-\cite{papa2003asymmetry}\ found that the spread of wild common bean populations may be facilitated by adaptive genes from domesticated common bean landraces through stable weedy hybrid populations, but little field evidence exists to support adaptive introgression in the direction of wild bean to domesticated.
+
+Though limited, there is also some evidence of natural introgression of wild genes into domesticated common bean.
 \cite{papa2003asymmetry}\ was unable to determine how much of the genetic variability of domesticated common bean landraces originated from weedy or wild types, but point to the allele for 'L' phaseolin as a possible example of an allele that has been introgressed from wild common beans into domesticated.
-For example, genetic material from cultivated common bean has been introgressed into wild populations, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
-This may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}\).
-These weedy intermediates are occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
-Common leaf blight and white mold resistances have experimentally bred into P.
-vulgaris L.
-from scarlet runner bean P.
-coccineus \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils Zabrotes subfasciatus and Acanthoscelides obtectus from Mexican wild beans \cite{kornegay1991inheritance}\.
-Between the Mesoamerican and Andean domesticated gene pools, there is little opportunity for admixture due to geographical isolation, though their respective ranges do meet in some places, such as Columbia \cite{gepts1986phaseolin}\.
-\cite{paredes1995extensive}\ suggests that common bean germplasm in Chile may be the result of introgression between the Mesoamerican and Andean gene pools, and breathes hope into the prospect of adaptive introgression in common bean breeding.
-(Actually, the above suggestion was confirmed by \cite{kwak2009structure]\, citing that as many as 70% of Chilean landraces may be hybrid in origin [[(and by hybrid, I'm pretty sure they mean hybrids between the two domesticated forms of common bean, so i'm not actually sure it "breathes hope" into natural adaptive introgressive hybridization.)]]
-\cite{kwak2009structure}\ also found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
-Furthermore, \cite{papa2003asymmetry}\ asks the reader to consider the view that the domestication of common bean was not a single event in the past, but is rather an ongoing "dynamic process resulting from selection, hybridization, and reselection over many years".
-phaseolin?
+Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (P.
+coccineus) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (Zabrotes subfasciatus and Acanthoscelides obtectus) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
 
 \subsection*{Sorghum}
-Anthropological evidence points to a domestication event of sorghum ((Sorghum bicolor subsp. bicolor) some 5,000 or 6,000 years ago Ethiopian Sudan \cite{smith2000sorghum}\.
-However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern savanna.
-The progenitor was S. arundinaceum \cite{doggett1988sorghum, harlan1971toward}\.
+Anthropological evidence points to a domestication event of sorghum (Sorghum bicolor subsp. bicolor) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
+However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
+The progenitor species was S. arundinaceum \cite{doggett1988sorghum, harlan1971toward}\.
 The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
-
+Within Eu-Sorghum, there are currently five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
 Members of Sorghum bicolor are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}\.
 
-fertility rates can be low due to incompatability and ploidy
-
+\cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution.
+Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\.
 Rates of outcrossing within sorghum range from 0-30% \cite{doggett1988sorghum}\.
-
-Within Eu-Sorghum, there are currently five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum)
-
-There are currently five races of S. bicolor subsp. bicolor currently cultivated: bicolor, guinea, kafir, caudatum, and durra \cite{smith2000sorghum}\.
-Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\.
 Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
 These hybrids express diminished fertility
-
-Members of Sorghum bicolor have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum}\.
-
-Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera, \cite{de1976cytogenetics}\) for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}\.
+Members of Sorghum bicolor have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum, CWR}\.
+Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera \cite{de1976cytogenetics}\) for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}\.
 Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}\.
-\cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution
-
-\subsection*{Other Notes}
-
-\cite{zohary2012domestication} has Map 1, showing the domestication center of einkorn wheat, emmer wheat, barley, chickpea, flax, lentil, pea, and bitter vetch in the Fertile Crescent.
-Most of the wild progenitors of founder crops (excluding barley and flax) have a distribution restricted to the Fertile Crescent \cite{zohary2012domestication}.
-However, such exceptions as barley, flax, foxtail millet, and oat have wild progenitors which inhabit a wide range of habitats beyond the Fertile Crescent.
-\cite{papa2003asymmetry}\ In order to persist, cultivated-weedy-wild breeding complexes require a conducive environment in which all three biological components can coexist in close proximity.
-Relevant environmental factors include biotic, climatic, and anthropogenic pressures.
-\cite{jarvis1999wild} Table 2 is an incredible resource, cited examples of farmer selection and/or use of introgressed types, and confirmation of introgression
-https://books.google.com/books?hl=en&lr=&id=CvcbUopfa54C&oi=fnd&pg=PA7&dq=common+bean+domestication&ots=ggAF-EoWG8&sig=dI9wOyATpzAelxLf_cg-gs0J2gI#v=onepage&q=common%20bean%20domestication&f=false
-
-\subsection*{non-crop systems}
-From {label3}
-Bacteria, malaria mosquitoes, blackflies, Darwin's finches, butterflies (each example has a cited source, not included in the bibliography of this outline).
-Also, non-crop plants, like trees (oak, larch).
-From {label15}
-Helianthus, Iris
-Mouse, Salamander
-From {label21}
-Iris, Helianthus, Cowania/Purshia
-Dacus (now Bactrocera), Anopheles
-Haemophilus influenza, Trypanosoma cruzi
-From {label23}
-165 proposed cases of introgression, 65 "deemed to be sufficiently documented" (plants)
-From {label28}
-Milkweeds
-From {label29}
-White clover
 

### 2015-12-18 crop_adaptation_gene_flow_gj_outline.tex
added figure.tex, a summary table

@@
 \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
-\emp{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
+\emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
 
@@
 
-The first domesticated wheat was einkorn wheat (Triticum monococcum), domesticated from wild einkorn (T.
+The first domesticated wheat was einkorn wheat (\emph{Tritium monococcum}), domesticated from wild einkorn (\emph{T.
 m.
-boeoticum) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
-Subsequently, emmer wheat (Triticum dicoccum) was domesticated from hybrids of diverse populations of wild emmer wheat (Triticum dioccoides), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}.
-Two distinct taxa of T.
-dioccoides exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
-The tetraploid durum wheat (Triticum turgidum) is the result of hybridization between red wild einkorn wheat (T.
-urartu) and wild goatgrass (Aegilops speltoides) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (Aegilops tauschii) produced the allohexaploid common bread wheat (T.
+boeoticum}) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
+Subsequently, emmer wheat (\emph{T.
+dicoccum}) was domesticated from hybrids of diverse populations of wild emmer wheat (\emph{T.
+dioccoides}), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}\.
+Two distinct taxa of \emph{T.
+dioccoides} exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
+The tetraploid durum wheat (\emph{T.
+turgidum}) is the result of hybridization between red wild einkorn wheat (\emph{T.
+urartu}) and wild goatgrass (\emph{Aegilops speltoides}) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (\emph{A.
+tauschii) produced the allohexaploid common bread wheat (T.
 aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
@@
 
-The progressive domestication of wheat has involved many introgressive acts between geographically dispersed lineages, evinced by identification of genomes' various sources in wheats with elevated ploidy levels.
-
-The convoluted domestication of wheat is beset by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
+The convoluted domestication of wheat is marked by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
 Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
 Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
-The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}.
+The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}\.
 Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
 Modern breeding efforts between wild and domesticated wheat have shown some success.
-For example, some researchers have experimentally bred resistance traits into durum wheat (stem rust resistance from wild einkorn, \cite{gerechter1971transfer}\, and powdery mildew resistance from T.
+For example, some researchers have experimentally bred resistance traits into durum wheat (stem rust resistance from wild einkorn, \cite{gerechter1971transfer}\, and powdery mildew resistance from \emph{T.
 turgidum var.
-dicoccoides, \cite{blanco2008molecular}\.
+dicoccoides}, \cite{blanco2008molecular}\.
 
@@
 
-Domesticated and wild barleys belong to the same species, Hordeum vulgare, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
-Barley (Hordeum vulgare subsp.
-vulgare) is believed to have been domesticated at least twice, once from wild subsp.
-spontaneum in the Fertile Crescent and once by subsp.
+Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
+Barley (\emph{Hordeum vulgare subsp.
+vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp.
+spontaneum} in the Fertile Crescent and once by \emph{subsp.
 spontaneum var.
-agriocrithon in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
+agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
 
-Variety agriocrithon (H. vulgare ssp. vulgare f. agriocrithon) is genetically diverse, and is found throughout much of the range of Barley.
-Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp.
-spontaneum, perhaps more than once (in Israel as well as in Tibet).
-Presently, the distribution of spontaneum consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
+Variety agriocrithon (\emph{H. vulgare ssp. vulgare f. agriocrithon}) is genetically diverse, and is found throughout much of the range of Barley.
+Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from \emph{ssp.
+spontaneum}, perhaps more than once (in Israel as well as in Tibet).
+Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
 These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}\.
@@
 Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
-Owing to this sparcity of evidence is the low rate of outcrossing of spontaneum, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
-There has been little genetic investigation into spontaneous barley/spontaneum hybrids \cite{ellstrand2003dangerous}.
-Barley/H.
-spontaneum hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+Owing to this sparcity of evidence is the low rate of outcrossing of \emph{spontaneum}, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
+There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}\.
+Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
 Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}\.
@@
 
-The common sunflower (Helianthus annuus) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
-Helianthus is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}\.
-[[This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.]] This sentence may not help much, actually.
-We're betting on genetic adaptability via introg hyb
-H.
-annuus adapted to the environment in Texas by hybridizing with H.
+The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
+\emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}\.
+This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.
+Domesticated sunflower lives in sympatry with wild relatives like \emph{H.
+petiolaris} and \emph{H.
+bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
+Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
+\emph{H.
+annuus} shows signs of persistent introgressive hybridization with \emph{H.
+petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
+\emph{H.
+annuus} adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
 debilis ssp.
-cucumerifolius, gaining advantageous alleles (these hybrids are now called H.
+cucumerifolius}), gaining advantageous alleles (these hybrids are now called \emph{H.
 annuus ssp.
-texanus, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
-H. annuus lives in sympatry with wild relatives like H. petiolaris and H. bolanderi and forms  stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
-H.
-annuus shows signs of persistent introgressive hybridization with H.
-petiolaris with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
-Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
-Weediness is a trait that has likely evolved many times \cite{kane2008genetics}\.
+texanus}, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}\.
 
-Helianthus has several genes for downy mildew resistance.
-Each imparts resistance to one or more races of P. halstedii, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
-Some of these downy mildew resistence genes were found in wild relatives (including H.
-argophyllus, H.
-tuberosus, and H.
-praecox) and have been successfully bred into modern H.
-annuus \cite{miller1991inheritance}\.
-PlArg, an allele found in wild silverleaf sunflowers (Helianthus argophyllus, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
-Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and Phomopsis resistance breeding efforts \cite{besnard1997specifying}\.
+\emph{Helianthus} has several genes for downy mildew resistance.
+Each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
+argophyllus}, \emph{H.
+tuberosus}, and \emph{H.
+praecox}) and have been successfully bred into modern \emph{H.
+annuus} \cite{miller1991inheritance}\.
+PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
+argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
+Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}\.
 
@@
 
-There are two main cultigens of rice; Asian rice (Oryza sativa) and African rice (O.
-glaberrima), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
-The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for O.
-sativa, 8,200-13,500 years ago, from wild O.
-rufipogon  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of O.
-perennis complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for O.
-glaberrima, 2-3,000 years ago, from a wild ancestor, perhaps O.
-barthii \cite{ricepedia}\ or O.
-breviligulata \cite{oka2012origin}\.
+There are two main cultigens of rice; Asian rice (\emph{Oryza sativa}) and African rice (\emph{O.
+glaberrima}), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
+The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O.
+sativa}, 8,200-13,500 years ago, from wild \emph{O.
+rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
+perennis} complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for \emph{O.
+glaberrima}, 2-3,000 years ago, from a wild ancestor, perhaps \emph{O.
+barthii} \cite{ricepedia}\ or \emph{O.
+breviligulata} \cite{oka2012origin}\.
 Asian rice has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
-Indica and Japonica are likely the result of independent domestications from separate O. rufipogon populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
+Indica and Japonica are likely the result of independent domestications from separate \emph{O.
+rufipogon} populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
 These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}\.
@@
 Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty, \cite{ricepedia}\), and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
-The greater genetic diversity within O.
-sativa is likely due to introgression with wild relatives both during domestication and upon the dispersal of O.
-sativa into new environments and sympatry with new relatives \cite{second1982origin}.
-Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into O.
-sativa L.
-by researchers \cite{brar1997alien, khush1974inheritance}.
+The greater genetic diversity within \emph{O.
+sativa} is likely due to introgression with wild relatives both during domestication and upon the dispersal of \emph{O.
+sativa} into new environments and sympatry with new relatives \cite{second1982origin}.
+Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
+sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
 Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
@@
 Rye domestication has recieved comparatively little academic interest.
-Although there is little evidence for or against the role of natural introgression in rye domestication, rye (Secale cereale) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
+Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
 
@@
 
-Modern potato (Solanum tuberosum) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
-The northern members of the polyphyletic S.
-brevicaule wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because s.
-tuberosum has a polyphyletic origin.
-Also, these northern members of the S.
-breviaule complex are not clearly defined, and may in fact be one singular species (in which case the species name would be S.
-bukasovii, \cite{spooner2005single}\).
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
+The northern members of the polyphyletic \emph{S.
+brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
+tuberosum} has a polyphyletic origin.
+Also, these northern members of the \emph{S.
+breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
+bukasovii}, \cite{spooner2005single}\).
 
@@
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
-Cultivated diploid potatoes have been shown introgressed with genes from S. sparsipilum \cite{rabinowitz1990high}\ and S. megistacrolobum \cite{johns1987relationships, huaman1980solanum}.
-Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the Central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
-Resistance genes have been experimentally introgressed from wild poptato relatives (S.
+Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
+sparsipilum} \cite{rabinowitz1990high}\ and \emph{S.
+megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
+Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
+Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
 tuberosum subsp.
-andigena, \cite{van1999tight}, S.
-bulbocastanum, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
+andigena}, \cite{van1999tight}, \emph{S.
+bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
 
@@
 
-The domestication center for tomato is not known with certainty, but \cite{sims1979history} quotes Dr.
-Charles Rick (University of California, August 1978) stating that tomato was domesticated from Lycopersicon esculentum v.
-cerasiformae [[solanum lycopersicum cerasiforme]] in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
-\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agrees.
-\cite{nesbitt2002comparative}, however, conclude that v.
-cerasiformae is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes L.
-esculentum and L.
-pimpinellifolium.
+The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
+Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{Solanum lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
+\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
+\cite{nesbitt2002comparative}, however, conclude that \emph{v.
+cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
+esculentum} and \emph{L.
+pimpinellifolium}.
 Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
@@
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of L.
+\cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
 hirsutum f.
-glabratum \cite{banerjee1990transfer}\ to an introgression on chromosome 11, which originated from the wild relative L.
-hirsutum.
-Similar resistance traits has also been shown to have been imparted from L. chilense, L. pimpinellifolium, and L. cheesmanii \cite{hanson2000mapping}\.
+glabratum} \cite{banerjee1990transfer}\ to an introgression on chromosome 11, which originated from the wild relative \emph{L.
+hirsutum}.
+Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}\.
 
@@
 
-Common bean (Phaseolus vulgaris) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
-The domesticated common bean progenitor species (P.
+Common bean (\emph{Phaseolus vulgaris}) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
+The domesticated common bean progenitor species (\emph{P.
 vulgaris var.
-aborigineus) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
+aborigineus}) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
 Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}\), rates of gene flow between it and wild relatives are expected to be lower.
@@
 \cite{papa2003asymmetry}\ was unable to determine how much of the genetic variability of domesticated common bean landraces originated from weedy or wild types, but point to the allele for 'L' phaseolin as a possible example of an allele that has been introgressed from wild common beans into domesticated.
-Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (P.
-coccineus) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (Zabrotes subfasciatus and Acanthoscelides obtectus) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
+Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (\emph{P.
+coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
 
 \subsection*{Sorghum}
-Anthropological evidence points to a domestication event of sorghum (Sorghum bicolor subsp. bicolor) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
+Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
 However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
-The progenitor species was S. arundinaceum \cite{doggett1988sorghum, harlan1971toward}\.
+The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
 The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
-Within Eu-Sorghum, there are currently five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
-Members of Sorghum bicolor are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}\.
+Within Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
+Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}\.
 
@@
 Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
-These hybrids express diminished fertility
-Members of Sorghum bicolor have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum, CWR}\.
+These hybrids express diminished fertility.
+Members of \emph{Sorghum bicolor} have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum, CWR}\.
 Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera \cite{de1976cytogenetics}\) for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}\.

### 2015-12-18 figure.tex
added figure.tex, a summary table

\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{llllll}
Crop                          & Center of Origin                                      & Wild Relatives                                                                    & Field Evidence of Gene Flow                          & Genetic Evidence of Gene Flow & Sources                                                                                                   \\
Barley                        & Multiple (Israel-Jordan, Tibet)                       & H. spontaneum, H. vulgare ssp. vulgare f. agriocrithon                            & Morphologically-intermediate hybrids in hybrid zones & not much                      & \cite\{ellstrand2003dangerous, badr2000origin, azhaguvel2007phylogenetic\}                                \\
Common Bean                   & Mesoamerica (Lerma-Santiago Basin) and Southern Andes & P. vulgaris var. aborigineus, others                                              & yes                                                  & yes                           & \cite\{papa2003asymmetry, kwak2009putative, debouck1993genetic\}\textbackslash                            \\
Maize                         & Balsas River Valley, Mexico                           & Zea mays subsp. mexicana                                                          & yes                                                  & yes                           &                                                                                                           \\
Potato                        & southern Peru                                         & many                                                                              & yes                                                  & yes                           & \cite\{spooner2005single, rabinowitz1990high, johns1987relationships, linder1987diversity\}\textbackslash \\
Rice (African, O. glaberrima) & Upper Niger River delta                               & O. barthii, many                                                                  & yes                                                  & yes                           & \cite\{oka2012origin, ricepedia\}                                                                         \\
Rice (Asian, O. sativa)       & unknown (China and/or India)                          & O. rufipogon, many                                                                & yes                                                  & yes                           & \cite\{oka2012origin, ricepedia\}                                                                         \\
Rye                           & unknown (Fertile Crescent)                            & several                                                                           & maybe?                                               & little/none                   & \cite\{vavilov1928geographical\}\textbackslash                                                            \\
Sorghum                       & Ethiopian Sudan                                       & S. bicolor ssp. drummondii, verticilliflorum, S. propinquum, S. halepense, others & yes                                                  & yes                           & \cite\{arriola1996crop, ellstrand2003dangerous                                                            \\
Sunflower                     & eastern United States                                 & H. petiolaris, H. bolanderi, H. debilis, H. argophyllus,                          & yes                                                  & yes                           & \cite\{harter2004origin, wills2006chloroplast\}\textbackslash                                             \\
Tomato                        & unknown (Mexico)                                      & many                                                                              & yes                                                  & yes                           & {[}{[}http://www.landscapeimagery.com/tomato.html{]}{]}                                                   \\
Wheat (Emmer)                 & Karacadag Mountains, southeastern Turkey              & T. araraticum, T. dioccoides                                                      & yes                                                  & yes                           & \cite\{lev2000cradle, civavn2013reticulated, luo2007structure\}                                           \\
Wheat (Einkorn)               & Fertile Crescent                                      & T. araraticum, T. dioccoides, T. monococccum boeoticum                            & yes                                                  & yes                           & \cite\{heun1997site, harlan1966distribution, ozkan2002aflp\}\textbackslash                               
\end{tabular}
\end{table}

### 2016-01-05 Li_outline.tex
made edits (mainly grammatical) to Li_outline.tex

@@
 \subsection{maize}
-Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it is sympatric with a lot of its wild relatives with a range of different divergence time \cite{hufford2013}.
-\emph{Zea mays} ssp. \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize is a well-studied pair.
+Maize (\emph{Zea mays} ssp.
+\emph{mays}) is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it is sympatric with a lot of its wild relatives with a range of different divergence time \cite{hufford2013}.
+The teosinte \emph{Zea mays} ssp. \emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize is a well-studied pair.
 Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
-However, \citep{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
+However, \citep{hufford2013} was the first to reveal evidence of adaptive introgression from \emph{mexicana} into Mexican highland maize.
 The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAMPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled.
-Among the nine regions, three spans the centromeres of chromosomes 5, 6 and 10 and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
+Among the nine regions, three spans the centromeres of chromosomes 5, 6 and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
 By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
-It was also revealed that in the growth chamber experiment maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
+It was also revealed that in the growth chamber experiment maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTLs controlling pigment density and macrohairs) and 9 (overlapped with QTLs for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
 To put it together, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
@@
 
-However, has the introgression from \emph{mexicana} spread to multiple highland populations? 
+Several questions regarding \emph{mexicana} yet remain.
+Has the introgression from \emph{mexicana} spread to multiple highland populations? 
 How does the adaptation differ among populations with and without introgression from \emph{mexicana}?
@@
 (2015) \citep{martin2015} were calculated to locate the introgressed loci.
-Figure \ref{fig:fd} showed the loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across two chromosomes (chr4 and chr5) in multiple comparisons. 
-On chromosome 4 (Figure \ref{fig:fd}) , Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
+Figure \ref{fig:fd} shows the loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across two chromosomes (chr4 and chr5) in multiple comparisons. 
+On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
 (2013) \citep{hufford2013}.
@@
 On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
-More details on the other chromosomes could be referred to \cite{Wang2015manuscript}.
-Overall, both analyses revealed that gene flow exist from \emph{mexicana} not only to MexHigh maize, but also to GuaHigh maize, as well as to some individuals in SW\_US, but not to the Andes and South American lowland (SA\_Low) populations, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations.
+More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
+Overall, both analyses revealed that gene flow from \emph{mexicana} exists, not only to MexHigh maize, but also to GuaHigh maize, as well as to some individuals in SW\_US, but not to the Andes and South American lowland (SA\_Low) populations, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations.
 The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
-%The stronger genetic drift introduced by the smaller historical population size improved the homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
+%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
 The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
-Even though it is clear that the absence of introgression from wild relatives provided less genetic resource for the highland adaptation and thus made the highland adaptation unique in the Andes, it is yet unknown that out of reach of wild relatives is a reason for reduced fitness in the Andes population.
-And the question whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+Even though it is clear that the absence of introgression from wild relatives provided fewer genetic resources for the highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
+And the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
 

### 2016-05-24 TiG_prospectus.tex
TiG_Files added

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage{natbib}
\bibliographystyle{plain}
\newcommand{\mbh}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments

\title{Review: Crop adaptation through post-domestication introgression from the wild}

\author[1]{Authors: Garrett M. Janzen}%author information
\author[1]{Li Wang}
\author[1,*]{Matthew B. Hufford}
\affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
\affil[*]{Correspondence: mhufford@iastate.edu (M.B. Hufford)}
\date{}

\begin{document}

\maketitle

The traditional paradigm in crop domestication studies has been origin from one or more defined geographic centers followed by dispersal and expansion to the modern-day extent of cultivation.
Absent from this paradigm are the genetic effects of hybridization between diffusing domesticates and closely-related, locally-adapted wild relatives outside the center of origin.
New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that introgression has conferred local adaptation to crops, 3) consider how the prevalence of post-domestication introgression alters traditional concepts of domestication, and 4) describe future advances in basic and applied genetics that can be made through the study of crop-wild introgression.

\section*{Introgression methods and recent applications}

\section*{The role of introgression in conferring local adaptation in crops}

\section*{The domestication process in light of introgression}

\section*{Future studies in crop-wild introgression}

\cite{Hufford2013}

\bibliography{TiG_references}

\end{document}

### 2016-05-24 TiG_prospectus.tex
edits to prospectus

@@
 
-The traditional paradigm in crop domestication studies has been origin from one or more defined geographic centers followed by dispersal and expansion to the modern-day extent of cultivation.
-Absent from this paradigm are the genetic effects of hybridization between diffusing domesticates and closely-related, locally-adapted wild relatives outside the center of origin.
+The traditional paradigm in crop domestication studies has been origin from a wild relative within one or more defined geographic centers followed by dispersal and expansion to the modern-day extent of cultivation.
+Absent from this paradigm are the effects of hybridization between diffusing domesticates and closely-related, locally-adapted wild relatives outside the center of origin.
 New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
-In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that introgression has conferred local adaptation to crops, 3) consider how the prevalence of post-domestication introgression alters traditional concepts of domestication, and 4) describe future advances in basic and applied genetics that can be made through the study of crop-wild introgression.
+In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation to crops, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
@@
 
-\section*{The role of introgression in conferring local adaptation in crops}
+\section*{The role of introgression in conferring local adaptation to crops}
 

### 2016-05-24 TiG_prospectus.tex
edits to prospectus

@@
 
-\title{Review: Crop adaptation through post-domestication introgression from the wild}
+\title{Review: Crop adaptation through post-domestication introgression from wild relatives}
 
@@
 
-The traditional paradigm in crop domestication studies has been origin from a wild relative within one or more defined geographic centers followed by dispersal and expansion to the modern-day extent of cultivation.
+The traditional paradigm in crop domestication has been origin from a wild relative within one or more defined geographic centers followed by expansion to the modern-day extent of cultivation.
 Absent from this paradigm are the effects of hybridization between diffusing domesticates and closely-related, locally-adapted wild relatives outside the center of origin.
@@
 \section*{Introgression methods and recent applications}
+The recent availability of genome-wide resequencing and reduced-representation genotyping data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species.
+\begin{enumerate}
+\item{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods}
+\item{}
+\end{enumerate}
 
-\section*{The role of introgression in conferring local adaptation to crops}
+\section*{Crop adaptation through introgression}
 
-\section*{The domestication process in light of introgression}
+\section*{Reevaluating concepts of domestication}
 

### 2016-05-24 TiG_prospectus.tex
edits to prospectus

@@
 \usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
-\usepackage{natbib}
-\bibliographystyle{plain}
+\usepackage[numbers,sort]{natbib}
+\bibliographystyle{ieeetr}
 \newcommand{\mbh}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
@@
 \section*{Introgression methods and recent applications}
-The recent availability of genome-wide resequencing and reduced-representation genotyping data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species.
+The recent availability of genome-wide resequencing and reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq) data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species.
 \begin{enumerate}
+\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin}
+\item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
+	\begin{enumerate}
+	\item{Introgression from Neanderthals and other archaic hominins into humans at loci controlling skin pigmentation, defense against pathogens, and tolerance of high altitude (reviewed in \cite{Racimo2015}})
+	\end{enumerate} 
 \item{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods}
@@
 
-\cite{Hufford2013}
+\cite{Hufford2013} \cite{Baute2015}
 

### 2016-05-24 TiG_prospectus.tex
edits to prospectus

@@
 \item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin}
-\item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
+\item{In several instances, introgression is conserved across individuals and populations suggesting an adaptive role:}
 	\begin{enumerate}
 	\item{Introgression from Neanderthals and other archaic hominins into humans at loci controlling skin pigmentation, defense against pathogens, and tolerance of high altitude (reviewed in \cite{Racimo2015}})
+	\item{Introgression across butterfly species at protective color-pattern loci conferring M\"{u}llerian mimicry \cite{Heliconius2012}}
+	\item{Adaptive introgression across mosquito species at a locus controlling insecticide resistance \cite{Norris2015}}
 	\end{enumerate} 
@@
 \section*{Crop adaptation through introgression}
+\cite{Hufford2013} \cite{Baute2015} \cite{Bredeson2015}
 
@@
 
-\cite{Hufford2013} \cite{Baute2015}
-
 \bibliography{TiG_references}

### 2016-05-24 crop_adaptation_gene_flow_gj_outline.tex
glitch fixes in Garrett's outline

@@
 \usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
-
+\usepackage{natbib}
+\bibliographystyle{ieeetr}
 \newcommand{\mbh}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
@@
 
-Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite {matsuoka2002single} from teosinte.
+Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite{matsuoka2002single} from teosinte.
 
@@
 urartu}) and wild goatgrass (\emph{Aegilops speltoides}) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (\emph{A.
-tauschii) produced the allohexaploid common bread wheat (T.
+tauschii} produced the allohexaploid common bread wheat (T.
 aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
 This hexaploidization event probably occured twice, once in southeast Turkey/northern Syria and, more recently, in Iran \cite{giles2006gludy}\.
-Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}\.
+Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}.
 
@@
 turgidum var.
-dicoccoides}, \cite{blanco2008molecular}\.
+dicoccoides}, \cite{blanco2008molecular}.
 
@@
 Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
-These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}\.
+These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
 
@@
 Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
-Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}\.
+Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.
 
@@
 The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
-\emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}\.
+The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
 This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.
@@
 Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
-\emph{H.
+The species \emph{H.
 annuus} shows signs of persistent introgressive hybridization with \emph{H.
 petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
-\emph{H.
+The species \emph{H.
 annuus} adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
@@
 annuus ssp.
-texanus}, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}\.
+texanus}, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
 
@@
 argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
-Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}\.
+Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
 
@@
 rufipogon} populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
-These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}\.
+These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}.
 
-Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty, \cite{ricepedia}\), and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
+Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia} and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
 The greater genetic diversity within \emph{O.
@@
 Rye domestication has recieved comparatively little academic interest.
-Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}\.
+Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}.
 
@@
 breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
-bukasovii}, \cite{spooner2005single}\).
+bukasovii} \cite{spooner2005single}.
 
@@
 However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
-Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}\.
+Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
 
@@
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
+Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
 hirsutum f.
-glabratum} \cite{banerjee1990transfer}\ to an introgression on chromosome 11, which originated from the wild relative \emph{L.
+glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
 hirsutum}.
-Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}\.
+Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
 
@@
 Wild common beans with this introgression are also occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
-\cite{kwak2009structure}\ also found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
-The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}\).
+Also, \cite{kwak2009structure} found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
+The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}.
 Farmer practices may also be diminishing introgressive gene flow.
@@
 Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (\emph{P.
-coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}\.
+coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}.
 
@@
 Within Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
-Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}\.
+Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}.
 
@@
 Members of \emph{Sorghum bicolor} have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum, CWR}\.
-Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera \cite{de1976cytogenetics}\) for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}\.
-Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}\.
+Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera \cite{de1976cytogenetics} for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}.
+Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}.
 
+\bibliography{bib_gj.bib}
 \end{document}

### 2016-05-25 TiG_prospectus.tex
edits to prospectus

--- 2016-05-24 TiG_prospectus.tex
@@
 New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
-In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation to crops, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
-\section*{Introgression methods and recent applications}
-The recent availability of genome-wide resequencing and reduced-representation genotyping data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species.
+\section*{Introgression methods and their recent applications}
+The recent availability of genome-wide resequencing and reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq) data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
 \begin{enumerate}
-\item{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods}
-\item{}
+\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa.}
+\item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
+	\begin{enumerate}
+	\item{Adaptive introgression has been discovered from Neanderthals and other archaic hominins into humans at loci controlling skin pigmentation, defense against pathogens, and tolerance of high altitude (reviewed in \cite{Racimo2015}}).
+	\item{Introgression across butterfly species at protective color-pattern loci has conferred M\"{u}llerian mimicry (\cite{Heliconius2012}; \textbf{Figure 1})}.
+	\item{Adaptive introgression across mosquito species has spread insecticide resistance \cite{Norris2015}}.
+	\item{Introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and has contributed to speciation \cite{Stankowski2015}}.
+	\end{enumerate}
+\item{Identification of introgression across a wide range of taxa at loci controlling a variety of adaptive traits suggests it has been an important evolutionary force.}
 \end{enumerate}
 
+\noindent{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods.}
+
+\noindent{\textbf{Figure 1}: Wing coloration patterns in \emph{Heliconius} and evidence for introgression across species based on Patterson's \emph{D}-statistic; adapted from \cite{Heliconius2012}.}
+
 \section*{Crop adaptation through introgression}
+\cite{Hufford2013} \cite{Baute2015} \cite{Bredeson2016} \cite{Poets2015} \cite{Huang2012} \cite{Diez2015}
 
@@
 
-\cite{Hufford2013}
-
 \bibliography{TiG_references}

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 \section*{Crop adaptation through introgression}
-\cite{Hufford2013} \cite{Baute2015} \cite{Bredeson2016} \cite{Poets2015} \cite{Huang2012} \cite{Diez2015}
+Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and wild relatives outside putative domestication centers.
+A history of introgression during diffusion appears to be rule for crops rather an exception.
+\begin{enumerate}
+	\item{Crops, given their frequent history of diffusion from defined centers of origin, are ideal recipients of adaptive introgression.
+Theory suggests that such colonizing species will overwhelmingly be recipients rather than donors of introgression \cite{Currat2008}.}
+	\item{Empirical studies are revealing that introgression has occurred in many of the world's most important crops.}
+	\begin{enumerate}
+		\item{Maize: As maize spread from a lowland center of origin it received introgression from a highland-adapted wild relative \cite{Hufford2013}}
+		\item{Sunflower: \cite{Baute2015}}
+		\item{Rice: \cite{Huang2012}}
+		\item{Barley: \cite{Poets2015}}
+		\item{Olive: \cite{Diez2015}}
+		\item{Cassava: \cite{Bredeson2016}}
+	\end{enumerate}
+\end{enumerate}
 

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 	\begin{enumerate}
-		\item{Maize: As maize spread from a lowland center of origin it received introgression from a highland-adapted wild relative \cite{Hufford2013}}
-		\item{Sunflower: \cite{Baute2015}}
+		\item{Maize: As maize spread from a lowland center of origin it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
+		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives post-domestication that has potentially reintroduced a branching morphology \cite{Baute2015}.}
 		\item{Rice: \cite{Huang2012}}
-		\item{Barley: \cite{Poets2015}}
+		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives suggesting introgression \cite{Poets2015}.}
 		\item{Olive: \cite{Diez2015}}

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 \newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
+\newcommand{\gmg}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
 
@@
 Theory suggests that such colonizing species will overwhelmingly be recipients rather than donors of introgression \cite{Currat2008}.}
-	\item{Empirical studies are revealing that introgression has occurred in many of the world's most important crops.}
+	\item{Empirical studies are increasingly revealing that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
 	\begin{enumerate}
@@
 	\end{enumerate}
+	\item{Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to their current extent of cultivation and distribution of wild relatives across environmental gradients (\textbf{Figure 2}).}
 \end{enumerate}
+
+\noindent{\textbf{Table 2:} Overview of literature evaluating potential for crop-wild introgression in the world's 15 most important crops including a column of references and a column summarizing major findings.}
+
+\noindent{\textbf{Figure 2}: Multi-panel figure showing maps of crop centers of domestication, distributions of wild relatives, and current extent of cultivation across gradients of (a) temperature, (b) precipitation, and (c) elevation.}
 

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 
-\noindent{\textbf{Figure 2}: Multi-panel figure showing maps of crop centers of domestication, distributions of wild relatives, and current extent of cultivation across gradients of (a) temperature, (b) precipitation, and (c) elevation.}
+\noindent{\textbf{Figure 2}: Multi-panel figure showing maps of crop centers of domestication, distributions of wild relatives, and current extent of cultivation across gradients of (a) temperature, (b) precipitation, and (c) elevation.
+This figure will illustrate the adaptive potential of wild-to-crop introgression.}
 

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 
-\title{Review: Crop adaptation through post-domestication introgression from wild relatives}
+\title{Review: Crop adaptation through historical introgression from wild relatives}
 
@@
 \begin{enumerate}
-\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa.}
+\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012}.}
 \item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
@@
 Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and wild relatives outside putative domestication centers.
-A history of introgression during diffusion appears to be rule for crops rather an exception.
+A history of introgression during diffusion appears to be the rule for crops rather than the exception.
 \begin{enumerate}
@@
 Theory suggests that such colonizing species will overwhelmingly be recipients rather than donors of introgression \cite{Currat2008}.}
-	\item{Empirical studies are increasingly revealing that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
+	\item{Empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
 	\begin{enumerate}
@@
 		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives post-domestication that has potentially reintroduced a branching morphology \cite{Baute2015}.}
-		\item{Rice: \cite{Huang2012}}
+		\item{Rice: Introgression appears to have played an important role in the history of rice.
+Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization and introgression from new wild rice populations in south and southeast Asia \cite{Huang2012}.}
 		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives suggesting introgression \cite{Poets2015}.}
-		\item{Olive: \cite{Diez2015}}
-		\item{Cassava: \cite{Bredeson2016}}
+		\item{Olive: Hybridization between wild and domesticated olive has been prevalent throughout the Mediterranean Basin \cite{Diez2015}.}
+		\item{Cassava: Substantial introgression from a closely-related species has been found in cassava, occurring both naturally and through a targeted breeding program for disease resistance \cite{Bredeson2016}.}
 	\end{enumerate}

### 2016-05-26 TiG_prospectus.tex
edits to prospectus

@@
 	\begin{enumerate}
-		\item{Maize: As maize spread from a lowland center of origin it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
+		\item{Maize: As maize spread from a lowland center of origin into the Mexican highlands it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
 		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives post-domestication that has potentially reintroduced a branching morphology \cite{Baute2015}.}
 		\item{Rice: Introgression appears to have played an important role in the history of rice.
-Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization and introgression from new wild rice populations in south and southeast Asia \cite{Huang2012}.}
+Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012}.}
 		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives suggesting introgression \cite{Poets2015}.}
 		\item{Olive: Hybridization between wild and domesticated olive has been prevalent throughout the Mediterranean Basin \cite{Diez2015}.}
-		\item{Cassava: Substantial introgression from a closely-related species has been found in cassava, occurring both naturally and through a targeted breeding program for disease resistance \cite{Bredeson2016}.}
+		\item{Cassava: Substantial introgression from a closely-related wild species has been found in cassava, occurring both naturally and through a targeted breeding program for disease resistance \cite{Bredeson2016}.}
 	\end{enumerate}
-	\item{Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to their current extent of cultivation and distribution of wild relatives across environmental gradients (\textbf{Figure 2}).}
+	\item{Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to both their current extent of cultivation and the distribution of wild relatives across environmental gradients (\textbf{Figure 2}).}
 \end{enumerate}
@@
 \section*{Reevaluating concepts of domestication}
+A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification.
+A demography incorporating introgression from additional sources appears to be more correct for many crops.
+With this in mind, certain aspects of crop evolution must be reevaluated:
+\begin{enumerate}
+	\item{The initial domestication bottleneck is most likely underestimated when introgression is not considered.
+Chromosomal regions experiencing introgression will potentially have increased effective population size relative to non-introgressed regions.
+Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}
+	\item{The time since domestication based on levels of sequence divergence may be inflated when introgressed haplotypes from divergent taxa are included in estimates.}
+	\item{The influx of diversity through introgression may belie selection occurring on these genomic regions during domestication or local adaptation since one of the hallmarks of selection is reduced diversity.}
+\end{enumerate}
 
 \section*{Future studies in crop-wild introgression}
+Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics:
+\begin{enumerate}
+	\item{Basic questions:}
+		\begin{enumerate}
+		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?}
+		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?}
+		 	\item{To what extent have domesticates served as bridges for gene flow between previously allopatric taxa?}
+		\end{enumerate} 	
+	\item{Applications:}
+		\begin{enumerate}
+		 	\item{Loci underlying the domesticated phenotype that are potentially beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.}
+		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}
+		 \end{enumerate}
+\end{enumerate}
 

### 2016-05-27 TiG_prospectus.tex
Li pushed comments

@@
 In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+\lwang{general comments: A.
+Could this be too long?
+I have no idea about the requirement (300-600 words).
+Is that okay if we listed the outline of each section?
+B.
+Some references are kind of old.
+They require references in the past 2-4 years.
+}
 
 \section*{Introgression methods and their recent applications}
+\lwang{The examples listed here are not crop-wild introgression, which is contradictory with the first point in the abstract part.
+"briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression".
+When I red here, I felt my attention was diverted.
+Maybe just one or two sentences.}
+
 The recent availability of genome-wide resequencing and reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq) data combined with new analytical methods have facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
 \begin{enumerate}
-\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012}.}
+\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.}
 \item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
@@
 \noindent{\textbf{Figure 1}: Wing coloration patterns in \emph{Heliconius} and evidence for introgression across species based on Patterson's \emph{D}-statistic; adapted from \cite{Heliconius2012}.}
+
+\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model.
+I put the three in the bib file.}
 
@@
 \end{enumerate}
+\lwang{I am afraid that most of the crop examples actually did show evidence of introgression, but few supplied strong evidence of adaptive introgression.
+I have not red all of these paper, maybe it is just my prejudice.
+}
 
@@
 Chromosomal regions experiencing introgression will potentially have increased effective population size relative to non-introgressed regions.
-Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}
-	\item{The time since domestication based on levels of sequence divergence may be inflated when introgressed haplotypes from divergent taxa are included in estimates.}
-	\item{The influx of diversity through introgression may belie selection occurring on these genomic regions during domestication or local adaptation since one of the hallmarks of selection is reduced diversity.}
+Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}
+	\item{The time since domestication based on levels of sequence divergence may be inflated when introgressed haplotypes from divergent taxa are included in estimates.}\lwang{It could be either inflated or underestimate,.
+It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
+	\item{The influx of diversity through introgression may belie selection occurring on these genomic regions during domestication or local adaptation since one of the hallmarks of selection is reduced diversity.} \lwang{I donot quite understand this point.
+So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right?
+But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
 \end{enumerate}
@@
 		\begin{enumerate}
-		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?}
+		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} \lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
 		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?}
-		 	\item{To what extent have domesticates served as bridges for gene flow between previously allopatric taxa?}
+		 	\item{To what extent have domesticates served as bridges for gene flow between previously allopatric taxa?} \lwang{I feel the second and the third points could be combined.
+How the wild species could introgress into the allopatric landraces?
+Does domesticates serve as bridges?}
 		\end{enumerate} 	
@@
 		\begin{enumerate}
-		 	\item{Loci underlying the domesticated phenotype that are potentially beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.}
+		 	\item{Loci underlying the domesticated phenotype that are potentially beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} \lwang{I am not quite clear with this point.
+It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow.
+Right?}
 		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}

### 2016-05-31 TiG_prospectus.tex
Garrett made some comments

@@
 \newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
-\newcommand{\gmg}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
+\newcommand{\gmj}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
 
@@
 Maybe just one or two sentences.}
+\gmj{I could not find the word count limit on the TiG website, but if we need to cut down, Item 2.
+of this section might be a place to do so.
+a-d could perhaps be reduced to a single sentence, for example.}
 
@@
 	\begin{enumerate}
-		\item{Maize: As maize spread from a lowland center of origin into the Mexican highlands it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
+		\item{Maize: As maize spread from a lowland center of origin into the Mexican highlands, it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
 		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives post-domestication that has potentially reintroduced a branching morphology \cite{Baute2015}.}
@@
 Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012}.}
-		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives suggesting introgression \cite{Poets2015}.}
+		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives, suggesting introgression \cite{Poets2015}.}
 		\item{Olive: Hybridization between wild and domesticated olive has been prevalent throughout the Mediterranean Basin \cite{Diez2015}.}
@@
 }
+\gmj{Not all crops that I have researched are included in this list.
+Are we only here mentioning the strongest cases, or will the weaker cases be excluded from the paper altogether?
+Also, regarding Li's comments above, many times there is indeed evidence that the introgression is adaptive.
+This might be most apparent when the introgression confers pest resistance alleles, but also in other cases as well.
+That being said, perhaps the caveat mentioning the distinction between adaptive and neutral introgression should be made more explicit?}
 
@@
 This figure will illustrate the adaptive potential of wild-to-crop introgression.}
+\gmj{Will we have to generate these images before we submit the prospectus?}
 
@@
 Chromosomal regions experiencing introgression will potentially have increased effective population size relative to non-introgressed regions.
-Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}
+Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?
+I just want to make sure I understand.}
 	\item{The time since domestication based on levels of sequence divergence may be inflated when introgressed haplotypes from divergent taxa are included in estimates.}\lwang{It could be either inflated or underestimate,.
@@
 But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
+	\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
+The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
 \end{enumerate}
@@
 		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} \lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
-		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?}
+		 	\item{At what geographic scale does adaptive introgression occur?
+Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?}\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g.
+temperature, precipitation, altitude, day lengh, etc.)?}
 		 	\item{To what extent have domesticates served as bridges for gene flow between previously allopatric taxa?} \lwang{I feel the second and the third points could be combined.
@@
 It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow.
-Right?}
-		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}
+Right?}\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?
+I hope that question made sense.}
+		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}\gmj{This point is very clear, but should it be fleshed out further?
+It wouldn't be difficult to include some examples.}
 		 \end{enumerate}

### 2016-06-01 TiG_prospectus.tex
submitted version of proposal

--- 2016-05-27 TiG_prospectus.tex
@@
 \bibliographystyle{ieeetr}
-\newcommand{\mbh}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
+\newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
 \newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
-\newcommand{\gmg}[1]{\textcolor{green}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
+\newcommand{\gmj}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
 
-\title{Review: Crop adaptation through historical introgression from wild relatives}
+\title{Review: Crop adaptation through introgression from wild relatives}
 
@@
 
-The traditional paradigm in crop domestication has been origin from a wild relative within one or more defined geographic centers followed by expansion to the modern-day extent of cultivation.
-Absent from this paradigm are the effects of hybridization between diffusing domesticates and closely-related, locally-adapted wild relatives outside the center of origin.
+The traditional paradigm in crop domestication has been origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
+Absent from this paradigm are the effects of hybridization between a diffusing nascent crop species and closely-related, locally-adapted wild relatives outside the center of origin.
 New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
-In this review, we will: 1) briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
-\lwang{general comments: A.
+In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+%\lwang{general comments: A.
 Could this be too long?
@@
 
-\section*{Introgression methods and their recent applications}
-\lwang{The examples listed here are not crop-wild introgression, which is contradictory with the first point in the abstract part.
+\section*{Introgression methods and their applications}
+%\lwang{The examples listed here are not crop-wild introgression, which is contradictory with the first point in the abstract part.
 "briefly describe these methods and provide a summary of their recent application for detection of crop-wild introgression".
@@
 Maybe just one or two sentences.}
+%\gmj{I could not find the word count limit on the TiG website, but if we need to cut down, Item 2.
+of this section might be a place to do so.
+a-d could perhaps be reduced to a single sentence, for example.}
 
@@
 \item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.}
-\item{In several instances, introgression has been found to be conserved across individuals and populations suggesting an adaptive role:}
-	\begin{enumerate}
-	\item{Adaptive introgression has been discovered from Neanderthals and other archaic hominins into humans at loci controlling skin pigmentation, defense against pathogens, and tolerance of high altitude (reviewed in \cite{Racimo2015}}).
-	\item{Introgression across butterfly species at protective color-pattern loci has conferred M\"{u}llerian mimicry (\cite{Heliconius2012}; \textbf{Figure 1})}.
-	\item{Adaptive introgression across mosquito species has spread insecticide resistance \cite{Norris2015}}.
-	\item{Introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and has contributed to speciation \cite{Stankowski2015}}.
-	\end{enumerate}
-\item{Identification of introgression across a wide range of taxa at loci controlling a variety of adaptive traits suggests it has been an important evolutionary force.}
+\item{Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; \textbf{Figure 1}), introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}}.
 \end{enumerate}
@@
 
-\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model.
+%\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model.
 I put the three in the bib file.}
@@
 \begin{enumerate}
-	\item{Crops, given their frequent history of diffusion from defined centers of origin, are ideal recipients of adaptive introgression.
-Theory suggests that such colonizing species will overwhelmingly be recipients rather than donors of introgression \cite{Currat2008}.}
-	\item{Empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
+	\item{Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
+Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.}
+	\item{Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
 	\begin{enumerate}
-		\item{Maize: As maize spread from a lowland center of origin into the Mexican highlands it received introgression from a highland-adapted wild relative \cite{Hufford2013}.}
-		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives post-domestication that has potentially reintroduced a branching morphology \cite{Baute2015}.}
+		\item{Maize: As maize spread from a lowland center of origin into the Mexican highlands, it received introgression from a highland-adapted wild relative.
+Introgression occurred predominantly at loci previously shown to underlie highland adaptation traits, strongly suggesting adaptive introgression \cite{Hufford2013}.}
+		\item{Cassava: Substantial introgression from a closely-related wild species has been found in cassava, occurring both naturally and through a targeted breeding program to introduce disease resistance \cite{Bredeson2016}.}
 		\item{Rice: Introgression appears to have played an important role in the history of rice.
-Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012}.}
-		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives suggesting introgression \cite{Poets2015}.}
-		\item{Olive: Hybridization between wild and domesticated olive has been prevalent throughout the Mediterranean Basin \cite{Diez2015}.}
-		\item{Cassava: Substantial introgression from a closely-related wild species has been found in cassava, occurring both naturally and through a targeted breeding program for disease resistance \cite{Bredeson2016}.}
+Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012}.
+Whether this wild introgression conferred local adaptation to indica has yet to be shown.}
+		\item{Barley: Landraces of barley show shared ancestry with geographically-proximate populations of wild relatives, suggesting introgression but not demonstrating conclusively that this was adaptive \cite{Poets2015}.}
+		\item{Olive: Hybridization between wild and domesticated olive has been prevalent throughout the Mediterranean Basin. \cite{Diez2015}.}
+		\item{Sunflower: Domesticated sunflower has received substantial introgression from wild relatives through breeding efforts to reintroduce a branching morphology.
+Natural introgression appears prominent in this system as well \cite{Baute2015}.}
 	\end{enumerate}
@@
 \end{enumerate}
-\lwang{I am afraid that most of the crop examples actually did show evidence of introgression, but few supplied strong evidence of adaptive introgression.
+%\lwang{I am afraid that most of the crop examples actually did show evidence of introgression, but few supplied strong evidence of adaptive introgression.
 I have not red all of these paper, maybe it is just my prejudice.
 }
+%\gmj{Not all crops that I have researched are included in this list.
+Are we only here mentioning the strongest cases, or will the weaker cases be excluded from the paper altogether?
+Also, regarding Li's comments above, many times there is indeed evidence that the introgression is adaptive.
+This might be most apparent when the introgression confers pest resistance alleles, but also in other cases as well.
+That being said, perhaps the caveat mentioning the distinction between adaptive and neutral introgression should be made more explicit?}
 
@@
 This figure will illustrate the adaptive potential of wild-to-crop introgression.}
+%\gmj{Will we have to generate these images before we submit the prospectus?}
 
-\section*{Reevaluating concepts of domestication}
+\section*{Re-evaluating concepts of domestication}
 A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification.
 A demography incorporating introgression from additional sources appears to be more correct for many crops.
-With this in mind, certain aspects of crop evolution must be reevaluated:
+With this in mind, certain aspects of crop evolution must be re-evaluated:
 \begin{enumerate}
-	\item{The initial domestication bottleneck is most likely underestimated when introgression is not considered.
-Chromosomal regions experiencing introgression will potentially have increased effective population size relative to non-introgressed regions.
-Inferences regarding the strength of domestication bottlenecks should be revisited with this in mind.}\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}
-	\item{The time since domestication based on levels of sequence divergence may be inflated when introgressed haplotypes from divergent taxa are included in estimates.}\lwang{It could be either inflated or underestimate,.
+	\item{Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
+Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
+For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.}%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?
+I just want to make sure I understand.}
+	\item{Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.}%\lwang{It could be either inflated or underestimate,.
 It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
-	\item{The influx of diversity through introgression may belie selection occurring on these genomic regions during domestication or local adaptation since one of the hallmarks of selection is reduced diversity.} \lwang{I donot quite understand this point.
+	\item{Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
+Introgression may alter these signatures and confound detection of domestication loci.} %\lwang{I donot quite understand this point.
 So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right?
 But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
+	%\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
+The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
 \end{enumerate}
@@
 		\begin{enumerate}
-		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} \lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
-		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?}
-		 	\item{To what extent have domesticates served as bridges for gene flow between previously allopatric taxa?} \lwang{I feel the second and the third points could be combined.
+		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} %\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
+		 	\item{At what geographic scale does adaptive introgression occur?
+Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?
+To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?}%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g.
+temperature, precipitation, altitude, day lengh, etc.)?}
+		 	\item{Can colonizing species served as bridges for gene flow between previously allopatric taxa?} %\lwang{I feel the second and the third points could be combined.
 How the wild species could introgress into the allopatric landraces?
@@
 		\begin{enumerate}
-		 	\item{Loci underlying the domesticated phenotype that are potentially beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} \lwang{I am not quite clear with this point.
+		 	\item{Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} %\lwang{I am not quite clear with this point.
 It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow.
-Right?}
-		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}
+Right?}\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?
+I hope that question made sense.}
+		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}%\gmj{This point is very clear, but should it be fleshed out further?
+It wouldn't be difficult to include some examples.}
 		 \end{enumerate}

### 2016-06-24 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 
+Sugarcane
+
+The genus Saccharum is commonly said to consist of six species (wild species S.
+spontaneum and S.
+robustum, cultivated species S.
+officinarum, S.
+barberi, S.
+sinense, and S.
+edule) \cite{brandes1958origin, jackson2005breeding}\.
+Others \cite{irvine1999saccharum}\ argue that there are fewer distinct species, as the six "species" are interfertile, and some appear to be of hybrid origins.
+Irvine  \cite{irvine1999saccharum}\ suggests that there are two species; S. spontaneum, the putatively ancestral form, and all of the others, adopting the name S. officinarum.
+These two are distinguished by differences in morphology and cytoplasmic DNA sequence.
+Sugarcane was cultivated in Papau, New Guinea about 8000 years ago \cite{keyroyalbotanicalgardens, daniels1987taxonomy, artschwager1958sugarcane}\.
+The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
+
+From \cite{ellstrand1999gene}\:
+33 - daniels1987taxonomy
+    "genetically based traits" evidence for hybrid origin of "many cultivars"
+    "chromosomal evidence" supports "spontaneous intergeneric hybridization" origin of "particular accessions of wild sugarcane relatives"
+    Erianthus maximus might be a hybrid of S. officinarium and Miscanthus (a wild relative genus), based on morphological and chromosomal evidence.
+104 - roach1995sugar
+    "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
+115 - sobral1994phylogenetic
+
+117 - stevenson1965genetics
+    "certain wild canes from Java and Mauritius are hybrids between" wild (S. spontaneum) and cultivated (S. officinarum)
+
+Cassava
+
+Although cultivated cassava (Manihot esculenta) is interfertile with some of its wild relatives (the putative progenitor species M.
+flabellifolia, for example, \cite{roa1997aflp}\), comparisons of single-copy nuclear gene G3pdh sequences provide little evidence to suggest that natural introgressive hybridization occurs, and if it does, it would appear to be in the direction of cultivar to wild relative \cite{olsen1999evidence}\.
+Clonal propogation and allopatry are listed among the barriers to gene flow between cassava and its wild relatives [[but see tovar2015diversity for discussion on elevated genetic diversity in M.
+esculenta, gene flow between species, occasional sexual reproduction, and trade between farmers]].
+Breeders have introgressed genes for protein and disease resistance genes into cassava from M. flabellifolia \cite{akinbo2008introgression}\.
+
+Sweet Potato
+
+Sweet potato (Ipomoea batatas) likely arose from multiple domestications, in Central America and in South America, which were secondarily hybridized upon introduction \cite{roullier2013disentangling}.
+[[see austin1988taxonomy for a summary of the genus Ipomoea, wild relatives and such.]]
+
+Olive
+
+Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to olive's genetic diversity, but it has not been shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+
 \subsection*{Sorghum}
@@
 
+/subsection*{Soybean}
+Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
+Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
+Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
+Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
+The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
+
 \bibliography{bib_gj.bib}

### 2016-06-28 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

--- 2016-06-24 crop_adaptation_gene_flow_gj_outline.tex
@@
 
-Supplement Li's work with the following ideas:
-
-Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it still grows in sympatry with many of its wild relatives \cite{hufford2013}.
-
-Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite{matsuoka2002single} from teosinte.
-
-Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
-
-In particular, \emph{Zea mays} ssp.
-\emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
-\emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
-
-Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
-[[Li's and my bib tags might be in conflict, e.g.
-wilkes1977 is likely my wilkes1977hybridization.]]
-.
- Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
-
-Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
-
-However, \cite{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
-The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled.
-Among the nine regions, three spans the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
-By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
-Growth chamber experiments with maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
-
-Gene flow between \emph{mexicana} and maize is low enough that only fitness-neutral or advantageous alleles are likely to be passed \cite{slatkin1987gene}.
-
-From \cite{baltazar2005pollination}
-Gene flow is probably greater from teosinte to maize.
-Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte and slow rates of introgression.
-Gene flow from maize to teosinte occurs most easily when teosinte pollinates maize
-[[what paper can i cite to mention the gene tb1 and its role in hybridization, viz a viz restricting directionality of teosinte pollen into maize
-
-From \cite{jiang1999genetic}, {bitocchi2009introgression}
-Maize-maize introgression
-Research on introgression between highland and lowland maize shows adaptive traits being transferred
-
-\cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
-\cite{hammer1987collection}\ cites an example in Italy of one domesticated maize adapting to coastal environment by introgressions from another domesticated maize.
-
 \subsection*{wheat}
-
-The first domesticated wheat was einkorn wheat (\emph{Tritium monococcum}), domesticated from wild einkorn (\emph{T.
-m.
-boeoticum}) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
-Subsequently, emmer wheat (\emph{T.
-dicoccum}) was domesticated from hybrids of diverse populations of wild emmer wheat (\emph{T.
-dioccoides}), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}\.
-Two distinct taxa of \emph{T.
-dioccoides} exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
-The tetraploid durum wheat (\emph{T.
-turgidum}) is the result of hybridization between red wild einkorn wheat (\emph{T.
-urartu}) and wild goatgrass (\emph{Aegilops speltoides}) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (\emph{A.
-tauschii} produced the allohexaploid common bread wheat (T.
-aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
-This hexaploidization event probably occured twice, once in southeast Turkey/northern Syria and, more recently, in Iran \cite{giles2006gludy}\.
-Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}.
 
@@
 Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
-Modern breeding efforts between wild and domesticated wheat have shown some success.
-For example, some researchers have experimentally bred resistance traits into durum wheat (stem rust resistance from wild einkorn, \cite{gerechter1971transfer}\, and powdery mildew resistance from \emph{T.
-turgidum var.
-dicoccoides}, \cite{blanco2008molecular}.
 
@@
 
-Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
-Barley (\emph{Hordeum vulgare subsp.
-vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp.
-spontaneum} in the Fertile Crescent and once by \emph{subsp.
-spontaneum var.
-agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
-
-Variety agriocrithon (\emph{H. vulgare ssp. vulgare f. agriocrithon}) is genetically diverse, and is found throughout much of the range of Barley.
-Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from \emph{ssp.
-spontaneum}, perhaps more than once (in Israel as well as in Tibet).
-Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
-These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
-
-The barley domestication process has reduced the number of alleles in the domesticate to only 40% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
 Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
 Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
-Owing to this sparcity of evidence is the low rate of outcrossing of \emph{spontaneum}, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
-There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}\.
 Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
-Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.
-
-\subsection*{sunflower}
-
-The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
-The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
-This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.
-Domesticated sunflower lives in sympatry with wild relatives like \emph{H.
-petiolaris} and \emph{H.
-bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
-Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
-The species \emph{H.
-annuus} shows signs of persistent introgressive hybridization with \emph{H.
-petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
-The species \emph{H.
-annuus} adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
-debilis ssp.
-cucumerifolius}), gaining advantageous alleles (these hybrids are now called \emph{H.
-annuus ssp.
-texanus}, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
-
-\emph{Helianthus} has several genes for downy mildew resistance.
-Each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
-Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
-argophyllus}, \emph{H.
-tuberosus}, and \emph{H.
-praecox}) and have been successfully bred into modern \emph{H.
-annuus} \cite{miller1991inheritance}\.
-PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
-argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
-Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
 
@@
 
-There are two main cultigens of rice; Asian rice (\emph{Oryza sativa}) and African rice (\emph{O.
-glaberrima}), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
-The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O.
-sativa}, 8,200-13,500 years ago, from wild \emph{O.
-rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
-perennis} complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for \emph{O.
-glaberrima}, 2-3,000 years ago, from a wild ancestor, perhaps \emph{O.
-barthii} \cite{ricepedia}\ or \emph{O.
-breviligulata} \cite{oka2012origin}\.
-Asian rice has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
-Indica and Japonica are likely the result of independent domestications from separate \emph{O.
-rufipogon} populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
-These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}.
-
-Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia} and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
-The greater genetic diversity within \emph{O.
-sativa} is likely due to introgression with wild relatives both during domestication and upon the dispersal of \emph{O.
-sativa} into new environments and sympatry with new relatives \cite{second1982origin}.
 Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
@@
 
-\subsection*{rye}
-
-Rye domestication has recieved comparatively little academic interest.
-Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}.
-
 \subsection*{potato}
 
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
-The northern members of the polyphyletic \emph{S.
-brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
-tuberosum} has a polyphyletic origin.
-Also, these northern members of the \emph{S.
-breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
-bukasovii} \cite{spooner2005single}.
-
-Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
-Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
-These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
-However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
-Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
-
 Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}\.
-The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
@@
 
-\subsection*{tomato}
+\subsection*{Sugarcane}
 
-The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
-Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{Solanum lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
-\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
-\cite{nesbitt2002comparative}, however, conclude that \emph{v.
-cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
-esculentum} and \emph{L.
-pimpinellifolium}.
-Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
-
-Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
-Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
-hirsutum f.
-glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
-hirsutum}.
-Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
-
-\subsection*{Common Bean}
-
-Common bean (\emph{Phaseolus vulgaris}) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
-The domesticated common bean progenitor species (\emph{P.
-vulgaris var.
-aborigineus}) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
-Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}\), rates of gene flow between it and wild relatives are expected to be lower.
-These expectations are supported by evidence provided by \cite{papa2003asymmetry}\.
-However, hybrid populations can still be found in sympatry with wild and domesticated common bean, forming domesticate-hybrid-weedy complexes.
-
-Some evidence for introgression exists in the direction of domesitcated to wild common bean, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
-Wild common beans with this introgression are also occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
-Also, \cite{kwak2009structure} found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
-The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}.
-Farmer practices may also be diminishing introgressive gene flow.
-Though some hybrids show heterosis, others (especially F1 hybrids) do not, expressing smaller plant size, smaller seed size, and increased mortality \cite{gutierrez1985heterosis, paredes1995extensive}\.
-Because F1 hybrids are intermediate in seed size and color, farmers may effectively identify and avoid hybrid stock.
-
-Though limited, there is also some evidence of natural introgression of wild genes into domesticated common bean.
-\cite{papa2003asymmetry}\ was unable to determine how much of the genetic variability of domesticated common bean landraces originated from weedy or wild types, but point to the allele for 'L' phaseolin as a possible example of an allele that has been introgressed from wild common beans into domesticated.
-Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (\emph{P.
-coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}.
-
-Sugarcane
-
-The genus Saccharum is commonly said to consist of six species (wild species S.
-spontaneum and S.
-robustum, cultivated species S.
-officinarum, S.
-barberi, S.
-sinense, and S.
-edule) \cite{brandes1958origin, jackson2005breeding}\.
-Others \cite{irvine1999saccharum}\ argue that there are fewer distinct species, as the six "species" are interfertile, and some appear to be of hybrid origins.
-Irvine  \cite{irvine1999saccharum}\ suggests that there are two species; S. spontaneum, the putatively ancestral form, and all of the others, adopting the name S. officinarum.
-These two are distinguished by differences in morphology and cytoplasmic DNA sequence.
-Sugarcane was cultivated in Papau, New Guinea about 8000 years ago \cite{keyroyalbotanicalgardens, daniels1987taxonomy, artschwager1958sugarcane}\.
 The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
@@
     "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
-115 - sobral1994phylogenetic
 
-117 - stevenson1965genetics
-    "certain wild canes from Java and Mauritius are hybrids between" wild (S. spontaneum) and cultivated (S. officinarum)
-
-Cassava
+\subsection*{Cassava}
 
@@
 
-Sweet Potato
+\subsection{Sweet Potato}
 

### 2016-06-28 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-\subsection*{Introduction}
+\subsection*{Sugarcane}
 
- "Whenever sympatric populations of crops and their wild relatives show a similar trait, there will be several possible explanations: (1) retention of an ancestral trait, (2) convergent or parallel evolution, and (3) introgression.
-Perhaps even more difficult is establishing the direction of introgression; does gene flow go from crop into wild species, the reverse, or in both directions?" and "For this reason, introgression may be best studied between crops and their more distant wild relatives where the probability of finding taxon-specific genetic markers is increased." \cite{doebley1990molecular}
+The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
+
+From \cite{ellstrand1999gene}\:
+33 - daniels1987taxonomy
+    "genetically based traits" evidence for hybrid origin of "many cultivars"
+    "chromosomal evidence" supports "spontaneous intergeneric hybridization" origin of "particular accessions of wild sugarcane relatives"
+    Erianthus maximus might be a hybrid of S. officinarium and Miscanthus (a wild relative genus), based on morphological and chromosomal evidence.
+104 - roach1995sugar
+    "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
 
 \subsection*{Maize}
+
+\subsection*{rice}
+
+Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
+sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
+Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
+Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
+Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
 
@@
 Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
-
-\subsection*{barley}
-
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
-Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
-Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
-
-\subsection*{rice}
-
-Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
-sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
-Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
-Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
-Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
 
@@
 
-\subsection*{Sugarcane}
-
-The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
-
-From \cite{ellstrand1999gene}\:
-33 - daniels1987taxonomy
-    "genetically based traits" evidence for hybrid origin of "many cultivars"
-    "chromosomal evidence" supports "spontaneous intergeneric hybridization" origin of "particular accessions of wild sugarcane relatives"
-    Erianthus maximus might be a hybrid of S. officinarium and Miscanthus (a wild relative genus), based on morphological and chromosomal evidence.
-104 - roach1995sugar
-    "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
-
 \subsection*{Cassava}
@@
 
+\subsection*{Soybean}
+Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
+Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
+Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\, to the degree that some researchers \cite{hymowitz1980taxonomy} consider the two to be conspecific.
+Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
+The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
+Presently, however, there is little or no evidence that natural wild-to-crop introgressive hybridization has had an adaptive impact on cultivated soybeans.
+
+\subsection*{Soybean}
+Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
+Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
+Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
+Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
+The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
+
+\subsection*{Palm}
+
+Oil palm (Elaeis guineensis) of western Africa has been crossed with the New-World species E. oleifera, the only other species in the genus.
+The two species differ in vegetative and reproductive morphology and in ecological adaptations.
+The hybrid appears to have partial resistance to pathogenic fungal Phytophthora palmivora, native to Latin America \cite{torres2010first}\.
+The F1 hybrids have moderately diminished viability and fertility, but the reproductive barrier is not fully developed \cite{hardon1969interspecific}\.
+Since there is only one wild relative in the same genus, and has not been sympatric with E.
+guineensis for a long time it seems unlikely that adaptive introgression has played a significant role in the domestication of this plant.
+
+\subsection*{Sugar Beet}
+
+Sugar beet was domesticated relatively recently
+Crop-wild weedy hybrids (wild alleles introgressing into crop populations) \cite{ellstrand2003dangerous}\
+Most introgression into sugar beet from wild relatives has been the intentional product of breeding efforts for the purpose of disease resistance (a review of these efforts and outcomes can be found at \cite{panella2007broadening})\.
+
+\subsection*{barley}
+
+Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
+Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
+Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+
 \subsection{Sweet Potato}
@@
 
-Olive
+\subsection*{Cotton}
+Cited (indirectly) in \cite{westengen2005genetic}:
+45 diploid species, 5 tetraploid
+monophyletic
+all tetraploids from New World
+two domestications, G. hirsutum (Mesoamerica/Caribbean) and G. barbadense (S. America/Caribbean)
 
-Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to olive's genetic diversity, but it has not been shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+Cited (indirectly) in \cite{ellstrand1999gene}\.
+Evidence of introgression from domesticated cottons into wild.
+
+Directly found by \cite[wendel1989genetic}\:
+Bidirectional interspecific introgression between cultivated and wild cottons account for a "significant" amount of allele diversity, in both domesticated species.
+This is not shown to be adaptive, though.
+
+\subsection*{Rapeseed}
+
+Domesticated rapeseed (Brassica napus) has many sympatric wild relatives, but reproductive barriers limit the degree of hybridization that occurs.
+\cite{scheffler1994opportunities}\ experimentally tested fertility rates between domesticated rapeseed and wild relatives.
+Also see \cite{ellstrand1999gene}\.
+
+\subsection*{Coconut}
 
 \subsection*{Sorghum}
-Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
-However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
-The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
-The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
-Within Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
-Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}.
 
@@
 
-/subsection*{Soybean}
-Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
-Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
-Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
-Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
-The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
-
 \bibliography{bib_gj.bib}

### 2016-06-28 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
+There are two types of cultivated coconut (Pacific and Indian/Atlantic), corresponding to two separate domestications, with a history of human-facilitated admixture between them \cite{gunn2011independent}\.
+[[See cited references in intro of "Population Genetics, Lethal Yellowing Disease, and Relationships among Mexican and
+Imported Coconut Ecotypes"]]
+
 \subsection*{Sorghum}

### 2016-06-29 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 
+\subsection*{Cotton}
+wendel1989genetic:
+bidirectinal gene flow between diploid species G. arboreum and G. herbaceum
+independent domestication events of G. arboreum and G. herbaceum
+
+The old-world A genome became part of the new-word genome in the polyploidization event; talk about this?
+
+abdalla2001genetic:
+Lots of interbreeding between cultivated lines of cotton, but what about wild relatives?
+
+Cotton tolerates crosses quite well.
+
+fryxell1979natural:
+Introgression from domesticate to wild cottons is prevalent enough to threaten the continued existence of two wild species, G. darwinii and G. tomentosum.
+
+wendel1990allozyme:
+confirms the suggestions (based on morphology) of fryxell1979natural (for G. darwinii)
+
 \bibliography{bib_gj.bib}

### 2016-06-29 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 \subsection*{Cotton}
+
+Gossypium hirsutum is the most important and most agronomically grown cotton today.
+G. hirsutum is from Mexico.
+There are many species in the genus, from around the world, and many have been cultivated.
+
+percy1990allozyme:
+G. barbadense is grown commercially in S. America, where it hybridizes with G. hirsutum as well
+
 wendel1989genetic:
@@
 
-Cotton tolerates crosses quite well.
+Cotton readily tolerates interspecific crossing.
+This leads to a reticulated phylogeny.
 

### 2016-06-29 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 
-Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to olive's genetic diversity, but it has not been shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+Olives (Olea europaea subsp.
+europaea var.
+sativa) were domesticated primarily in the northeastern Levant and then dispersed west along the length of the Mediterranean range of wild (var.
+oleaster) olives \cite{besnard2013complex}\.
+The prevalence of wild olives (var. oleaster) has since been diminished.
+
+Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to cultivated olive's genetic diversity, but it has not been specifically shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+
+Olives have been propogated clonally for much of their history, limiting gene flow, but given the chance, cultivated and wild olives are fully interfertile [[green and wickens 1989]].
+Feral cultivated/wild hybrid oleasters are evidence of such outcrossing events.
 

### 2016-06-30 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 
+Direct evidence of adaptive wild-to-crop introgressive hybridization is at present absent, however.
+
 \subsection*{Sorghum}
@@
 Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
+Wild soybeans outcross at a rate between 9.3 and 19% \cite{fujita1997extent}\.
 Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
@@
 The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
+QTLs for adaptive traits (pest resistance, disease resistance, seed quality, etc.) have been identified and described [[see the highlighted intro paragraph in concibido2003introgression]]
 

### 2016-07-01 crop_adaptation_gene_flow_gj_outline.tex
Update crop_adaptation_gene_flow_gj_outline.tex

@@
 
+And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
+It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
 Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
+\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
 Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
@@
 [[see austin1988taxonomy for a summary of the genus Ipomoea, wild relatives and such.]]
+
+From disentangling:
+domesticate/wild hybridization is still possible, but with diminished vigor [[50,55]]
+Gene flow would be diminished during cultivation
+
+http://link.springer.com/chapter/10.1007/978-0-387-92765-7_3
 

### 2016-07-06 TiG_prospectus.tex
Li added methods part and some references

@@
 \usepackage[numbers,sort]{natbib}
+\usepackage{amsmath}
 \bibliographystyle{ieeetr}
@@
 \item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.}
+
+The methods introduced here do not include those marginally estimating introgression/migration rate as a component in the complete demographic history, such as Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference (dadi) \cite {gutenkunst2009}, sorts of isolation with migration model \cite{hey2004} and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}.
+The collection of methods described different layers of genetic variation: diversity and divergence, haplotype variation and phylogenetic relationships.
+We focus here on analytical prospected and limitations and introduce only a few representatives of each category.
+
+First, genomic regions of introgression were expected to show a low level of divergence.
+ F_{st} and d_{XY} as well as its derivates, such as G_{min} \cite{geneva2015} and RND_{min}\cite{rosenzweig2016}, are representatives of such statistics. 
+ The former two are insensitive to rare migrants, thus limited in detecting recent introgression; while the latter two  overcame the disadvantages.
+
+ \begin{equation}
+    G_{min} = \frac{d_{min}}{d_{XY}}
+ \end{equation}
+ where d_{min} is the minimum sequence distance between haplotypes in species X and Y.
+
+ \begin{equation}
+ 	RND_{min} = \frac{d_{min}}{d_{out}}
+ \end{equation}
+where d_{out} equals to $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup (O).
+In addition, RND_{min} is sensitive to variable mutation rate, which reflects on the branch length to the outgroup taxa.
+
+Second, local ancestry reconvolution (also known as chromosome painting) can identify the genomic regions that come from different source populations \cite{schraiber2015}. 
+Such method usually takes phased haplotypes as input and some can even take multiple source populations, such as MULTIMIX \cite{churchhouse2013}.
+
+Third, ABBA-BABA statistics (also known as D-statistics) and its derivatives are widely applied to search for signs of introgression in genomic patterns of shared derived variants between populations or species.
+
 \item{Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; \textbf{Figure 1}), introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}}.
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}}.
 \end{enumerate}

### 2016-07-09 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

--- 2016-06-28 Crop_adaptation_gene_flow_gj_outline_shortened.tex
@@
 
-\subsection*{Introduction}
-
- "Whenever sympatric populations of crops and their wild relatives show a similar trait, there will be several possible explanations: (1) retention of an ancestral trait, (2) convergent or parallel evolution, and (3) introgression.
-Perhaps even more difficult is establishing the direction of introgression; does gene flow go from crop into wild species, the reverse, or in both directions?" and "For this reason, introgression may be best studied between crops and their more distant wild relatives where the probability of finding taxon-specific genetic markers is increased." \cite{doebley1990molecular}
-
 \subsection*{Maize}
 
-\subsection*{wheat}
+Supplement Li's work with the following ideas:
 
-The convoluted domestication of wheat is marked by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
-Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
-Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
-The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}\.
-Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
+Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it still grows in sympatry with many of its wild relatives \cite{hufford2013}.
 
-\subsection*{barley}
+Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite{matsuoka2002single} from teosinte.
 
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
-Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
-Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
 
-\subsection*{rice}
+In particular, \emph{Zea mays} ssp.
+\emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
+\emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
 
-Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
-sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
-Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
-Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
-Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
+Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
+[[Li's and my bib tags might be in conflict, e.g.
+wilkes1977 is likely my wilkes1977hybridization.]]
+.
+ Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
+
+Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
+
+However, \cite{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
+The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled.
+Among the nine regions, three spans the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
+By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
+Growth chamber experiments with maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
+
+Gene flow between \emph{mexicana} and maize is low enough that only fitness-neutral or advantageous alleles are likely to be passed \cite{slatkin1987gene}.
+
+From \cite{baltazar2005pollination}
+Gene flow is probably greater from teosinte to maize.
+Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte and slow rates of introgression.
+Gene flow from maize to teosinte occurs most easily when teosinte pollinates maize
+[[what paper can i cite to mention the gene tb1 and its role in hybridization, viz a viz restricting directionality of teosinte pollen into maize
+
+From \cite{jiang1999genetic}, {bitocchi2009introgression}
+Maize-maize introgression
+Research on introgression between highland and lowland maize shows adaptive traits being transferred
+
+\cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
+\cite{hammer1987collection}\ cites an example in Italy of one domesticated maize adapting to coastal environment by introgressions from another domesticated maize.
 
@@
 
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
+The northern members of the polyphyletic \emph{S.
+brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
+tuberosum} has a polyphyletic origin.
+Also, these northern members of the \emph{S.
+breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
+bukasovii} \cite{spooner2005single}.
+
+Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
+Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
+These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
+However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
+Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
+
 Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}\.
+The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
@@
 
-\subsection*{Sugarcane}
+\subsection*{sunflower}
 
-The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
+The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
+The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
+This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.
+Domesticated sunflower lives in sympatry with wild relatives like \emph{H.
+petiolaris} and \emph{H.
+bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
+Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
+The species \emph{H.
+annuus} shows signs of persistent introgressive hybridization with \emph{H.
+petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
+The species \emph{H.
+annuus} adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
+debilis ssp.
+cucumerifolius}), gaining advantageous alleles (these hybrids are now called \emph{H.
+annuus ssp.
+texanus}, \cite{kim1999genetic, heiser1951hybridization, rieseberg1999hybrid, rieseberg1990helianthus}.
 
-From \cite{ellstrand1999gene}\:
-33 - daniels1987taxonomy
-    "genetically based traits" evidence for hybrid origin of "many cultivars"
-    "chromosomal evidence" supports "spontaneous intergeneric hybridization" origin of "particular accessions of wild sugarcane relatives"
-    Erianthus maximus might be a hybrid of S. officinarium and Miscanthus (a wild relative genus), based on morphological and chromosomal evidence.
-104 - roach1995sugar
-    "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
-
-\subsection*{Cassava}
-
-Although cultivated cassava (Manihot esculenta) is interfertile with some of its wild relatives (the putative progenitor species M.
-flabellifolia, for example, \cite{roa1997aflp}\), comparisons of single-copy nuclear gene G3pdh sequences provide little evidence to suggest that natural introgressive hybridization occurs, and if it does, it would appear to be in the direction of cultivar to wild relative \cite{olsen1999evidence}\.
-Clonal propogation and allopatry are listed among the barriers to gene flow between cassava and its wild relatives [[but see tovar2015diversity for discussion on elevated genetic diversity in M.
-esculenta, gene flow between species, occasional sexual reproduction, and trade between farmers]].
-Breeders have introgressed genes for protein and disease resistance genes into cassava from M. flabellifolia \cite{akinbo2008introgression}\.
-
-\subsection{Sweet Potato}
-
-Sweet potato (Ipomoea batatas) likely arose from multiple domestications, in Central America and in South America, which were secondarily hybridized upon introduction \cite{roullier2013disentangling}.
-[[see austin1988taxonomy for a summary of the genus Ipomoea, wild relatives and such.]]
-
-Olive
-
-Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to olive's genetic diversity, but it has not been shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+\emph{Helianthus} has several genes for downy mildew resistance.
+Each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
+argophyllus}, \emph{H.
+tuberosus}, and \emph{H.
+praecox}) and have been successfully bred into modern \emph{H.
+annuus} \cite{miller1991inheritance}\.
+PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
+argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
+Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
 
@@
 
+\subsection*{rice}
+
+There are two main cultigens of rice; Asian rice (\emph{Oryza sativa}) and African rice (\emph{O.
+glaberrima}), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
+The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O.
+sativa}, 8,200-13,500 years ago, from wild \emph{O.
+rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
+perennis} complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for \emph{O.
+glaberrima}, 2-3,000 years ago, from a wild ancestor, perhaps \emph{O.
+barthii} \cite{ricepedia}\ or \emph{O.
+breviligulata} \cite{oka2012origin}\.
+Asian rice has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
+Indica and Japonica are likely the result of independent domestications from separate \emph{O.
+rufipogon} populations in India/Indochina and southern China, respectively \cite{londo2006phylogeography}\.
+These subspecies display adaptations to the environmental coniditions corresponding to the differentiated geographical locations they inhabit \cite{khush2003classifying}.
+
+Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia} and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
+The greater genetic diversity within \emph{O.
+sativa} is likely due to introgression with wild relatives both during domestication and upon the dispersal of \emph{O.
+sativa} into new environments and sympatry with new relatives \cite{second1982origin}.
+Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
+sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
+Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
+Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
+Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
+
+\subsection*{wheat}
+
+The first domesticated wheat was einkorn wheat (\emph{Tritium monococcum}), domesticated from wild einkorn (\emph{T.
+m.
+boeoticum}) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
+Subsequently, emmer wheat (\emph{T.
+dicoccum}) was domesticated from hybrids of diverse populations of wild emmer wheat (\emph{T.
+dioccoides}), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}\.
+Two distinct taxa of \emph{T.
+dioccoides} exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
+The tetraploid durum wheat (\emph{T.
+turgidum}) is the result of hybridization between red wild einkorn wheat (\emph{T.
+urartu}) and wild goatgrass (\emph{Aegilops speltoides}) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (\emph{A.
+tauschii} produced the allohexaploid common bread wheat (T.
+aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
+This hexaploidization event probably occured twice, once in southeast Turkey/northern Syria and, more recently, in Iran \cite{giles2006gludy}\.
+Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}.
+
+The convoluted domestication of wheat is marked by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
+Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
+Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
+The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}\.
+Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
+Modern breeding efforts between wild and domesticated wheat have shown some success.
+For example, some researchers have experimentally bred resistance traits into durum wheat (stem rust resistance from wild einkorn, \cite{gerechter1971transfer}\, and powdery mildew resistance from \emph{T.
+turgidum var.
+dicoccoides}, \cite{blanco2008molecular}.
+
+\subsection*{Cotton}
+
+Gossypium hirsutum is the most important and most agronomically grown cotton today.
+G. hirsutum is from Mexico.
+There are many species in the genus, from around the world, and many have been cultivated.
+
+percy1990allozyme:
+G. barbadense is grown commercially in S. America, where it hybridizes with G. hirsutum as well
+
+wendel1989genetic:
+bidirectinal gene flow between diploid species G. arboreum and G. herbaceum
+independent domestication events of G. arboreum and G. herbaceum
+
+The old-world A genome became part of the new-word genome in the polyploidization event; talk about this?
+
+abdalla2001genetic:
+Lots of interbreeding between cultivated lines of cotton, but what about wild relatives?
+
+Cotton readily tolerates interspecific crossing.
+This leads to a reticulated phylogeny.
+
+fryxell1979natural:
+Introgression from domesticate to wild cottons is prevalent enough to threaten the continued existence of two wild species, G. darwinii and G. tomentosum.
+
+wendel1990allozyme:
+confirms the suggestions (based on morphology) of fryxell1979natural (for G. darwinii)
+
+\subsection*{tomato}
+
+The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
+Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{Solanum lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
+\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
+\cite{nesbitt2002comparative}, however, conclude that \emph{v.
+cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
+esculentum} and \emph{L.
+pimpinellifolium}.
+Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
+
+And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
+It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
+Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
+Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
+\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
+Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
+hirsutum f.
+glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
+hirsutum}.
+Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
+
+\subsection*{Common Bean}
+
+Common bean (\emph{Phaseolus vulgaris}) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
+The domesticated common bean progenitor species (\emph{P.
+vulgaris var.
+aborigineus}) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
+Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}\), rates of gene flow between it and wild relatives are expected to be lower.
+These expectations are supported by evidence provided by \cite{papa2003asymmetry}\.
+However, hybrid populations can still be found in sympatry with wild and domesticated common bean, forming domesticate-hybrid-weedy complexes.
+
+Some evidence for introgression exists in the direction of domesitcated to wild common bean, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
+Wild common beans with this introgression are also occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
+Also, \cite{kwak2009structure} found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
+The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}.
+Farmer practices may also be diminishing introgressive gene flow.
+Though some hybrids show heterosis, others (especially F1 hybrids) do not, expressing smaller plant size, smaller seed size, and increased mortality \cite{gutierrez1985heterosis, paredes1995extensive}\.
+Because F1 hybrids are intermediate in seed size and color, farmers may effectively identify and avoid hybrid stock.
+
+Though limited, there is also some evidence of natural introgression of wild genes into domesticated common bean.
+\cite{papa2003asymmetry}\ was unable to determine how much of the genetic variability of domesticated common bean landraces originated from weedy or wild types, but point to the allele for 'L' phaseolin as a possible example of an allele that has been introgressed from wild common beans into domesticated.
+Experimentally, common leaf blight and white mold resistances have been bred into common bean from scarlet runner bean (\emph{P.
+coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}.
+
+Olive
+
+Olives (Olea europaea subsp.
+europaea var.
+sativa) were domesticated primarily in the northeastern Levant and then dispersed west along the length of the Mediterranean range of wild (var.
+oleaster) olives \cite{besnard2013complex}\.
+The prevalence of wild olives (var. oleaster) has since been diminished.
+
+Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to cultivated olive's genetic diversity, but it has not been specifically shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.
+
+Olives have been propogated clonally for much of their history, limiting gene flow, but given the chance, cultivated and wild olives are fully interfertile [[green and wickens 1989]].
+Feral cultivated/wild hybrid oleasters are evidence of such outcrossing events.
+
+Direct evidence of adaptive wild-to-crop introgressive hybridization is at present absent, however.
+
+[[Palm]]
+
+Sugarcane
+
+The genus Saccharum is commonly said to consist of six species (wild species S.
+spontaneum and S.
+robustum, cultivated species S.
+officinarum, S.
+barberi, S.
+sinense, and S.
+edule) \cite{brandes1958origin, jackson2005breeding}\.
+Others \cite{irvine1999saccharum}\ argue that there are fewer distinct species, as the six "species" are interfertile, and some appear to be of hybrid origins.
+Irvine  \cite{irvine1999saccharum}\ suggests that there are two species; S. spontaneum, the putatively ancestral form, and all of the others, adopting the name S. officinarum.
+These two are distinguished by differences in morphology and cytoplasmic DNA sequence.
+Sugarcane was cultivated in Papau, New Guinea about 8000 years ago \cite{keyroyalbotanicalgardens, daniels1987taxonomy, artschwager1958sugarcane}\.
+The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
+
+From \cite{ellstrand1999gene}\:
+33 - daniels1987taxonomy
+    "genetically based traits" evidence for hybrid origin of "many cultivars"
+    "chromosomal evidence" supports "spontaneous intergeneric hybridization" origin of "particular accessions of wild sugarcane relatives"
+    Erianthus maximus might be a hybrid of S. officinarium and Miscanthus (a wild relative genus), based on morphological and chromosomal evidence.
+104 - roach1995sugar
+    "spontaneous hybridization between cultivated and wild Saccharum in Australasia and islands of the Indian Ocean" in origin of "many cultivars"
+115 - sobral1994phylogenetic
+
+117 - stevenson1965genetics
+    "certain wild canes from Java and Mauritius are hybrids between" wild (S. spontaneum) and cultivated (S. officinarum)
+
+ from roach1995sugar:
+ Spontaneous hybridization in Australasia with wild relatives.
+ No evidence yet of adaptive introgression, though.
+
+Cassava
+
+Although cultivated cassava (Manihot esculenta) is interfertile with some of its wild relatives (the putative progenitor species M.
+flabellifolia, for example, \cite{roa1997aflp}\), comparisons of single-copy nuclear gene G3pdh sequences provide little evidence to suggest that natural introgressive hybridization occurs, and if it does, it would appear to be in the direction of cultivar to wild relative \cite{olsen1999evidence}\.
+Clonal propogation and allopatry are listed among the barriers to gene flow between cassava and its wild relatives [[but see tovar2015diversity for discussion on elevated genetic diversity in M.
+esculenta, gene flow between species, occasional sexual reproduction, and trade between farmers]].
+Breeders have introgressed genes for protein and disease resistance genes into cassava from M. flabellifolia \cite{akinbo2008introgression}\.
+
+Sweet Potato
+
+Sweet potato (Ipomoea batatas) likely arose from multiple domestications, in Central America and in South America, which were secondarily hybridized upon introduction \cite{roullier2013disentangling}.
+[[see austin1988taxonomy for a summary of the genus Ipomoea, wild relatives and such.]]
+
+From disentangling:
+domesticate/wild hybridization is still possible, but with diminished vigor [[50,55]]
+Gene flow would be diminished during cultivation
+
+http://link.springer.com/chapter/10.1007/978-0-387-92765-7_3
+
 /subsection*{Soybean}
@@
 The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
+QTLs for adaptive traits (pest resistance, disease resistance, seed quality, etc.) have been identified and described [[see the highlighted intro paragraph in concibido2003introgression]]
+
+\subsection*{barley}
+
+Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
+Barley (\emph{Hordeum vulgare subsp.
+vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp.
+spontaneum} in the Fertile Crescent and once by \emph{subsp.
+spontaneum var.
+agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
+
+Variety agriocrithon (\emph{H. vulgare ssp. vulgare f. agriocrithon}) is genetically diverse, and is found throughout much of the range of Barley.
+Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from \emph{ssp.
+spontaneum}, perhaps more than once (in Israel as well as in Tibet).
+Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
+These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
+
+The barley domestication process has reduced the number of alleles in the domesticate to only 40% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
+Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield,  \cite{von2006ab, handley1994chromosome}\.
+Although the conditions of barley domestication would seem to allow (if not promote) natural adaptive introgression between barley and its wild relatives, there is little evidence of this genetic interaction at present.
+Owing to this sparcity of evidence is the low rate of outcrossing of \emph{spontaneum}, estimated by \cite{brown1978outcrossing} to be 1.6% in Israeli populations.
+There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}\.
+Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.
+
+\subsection*{rye}
+
+Rye domestication has recieved comparatively little academic interest.
+Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}.
 

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 \subsection*{Maize}
-
-Supplement Li's work with the following ideas:
 
@@
 
-Gene flow between \emph{mexicana} and maize is low enough that only fitness-neutral or advantageous alleles are likely to be passed \cite{slatkin1987gene}.
-
-From \cite{baltazar2005pollination}
-Gene flow is probably greater from teosinte to maize.
-Certain morphological and flower timing characterists promote genetic isolation between maize and teosinte and slow rates of introgression.
-Gene flow from maize to teosinte occurs most easily when teosinte pollinates maize
-[[what paper can i cite to mention the gene tb1 and its role in hybridization, viz a viz restricting directionality of teosinte pollen into maize
-
-From \cite{jiang1999genetic}, {bitocchi2009introgression}
-Maize-maize introgression
-Research on introgression between highland and lowland maize shows adaptive traits being transferred
-
 \cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
-\cite{hammer1987collection}\ cites an example in Italy of one domesticated maize adapting to coastal environment by introgressions from another domesticated maize.
 

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-Maize is a classical model in crop species to study the issue of adaptive introgresssion from wild relatives, as it still grows in sympatry with many of its wild relatives \cite{hufford2013}.
-
-Maize was domesticated about 10,000 years ago in Mexico \cite{smith1997initial} \cite{hufford2012comparative} in the Balsas River valley \cite{matsuoka2002single} from teosinte.
-
+Maize is an ideal model system for the study of adaptive introgresssion from wild relatives.
 Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
-
+ Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
+ Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
 In particular, \emph{Zea mays} ssp.
@@
 \emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
-
 Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
@@
 wilkes1977 is likely my wilkes1977hybridization.]]
-.
- Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
-
-Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
 
@@
 Growth chamber experiments with maize populations with introgression from \emph{mexicana} on chromosome 4 (associated with QTL controlling pigment density and macrohairs) and 9 (overlapped with QTL for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
-
-\cite{bitocchi2009introgression} has a good review of the history of domestication, introduction to Europe, and further development.
 

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
 The northern members of the polyphyletic \emph{S.
@@
 However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
-Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
 
 Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}\.
-The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
+The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\ (fueling the assertion that these potatoes comprise a single species).
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
@@
 
-The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
-The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
-This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}\.
 Domesticated sunflower lives in sympatry with wild relatives like \emph{H.
@@
 
-\emph{Helianthus} has several genes for downy mildew resistance.
-Each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
 Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
@@
 Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
-However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
+
+However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.s
 The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
 The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
-Within Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
+
 Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}.

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.s
+However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
 The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
+Within the sub-genus Eu-Sorghum, there are currently all domesticated, weedy, and closely-related wild relatives, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
 
@@
 The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
-The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
+The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that  \cite{USDAARS2007, garber1950cytotaxonomic}.
 
@@
 \cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution.
-Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\.
+Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\, but these hybrids express diminished fertility.
 Rates of outcrossing within sorghum range from 0-30% \cite{doggett1988sorghum}\.
 Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
-These hybrids express diminished fertility.
-Members of \emph{Sorghum bicolor} have adapted to a broad range of altitudinal, precipitation, and temperature clines across Africa and around the world \cite{po1982sorghum, CWR}\.
-Breeding efforts have focused on incorporating exotic germplasm (wild sorghums as well as relatives from other genera \cite{de1976cytogenetics} for its adaptations to biotic and abiotic stresses \cite{reddy2006current, po1982sorghum, johnson1979breeding}.
+
+Members of \emph{Sorghum bicolor} have adapted to a broad range of altitudinal, precipitation, and temperature clines
 Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}.
@@
 
-The first domesticated wheat was einkorn wheat (\emph{Tritium monococcum}), domesticated from wild einkorn (\emph{T.
-m.
-boeoticum}) in or around the Karacadag mountains of southeast Turkey \cite{heun1997site, harlan1966distribution, ozkan2002aflp}\.
-Subsequently, emmer wheat (\emph{T.
-dicoccum}) was domesticated from hybrids of diverse populations of wild emmer wheat (\emph{T.
-dioccoides}), also in the Fertile Crescent \cite{lev2000cradle, civavn2013reticulated, luo2007structure, ozkan2002aflp}\.
-Two distinct taxa of \emph{T.
-dioccoides} exist (a western one and a central-eastern one), but the central-eastern one was the progenitor species from which modern wheat was domesticated \cite{ozkan2005reconsideration}
-The tetraploid durum wheat (\emph{T.
-turgidum}) is the result of hybridization between red wild einkorn wheat (\emph{T.
-urartu}) and wild goatgrass (\emph{Aegilops speltoides}) \cite{mcfadden1946origin, petersen2006phylogenetic}\, and a further hybridization with Tausch's goatgrass (\emph{A.
-tauschii} produced the allohexaploid common bread wheat (T.
-aestivum), the most agronomically important race of wheat \cite{salamini2002genetics, hancock2012plant, dvorak2006molecular}\.
-This hexaploidization event probably occured twice, once in southeast Turkey/northern Syria and, more recently, in Iran \cite{giles2006gludy}\.
 Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}.
+[[I should really go into more detail here.]]
 

### 2016-07-10 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
+\subsection*{tomato}
+
+And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
+It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
+Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
+Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
+\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
+Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
+hirsutum f.
+glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
+hirsutum}.
+Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
+[[Dig deeper into these sources - evidence?  and naturally?]]
+
 \subsection*{Cotton}
@@
 
-fryxell1979natural:
-Introgression from domesticate to wild cottons is prevalent enough to threaten the continued existence of two wild species, G. darwinii and G. tomentosum.
-
-wendel1990allozyme:
-confirms the suggestions (based on morphology) of fryxell1979natural (for G. darwinii)
-
-\subsection*{tomato}
-
-The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
-Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{Solanum lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
-\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
-\cite{nesbitt2002comparative}, however, conclude that \emph{v.
-cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
-esculentum} and \emph{L.
-pimpinellifolium}.
-Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
-
-And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
-It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
-Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
-Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
-Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
-hirsutum f.
-glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
-hirsutum}.
-Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
-
 \subsection*{Common Bean}
 
-Common bean (\emph{Phaseolus vulgaris}) was domesticated from a wild common bean at least twice, once in the Lerma-Santiago Basin of Mesoamerica and once in the southern Andes \cite{kwak2009putative}\.
-The domesticated common bean progenitor species (\emph{P.
-vulgaris var.
-aborigineus}) and other wild relative species can be found across the breadth of this region and beyond, from Mexico to Argentina, inhabiting a broad range of environments \cite{debouck1988implications, CWR}\.
 Because common bean is generally self-pollinating (outcrossing at an estimated rate under 3%, \cite{ramalho2006}\), rates of gene flow between it and wild relatives are expected to be lower.
 These expectations are supported by evidence provided by \cite{papa2003asymmetry}\.
-However, hybrid populations can still be found in sympatry with wild and domesticated common bean, forming domesticate-hybrid-weedy complexes.
+However, hybrid populations can still be found in sympatry with wild and domesticated common bean, forming domesticate-hybrid-weedy complexes [[perhaps use kwak2009structure?
+would have to check again]].
 
 Some evidence for introgression exists in the direction of domesitcated to wild common bean, leading to increased seed size, an adaptation that permitted their spread to higher elevations with cooler climates \cite{debouck1993genetic}\.
-Wild common beans with this introgression are also occasionally harvested along with cultivated common bean and consumed by subsistence farmers \cite{papa2003asymmetry}\.
 Also, \cite{kwak2009structure} found via STRUCTURE analysis that certain Andean wild common bean accessions were possibly the result of hybridizations with domesticated common bean.
-The disparity in gene flow directionality may be due in part to the large population size of common beans in a cultivated field relative to the population size of wild common beans within pollen dispersal range (a distance that may be quite small, \cite{papa2003asymmetry}.
-Farmer practices may also be diminishing introgressive gene flow.
-Though some hybrids show heterosis, others (especially F1 hybrids) do not, expressing smaller plant size, smaller seed size, and increased mortality \cite{gutierrez1985heterosis, paredes1995extensive}\.
-Because F1 hybrids are intermediate in seed size and color, farmers may effectively identify and avoid hybrid stock.
 
@@
 coccineus}) \cite{park1987transfer, schwartz2006inheritance}\, as was resistance to weevils (\emph{Zabrotes subfasciatus} and \emph{Acanthoscelides obtectus}) from Mexican wild bean species \cite{kornegay1991inheritance}\, but it may take careful, larger-scale breeding programs to exploit heterotic hybrids for agronomic gain \cite{paredes1995extensive}.
+
+[[Olive, Sugarcane, Soybeans, Barley show similar patterns (ie promising conditions but no direct evidence).]]
 
@@
 
-[[Palm]]
-
 Sugarcane
-
-The genus Saccharum is commonly said to consist of six species (wild species S.
-spontaneum and S.
-robustum, cultivated species S.
-officinarum, S.
-barberi, S.
-sinense, and S.
-edule) \cite{brandes1958origin, jackson2005breeding}\.
-Others \cite{irvine1999saccharum}\ argue that there are fewer distinct species, as the six "species" are interfertile, and some appear to be of hybrid origins.
-Irvine  \cite{irvine1999saccharum}\ suggests that there are two species; S. spontaneum, the putatively ancestral form, and all of the others, adopting the name S. officinarum.
-These two are distinguished by differences in morphology and cytoplasmic DNA sequence.
-Sugarcane was cultivated in Papau, New Guinea about 8000 years ago \cite{keyroyalbotanicalgardens, daniels1987taxonomy, artschwager1958sugarcane}\.
-The history of sugarcane is marked with phases of concerted interspecific breeding efforts, primarily but not exclusively between S. officinarum and S. spontaneum \cite{roach1989origin}\.
 
@@
 115 - sobral1994phylogenetic
-
 117 - stevenson1965genetics
@@
  No evidence yet of adaptive introgression, though.
-
-Cassava
-
-Although cultivated cassava (Manihot esculenta) is interfertile with some of its wild relatives (the putative progenitor species M.
-flabellifolia, for example, \cite{roa1997aflp}\), comparisons of single-copy nuclear gene G3pdh sequences provide little evidence to suggest that natural introgressive hybridization occurs, and if it does, it would appear to be in the direction of cultivar to wild relative \cite{olsen1999evidence}\.
-Clonal propogation and allopatry are listed among the barriers to gene flow between cassava and its wild relatives [[but see tovar2015diversity for discussion on elevated genetic diversity in M.
-esculenta, gene flow between species, occasional sexual reproduction, and trade between farmers]].
-Breeders have introgressed genes for protein and disease resistance genes into cassava from M. flabellifolia \cite{akinbo2008introgression}\.
-
-Sweet Potato
-
-Sweet potato (Ipomoea batatas) likely arose from multiple domestications, in Central America and in South America, which were secondarily hybridized upon introduction \cite{roullier2013disentangling}.
-[[see austin1988taxonomy for a summary of the genus Ipomoea, wild relatives and such.]]
-
-From disentangling:
-domesticate/wild hybridization is still possible, but with diminished vigor [[50,55]]
-Gene flow would be diminished during cultivation
-
-http://link.springer.com/chapter/10.1007/978-0-387-92765-7_3
 
@@
 Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
-Barley (\emph{Hordeum vulgare subsp.
-vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp.
-spontaneum} in the Fertile Crescent and once by \emph{subsp.
-spontaneum var.
-agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
 
 Variety agriocrithon (\emph{H. vulgare ssp. vulgare f. agriocrithon}) is genetically diverse, and is found throughout much of the range of Barley.
-Some have suggested that agriocrithon may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from \emph{ssp.
-spontaneum}, perhaps more than once (in Israel as well as in Tibet).
-Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and West-Central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
-These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
 
@@
 
-\subsection*{rye}
-
-Rye domestication has recieved comparatively little academic interest.
-Although there is little evidence for or against the role of natural introgression in rye domestication, rye (\emph{Secale cereale}) has wild relatives outside its center of domestication in or around the Fertile Crescent \cite{vavilov1928geographical}\, and the species contains domesticated, weedy, and wild rye subspecies \cite{khush1961cytogenetic}.
-
 \bibliography{bib_gj.bib}

### 2016-07-11 TiG_prospectus.tex
Li added methods

@@
 Second, local ancestry reconvolution (also known as chromosome painting) can identify the genomic regions that come from different source populations \cite{schraiber2015}. 
-Such method usually takes phased haplotypes as input and some can even take multiple source populations, such as MULTIMIX \cite{churchhouse2013}.
+Such method usually takes phased haplotypes as input and an exception is MULTIMIX \cite{churchhouse2013}, which can take both phased and unphased haplotypes as input.
+and unveil the genomic ancestry from multiple source populations.
 
 Third, ABBA-BABA statistics (also known as D-statistics) and its derivatives are widely applied to search for signs of introgression in genomic patterns of shared derived variants between populations or species.
+Based on D-statistics, $\hat{f_{d}}$ \cite{Martin2015} and five-taxon D statistics \cite{pease2015} are developed to localize the genomic regions under introgression. 
+The former took allele frequencies from each population/species and the latter detected introgression from the localized phylogenetic pattern. 
+Their strength in identifying specific regions enabled us to explore the relationships between introgression and recombination rate / gene density / distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
 

### 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-Maize is an ideal model system for the study of adaptive introgresssion from wild relatives.
-Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have observed desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
- Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
- Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
+Maize (\emph{Zea mays} ssp. \emph{mays}) is an ideal model system for the study of adaptive introgresssion from wild relatives.
+The wild relatives and progenitors of maize are collectively known as teosinte (\emph{Zea mays}).
 In particular, \emph{Zea mays} ssp.
-\emph{mexicana} (referred to \emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
-\emph{parvaglumis} has also been shown to hybridize with maize, \cite{wilkes1977hybridization}).
+\emph{mexicana} (\emph{mexicana} hereafter) and Mexican highland maize have become a well-studied pair (although ssp.
+\emph{parvaglumis} has also been shown to hybridize with maize \cite{wilkes1977hybridization}).
 Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite{wilkes1977, lauter2004, doebley1984} and molecular analysis \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
@@
 wilkes1977 is likely my wilkes1977hybridization.]]
+Mexican farmers have long been aware of maize-teosinte hybrids, and in some places, such as the Nobogame valley, farmers have reported observing desirable traits passed to their crops through the hybrid plants \cite{wilkes1977hybridization} \cite{lumholtz1902unknown} \cite{wilkes1970teosinte}.
+John Doebley (\cite{doebley1987patterns}) found that north-central high-elevation Mexican landraces (Apachito, Arrocillo Amarillo, Azul, Celaya, Chaloqueno, Conico, Conico Norteno, Gordo, Harinoso de Ocho Occidentales)[[special characters]] show evidence of introgression of two alleles from \emph{mexicana}, and suspected that they might have fitness advantages for adapting to highland environmental conditions.
+Population structure analyses by \cite{matsuoka2002single} estimate that \emph{mexicana} gene flow has contributed to the gene pools of the Cacahuacintle, Palomero de Jalisco, and Palomero Taloqueno landraces, but also into (non-landrace) Mexican maize at the same elevation.
 
-However, \cite{hufford2013} for the first time revealed the evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
+However, \cite{hufford2013} was the first to reveal genomic evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize.
 The authors identified nine genomic regions, which showed evidence of introgression from \emph{mexicana} to maize in both the HAPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the total nine pairs sampled.

### 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 The northern members of the polyphyletic \emph{S.
-brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
+brevicaule} wild potato complex have been identified as likely progenitors of \emph{Solanum tuberosum}, the modern potato \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
 tuberosum} has a polyphyletic origin.
@@
 Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
-Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
+Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridization \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
 These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
@@
 Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
-Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
+Cultivated diploid potatoes have been shown to be introgressed with genes from \emph{S.
 sparsipilum} \cite{rabinowitz1990high}\ and \emph{S.
 megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
-Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
+Drought- and frost-resistance genes introgressed from wild relatives facilitated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
 Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.

### 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-\subsection*{potato}
+\subsection*{Potato}
 
@@
 
-\subsection*{sunflower}
+\subsection*{Sunflower}
 
-Domesticated sunflower lives in sympatry with wild relatives like \emph{H.
+Domesticated sunflower \emph{Helianthus annuus} lives in sympatry with wild relatives like \emph{H.
 petiolaris} and \emph{H.
@@
 Wild sunflowers are locally adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}\.
-The species \emph{H.
-annuus} shows signs of persistent introgressive hybridization with \emph{H.
+Domesticated sunflower shows signs of persistent introgressive hybridization with \emph{H.
 petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}\.
-The species \emph{H.
-annuus} adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
+Domesticated sunflower adapted to the environment in Texas by hybridizing with cucumberleaf sunflower (\emph{H.
 debilis ssp.
@@
 
-\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+There are several downy mildew resistance genes in \emph{Helianthus}, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
 Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
@@
 PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
-argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
+argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others alleles (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
 Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
@@
 \subsection*{Sorghum}
-Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
-Within the sub-genus Eu-Sorghum, there are currently all domesticated, weedy, and closely-related wild relatives, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
-
+The sub-genus Eu-Sorghum contains all current domesticated, weedy, and closely-related wild relatives, five cultivated species (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
 However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.

### 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 \subsection*{Sorghum}
-The sub-genus Eu-Sorghum contains all current domesticated, weedy, and closely-related wild relatives, five cultivated species (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
-However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
-The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
-The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that  \cite{USDAARS2007, garber1950cytotaxonomic}.
-
+The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
+Within the sub-genus Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
 Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}.
@@
 \cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution.
-Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\, but these hybrids express diminished fertility.
-Rates of outcrossing within sorghum range from 0-30% \cite{doggett1988sorghum}\.
-Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
-
-Members of \emph{Sorghum bicolor} have adapted to a broad range of altitudinal, precipitation, and temperature clines
+Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\, but these hybrids exhibit diminished fertility.
+Rates of outcrossing within sorghum range from 0-30% \cite{doggett1988sorghum}\, and Natural introgression between wild and domesticated sorghum has been documented in both directions \cite{kuhlman2006interspecific, aldrich1992restriction, aldrich1992patterns, doggett1988sorghum, baker1972migrations}\.
 Resistance to greenbug has been introgressed into cultivated sorghum through modern breeding efforts \cite{johnson1979breeding}.
 
-\subsection*{rice}
+\subsection*{Rice}
 
-There are two main cultigens of rice; Asian rice (\emph{Oryza sativa}) and African rice (\emph{O.
-glaberrima}), which are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
-The centers of rice domestication are not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O.
-sativa}, 8,200-13,500 years ago, from wild \emph{O.
-rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
-perennis} complex" \cite{oka2012origin}\ and the Upper Niger River delta in Mali, Africa for \emph{O.
-glaberrima}, 2-3,000 years ago, from a wild ancestor, perhaps \emph{O.
-barthii} \cite{ricepedia}\ or \emph{O.
-breviligulata} \cite{oka2012origin}\.
-Asian rice has two main subspecies, Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
+The two main cultigens of rice are Asian rice (\emph{Oryza sativa}) and African rice (\emph{O.
+glaberrima}), and they are easily distinguish by ligule length, number of secondary panicle branches, panicle axis thickness, and differences in life cycles \cite{oka2012origin}.
+Within \emph{O.
+sativa}, there are two main subspecies; Indica (with subpopulations indica and aus) and Japonica (with subpopulations temperate japonica, tropical japonica known as javanica, and aromatic) \cite{chang2003origin, glaszmann1987isozymes, ricepedia}.
 Indica and Japonica are likely the result of independent domestications from separate \emph{O.
@@
 
-Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia} and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
+Both Asian and African rice naturally hybridize with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}), and introgression is common \cite{oka2012origin, second1982origin, zhao2010genomic}\.
 The greater genetic diversity within \emph{O.
@@
 sativa} into new environments and sympatry with new relatives \cite{second1982origin}.
-Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
+Resistance alleles to several pests (grassy stunt virus, bacterial blight, brown planthopper, blast) have been introgressed from wild relatives into \emph{O.
 sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
-Beyond investigative experiments, gene flow from wild relatives has been used to produce agronomic rice varieties.
+Beyond investigative endeavours, gene flow from wild relatives has been used to produce agronomic rice varieties.
 Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.

### 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-\subsection*{wheat}
+\subsection*{Tomato}
 
-Each of the three genomes that comprise bread wheat shows genomic signatures of ancestral hybridizations as well \cite{marcussen2014ancient}.
-[[I should really go into more detail here.]]
-
-The convoluted domestication of wheat is marked by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
-Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
-Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
-The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}\.
-Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
-
-\subsection*{tomato}
-
+The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
+Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{S.
+lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
+\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
+\cite{nesbitt2002comparative}, however, conclude that \emph{v.
+cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
+esculentum} and \emph{L.
+pimpinellifolium}.
+Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
 And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
 It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
+
 Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
@@
 
+\subsection*{Wheat}
+
+The convoluted domestication of wheat is marked by many instances of hybridization between morphologically-distinct species within a restrained geographical range.
+Throughout wheat domestication, there has been gene flow between species and between wild and domesticated wheats.
+Emmer wheat, for example, has high genetic diversity, likely due to gene flow from wild emmer \cite{luo2007structure, dvorak2006molecular}\.
+The a, b, c, and f alleles of mildew resistance gene Pm3 may have been introgressed into wheat from wild emmer wheat shortly after domestication (although the e, d, and g alleles were probably formed by de novo mutations) \cite{TPJ:TPJ2772}\.
+Furthermore, wild wheat relatives are can be found in a much broader swath than the domesication center in the Fertile Crescent \cite{CWR}\.
+
 \subsection*{Cotton}
@@
 percy1990allozyme:
-G. barbadense is grown commercially in S. America, where it hybridizes with G. hirsutum as well
+G. barbadense is grown commercially in S. America, where it hybridizes with G. hirsutum as well.
 
 wendel1989genetic:
-bidirectinal gene flow between diploid species G. arboreum and G. herbaceum
-independent domestication events of G. arboreum and G. herbaceum
-
+Bidirectinal gene flow between diploid species G. arboreum and G. herbaceum.
+Independent domestication events of G. arboreum and G. herbaceum.
 The old-world A genome became part of the new-word genome in the polyploidization event; talk about this?
@@
 
+[[I don't suspect that this section will go below this point (I've ordered the subsections in order of decreasing evidence of adaptive natural introgression.]]
 [[Olive, Sugarcane, Soybeans, Barley show similar patterns (ie promising conditions but no direct evidence).]]
 
-Olive
+\subsection*{Olive}
 
@@
 
-Sugarcane
+\subsection*{Sugarcane}
 
@@
 
-/subsection*{Soybean}
+\subsection*{Soybean}
+
 Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
@@
 
-\subsection*{barley}
+\subsection*{Barley}
 

### 2016-07-13 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
+The tomato (\emph{Solanum lycopersicum}) center of domestication is not known with certainty.
+Various authors have proposed Mexico and/or Peru \cite{robertson2007genetic}.
+
 The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
-Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{S.
-lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
+Charles Rick (University of California, August 1978) stating that tomato was domesticated from cherry tomato \emph{S.
+lycopersicum} var.
+\emph{cerasiforme} in the New World (most likely Mexico or Central America) [[http://www.landscapeimagery.com/tomato.html]].
 \cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
-\cite{nesbitt2002comparative}, however, conclude that \emph{v.
-cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
-esculentum} and \emph{L.
-pimpinellifolium}.
+\cite{nesbitt2002comparative}, however, conclude that \emph{v. cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid of wild and cultivated tomatoes.
 Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
@@
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
+\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives (specifically \emph{Lycopersicon pimpinellifolium}) in South America, despite cultivated tomato's proclivity to self-pollinate.
 Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
@@
 hirsutum}.
-Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
+Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.e
 [[Dig deeper into these sources - evidence?  and naturally?]]

### 2016-07-13 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

--- 2016-07-12 Crop_adaptation_gene_flow_gj_outline_shortened.tex
@@
 
-The domestication center for tomato (\emph{Solanum lycopersicum}) is not known with certainty, but \cite{sims1979history} quotes Dr.
-Charles Rick (University of California, August 1978) stating that tomato was domesticated from \emph{S.
-lycopersicum cerasiforme} in the New World (most likely Mexico) [[http://www.landscapeimagery.com/tomato.html]].
-\cite{robertson2007genetic, bai2007domestication} and [[http://www.landscapeimagery.com/tomato.html]] agree with this hypothesis.
-\cite{nesbitt2002comparative}, however, conclude that \emph{v.
-cerasiformae} is not the direct ancestor of the domesticated tomato, but rather a hybrid between wild tomatoes \emph{L.
-esculentum} and \emph{L.
-pimpinellifolium}.
-Widescale recent introgression between wild relatives confound attempts at elucidating phylogenies.
-And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
-It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.
+The tomato (\emph{Solanum lycopersicum}, alternatively \emph{Lycopersicon lycopersicum} or \emph{L. esculentum}) center of domestication is not known with certainty.
+Various authors have proposed Mexico and/or Peru \cite{robertson2007genetic}, but widescale recent introgression between wild relatives confound attempts at elucidating phylogenies \cite{nesbitt2002comparative}.
+[[And yet, \cite{bai2007domestication} claim that "[m]ost likely, no exhange of genetic information with the wild germplasm took place until 1940."
+It seems more likely that such genetic exchange would simply have been unintentional, rather than strictly nonexistent.]]
+
+\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives (specifically \emph{Solanum pimpinellifolium}) in South America, despite cultivated tomato's proclivity to self-pollinate.
+However, these natural introgressions have yet be shown to be adaptive for the plant.
+Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{S.
+habrochaites} f.
+\emph{glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated through breeding with the wild relative \emph{S.
+habrochaites}.
+
+Similar resistance traits has also been shown to have been imparted from \emph{S. chilense}, \emph{S. pimpinellifolium}, and \emph{S. cheesmanii} \cite{hanson2000mapping}.
+[[Dig deeper into these sources - evidence?  and naturally?]]
 
@@
 Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
-\cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives in South America, despite cultivated tomato's proclivity to self-pollinate.
-Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{L.
-hirsutum f.
-glabratum} \cite{banerjee1990transfer} to an introgression on chromosome 11, which originated from the wild relative \emph{L.
-hirsutum}.
-Similar resistance traits has also been shown to have been imparted from \emph{L. chilense}, \emph{L. pimpinellifolium}, and \emph{L. cheesmanii} \cite{hanson2000mapping}.
-[[Dig deeper into these sources - evidence?  and naturally?]]
 

### 2016-07-13 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 \cite{rick1958role} summarizes morphological evidence of natural introgressive hybridization with wild relatives (specifically \emph{Solanum pimpinellifolium}) in South America, despite cultivated tomato's proclivity to self-pollinate.
-However, these natural introgressions have yet be shown to be adaptive for the plant.
-Also, \cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{S.
+However, these natural introgressions have yet been shown to be adaptive for the plant.
+Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
+Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
+\cite{hanson2000mapping} mapped a tomato yellow leaf curl virus resistance gene in an accession of \emph{S.
 habrochaites} f.
@@
 habrochaites}.
-
 Similar resistance traits has also been shown to have been imparted from \emph{S. chilense}, \emph{S. pimpinellifolium}, and \emph{S. cheesmanii} \cite{hanson2000mapping}.
-[[Dig deeper into these sources - evidence?  and naturally?]]
-
-Since the work of Charles Rick in 1940s and 50s \cite{rick1953novel}\, breeding efforts have focused on incorporating and utilizing the relatively diverse genetic germplasm of wild tomato relatives \cite{rick1988tomato, miller1990rflp, rick1982potential}\.
-Wild tomato relatives have been shown to harbor resistance genes for more than 40 agronomically important diseases, and breeders are working to introgress them into cultivated tomato \cite{rick1995utilization}\.
 

### 2016-07-13 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

@@
 
-\cite{po1982sorghum}\ states that natural introgression has been an important force in sorghum evolution.
+\cite{gebrekidan1982utilization}\ states that natural introgression of desirable alleles into domesticated sorghum from sympatric weedy populations occurs and has been occuring for "a long time", and that breeding probrams should be developed to accelerate this gene flow.
 Domesticate-weedy-wild complexes are common when sorghum cultivars are grown in sympatry with wild members of the Eu-Sorghum subgenus \cite{de1978systematics, doggett1968disruptive, baker1972human}\, but these hybrids exhibit diminished fertility.

### 2016-07-27 Table1.tex
adding figure 1 and table 1

\begin{table}
\rowcolors{2}{white}{gray!25}
\begin{center}
\caption{Programs for identifying genomic regions under introgression} \label{tab:tools}
\begin{tabular}{llll}
\\\toprule  
\rowcolor{white}
{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule

\rowcolor{gray!25}
{\emph{\bf chromosome paiting}} &   &   &   \\
\rowcolor{gray!25}
Hapmix	& phased haplotype; reference panel		& Price et al. 2009	&  Hufford et al. 2013;  Suarez et al. 2016 \\ 
\rowcolor{gray!25}
RASPberry &	phased haplotype &	Wegmann et al. 2011	 & Christe et al. 2016 \\
\rowcolor{gray!25}
MultiMix & phased/unphased genotype; reference panel &	Churchhouse and Marchini 2013 &	Eyheramendy et al. 2015 \\
\rowcolor{gray!25}
PCAdmix	 & phased haplotype	 & Brisbin et al. 2012	 & Moreno et al. 2014; Pugach et al. 2016 \\
\rowcolor{gray!25}
LAMP  &	phased haplotypes; reference panel	 & Sankararaman et al. 2008	 & Patterson et al. 2012 \\

\rowcolor{white}
{\emph{\bf phylogenetic relationship}} &   &   &   \\
\rowcolor{white}
ABBA-BABA/D-statistics	 & biallelic SNP  &	Durand et al. 2011	 &  Heliconius Genome Consortium 2012 \\
\rowcolor{white}
fd statistic &	biallelic SNP &	Martin et al. 2015  &	Malinsky et al. 2015; Zhang et al. 2016 \\ 
\rowcolor{white}
five taxon D statistics	& biallelic SNP	&  Pease and Hahn 2015	& Fontaine et al. 2015; Pease et al. 2016 \\

\rowcolor{gray!25}
{\emph{\bf divergence}} &   &   &   \\
\rowcolor{gray!25}
Gmin &	biallelic SNP	&  Geneva et al. 2015	&  Kingan et al. 2015 \\
\rowcolor{gray!25}
RNDmin	& phased haplotype	& Rosenzweig et al. 2016	&  NA \\

\rowcolor{white}
{\emph{\bf Bayesian approaches}} &   &   &   \\
\rowcolor{white}
\rowcolor{white}

\end{tabular}
\end{center}
\end{table}

### 2016-07-27 FigsAndTables/figure1.R

@@ -1,0 +1,23 @@
+library(data.table)
+library(cowplot)
+library(dplyr)
+library(tidyr)
+df <- read.delim("HapMix_vectors.txt")
+head(df)
+chr4 <- subset(df, df$chr==4)
+head(chr4)
+names(chr4)
+
+chr4.x2z <- chr4 %>% select(1:13)
+names(chr4.x2z)
+names(chr4.x2z)[5:13] <- gsub("_x2z", "", names(chr4.x2z)[5:13])
+names(chr4.x2z)
+chr4.long <- gather(chr4.x2z, pop, introgression, ELP:XOC)
+head(chr4.long)
+ggplot(data=chr4.long, aes(x=gen, y=introgression, color=pop, group=pop)) + geom_line() + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
+ggsave("z2mIntro.pdf")
+
+
+ggplot(data=chr4.long, aes(x=gen, y=introgression, color=pop, group=pop)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
+ggsave("z2mIntro.loessRegression.pdf")
+

### 2016-08-01 TiG_prospectus.tex
Li added modification

@@
 Second, local ancestry reconvolution (also known as chromosome painting) can identify the genomic regions that come from different source populations \cite{schraiber2015}. 
-Such method usually takes phased haplotypes as input and an exception is MULTIMIX \cite{churchhouse2013}, which can take both phased and unphased haplotypes as input.
-and unveil the genomic ancestry from multiple source populations.
+Such method usually takes phased haplotypes as input and an exception is MULTIMIX \cite{churchhouse2013}, which can take both phased and unphased haplotypes as input, and unveil the genomic ancestry from multiple source populations.
 
@@
 Their strength in identifying specific regions enabled us to explore the relationships between introgression and recombination rate / gene density / distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
+
+Fourth, structure related approaches are based on the popular software STRUCTURE \cite{pritchard2000}, which provided an estimate of fraction of each individual's genome coming from each population. 
+Many approaches in this respect have been developed.
+Two of them, fineStructure \cite{Lawson2012} and Globetrotter \cite{hellenthal2014}, were selected as representations.
+Different from STRUCTURE, fineStructure not only identifies the number of populations but also estimates the probability that genomic regions belong to one or the other populations. 
+Globotrotter is a development of fineStructure at tracing back to an unsampled source population and at dating the admixture events.
 

### 2017-02-06 TiG_prospectus.tex
update methods part

@@
 The collection of methods described different layers of genetic variation: diversity and divergence, haplotype variation and phylogenetic relationships.
-We focus here on analytical prospected and limitations and introduce only a few representatives of each category.
+We focus here on analytical prospect and limitations and introduce only a few representatives of each category.
 
-First, genomic regions of introgression were expected to show a low level of divergence.
+First, genomic regions of introgression in admixed populations were expected to show a low level of divergence from source population.
  F_{st} and d_{XY} as well as its derivates, such as G_{min} \cite{geneva2015} and RND_{min}\cite{rosenzweig2016}, are representatives of such statistics. 
@@
 In addition, RND_{min} is sensitive to variable mutation rate, which reflects on the branch length to the outgroup taxa.
+
+ These statistics were recently developed into several new statistics, such as $U_{A,B,C(w,x,y)}$ by adding the divergence between non-admixed populations (A) and source population (C) as a comparison to that between admixed populations (B) and source population (C)\cite{racimo2016}.
+ This statistic summarizes number of sites, where one particular allele at frequency $y$ in source population has a frequency bigger than $x$ in admixed population (B) and a frequency smaller than $w$ in non-admixed population (C).
+ Similarly, $Q95_{A,B,C(w,y)}$ just set a hard cutoff as $95^th$ percentile of allele frequencies in panel B \cite{racimo2016}. 
+ These two statistics were also developed into incorporating two source populations (see details in \cite{racimo2016}).
 

### 2017-04-04 paper.tex
A new rough draft of the paper.

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage[numbers,sort]{natbib}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{adjustbox}
\bibliographystyle{ieeetr}
\newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
\newcommand{\gmj}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments

\title{Review: Crop adaptation through introgression from wild relatives}

\author[1]{Authors: Garrett M. Janzen}%author information
\author[1]{Li Wang}
\author[1,*]{Matthew B. Hufford}
\affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
\affil[*]{Correspondence: mhufford@iastate.edu (M.B. Hufford)}
\date{}

\begin{document}

\maketitle

%Short Research reviews should be in the range 3500–4000 words, with up to 40 references and 6 figures/tables.

The traditional paradigm in crop domestication has been origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
Absent from this paradigm are the effects of hybridization between a diffusing nascent crop species and closely-related, locally-adapted wild relatives outside the center of origin.
New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.

\section*{Introgression methods and their applications}

The recent availability of genome-wide resequencing and reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq) data, combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).

\begin{enumerate}

\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.}

The methods introduced here do not include those marginally estimating introgression\slash migration rate as a component in the complete demographic history, such as Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference (dadi) \cite {gutenkunst2009}, sorts of isolation with migration model \cite{hey2004}, and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}. 
The collection of methods described different layers of genetic variation: diversity and divergence, haplotype variation and phylogenetic relationships.
We focus here on analytical prospects and limitations and introduce only a few representatives of each category.

First, genomic regions of introgression in admixed populations were expected to show a low level of divergence from source population.
$F_{st}$ and $d_{XY}$, as well as derivates such as $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016}, are representatives of such statistics. 
The former two are insensitive to rare migrants, and are therefore poorer at detecting recent introgression, while the latter two overcome these disadvantages.

 \begin{equation}
    G_{min} = \frac{d_{min}}{d_{XY}}
 \end{equation}
 where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y.

 \begin{equation}
 	RND_{min} = \frac{d_{min}}{d_{out}}
 \end{equation}
where $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
In addition, $RND_{min}$ is sensitive to variable mutation rate, which is indicated by branch length to the outgroup taxa.

These statistics were recently developed into several new statistics, such as $U_{A,B,C(w,x,y)}$, by adding the divergence between non-admixed populations ($A$) and source population ($C$) as a comparison to that between admixed populations ($B$) and source population ($C$)\cite{racimo2016}. 
This statistic summarizes number of sites, where one particular allele at frequency $y$ in source population has a frequency bigger than $x$ in admixed population ($B$) and a frequency smaller than $w$ in non-admixed population ($C$).
Similarly, $Q95_{A,B,C(w,y)}$ just sets a hard cutoff as $95^th$ percentile of allele frequencies in panel B \cite{racimo2016}. 
These two statistics were also designed to facilitate incorporatioon of two source populations (see details in \cite{racimo2016}).

Second, local ancestry reconvolution (also known as chromosome painting) can identify the genomic regions that come from different source populations \cite{schraiber2015}. 
Such methods usually take phased haplotypes as input, but MULTIMIX \cite{churchhouse2013} can take both phased and unphased haplotypes as input, unveiling the genomic ancestry from multiple source populations.

Third, ABBA-BABA statistics (also known as D-statistics) and their derivatives are widely applied to the search for signs of introgression in genomic patterns of shared derived variants between populations or species. 
Based on D-statistics, $\hat{f_{d}}$ \cite{Martin2015} and five-taxon D statistics \cite{pease2015} were developed to localize genomic regions under introgression. 
The former took allele frequencies from each population/species, and the latter detected introgression from the localized phylogenetic pattern.
Their strength in identifying specific regions enables one to explore the relationships between introgression and recombination rate/gene density/distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.

Fourth, structure-related approaches are based on the popular software STRUCTURE \cite{pritchard2000}, which provides an estimate of the fraction of each individual's genome coming from each population. 
Many approaches in this respect have been developed.
Two of these, fineSTRUCTURE \cite{Lawson2012} and GLOBETROTTER \cite{hellenthal2014}, were selected as representations.
Different from STRUCTURE, fineSTRUCTURE not only identifies the number of populations, but also estimates the probability that genomic regions belong to one or the other populations. 
GLOBETROTTER is a development of fineSTRUCTURE that allows for tracing back to an unsampled source population and dating admixture events.

\item{Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role. For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}}.
\end{enumerate}

\noindent{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods.}

\noindent{\textbf{Figure 1}: Wing coloration patterns in \emph{Heliconius} and evidence for introgression across species based on Patterson's \emph{D}-statistic; adapted from \cite{Heliconius2012}.}

%\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model. I put the three in the bib file.}

\section*{Crop adaptation through introgression}
Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and wild relatives outside putative domestication centers.
A history of introgression during diffusion appears to be the rule for crops rather than the exception.
\begin{enumerate}
	\item{Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}. Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.}
	\item{Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}

\begin{enumerate}
\item{Maize:}

Maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter referred to as \emph{mexicana}) have long been the subject of research into adaptive crop/wild introgression.
Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis})approximately 9,000 years ago in the Balsas River Valley \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican central plateau, where it came into sympatry with wild \emph{mexicana}.
Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analyses \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
However, \citep{hufford2013} was the first to reveal evidence of adaptive introgression from \emph{mexicana} into Mexican highland maize.
The authors identified nine genomic regions which showed evidence of introgression from \emph{mexicana} to maize in both the HAMPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the nine pairs sampled.
Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
It was also revealed that in the growth chamber experiment, maize populations with introgression from \emph{mexicana} on chromosomes 4 (associated with QTLs controlling pigment density and macrohairs) and 9 (overlapped with QTLs for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
In summation, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations which are allopatric to \emph{mexicana}. 
A recent study \cite{Takuno2015} found little empirical or theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly by the difference in potential for adaptive introgression from wild relatives.
Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.
Furthermore, the adaptive SNPs in the Mexican highland population were more likely also showing signatures of local adaptation in \emph{mexicana} and \emph {pulviglumis} populations than those South American highland population.
For these reasons, adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.

Several questions regarding \emph{mexicana} introgression yet remain.
Has the introgression from \emph{mexicana} spread to multiple highland populations? 
How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
A recent study \cite{Wang2015manuscript} first utilized the ABBA-BABA statistics to evaluate the existence or absence of introgression from \emph{mexicana} to multiple highland populations, and then $\hat{f_{d}}$ statistic proposed by Martin et al. (2015) \citep{martin2015} were calculated to locate the introgressed loci. 
Figure \ref{fig:fd} shows the loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across chromosomes 4 and 5 in multiple comparisons. 
On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al. (2013) \citep{hufford2013}.
The signal of introgression is absent for the other three populations.
On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
Overall, both analyses revealed that gene flow from \emph{mexicana} exists, not only to MexHigh maize, but also to GuaHigh maize, as well as to some individuals in SW\_US, but not to the Andes and South American lowland (SA\_Low) populations, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations. 
The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects. 
%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.

\item{Asian Rice:}

Introgression appears to have played an important role in the domestication history of rice.
The center of Asian rice (\emph{Oryza sativa}) domestication is not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O. sativa}, 8,200-13,500 years ago, from wild \emph{O. rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O. perennis} complex" \cite{oka2012origin}\.
Asian rices readily hybridize both with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}).
The high genetic diversity within \emph{O. sativa} is likely due to introgression with wild relatives both during domestication and upon the dispersal of \emph{O. sativa} into new environments and sympatry with new relatives \cite{second1982origin}.

In an attempt to categorize the varieties of Asian rice, \cite{khush2003classifying} performed isozyme polymorphism analysis (11 enzymes, 20 loci, 79 alleles) on 25,519 rice varieties.
The vast majority of the rice varieties fell into two groups; indica (73.45\%) and japonicas (23.02\%).
Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
However,about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
Introgression of alleles between cultivated rices, as well as between cultivated and wild rices, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticate is suspected (\cite{zhao2010genomic} calls for research into this possibility).
These rices often have blight resistance gene, likley imparted from the wild relative parent.
Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) are known to have been introgressed from wild relatives into \emph{O. sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
Some aus and rayada varieties retain characteristics of the putative wild parent.
Aus exhibits a sprawling growth pattern and easy-threshing grain.
Rayada is adapted to areas with longer periods of flooding, with heightened seed dormancy and photoperiod sensitivity (permitting harvest during times without standing water).
Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between O. sativa and deepwater-adapted O. rufipogon populations.

In addition to investigative experiments, gene flow from wild relatives has been utilized to produce agronomic rice varieties.
Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.

\item{Barley:}

Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp. spontaneum} in the Fertile Crescent and once from subsp. \emph{spontaneum} var. \emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.

Variety \emph{agriocrithon} is genetically diverse, and is found throughout much of the range of barley.
Some have suggested that \emph{agriocrithon} may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp. \emph{spontaneum}, perhaps more than once (in Israel as well as in Tibet).
Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and west-central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.

There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}\.
Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.

The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}\.
The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces.
The authors found a high amount variability in the amount of contribution from wild relatives, as well as its location in the genome, within barley populations.
This is indicative of contribution from numerous wild populations.
Furthermore, the authors find that wild introgression contribution is generally greatest from geographically-proximate populations, and that introgressed regions might be combined from geographically-separate wild populations.
Low linkage disequilibrium and small blocks of identity by state indicate that these introgressed regions are old, perhaps dating back to the beginning of domestication.
As landraces and nearby wild relatives share similar genomic sequences, the introgressed regions that are exclusive to that landrace are more more likely to contain adaptive alleles.

The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}.

\item{Sunflower: }

The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
Pre-Columbian H. annuus distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H. petiolaris} and \emph{H. bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
However, the most striking example of adaptive introgression within Helianthus is that of the cucumberleaf sunflower, H. debilis ssp. cucumerifolius.
Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
Introgressive hybridization imparted locally-adapted alleles from H. debilis to H. annuus via introgressive hybridization. 
These introgressed hybrids formed a new lineage of sunflower (H. annuus ssp. texanus, H. a. texanus hereafter) which displays H. debilis-like traits adaptive to south-central Texas climate and ecology.
These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than H. annuus (as measured by higher seed production)\cite{whitney2006adaptive}.
Although H. annuus and H. a. texanus are interfertile, H. a. texanus displays persistent phenotypic differences from H. annuus.

The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
Some of these downy mildew resistence genes were found in wild relatives (including \emph{H. argophyllus}, \emph{H. tuberosus}, and \emph{H. praecox}) and have been successfully bred into modern \emph{H. annuus} \cite{miller1991inheritance}\.
PlArg, an allele found in wild silverleaf sunflowers (\emph{H. argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
\emph{H. annuus} shows signs of persistent introgressive hybridization with \emph{H. petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}.

Recent investigations into the history of Helianthus introgression have implemented genomic methods.
\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild H. annuus, H. petiolaris, and H. argophyllus.
Using STRUCTURE, these authors found that introgressions from wild relatives can be found on every chromosome in at least one modern line, covering over 10\% of the genome.
Of particular note is the modern line RHA 274, a modern line which was bred with H. a. texanus in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
RHA 274 has several large introgression from H. a. texanus, including one at the site of HaGNAT, the domestication gene associated with branching.
These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the H. a. texanus introgressed regions are causative.

%The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
%This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}.

%whitney2006adaptive
%"these results suggest that introgression of biotic resistance traits was important in the adaptation of H. annuus to central and southern Texas."
%fitness (seed production) high in texanus than in annuus
%identified two pests to which resistance was imparted and important
%Discussion points: 1. texanus has higher fitness than annuus in central texas.
%2. texanus shared traits more in common with debilis than with annuus (herbivore resistance)
%3. 2 of the three traits from point two above are important for adaptation to life in central south Texas.
%The authors ask why these pest-resistance genes have not spread further north beyond this range.  They are uncertain, but conjecture that negative pleiotrophic effects are at play.

%whitney2010adaptive
%"We demonstrate that introgression has altered multiple aspects of the H. annuus phenotype in an adaptive manner, has affected traits relevant to both biotic and abiotic environments, and may have aided expansion of the H. annuus range into central Texas, USA."
%This paper really does show that this case of natural introgression was adaptive.
%The companion paper (2006 Whitney) will likely show the same thing, but I haven't read it yet.

%scascitelli2010genome
%"long-term migration rates were high, genome-wide and asymmetric, with higher migration rates from H. annuus texanus into the two parental taxa than vice versa."
%"H. annuus texanus may serve as a bridge for the transfer of alleles between its parental taxa."
%"contradict recent theory suggesting that introgression should predominantly be in the direction of the colonizing species."

%rieseberg2007hybridization
%this is a very good review.  i think i should use it for the table.
%they also show that traits were imparted from debilis to texanus, and that these are adaptive.  i think they also show genetic evidence, but i'm not sure.

	\end{enumerate}
	\item{Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to both their current extent of cultivation and the distribution of wild relatives across environmental gradients (\textbf{Figure 2}).}

\end{enumerate}

\noindent{\textbf{Table 2:} Overview of literature evaluating potential for crop-wild introgression in the world's 15 most important crops including a column of references and a column summarizing major findings.}

\begin{table}
\centering
\begin{adjustbox}{width=1\textwidth}
\small
\label{my-label}
\begin{tabular}{llllll}
\hline
Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Introgression into Crop & Evidence of Adaptiveness & Source \\
\hline
Maize (Zea mays subsp. mays) & Zea mays subsp. mexicana, Zea mays subsp. parviglumis & X & X & X & \cite{hufford2013genomic} \\
Rice (Oryza sativa, Oryza Glaberrima) & Oryza rufipogon & X & X & X & \cite{Huang2012} \\
Barley (Hordeum vulgare) & Hordeum vulgare subsp. spontaneum & X & X & X & \cite{Poets2015} \\
Cassava (Manihot esculenta) & Manihot glaziovii & X & X & X & \cite{bredeson2016sequencing} \\
Olive (Olea europaea ssp. europaea var. sativa) & [[]] & X & X & & \cite{diez2015olive} \\
Sunflower (Helianthus annuus) & wild H. annuus, H. argophyllus, H. petiolaris & X & X & X & [[]]\\
Potato (Solanum tuberosum) & many & X & X & X & \cite{johns1986ongoing} \\
Soybeans (Glycine max) & G. soja & [[]] & X & X & \cite{lam2010resequencing} \\
Grapes (Vitis vinifera subsp. vinifera) & Vitis vinifera subsp. sylvestris & [[]] & X & [[]] &  \cite{myles2011genetic} \\
\hline
\end{tabular}
\end{adjustbox}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[width=17.35cm]{map.png}
	\caption{Map of the domestication centers of four crops and the distribution of their interfertile wild relatives, overlaying average annual precipitation.}
	\label{boxplot:map}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=17.35cm]{boxplot.png}
	\caption{The distribution of average annual precipitations experienced in the geographic home ranges of wild relatives interfertile with four crops}
	\label{fig:map}
\end{figure}

\noindent{\textbf{Figure 2}: Multi-panel figure showing maps of crop centers of domestication, distributions of wild relatives, and current extent of cultivation across gradient of precipitation. This figure will illustrate the adaptive potential of wild-to-crop introgression.}
%\gmj{Will we have to generate these images before we submit the prospectus?}

\section*{Re-evaluating concepts of domestication}
A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification. A demography incorporating introgression from additional sources appears to be more correct for many crops. With this in mind, certain aspects of crop evolution must be re-evaluated:
\begin{enumerate}
	\item{Estimates of the initial domestication bottleneck may be skewed when introgression is not considered. Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.  For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.}%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?  I just want to make sure I understand.}
	\item{Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.}%\lwang{It could be either inflated or underestimate,. It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
	\item{Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa. Introgression may alter these signatures and confound detection of domestication loci.} %\lwang{I donot quite understand this point. So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right? But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
	%\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.  The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
\end{enumerate}

\section*{Future studies in crop-wild introgression}
Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics:
\begin{enumerate}
	\item{Basic questions:}
		\begin{enumerate}
		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} %\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges? To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?}%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g. temperature, precipitation, altitude, day lengh, etc.)?}
		 	\item{Can colonizing species served as bridges for gene flow between previously allopatric taxa?} %\lwang{I feel the second and the third points could be combined. How the wild species could introgress into the allopatric landraces? Does domesticates serve as bridges?}
		\end{enumerate} 	
	\item{Applications:}
		\begin{enumerate}
		 	\item{Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} %\lwang{I am not quite clear with this point. It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow. Right?}\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?  I hope that question made sense.}
		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}%\gmj{This point is very clear, but should it be fleshed out further?  It wouldn't be difficult to include some examples.}
		 \end{enumerate}
\end{enumerate}

\bibliography{bib_gj}

\end{document}

### 2017-04-06 paper.tex
Update paper.tex

@@
 Rayada is adapted to areas with longer periods of flooding, with heightened seed dormancy and photoperiod sensitivity (permitting harvest during times without standing water).
-Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between O. sativa and deepwater-adapted O. rufipogon populations.
+Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between \emph{O.
+sativa} and deepwater-adapted \emph{O.
+rufipogon} populations.
 
@@
 Barley (\emph{Hordeum vulgare} subsp.
-\emph{vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp.
-spontaneum} in the Fertile Crescent and once from subsp.
+\emph{vulgare}) is believed to have been domesticated at least twice, once from wild subsp.
+\emph{spontaneum} in the Fertile Crescent and once from subsp.
 \emph{spontaneum} var.
@@
 The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
-Pre-Columbian H. annuus distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
+Pre-Columbian \emph{H.
+annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
 Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H.
@@
 Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
-However, the most striking example of adaptive introgression within Helianthus is that of the cucumberleaf sunflower, H. debilis ssp. cucumerifolius.
+However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
 Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
-Introgressive hybridization imparted locally-adapted alleles from H. debilis to H. annuus via introgressive hybridization. 
-These introgressed hybrids formed a new lineage of sunflower (H.
-annuus ssp.
-texanus, H.
+Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
+These introgressed hybrids formed a new lineage of sunflower (\emph{H.
+annuus} ssp.
+\emph{texanus}, \emph{H.
 a.
-texanus hereafter) which displays H.
-debilis-like traits adaptive to south-central Texas climate and ecology.
-These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than H.
-annuus (as measured by higher seed production)\cite{whitney2006adaptive}.
-Although H. annuus and H. a. texanus are interfertile, H. a. texanus displays persistent phenotypic differences from H. annuus.
+texanus} hereafter) which displays \emph{H.
+debilis}-like traits adaptive to south-central Texas climate and ecology.
+These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
+annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
+Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
 
@@
 
-Recent investigations into the history of Helianthus introgression have implemented genomic methods.
-\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild H. annuus, H. petiolaris, and H. argophyllus.
+Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
+\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
 Using STRUCTURE, these authors found that introgressions from wild relatives can be found on every chromosome in at least one modern line, covering over 10\% of the genome.
@@
 texanus in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
-RHA 274 has several large introgression from H. a. texanus, including one at the site of HaGNAT, the domestication gene associated with branching.
-These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the H. a. texanus introgressed regions are causative.
+RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
+These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
 
@@
 	\centering
-	\includegraphics[width=17.35cm]{map.png}
+	\includegraphics[width=17.35cm]{figure_outline.png}
 	\caption{Map of the domestication centers of four crops and the distribution of their interfertile wild relatives, overlaying average annual precipitation.}

### 2017-04-07 paper.tex
Update paper.tex

@@
 %Short Research reviews should be in the range 3500–4000 words, with up to 40 references and 6 figures/tables.
+
+Whereas the process of domestication used to be concieved of as rapid and geographically-constrained, archaeological and genetic evidence has recently revealed that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
+This new paradigm emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
+Adaptive introgression has two steps: hybridization between two genomes (in the case of this review, a domesticated crop and its wild relative), followed by recombination and selection for progressively smaller genomic segments that harbor the adaptive locus/loci with fewer maladaptive alleles being dragged along \cite{Feuillet200824}.
 
@@
 The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}\.
+Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including powdery mildew resistance \cite{dreiseitl2017heterogeneity}, brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}\.
 The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces.

### 2017-04-08 paper.tex
Update paper.tex

@@
 This new paradigm emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
-Adaptive introgression has two steps: hybridization between two genomes (in the case of this review, a domesticated crop and its wild relative), followed by recombination and selection for progressively smaller genomic segments that harbor the adaptive locus/loci with fewer maladaptive alleles being dragged along \cite{Feuillet200824}.
+Adaptive introgression has three components: hybridization between two genomes (in the case of this review, a domesticated crop and its wild relative), backcrossing to one of the parents (in this case, the domesticate), recombination, and selection on different recombinant genotypes with progressively smaller linkage \cite{barton2001role}.
+
+followed by recombination and selection for progressively smaller genomic segments that harbor the adaptive locus/loci with fewer maladaptive alleles being dragged along \cite{Feuillet200824}.
+Much attention has been given to the risk of natural introgression of transgenes from domessticated crops into wild relatives (for a review, \cite{stewart2003transgene}), or the history, utility, and methods of breeding programs designed to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use})
+, but what
 
@@
 \end{document}
+
+\end{document}

### 2017-04-10 paper.tex
paste Table1 and Figure1 into the paper.tex

@@
 
-\noindent{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods.}
+%Table1
+\begin{figure}
+\begin{tabular}{l|l|l|l}
+\bf methods	& \bf data type  &	\bf reference &  \bf empirical studies  \\ \hline
 
-\noindent{\textbf{Figure 1}: Wing coloration patterns in \emph{Heliconius} and evidence for introgression across species based on Patterson's \emph{D}-statistic; adapted from \cite{Heliconius2012}.}
+\emph{\bf chromosome paiting} &   &   &   \\
+Hapmix	& phased haplotype; reference panel		& Price et al. 2009	&  Hufford et al. 2013;  Suarez et al. 2016 \\ 
+RASPberry &	phased haplotype &	Wegmann et al. 2011	 & Christe et al. 2016 \\
+MultiMix & phased/unphased genotype; reference panel &	Churchhouse and Marchini 2013 &	Eyheramendy et al. 2015 \\
+PCAdmix	 & phased haplotype	 & Brisbin et al. 2012	 & Moreno et al. 2014; Pugach et al. 2016 \\
+LAMP  &	phased haplotypes; reference panel	 & Sankararaman et al. 2008	 & Patterson et al. 2012 \\
 
-%\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model.
-I put the three in the bib file.}
+\emph{\bf phylogenetic relationship} &   &   &   \\
+
+ABBA-BABA/D-statistics	 & biallelic SNP  &	Durand et al. 2011	 &  Heliconius Genome Consortium 2012 \\
+fd statistic &	biallelic SNP &	Martin et al. 2015  &	Malinsky et al. 2015; Zhang et al. 2016 \\ 
+five taxon D statistics	& biallelic SNP	&  Pease and Hahn 2015	& Fontaine et al. 2015; Pease et al. 2016 \\
+
+\emph{\bf divergence} &   &   &   \\
+Gmin &	biallelic SNP	&  Geneva et al. 2015	&  Kingan et al. 2015 \\
+RNDmin	& phased haplotype	& Rosenzweig et al. 2016	&  NA \\
+U and Q95	& biallelic SNP	& Racimo et al. 2016	 &  Sams et al. 2016 \\
+
+\emph{\bf population structure related} &   &   &   \\
+fineStructure &	phased haplotype &	Lawson et al. 2012	& Skoglund et al. 2015 \\
+Globetrotter &	phased haplotype &	Hellenthal et al. 2014 & Scott et al. 2016 
+\end{tabular}
+\caption{Programs for identifying genomic regions under introgression \label{tab:tools}}
+\end{figure}
+
+%figure 1
+\begin{figure*}[b]
+\centering
+\includegraphics{FigsAndTables/z2mIntroLoessRegression.pdf}
+\caption{Loess regression of introgression level from \emph{mexicana} to Mexican Highland maize populations on chromosome 4 adapted from \cite{hufford2013genomic}.
+\label{fig:maizeChr4Intro}}
+\end{figure*}
 

### 2017-04-11 paper.tex
Update paper.tex

--- 2017-04-07 paper.tex
@@
 \usepackage{adjustbox}
+\usepackage[english]{babel}
 \bibliographystyle{ieeetr}
@@
 
-Whereas the process of domestication used to be concieved of as rapid and geographically-constrained, archaeological and genetic evidence has recently revealed that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
-This new paradigm emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
-Adaptive introgression has two steps: hybridization between two genomes (in the case of this review, a domesticated crop and its wild relative), followed by recombination and selection for progressively smaller genomic segments that harbor the adaptive locus/loci with fewer maladaptive alleles being dragged along \cite{Feuillet200824}.
+The process of domestication was one thought to be rapid and geographically-constrained, with origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
+However, archaeological and genetic evidence are revealing that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
+This new conception of domestication emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
 
-The traditional paradigm in crop domestication has been origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
-Absent from this paradigm are the effects of hybridization between a diffusing nascent crop species and closely-related, locally-adapted wild relatives outside the center of origin.
-New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
+Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
+In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the domesticate and adaptive genomic regions are retained and increase in frequency in the population as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and lost.
+Much attention has been given to the risk of natural introgression of transgenes from domessticated crops into wild relatives (for a review, \cite{stewart2003transgene}), or the history, utility, and methods of breeding programs designed to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but recent tools and methods have been employed to detect genome-wide patterns of introgression in a number of species, granting insights into the prominence of adaptive introgression in crop domestication.
+
 In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
@@
 debilis}-like traits adaptive to south-central Texas climate and ecology.
-These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
+These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
 annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
@@
 \label{my-label}
-\begin{tabular}{llllll}
+\begin{tabular}{|p{5cm}|p{5cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|l|}
 \hline
-Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Introgression into Crop & Evidence of Adaptiveness & Source \\
+Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
+Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{hufford2013genomic} \\ 
 \hline 
-Maize (Zea mays subsp. mays) & Zea mays subsp. mexicana, Zea mays subsp. parviglumis & X & X & X & \cite{hufford2013genomic} \\
-Rice (Oryza sativa, Oryza Glaberrima) & Oryza rufipogon & X & X & X & \cite{Huang2012} \\
-Barley (Hordeum vulgare) & Hordeum vulgare subsp. spontaneum & X & X & X & \cite{Poets2015} \\
-Cassava (Manihot esculenta) & Manihot glaziovii & X & X & X & \cite{bredeson2016sequencing} \\
-Olive (Olea europaea ssp. europaea var. sativa) & [[]] & X & X & & \cite{diez2015olive} \\
-Sunflower (Helianthus annuus) & wild H. annuus, H. argophyllus, H. petiolaris & X & X & X & [[]]\\
-Potato (Solanum tuberosum) & many & X & X & X & \cite{johns1986ongoing} \\
-Soybeans (Glycine max) & G. soja & [[]] & X & X & \cite{lam2010resequencing} \\
-Grapes (Vitis vinifera subsp. vinifera) & Vitis vinifera subsp. sylvestris & [[]] & X & [[]] &  \cite{myles2011genetic} \\
+Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
+\hline
+Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
+Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X & X & X & \cite{rieseberg2007hybridization}\\ 
+\hline
+Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\ 
+\hline
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing} \\ \hline
+Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
+\hline
+Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\ 
+\hline
+Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
 \hline
@@
 With this in mind, certain aspects of crop evolution must be re-evaluated:
-\begin{enumerate}
-	\item{Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
+
+Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
 Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
-For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.}%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?
+For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.
+
+%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?
 I just want to make sure I understand.}
-	\item{Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.}%\lwang{It could be either inflated or underestimate,.
-It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
-	\item{Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
-Introgression may alter these signatures and confound detection of domestication loci.} %\lwang{I donot quite understand this point.
+Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
+%\lwang{It could be either inflated or underestimate,. It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
+
+Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
+Introgression may alter these signatures and confound detection of domestication loci.
+%\lwang{I donot quite understand this point.
 So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right?
@@
 The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
-\end{enumerate}
 
 \section*{Future studies in crop-wild introgression}
-Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics:
-\begin{enumerate}
-	\item{Basic questions:}
-		\begin{enumerate}
-		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} %\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
-		 	\item{At what geographic scale does adaptive introgression occur?
+
+New insights into domestication histories of crops are of potential value for the continued improvement of modern crops.
+Continued research into this crop/wild introgression will prove valuable
+
+Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics.
+Basic questions:
+To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?
+%\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
+At what geographic scale does adaptive introgression occur?
 Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?
-To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?}%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g.
+To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g.
 temperature, precipitation, altitude, day lengh, etc.)?}
-		 	\item{Can colonizing species served as bridges for gene flow between previously allopatric taxa?} %\lwang{I feel the second and the third points could be combined.
+Can colonizing species served as bridges for gene flow between previously allopatric taxa?
+%\lwang{I feel the second and the third points could be combined.
 How the wild species could introgress into the allopatric landraces?
-Does domesticates serve as bridges?}
-		\end{enumerate} 	
-	\item{Applications:}
-		\begin{enumerate}
-		 	\item{Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} %\lwang{I am not quite clear with this point.
+Does domesticates serve as bridges?
+
+Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.
+%\lwang{I am not quite clear with this point.
 It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow.
-Right?}\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?
+Right?}
+%\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?
 I hope that question made sense.}
-		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}%\gmj{This point is very clear, but should it be fleshed out further?
-It wouldn't be difficult to include some examples.}
-		 \end{enumerate}
-\end{enumerate}
+Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
 

### 2017-04-12 paper.tex
Update paper.tex

@@
 \hline 
-Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
+Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
 \hline
@@
 \hline
-Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing} \\ \hline
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing} \\
+\hline
+Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
+\hline
 Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
@@
 \hline
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X[[?not much]] &  & \cite{papa2003asymmetry} \\
+\hline
 Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
+\hline
+Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
+\hline
+Wheat (\emph{Tritium monococcum, T.
+dicoccum, T.
+aestivum}) & \emph{T.
+m.
+boeoticum, T.
+dioccoides, T.
+urartu, Aegilops speltoides, A.
+tauschii} & X & X[[?hybrid origins]] &  & \cite{zohary1969wild} \\
 \hline

### 2017-04-13 paper.tex
Update paper.tex

@@
 
-\begin{enumerate}
-
-\item{High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.}
+High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
 
@@
 
-\item{Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}}.
-\end{enumerate}
+Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
 
@@
 \section*{Crop adaptation through introgression}
+
 Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and wild relatives outside putative domestication centers.
 A history of introgression during diffusion appears to be the rule for crops rather than the exception.
-\begin{enumerate}
-	\item{Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
-Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.}
-	\item{Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).}
+Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
+Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.
+Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).
 
@@
 	\end{enumerate}
-	\item{Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to both their current extent of cultivation and the distribution of wild relatives across environmental gradients (\textbf{Figure 2}).}
-
-\end{enumerate}
-
-\noindent{\textbf{Table 2:} Overview of literature evaluating potential for crop-wild introgression in the world's 15 most important crops including a column of references and a column summarizing major findings.}
+	Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to both their current extent of cultivation and the distribution of wild relatives across environmental gradients (\textbf{Figure 2}).
 
@@
 
-\noindent{\textbf{Figure 2}: Multi-panel figure showing maps of crop centers of domestication, distributions of wild relatives, and current extent of cultivation across gradient of precipitation.
-This figure will illustrate the adaptive potential of wild-to-crop introgression.}
-%\gmj{Will we have to generate these images before we submit the prospectus?}
-
 \section*{Re-evaluating concepts of domestication}

### 2017-05-24 paper.tex
Update paper.tex
rough draft complete

@@
 
-The process of domestication was one thought to be rapid and geographically-constrained, with origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
+The process of domestication was once thought to be rapid and geographically-constrained, with crops originating from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
 However, archaeological and genetic evidence are revealing that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
@@
 
-Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
-In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the domesticate and adaptive genomic regions are retained and increase in frequency in the population as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and lost.
-Much attention has been given to the risk of natural introgression of transgenes from domessticated crops into wild relatives (for a review, \cite{stewart2003transgene}), or the history, utility, and methods of breeding programs designed to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but recent tools and methods have been employed to detect genome-wide patterns of introgression in a number of species, granting insights into the prominence of adaptive introgression in crop domestication.
+Adaptive introgression has three components: Hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
+In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the crop and adaptive genomic regions are retained and increase in frequency as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and potentially purged from the population.
+Much attention has been given to the risk of natural introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and to the history, utility, and methods of breeding programs designed for the introgressiin of desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but the history of natural introgression of wild alleles into domesticated crops thoughout evolutionary timescales has recieved considerably less attention.
+Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prominence of adaptive introgression in crop domestication.
 
-In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe these recent methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
@@
 
-High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
-
+High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgressions across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
 The methods introduced here do not include those marginally estimating introgression\slash migration rate as a component in the complete demographic history, such as Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference (dadi) \cite {gutenkunst2009}, sorts of isolation with migration model \cite{hey2004}, and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}.
@@
 
-Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and wild relatives outside putative domestication centers.
-A history of introgression during diffusion appears to be the rule for crops rather than the exception.
-Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
-Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.
+Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and their wild relatives outside putative domestication centers.
 Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).
@@
 
-Maize (\emph{Zea mays} ssp.
+The relationship between maize (\emph{Zea mays} ssp.
 \emph{mays}) and the teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter referred to as \emph{mexicana}) have long been the subject of research into adaptive crop/wild introgression.
-Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis})approximately 9,000 years ago in the Balsas River Valley \cite{matsuoka2002single}.
+\emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive crop/wild introgression.
+Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the Balsas River Valley in Mexico \cite{matsuoka2002single}.
 From this domestication center, maize spread into the highlands of the Mexican central plateau, where it came into sympatry with wild \emph{mexicana}.
 Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analyses \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
-However, \citep{hufford2013} was the first to reveal evidence of adaptive introgression from \emph{mexicana} into Mexican highland maize.
+However, \citep{hufford2013} was the first to reveal evidence that introgression from \emph{mexicana} into Mexican highland maize is adaptive.
+%The following is not fully clear.  Cut or revise.
 The authors identified nine genomic regions which showed evidence of introgression from \emph{mexicana} to maize in both the HAMPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the nine pairs sampled.
@@
 It was also revealed that in the growth chamber experiment, maize populations with introgression from \emph{mexicana} on chromosomes 4 (associated with QTLs controlling pigment density and macrohairs) and 9 (overlapped with QTLs for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
-In summation, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
+In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
 
@@
 rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
-perennis} complex" \cite{oka2012origin}\.
+perennis} complex" \cite{oka2012origin}.
 Asian rices readily hybridize both with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}).
@@
 In an attempt to categorize the varieties of Asian rice, \cite{khush2003classifying} performed isozyme polymorphism analysis (11 enzymes, 20 loci, 79 alleles) on 25,519 rice varieties.
-The vast majority of the rice varieties fell into two groups; indica (73.45\%) and japonicas (23.02\%).
+The vast majority of the rice varieties fell into the group indica (73.45\%) or japonica (23.02\%).
 Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
-However,about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
+However, about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
 Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
@@
 In addition to investigative experiments, gene flow from wild relatives has been utilized to produce agronomic rice varieties.
-Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}\.
+Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}.
 Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
@@
 
-Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
+Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
 Barley (\emph{Hordeum vulgare} subsp.
@@
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.
+\emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\.
+However, many of the details of barley domestication are still disputed.
 
-Variety \emph{agriocrithon} is genetically diverse, and is found throughout much of the range of barley.
-Some have suggested that \emph{agriocrithon} may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp.
+%Variety \emph{agriocrithon} is genetically diverse, and is found throughout much of the range of barley.
+%Some have suggested that \emph{agriocrithon} may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp.
 \emph{spontaneum}, perhaps more than once (in Israel as well as in Tibet).
-Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and west-central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}\.
-These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
+%Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and west-central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}.
+%These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
 
-There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}\.
-Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}\.
+There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}.
+Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}.
 Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.
 
-The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including powdery mildew resistance \cite{dreiseitl2017heterogeneity}, brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}\.
-The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces.
-The authors found a high amount variability in the amount of contribution from wild relatives, as well as its location in the genome, within barley populations.
+The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}.
+The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces, and found a high amount variability in the amount of contribution from wild relatives, as well as its location in the genome, within barley populations.
 This is indicative of contribution from numerous wild populations.
-Furthermore, the authors find that wild introgression contribution is generally greatest from geographically-proximate populations, and that introgressed regions might be combined from geographically-separate wild populations.
-Low linkage disequilibrium and small blocks of identity by state indicate that these introgressed regions are old, perhaps dating back to the beginning of domestication.
-As landraces and nearby wild relatives share similar genomic sequences, the introgressed regions that are exclusive to that landrace are more more likely to contain adaptive alleles.
-
-The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}\.
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including brittleness, flowering time, plant height, lodging, and yield \cite{von2006ab,handley1994chromosome}.
+Furthermore, the authors found that wild introgression contribution is generally greatest from geographically-proximate populations, and that introgressed regions might be combined from geographically-separate wild populations.
+Low linkage disequilibrium and small blocks of identity by state indicate that these introgressed regions are old, perhaps dating back to the beginning of barley domestication.
+As landraces and nearby wild relatives share similar genomic sequences, the introgressed regions that are exclusive to that landrace are more likely to contain adaptive alleles. 
+Such alleles were not identified specifically, though wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}.
 
@@
 
-The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
+The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}.
 Pre-Columbian \emph{H.
 annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
+
 Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H.
 petiolaris} and \emph{H.
-bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
+bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
 Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
@@
 These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
-annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
+annuus} (as measured by higher seed production \cite{whitney2006adaptive}).
 Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
@@
 The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
-\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}\.
+\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
 Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
@@
 praecox}) and have been successfully bred into modern \emph{H.
-annuus} \cite{miller1991inheritance}\.
+annuus} \cite{miller1991inheritance}.
 PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
-argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}\.
+argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}.
 Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
@@
 \cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
-Using STRUCTURE, these authors found that introgressions from wild relatives can be found on every chromosome in at least one modern line, covering over 10\% of the genome.
-Of particular note is the modern line RHA 274, a modern line which was bred with H.
+Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
+Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
 a.
-texanus in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
+texanus} in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
 RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
@@
 \end{enumerate}
-	Predictions regarding the likelihood of adaptive introgression in additional crops can be made through comparison of their centers of origin to both their current extent of cultivation and the distribution of wild relatives across environmental gradients (\textbf{Figure 2}).
 
@@
 \hline
-Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X[[?not much]] &  & \cite{papa2003asymmetry} \\
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X[[present, but low]] &  & \cite{papa2003asymmetry} \\
 \hline
@@
 	\centering
-	\includegraphics[width=17.35cm]{figure_outline.png}
-	\caption{Map of the domestication centers of four crops and the distribution of their interfertile wild relatives, overlaying average annual precipitation.}
+	\includegraphics[width=17.35cm]{par_figure.png}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}
 	\label{boxplot:map}
@@
 	\includegraphics[width=17.35cm]{boxplot.png}
-	\caption{The distribution of average annual precipitations experienced in the geographic home ranges of wild relatives interfertile with four crops}
+	\caption{The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
 	\label{fig:map}
@@
 \section*{Re-evaluating concepts of domestication}
+
 A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification.
-A demography incorporating introgression from additional sources appears to be more correct for many crops.
+An history of introgression during diffusion appears to be the rule for crops rather than the exception.
+Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
+Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.
+
 With this in mind, certain aspects of crop evolution must be re-evaluated:
 
-Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
+* Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
 Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
 For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.
+Conversely, if the donor population has a relatively small effective population size, the opposite bias may be imposed upon bottleneck estimations.
+In most cases, the effective population size of the domesticated crop will be lower than that of the wild progenitor and wild relative populations.
 
-%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?
-I just want to make sure I understand.}
-Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
-%\lwang{It could be either inflated or underestimate,. It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
+* Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
+The directionality of this effect is likewise dependent on $N_e$ of the donor population.
 
-Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
+* Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
 Introgression may alter these signatures and confound detection of domestication loci.
+
 %\lwang{I donot quite understand this point.
@@
 But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
-%\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
+
+* When  highly-introgressed or hybrid populations are selected for domestication (as in the case of potato and tomato), identification of original progenitor(s) and domestication centers is difficult.
+Determining whether a crop 
+Crops arising from highly-interfertile compleces of wild relatives
+
+%\gmj{It may not fit here perfectly, but we might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
 The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
@@
 
-New insights into domestication histories of crops are of potential value for the continued improvement of modern crops.
-Continued research into this crop/wild introgression will prove valuable
+Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication histories of many agronomically-important crops.
+However, the dynamics of the process in these cases are not yet fully understood.
+To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
+Can colonizing species and/or hybrid swarms serve as bridges for gene flow between previously allopatric taxa?
+At what geographic scale does adaptive introgression occur?
+Is introgression frequently restricted to very local populations, or is it often seen over broad geographic ranges?
+To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?
+How did the conscious, subconscious, and unconscious decisions of early farmers facilitate or hinder adaptive introgression into their crops during early domestication?
+How do the practices of contemporary farmers affect the process of adaptive introgression today?
 
-Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics.
-Basic questions:
-To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?
-%\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
-At what geographic scale does adaptive introgression occur?
-Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges?
-To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g.
-temperature, precipitation, altitude, day lengh, etc.)?}
-Can colonizing species served as bridges for gene flow between previously allopatric taxa?
-%\lwang{I feel the second and the third points could be combined.
-How the wild species could introgress into the allopatric landraces?
-Does domesticates serve as bridges?
+Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
+Loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signal of introgression.
+These loci are potentially beneficial targets for crop improvement.
+Furthermore, adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
 
-Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.
-%\lwang{I am not quite clear with this point.
-It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow.
-Right?}
-%\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?
-I hope that question made sense.}
-Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
+\section*{Conclusion}
+
+The study of crop domestication has been revolutionized by the advent and application of genomic tools.
+The genomes of crops and their wild relatives tell a story of give-and-take that extends well beyond the initial stages of domestication.
+Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adapted alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
 

### 2017-05-24 paper.tex
Update paper.tex

@@
 \usepackage[english]{babel}
+\usepackage{colortbl}
+\usepackage{booktabs}
+\usepackage[x11names,dvipsnames,table]{xcolor}
 \bibliographystyle{ieeetr}
@@
 \begin{table}
+\rowcolors{2}{white}{gray!25}
+\begin{center}
+\caption{Programs for identifying genomic regions under introgression} \label{tab:tools}
+\begin{tabular}{llll}
+\\\toprule  
+\rowcolor{white}
+{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule
+
+\rowcolor{gray!25}
+{\emph{\bf chromosome paiting}} &   &   &   \\
+\rowcolor{gray!25}
+Hapmix	& phased haplotype; reference panel		& Price et al. 2009	&  Hufford et al. 2013;  Suarez et al. 2016 \\ 
+\rowcolor{gray!25}
+RASPberry &	phased haplotype &	Wegmann et al. 2011	 & Christe et al. 2016 \\
+\rowcolor{gray!25}
+MultiMix & phased/unphased genotype; reference panel &	Churchhouse and Marchini 2013 &	Eyheramendy et al. 2015 \\
+\rowcolor{gray!25}
+PCAdmix	 & phased haplotype	 & Brisbin et al. 2012	 & Moreno et al. 2014; Pugach et al. 2016 \\
+\rowcolor{gray!25}
+LAMP  &	phased haplotypes; reference panel	 & Sankararaman et al. 2008	 & Patterson et al. 2012 \\
+
+\rowcolor{white}
+{\emph{\bf phylogenetic relationship}} &   &   &   \\
+\rowcolor{white}
+ABBA-BABA/D-statistics	 & biallelic SNP  &	Durand et al. 2011	 &  Heliconius Genome Consortium 2012 \\
+\rowcolor{white}
+fd statistic &	biallelic SNP &	Martin et al. 2015  &	Malinsky et al. 2015; Zhang et al. 2016 \\ 
+\rowcolor{white}
+five taxon D statistics	& biallelic SNP	&  Pease and Hahn 2015	& Fontaine et al. 2015; Pease et al. 2016 \\
+
+\rowcolor{gray!25}
+{\emph{\bf divergence}} &   &   &   \\
+\rowcolor{gray!25}
+Gmin &	biallelic SNP	&  Geneva et al. 2015	&  Kingan et al. 2015 \\
+\rowcolor{gray!25}
+RNDmin	& phased haplotype	& Rosenzweig et al. 2016	&  NA \\
+\rowcolor{gray!25}
+(see .tex file for comment) & biallelic SNP & Racimo et al. 2016 & Sams et al. 2016 \\
+
+%I was not able to add Li's entry to this cell without errors.  I'll put her entry in the line below:
+%U_{A,B,C(w,x,y)} and Q95_{A,B,C(w,y)}
+
+\rowcolor{white}
+{\emph{\bf population structure related}} &   &   &   \\
+\rowcolor{white}
+fineStructure &	phased haplotype &	Lawson et al. 2012	& Skoglund et al. 2015 \\
+\rowcolor{white}
+Globetrotter &	phased haplotype &	Hellenthal et al. 2014 & Scott et al. 2016 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
 \centering
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=17.35cm]{li_figure.png}
+	\caption{Li's caption here.}
+	\label{fig:map}
+\end{figure}
+
 \section*{Re-evaluating concepts of domestication}
@@
 
-\section*{Conclusion}
+\section*{Conclusions}
 

### 2017-06-08 paper.tex
Update paper.tex
minor edits

@@
 However, archaeological and genetic evidence are revealing that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
-This new conception of domestication emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
+This new conception of domestication emphasizes the role of gene flow of beneficial alleles into populations in the process of domestication from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
 
@@
 In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the crop and adaptive genomic regions are retained and increase in frequency as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and potentially purged from the population.
-Much attention has been given to the risk of natural introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and to the history, utility, and methods of breeding programs designed for the introgressiin of desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but the history of natural introgression of wild alleles into domesticated crops thoughout evolutionary timescales has recieved considerably less attention.
-Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prominence of adaptive introgression in crop domestication.
+Much attention has been given to the risk of natural introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and to the history, utility, and methods of breeding programs designed for the introgression of desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but the history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
+Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prominence of adaptive introgression in crop histories and suggesting the need for a reevaluation of the domestication paradigm.
 
-In this review, we will: 1) briefly describe these recent methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe these recent methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, cnn3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 

### 2017-06-09 paper.tex
edits to intro and first part of introgression methods

--- 2017-05-24 paper.tex
@@
 
-\title{Review: Crop adaptation through introgression from wild relatives}
+\title{Review: Redefining Domestication: Adaptive Introgression during Crop Expansion}
 
@@
 
-The process of domestication was once thought to be rapid and geographically-constrained, with crops originating from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
-However, archaeological and genetic evidence are revealing that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
-This new conception of domestication emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
+The process of domestication was once thought to be rapid and geographically constrained, with crops originating from a wild progenitor within one or more geographically defined centers followed by expansion to the modern-day extent of cultivation.
+\mbh{will need citation for this}
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex}.
+\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper}
+This new conception of domestication emphasizes the role of beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion after initial domestication.
 
-Adaptive introgression has three components: Hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
-In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the crop and adaptive genomic regions are retained and increase in frequency as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and potentially purged from the population.
-Much attention has been given to the risk of natural introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and to the history, utility, and methods of breeding programs designed for the introgressiin of desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but the history of natural introgression of wild alleles into domesticated crops thoughout evolutionary timescales has recieved considerably less attention.
-Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prominence of adaptive introgression in crop domestication.
+Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
+In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and removal of undesirable wild background.
+Literature on crop-wild gene flow has focused on the risk of introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
+The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
+Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
+Emerging results suggest a need for reevaluation of the existing domestication paradigm.
 
-In this review, we will: 1) briefly describe these recent methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
-\section*{Introgression methods and their applications}
+\section*{Introgression methods and their application}
 
-The recent availability of genome-wide resequencing and reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq) data, combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
+High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
+The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}).
+Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity and differentiation and on phylogenetic relationships.
+%We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
 
-High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and to identify introgressions across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
-The methods introduced here do not include those marginally estimating introgression\slash migration rate as a component in the complete demographic history, such as Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference (dadi) \cite {gutenkunst2009}, sorts of isolation with migration model \cite{hey2004}, and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}.
-The collection of methods described different layers of genetic variation: diversity and divergence, haplotype variation and phylogenetic relationships.
-We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
+First, introgressed segments are expected to show low differentiation from their source population.
+The $F_{st}$ and $d_{XY}$ statistics and their derivates such as $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
+The former two are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
+Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup.
 
-First, genomic regions of introgression in admixed populations were expected to show a low level of divergence from source population.
-$F_{st}$ and $d_{XY}$, as well as derivates such as $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016}, are representatives of such statistics. 
-The former two are insensitive to rare migrants, and are therefore poorer at detecting recent introgression, while the latter two overcome these disadvantages.
+\mbh{abrupt transition here to Gmin equation...can you smooth this into the text?}
 
@@
 where $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
-In addition, $RND_{min}$ is sensitive to variable mutation rate, which is indicated by branch length to the outgroup taxa.
 
-These statistics were recently developed into several new statistics, such as $U_{A,B,C(w,x,y)}$, by adding the divergence between non-admixed populations ($A$) and source population ($C$) as a comparison to that between admixed populations ($B$) and source population ($C$)\cite{racimo2016}.
-This statistic summarizes number of sites, where one particular allele at frequency $y$ in source population has a frequency bigger than $x$ in admixed population ($B$) and a frequency smaller than $w$ in non-admixed population ($C$).
-Similarly, $Q95_{A,B,C(w,y)}$ just sets a hard cutoff as $95^th$ percentile of allele frequencies in panel B \cite{racimo2016}. 
-These two statistics were also designed to facilitate incorporatioon of two source populations (see details in \cite{racimo2016}).
+These statistics have recently been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
+For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites, where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($C$).
+A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^th$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
+\mbh{not quite sure I follow here, needs to be made more clear} 
+These statistics were further developed to allow specification two source populations (see details in \cite{racimo2016}).
 
@@
 \hline
-Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X[[present, but low]] &  & \cite{papa2003asymmetry} \\
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
 \hline
@@
 \hline
-Wheat (\emph{Tritium monococcum, T.
-dicoccum, T.
-aestivum}) & \emph{T.
-m.
-boeoticum, T.
-dioccoides, T.
-urartu, Aegilops speltoides, A.
-tauschii} & X & X[[?hybrid origins]] &  & \cite{zohary1969wild} \\
+Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
+\hline
+Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
 \hline

### 2017-06-09 paper.tex
edits through introgression methods

@@
 In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and removal of undesirable wild background.
-Literature on crop-wild gene flow has focused on the risk of introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
+To date, literature on crop-wild gene flow has focused on the risk of introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
+\mbh{cite one paper here...the best, most recent and comprehensive review}
 The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
-Recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
+However, recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
 Emerging results suggest a need for reevaluation of the existing domestication paradigm.
@@
 First, introgressed segments are expected to show low differentiation from their source population.
-The $F_{st}$ and $d_{XY}$ statistics and their derivates such as $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
-The former two are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
+The $F_{st}$ and $d_{XY}$ statistics and their derivates including $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
+The former two statistics are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
 Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup.
@@
 For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites, where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($C$).
-A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^th$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
-\mbh{not quite sure I follow here, needs to be made more clear} 
-These statistics were further developed to allow specification two source populations (see details in \cite{racimo2016}).
+A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
+\mbh{Did I get this right, Li?} 
+Further modifications have allowed specification of two source populations (see details in \cite{racimo2016}).
 
-Second, local ancestry reconvolution (also known as chromosome painting) can identify the genomic regions that come from different source populations \cite{schraiber2015}. 
-Such methods usually take phased haplotypes as input, but MULTIMIX \cite{churchhouse2013} can take both phased and unphased haplotypes as input, unveiling the genomic ancestry from multiple source populations.
+Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations \cite{schraiber2015}.
+\mbh{How?
+How do these methods work?
+I know at least some are based on Hidden Markov Models...need a sentence or two on the nuts and bolts}
+Such methods typically require phased haplotypes as input (but see \cite{churchhouse2013}).
 
-Third, ABBA-BABA statistics (also known as D-statistics) and their derivatives are widely applied to the search for signs of introgression in genomic patterns of shared derived variants between populations or species.
-Based on D-statistics, $\hat{f_{d}}$ \cite{Martin2015} and five-taxon D statistics \cite{pease2015} were developed to localize genomic regions under introgression. 
-The former took allele frequencies from each population/species, and the latter detected introgression from the localized phylogenetic pattern.
-Their strength in identifying specific regions enables one to explore the relationships between introgression and recombination rate/gene density/distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
+Third, the ABBA-BABA statistic (also known as the D-statistic) and its derivatives are widely applied to introgression detection and make inferences based on genomic patterns of shared derived variants between populations or species in a phylogenetic context.
+\mbh{need sentence here explaining what D-statistic does}
+Elaborations of the D-statistic include $\hat{f_{d}}$ \cite{Martin2015} and the five-taxon D-statistic \cite{pease2015}. 
+The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern.
+\mbh{need to clarify here how this is different than $\hat{f_{d}}$}
+%Their strength in identifying specific regions enables one to explore the relationships between introgression and recombination rate/gene density/distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
 
-Fourth, structure-related approaches are based on the popular software STRUCTURE \cite{pritchard2000}, which provides an estimate of the fraction of each individual's genome coming from each population.
-Many approaches in this respect have been developed.
-Two of these, fineSTRUCTURE \cite{Lawson2012} and GLOBETROTTER \cite{hellenthal2014}, were selected as representations.
-Different from STRUCTURE, fineSTRUCTURE not only identifies the number of populations, but also estimates the probability that genomic regions belong to one or the other populations. 
-GLOBETROTTER is a development of fineSTRUCTURE that allows for tracing back to an unsampled source population and dating admixture events.
+Fourth, approaches that assign genome-wide genetic ancestry are based on the popular software STRUCTURE \cite{pritchard2000}. 
+For example, fineSTRUCTURE \cite{Lawson2012} calculates the probability that genomic regions belong to genetic groups by \mbh{need details on method here}. 
+GLOBETROTTER \cite{hellenthal2014} is a further development of fineSTRUCTURE that allows for identification of an unsampled source population and for dating of admixture events.
 
 Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry across butterfly species (\cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry \mbh{I would explain a bit more here...wing coloration loci, protects against predation...} across butterfly species \cite{Heliconius2012} introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
 

### 2017-06-09 paper.tex
edits to maize and rice

@@
 \emph{mays}) and the teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive crop/wild introgression.
-Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the Balsas River Valley in Mexico \cite{matsuoka2002single}.
-From this domestication center, maize spread into the highlands of the Mexican central plateau, where it came into sympatry with wild \emph{mexicana}.
-Introgression between \emph{mexicana} and Mexican highland maize has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analyses \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
-However, \citep{hufford2013} was the first to reveal evidence that introgression from \emph{mexicana} into Mexican highland maize is adaptive.
+\emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
+Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with wild \emph{mexicana}.
+Introgression from \emph{mexicana} to maize in the highlands of Mexico has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analyses \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
+However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions which showed evidence of introgression from \emph{mexicana} to maize in both the HAMPMIX and the linkage model of STRUCTURE analyses with over seven sympatric population pairs among the nine pairs sampled.
-Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
-By further characterizing the nine introgression regions, it is found that most regions contain long tracts of zero diversity, enriched with QTL linked with anthocyanin content and leaf macrohairs \cite{lauter2004} and over-represented with the SNPs demonstrating high association with temperature seasonality.
-It was also revealed that in the growth chamber experiment, maize populations with introgression from \emph{mexicana} on chromosomes 4 (associated with QTLs controlling pigment density and macrohairs) and 9 (overlapped with QTLs for macrohairs) exhibited more macrohairs and greater pigmentation under the highland environmental settings than the populations with absence of introgression from \emph{mexicana}.
-In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
+The authors identified nine genomic regions in several maize populations which showed evidence of\emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE.
+These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be important in adaptation to high elevation.
+In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental settings than populations that lacked introgression.
+Height differences were not detected under lowland conditions.
+%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
+%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
 
-Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations which are allopatric to \emph{mexicana}.
-A recent study \cite{Takuno2015} found little empirical or theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly by the difference in potential for adaptive introgression from wild relatives.
-Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.
-Furthermore, the adaptive SNPs in the Mexican highland population were more likely also showing signatures of local adaptation in \emph{mexicana} and \emph {pulviglumis} populations than those South American highland population.
-For these reasons, adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.
+%Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations which are allopatric to \emph{mexicana}.
+%A recent study \cite{Takuno2015} found little empirical or theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly by the difference in potential for adaptive introgression from wild relatives.
+%Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.
+%Furthermore, the adaptive SNPs in the Mexican highland population were more likely also showing signatures of local adaptation in \emph{mexicana} and \emph {pulviglumis} populations than those South American highland population.
+%For these reasons, adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.
 
-Several questions regarding \emph{mexicana} introgression yet remain.
-Has the introgression from \emph{mexicana} spread to multiple highland populations? 
-How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
-A recent study \cite{Wang2015manuscript} first utilized the ABBA-BABA statistics to evaluate the existence or absence of introgression from \emph{mexicana} to multiple highland populations, and then $\hat{f_{d}}$ statistic proposed by Martin et al.
-(2015) \citep{martin2015} were calculated to locate the introgressed loci.
-Figure \ref{fig:fd} shows the loess regression of $\hat{f_{d}}$ in $10kb$ non\-overlapping windows across chromosomes 4 and 5 in multiple comparisons. 
-On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
+Populations of \emph{mexicana} cannot be found outside of Mexico, yet maize has colonized and adapted to high elevation in a number of additional regions.
+%Several questions regarding \emph{mexicana} introgression yet remain.
+%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
+%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
+A recent study \cite{Wang2015manuscript} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico (Figure \ref{fig:fd}).
+%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
 (2013) \citep{hufford2013}.
-The signal of introgression is absent for the other three populations.
-On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
-More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
-Overall, both analyses revealed that gene flow from \emph{mexicana} exists, not only to MexHigh maize, but also to GuaHigh maize, as well as to some individuals in SW\_US, but not to the Andes and South American lowland (SA\_Low) populations, suggesting different extent contributions \emph{mexicana} made to multiple maize highland landrace populations.
-The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
+%The signal of introgression is absent for the other three populations.
+%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
+%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
+Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but that more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+\mbh{on the fence about including deleterious allele results from introgressed regions}
+%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
 %The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
-The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
-Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
-Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
+%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
+%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
 
@@
 Introgression appears to have played an important role in the domestication history of rice.
-The center of Asian rice (\emph{Oryza sativa}) domestication is not known with complete confidence, but genetic and archaeobotanical evidence points towards both the Yangzee Basin in China and the Ganges plains in India for \emph{O.
-sativa}, 8,200-13,500 years ago, from wild \emph{O.
-rufipogon}  \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}\ or "the Asian form of \emph{O.
-perennis} complex" \cite{oka2012origin}.
-Asian rices readily hybridize both with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}).
+The center of Asian rice (\emph{Oryza sativa}) domestication is not known with complete confidence, but genetic and archaeobotanical evidence point toward both the Yangzee Basin in China and the Ganges plains in India.
+\mbh{there is a serious debate over whether there were one or two domestication centers of rice...I'd be sure to read up on the most current thoughts about this so that we don't upset potential reviewers}
+Domestication occurred 8,200-13,500 BP from the wild species \emph{O. rufipogon} \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}. \mbh{here limit to one or two citations}
+Asian rice readily hybridizes with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}).
+\mbh{probably best to find a primary literature citation for this (journal article rather than ricepedia).
+What other domesticated subspecies are referred to here?
+Also, when referring to Asian rice, is this both japonica and indica?
+Finally, what are the twenty relatives?
+My understanding is gene flow has mostly been with rufipogon and nivara}
 The high genetic diversity within \emph{O.
-sativa} is likely due to introgression with wild relatives both during domestication and upon the dispersal of \emph{O.
-sativa} into new environments and sympatry with new relatives \cite{second1982origin}.
+sativa} is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed following domestication \cite{second1982origin}.
 
-In an attempt to categorize the varieties of Asian rice, \cite{khush2003classifying} performed isozyme polymorphism analysis (11 enzymes, 20 loci, 79 alleles) on 25,519 rice varieties.
-The vast majority of the rice varieties fell into the group indica (73.45\%) or japonica (23.02\%).
+Asian rice can be broadly clustered into two groups, japonica and indica \cite{khush2003classifying}.
+%The vast majority of the rice varieties fell into the group indica (73.45\%) or japonica (23.02\%).
 Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
-However, about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
-Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
-Introgression of alleles between cultivated rices, as well as between cultivated and wild rices, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticate is suspected (\cite{zhao2010genomic} calls for research into this possibility).
-These rices often have blight resistance gene, likley imparted from the wild relative parent.
+\mbh{Is this thought to be due to adaptation?
+Did the wild rice populations confer local adaptation?
+If so, I'd state this and explain what type of adaptation this might have been if it's known}
+%However, about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
+%Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
+Introgression of alleles between cultivated rice populations, as well as between cultivated and wild rice, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticated rice is suspected (\cite{zhao2010genomic} calls for research into this possibility).
+\mbh{strange sentence here...says introgression between cultivated and wild rice is common but then says introgression from wild to cultivated is just suspected...need to clarify}
+%These rices often have blight resistance gene, likely imparted from the wild relative parent.
 Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) are known to have been introgressed from wild relatives into \emph{O.
 sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
-Some aus and rayada varieties retain characteristics of the putative wild parent.
+\mbh{don't think this is what we're after in these case studies.
+We want to narrow our focus to natural introgression from wild to crop species during the expansion of the crop.
+Wild introgression during breeding is interesting, but not the focus of the paper}.
+Some aus and rayada varieties retain characteristics of the putative wild parent. \mbh{more background on aus, rayada and aswina...what is there origin? how doe they relate to japonica and indica?}
 Aus exhibits a sprawling growth pattern and easy-threshing grain.
@@
 rufipogon} populations.
+\mbh{need citations regarding these varieties}
 
@@
 Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}.
-Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China.
+Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China. \mbh{again, don't think this is what we're after with these case studies}
 
@@
 
-Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
+Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
 Barley (\emph{Hordeum vulgare} subsp.
-\emph{vulgare}) is believed to have been domesticated at least twice, once from wild subsp.
+\emph{vulgare}) is believed to have been domesticated at least twice roughly 10,000 BP, once from wild subsp.
 \emph{spontaneum} in the Fertile Crescent and once from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\.
-However, many of the details of barley domestication are still disputed.
+\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}.
+However, details of barley domestication are still disputed.
 

### 2017-06-12 paper.tex
Update paper.tex

@@
 \mbh{will need citation for this}
+\gmj{ the paper "Domestication rates in wild-type wheats and
+barley under primitive cultivation" states that domestication could have taken place in as little as 20 or 30 years.}
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex}.
-\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper}
+\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper} \gmj{also include the paper I just read, "ARCHAEOLOGICAL DATA REVEAL SLOWRATES OF EVOLUTION DURING PLANTDOMESTICATION".
+This paper suggests that the rate of phenotypic change in domestication is slower than expected, nearly that of natural selection in wild populations.
+Also, see the first paragraph of the Discussion for a list of older papers supporting rapid domestication.}
 This new conception of domestication emphasizes the role of beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion after initial domestication.

### 2017-06-12 paper.tex
Create paper.tex

@@
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex}.
-\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper} \gmj{also include the paper I just read, "ARCHAEOLOGICAL DATA REVEAL SLOWRATES OF EVOLUTION DURING PLANTDOMESTICATION".
+\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper} \gmj{the rice paper you mentioned, I think, is "ARCHAEOLOGICAL DATA REVEAL SLOWRATES OF EVOLUTION DURING PLANTDOMESTICATION".
 This paper suggests that the rate of phenotypic change in domestication is slower than expected, nearly that of natural selection in wild populations.

### 2017-06-13 paper.tex
partial comments addressed

--- 2017-06-09 paper.tex
@@
 \mbh{will need citation for this}
-However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex}.
-\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper}
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017}.
+\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper} \lwang{I have added these three papers.}
 This new conception of domestication emphasizes the role of beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion after initial domestication.
@@
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
-The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and the multiple sequentially Markovian coalescent (MSMC) \cite{schiffels2014}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity and differentiation and on phylogenetic relationships.
+The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of method utilizing sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
+Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity, level of differentiation and on phylogenetic relationships.
 %We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
@@
 \mbh{abrupt transition here to Gmin equation...can you smooth this into the text?}
-
- \begin{equation}
-    G_{min} = \frac{d_{min}}{d_{XY}}
- \end{equation}
- where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y.
+\lwang{I am not much a fan of Gmin, as we applied this method and it is not good. I am wondering if we just give the equation of RNDmin, in order to focus readers attention on the later.}
+% \begin{equation}
+%    G_{min} = \frac{d_{min}}{d_{XY}}
+% \end{equation}
+% where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y.
 
@@
  \end{equation}
-where $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
+where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y; $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
 
@@
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
-\mbh{Did I get this right, Li?} 
+\mbh{Did I get this right, Li?} \lwang{yes}
 Further modifications have allowed specification of two source populations (see details in \cite{racimo2016}).
@@
 
-Fourth, approaches that assign genome-wide genetic ancestry are based on the popular software STRUCTURE \cite{pritchard2000}. 
+Fourth, approaches that assign genome-wide genetic ancestry based on the popular software STRUCTURE \cite{pritchard2000} are another category to detect introgression. 
 For example, fineSTRUCTURE \cite{Lawson2012} calculates the probability that genomic regions belong to genetic groups by \mbh{need details on method here}. 
@@
 Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}), introgression has conferred M\"{u}llerian mimicry \mbh{I would explain a bit more here...wing coloration loci, protects against predation...} across butterfly species \cite{Heliconius2012} introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
-
-\noindent{\textbf{Table 1:} List and brief description of recently developed methods and examples of empirical studies employing these methods.}
-
-\noindent{\textbf{Figure 1}: Wing coloration patterns in \emph{Heliconius} and evidence for introgression across species based on Patterson's \emph{D}-statistic; adapted from \cite{Heliconius2012}.}
-
-%\lwang{In this part, maybe better to incorporate this reference, Detection and Polarization of Introgression in a Five-Taxon Phylogeny; and a new method published this year RND\_min Powerful methods for detecting introgressed regions from population genomic data; I donot know if we should also include G\_min, that is the statistic we tried that didnot work well for us: A new method to scan genomes for introgression in a secondary contact model.
-I put the three in the bib file.}
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}); introgression has conferred M\"{u}llerian mimicry \mbh{I would explain a bit more here...wing coloration loci, protects against predation...} across butterfly species \cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
 
@@
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions in several maize populations which showed evidence of\emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE.
+The authors identified nine genomic regions in several maize populations which showed evidence of\emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
 These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be important in adaptation to high elevation.
@@
 %How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
-A recent study \cite{Wang2015manuscript} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico (Figure \ref{fig:fd}).
+A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
 %On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
@@
 %More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
-Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but that more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
 \mbh{on the fence about including deleterious allele results from introgressed regions}
@@
 %Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=17.35cm]{li_figure.png}
+	\caption{Li's caption here.}
+	\label{fig:introgressionMaize}
+\end{figure}
 
@@
 \begin{center}
-\caption{Programs for identifying genomic regions under introgression} \label{tab:tools}
+\caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
 \begin{tabular}{llll}
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=17.35cm]{li_figure.png}
-	\caption{Li's caption here.}
-	\label{fig:map}
-\end{figure}
-
 \section*{Re-evaluating concepts of domestication}

### 2017-06-13 paper.tex
edits to intro.

@@
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017}.
-\mbh{citations here to Li's paper, Purugganan rice paper, Brandon Gaut's recent grape paper} \lwang{I have added these three papers.}
 This new conception of domestication emphasizes the role of beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion after initial domestication.
@@
 The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
-However, recent tools and methods have been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
+However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
 Emerging results suggest a need for reevaluation of the existing domestication paradigm.
@@
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
-The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of method utilizing sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity, level of differentiation and on phylogenetic relationships.
+The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
+Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity, the extent of differentiation, and phylogenetic relationships.
 %We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
@@
  \end{equation}
-where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y; $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
+where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
 
@@
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
-\mbh{Did I get this right, Li?} \lwang{yes}
 Further modifications have allowed specification of two source populations (see details in \cite{racimo2016}).
@@
 
-Third, the ABBA-BABA statistic (also known as the D-statistic) and its derivatives are widely applied to introgression detection and make inferences based on genomic patterns of shared derived variants between populations or species in a phylogenetic context.
-\mbh{need sentence here explaining what D-statistic does}
-Elaborations of the D-statistic include $\hat{f_{d}}$ \cite{Martin2015} and the five-taxon D-statistic \cite{pease2015}. 
+Third, the ABBA-BABA statistic (also known as the D-statistic) and its derivatives are widely applied to introgression detection.
+These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
+Patterns of allele sharing are interpreted in a phylogenetic context and the method is best suited to detection of introgression at the genome level.
+Elaborations of the D-statistic capable of localizing introgression to specific genomic regions include $\hat{f_{d}}$ \cite{Martin2015} and the five-taxon D-statistic \cite{pease2015}. 
 The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern.
-\mbh{need to clarify here how this is different than $\hat{f_{d}}$}
 %Their strength in identifying specific regions enables one to explore the relationships between introgression and recombination rate/gene density/distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
 
-Fourth, approaches that assign genome-wide genetic ancestry based on the popular software STRUCTURE \cite{pritchard2000} are another category to detect introgression. 
+The final category to detect introgression include approaches assigning genetic ancestry based on the popular software STRUCTURE \cite{pritchard2000}. 
 For example, fineSTRUCTURE \cite{Lawson2012} calculates the probability that genomic regions belong to genetic groups by \mbh{need details on method here}. 
@@
 
-Introgression appears to have played an important role in the domestication history of rice.
-The center of Asian rice (\emph{Oryza sativa}) domestication is not known with complete confidence, but genetic and archaeobotanical evidence point toward both the Yangzee Basin in China and the Ganges plains in India.
-\mbh{there is a serious debate over whether there were one or two domestication centers of rice...I'd be sure to read up on the most current thoughts about this so that we don't upset potential reviewers}
-Domestication occurred 8,200-13,500 BP from the wild species \emph{O. rufipogon} \cite{oka2012origin, fuller2010consilience, ricepedia, molina2011molecular}. \mbh{here limit to one or two citations}
-Asian rice readily hybridizes with other domesticated subspecies and with wild relatives (of which there are about twenty \cite{ricepedia}).
-\mbh{probably best to find a primary literature citation for this (journal article rather than ricepedia).
-What other domesticated subspecies are referred to here?
-Also, when referring to Asian rice, is this both japonica and indica?
-Finally, what are the twenty relatives?
-My understanding is gene flow has mostly been with rufipogon and nivara}
-The high genetic diversity within \emph{O.
-sativa} is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed following domestication \cite{second1982origin}.
+The story of the domestication of asian rice is still under debate.
+A recent model utilizing molecular clocks points to single domestication of asian rice that occurred 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{O.
+rufipogon}, with later divergence of the two prominent varietal groups indica and japonica lines \cite{molina2011molecular}.
+Sequence data from indica, japonica, and wild O.
+rufipogon accessions points at japonica domestication in the Pearl Valley in China, and indica being formed from hybridization between japonica and rufipogon in Southern and South-eastern Asia \cite{Huang2012}.
+On the other hand, genetic and archaeobotanical evidence points towards independent domestications of the indica and japonica subspecies in the Yangzee Basin and the Ganges plains in India \cite{fuller2010consilience}.
 
-Asian rice can be broadly clustered into two groups, japonica and indica \cite{khush2003classifying}.
-%The vast majority of the rice varieties fell into the group indica (73.45\%) or japonica (23.02\%).
-Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
-\mbh{Is this thought to be due to adaptation?
-Did the wild rice populations confer local adaptation?
-If so, I'd state this and explain what type of adaptation this might have been if it's known}
-%However, about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
-%Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
-Introgression of alleles between cultivated rice populations, as well as between cultivated and wild rice, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticated rice is suspected (\cite{zhao2010genomic} calls for research into this possibility).
-\mbh{strange sentence here...says introgression between cultivated and wild rice is common but then says introgression from wild to cultivated is just suspected...need to clarify}
-%These rices often have blight resistance gene, likely imparted from the wild relative parent.
-Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) are known to have been introgressed from wild relatives into \emph{O.
-sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
-\mbh{don't think this is what we're after in these case studies.
-We want to narrow our focus to natural introgression from wild to crop species during the expansion of the crop.
-Wild introgression during breeding is interesting, but not the focus of the paper}.
-Some aus and rayada varieties retain characteristics of the putative wild parent. \mbh{more background on aus, rayada and aswina...what is there origin? how doe they relate to japonica and indica?}
-Aus exhibits a sprawling growth pattern and easy-threshing grain.
-Rayada is adapted to areas with longer periods of flooding, with heightened seed dormancy and photoperiod sensitivity (permitting harvest during times without standing water).
-Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between \emph{O.
-sativa} and deepwater-adapted \emph{O.
-rufipogon} populations.
-\mbh{need citations regarding these varieties}
-
-In addition to investigative experiments, gene flow from wild relatives has been utilized to produce agronomic rice varieties.
-Yatsen No. 1, for example, showed resistance to pests and diseases and adapted well to environmental conditions \cite{ting1933wild}.
-Several lines were derived from Yatsen No. 1, and went on to be utilized extensively in parts of China. \mbh{again, don't think this is what we're after with these case studies}
+The authors of \cite{vaughan2008evolving} review the evidence for both hypotheses and support the single domestication event hypothesis, contributing the contrary evidence to the effects of a history of introgression, selection (primarily that of the non-shattering trait allele sh4), and diversification.
+In their defense of the single-domestication hypothesis, the authors review the case that, during the spread of domesticated rice into new environments, gene flow from wild relatives would have imparted adaptive traits via introgression.
+The wild relatives rufipogon and nivara are both phenotypically diverse and show evidence of local adaptation.
+Rufipogon reproduction can be either primarily sexual or vegetative.
+Whereas rufipogon is adapted to forested wetland environments, nivara is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
+Likewise, cultivated rice varieties display phenotypic differences important for fitness.
+In particular, two of the cultivated rice deepwater varieties (rayada and ashwina) are said to be selected for the environment along the Ganges river.
+Gene flow between these wild and cultivated rices is common, producing nuisance weedy hybrids in when the two are grown in sympatry.
+However, during early domestication, introgression may have been even more prevalent because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established.
+Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
+Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
+The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}.
 

### 2017-06-13 paper.tex

@@
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based on patterns of nucleotide/haplotype diversity, the extent of differentiation, and phylogenetic relationships.
+Rather, we focus on methods that explicitly identify introgressed genomic segments based the extent of differentiation, on patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 %We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
@@
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
-Further modifications have allowed specification of two source populations (see details in \cite{racimo2016}).
+Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
 
-Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations \cite{schraiber2015}.
-\mbh{How?
-How do these methods work?
-I know at least some are based on Hidden Markov Models...need a sentence or two on the nuts and bolts}
-Such methods typically require phased haplotypes as input (but see \cite{churchhouse2013}).
+Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations based on patterns of allele/haplotype sharing \cite{schraiber2015}. 
+One form of chromosome painting utilizes hidden Markov models to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
+Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
+And finally, introgression can be detected through chromosome painting by using a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
 
@@
 Elaborations of the D-statistic capable of localizing introgression to specific genomic regions include $\hat{f_{d}}$ \cite{Martin2015} and the five-taxon D-statistic \cite{pease2015}. 
-The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern.
-%Their strength in identifying specific regions enables one to explore the relationships between introgression and recombination rate/gene density/distribution of deleterious alleles and to better understand the genomic regions penetrable to foreign gene flow.
-
-The final category to detect introgression include approaches assigning genetic ancestry based on the popular software STRUCTURE \cite{pritchard2000}. 
-For example, fineSTRUCTURE \cite{Lawson2012} calculates the probability that genomic regions belong to genetic groups by \mbh{need details on method here}. 
-GLOBETROTTER \cite{hellenthal2014} is a further development of fineSTRUCTURE that allows for identification of an unsampled source population and for dating of admixture events.
+The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern and is capable of determining introgression directionality.
 
@@
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with wild \emph{mexicana}.
-Introgression from \emph{mexicana} to maize in the highlands of Mexico has been reported based on evidence from both morphological data \cite {wilkes1977, lauter2004, doebley1984} and molecular analyses \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005}.
+Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
 However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions in several maize populations which showed evidence of\emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
-These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be important in adaptation to high elevation.
-In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental settings than populations that lacked introgression.
+The authors identified nine genomic regions in several maize populations which showed evidence of \emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
+These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be important in high elevation adaptation.
+In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
 Height differences were not detected under lowland conditions.
@@
 
-Populations of \emph{mexicana} cannot be found outside of Mexico, yet maize has colonized and adapted to high elevation in a number of additional regions.
+Populations of \emph{mexicana} cannot be found outside of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
 %Several questions regarding \emph{mexicana} introgression yet remain.
@@
 Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-\mbh{on the fence about including deleterious allele results from introgressed regions}
 %The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
@@
 
-The story of the domestication of asian rice is still under debate.
-A recent model utilizing molecular clocks points to single domestication of asian rice that occurred 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{O.
-rufipogon}, with later divergence of the two prominent varietal groups indica and japonica lines \cite{molina2011molecular}.
-Sequence data from indica, japonica, and wild O.
-rufipogon accessions points at japonica domestication in the Pearl Valley in China, and indica being formed from hybridization between japonica and rufipogon in Southern and South-eastern Asia \cite{Huang2012}.
-On the other hand, genetic and archaeobotanical evidence points towards independent domestications of the indica and japonica subspecies in the Yangzee Basin and the Ganges plains in India \cite{fuller2010consilience}.
+The story of Asian rice domestication is still debated.
+A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon}, with later divergence of the two prominent varietal groups, indica and japonica \cite{molina2011molecular}.
+On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of japonica and indica in the Yangtze Basin and the Indian Ganges plain respectively \cite{fuller2010consilience}.
+An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of japonica in China followed by formation of the indica group through hybridization of japonica and local \emph{O.
+rufipogon} populations in Southern and South-eastern Asia.
+As suggested by \cite{vaughan2008evolving}, gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits.
+Adaptive introgression may in fact be the signal interpreted as independent domestications.
 
-The authors of \cite{vaughan2008evolving} review the evidence for both hypotheses and support the single domestication event hypothesis, contributing the contrary evidence to the effects of a history of introgression, selection (primarily that of the non-shattering trait allele sh4), and diversification.
-In their defense of the single-domestication hypothesis, the authors review the case that, during the spread of domesticated rice into new environments, gene flow from wild relatives would have imparted adaptive traits via introgression.
 The wild relatives rufipogon and nivara are both phenotypically diverse and show evidence of local adaptation.
@@
 	\centering
-	\includegraphics[width=17.35cm]{par_figure.png}
+	\includegraphics[width=17.35cm]{par_figure_stars.png}
 	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}

### 2017-06-13 paper.tex

@@
 \emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}.
+\gmj{the source morrell2007genetic will be used to update this paragraph once we've worked out for sure that the second domestication of barley was in the Zargos mountains, rather than Tibet.
+Also, this paper puts the earliest archaeological samples of barley at closer to 8500 calibrated years ago.}
 However, details of barley domestication are still disputed.

### 2017-06-14 paper.tex
re-worked the rice section.

@@
 
-\item{Asian Rice:}
-
-The story of Asian rice domestication is still debated.
-A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon}, with later divergence of the two prominent varietal groups, indica and japonica \cite{molina2011molecular}.
-On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of japonica and indica in the Yangtze Basin and the Indian Ganges plain respectively \cite{fuller2010consilience}.
-An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of japonica in China followed by formation of the indica group through hybridization of japonica and local \emph{O.
-rufipogon} populations in Southern and South-eastern Asia.
-As suggested by \cite{vaughan2008evolving}, gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits.
-Adaptive introgression may in fact be the signal interpreted as independent domestications.
-
-The wild relatives rufipogon and nivara are both phenotypically diverse and show evidence of local adaptation.
-Rufipogon reproduction can be either primarily sexual or vegetative.
-Whereas rufipogon is adapted to forested wetland environments, nivara is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
-Likewise, cultivated rice varieties display phenotypic differences important for fitness.
-In particular, two of the cultivated rice deepwater varieties (rayada and ashwina) are said to be selected for the environment along the Ganges river.
-Gene flow between these wild and cultivated rices is common, producing nuisance weedy hybrids in when the two are grown in sympatry.
-However, during early domestication, introgression may have been even more prevalent because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established.
-Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
-Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
-The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}.
-
 \item{Barley:}
@@
 %they also show that traits were imparted from debilis to texanus, and that these are adaptive.  i think they also show genetic evidence, but i'm not sure.
+
+\item{Asian Rice:}
+
+The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
+A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
+On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+Adaptive introgression may in fact be the signal interpreted as independent domestications.
+
+The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
+The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high phenotypic diversity and exhibit locally-adaptive traits.
+\emph{Rufipogon} reproduction can be either primarily sexual or vegetative, and whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
+Likewise, domesticated rice varieties display patterns of local adaptation.
+For examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+Gene flow between these wild and cultivated rices, though asymmetrical, is exceedingly common, producing nuisance weedy hybrids when the two are grown in sympatry.
+
+To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples.
+\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgrogression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
+There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
+As with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
+This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
+However, during early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
+Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
+
+%Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
 

### 2017-06-14 paper.tex
Update paper.tex

@@
 A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
+\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
 On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.

### 2017-06-15 paper.tex
edits to case studies
continued edits to case studies

@@
 
-The process of domestication was once thought to be rapid and geographically constrained, with crops originating from a wild progenitor within one or more geographically defined centers followed by expansion to the modern-day extent of cultivation.
+The process of domestication was once thought to be rapid and geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation.
 \mbh{will need citation for this}
@@
 In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and removal of undesirable wild background.
-To date, literature on crop-wild gene flow has focused on the risk of introgression of transgenes from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
+To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
 \mbh{cite one paper here...the best, most recent and comprehensive review}
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Emerging results suggest a need for reevaluation of the existing domestication paradigm.
+Emerging results suggest a need for reevaluating the existing domestication paradigm.
 
-In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of their application for detecting crop-wild introgression, 2) review evidence suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
 \section*{Introgression methods and their application}
+
+\mbh{In this section, I think the overall content is good, but we need to edit to make it more accessible and more explicit about how methods are implemented to detect adaptive introgression}
 
@@
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based the extent of differentiation, on patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+Rather, we focus on methods that explicitly identify introgressed genomic segments based on the extent of differentiation, on patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 %We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
@@
 The former two statistics are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
-Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup.
+Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup:
 
-\mbh{abrupt transition here to Gmin equation...can you smooth this into the text?}
-\lwang{I am not much a fan of Gmin, as we applied this method and it is not good. I am wondering if we just give the equation of RNDmin, in order to focus readers attention on the later.}
 % \begin{equation}
@@
 These statistics have recently been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
-For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites, where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($C$).
+For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($C$).
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
@@
 Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
-From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with wild \emph{mexicana}.
+From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
 Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
@@
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions in several maize populations which showed evidence of \emph{mexicana} introgression based on local ancestry deconvolution using HAPMIX and results from the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
-These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be important in high elevation adaptation.
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
+These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
@@
 
-Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
+%Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
 Barley (\emph{Hordeum vulgare} subsp.
-\emph{vulgare}) is believed to have been domesticated at least twice roughly 10,000 BP, once from wild subsp.
+\emph{vulgare}) was domesticated at least twice roughly 10,000 BP: once from the wild subsp.
 \emph{spontaneum} in the Fertile Crescent and once from subsp.
@@
 Also, this paper puts the earliest archaeological samples of barley at closer to 8500 calibrated years ago.}
-However, details of barley domestication are still disputed.
-
+%However, details of barley domestication are still somewhat unclear.
 %Variety \emph{agriocrithon} is genetically diverse, and is found throughout much of the range of barley.
@@
 \emph{spontaneum}, perhaps more than once (in Israel as well as in Tibet).
-%Presently, the distribution of \emph{spontaneum} consists of the Mediterranean, the Middle-East, and west-central Asia, while other barley wild relatives inhabit Asiatic regions (including Tibet) more broadly \cite{nevo2010drought, harlan1995living, CWR}.
-%These wild relatives inhabit regions spanning such abiotic clines as temperature, precipitation, soil type, and altitude, as well as biotic clines \cite{nevo2010drought}.
+Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
+%While little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}.
+Barley/\emph{spontaneum} hybrids are known to be fertile and are found spontaneously when wild and domesticated barleys co-occur.
+In some cases, wild-to-crop introgression has been shown to occur over distances of more than a kilometer \cite{hillman2001new}.
 
-There has been little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}.
-Barley/\emph{spontaneum} hybrids are fertile, and morphologically intermediate (putatively hybrid) barleys are found when wild and domesticated barleys are grown in sympatry, but hybrids of other wild relatives generally exhibit greatly diminished fertility \cite{ellstrand2003dangerous, harlan1995living}.
-Even when the two are not grown immediately adjacent to one another, introgression from wild to domesticate has been shown to happen over distances of more than a kilometer \cite{hillman2001new}.
-
-The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}.
+%The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}.
+\mbh{I think here a more focused discussion of potential for adaptive introgression based on the Poets paper is needed}
 The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces, and found a high amount variability in the amount of contribution from wild relatives, as well as its location in the genome, within barley populations.
@@
 
-The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}.
-Pre-Columbian \emph{H.
-annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
-
-Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H.
+The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
+The pre-Columbian \emph{H.
+annuus} distribution of cultivated sunflower spanned much of the Great Plains, from what is now north-central Texas to Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
+Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
 petiolaris} and \emph{H.
-bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
-Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
+bolanderi} and forms stable hybrid populations with these taxa \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
+Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
 However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
+\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
 Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.

### 2017-06-16 paper.tex
Update paper.tex

@@
 And finally, introgression can be detected through chromosome painting by using a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+
+\gmj{Li, are you familiar with the analytical tools MIGRATE-N and BAYESASS?  Rieseberg comapres these two to STRUCTURE at some length in the discussion of this paper:
+https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
+Should we include these two methods, even in passing, in this portion of the paper?}
 

### 2017-06-16 paper.tex
Update paper.tex

@@
 These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
+
+Notes for review:
+
+Genome scan of hybridizing sunflowers from Texas
+(Helianthus annuus and H. debilis) reveals asymmetric
+patterns of introgression and small islands of genomic
+differentiation:
+
+In contrast, coalescent analyses of long-term migration
+rates imply that interspecific migration has been
+more important than mutation in providing new genetic
+variation in ssp. texanus populations. Immigration from
+H. debilis cucumerifolius into ssp. texanus (M = 4.05;
+Nem = 1.74) is higher than median estimates of intraspecific
+migration rates in plants (Morjan and Rieseberg 2004) 
+and only slightly lower than immigration from
+ssp. annuus into ssp. texanus (M = 5.00). Immigration
+rates from ssp. texanus into H. debilis cucumerifolius and
+ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
+implying that the putative hybrid lineage serves
+as a bridge for migration between the parental species.
+
+One puzzle is
+that the estimated immigration rate between the allopatric
+taxa (H. debilis cucumerifolius and ssp. annuus) is
+approximately comparable with that between the sympatric
+taxa (H. debilis cucumerifolius and ssp. texanus ssp.
+texanus). This observation may point to the effectiveness
+of ssp. texanus as a bridge for the transfer of alleles
+between H. debilis and ssp. annuus. An alternative explanation
+is that the BAYESASS estimates are unreliable
+because homoplasy of microsatellite alleles is high in
+taxa with large-effective population sizes (Estoup et al.
+2002) such as the annual sunflowers studied here.
+
+Despite the apparent discrepancy between the results
+from STRUCTURE and BAYEASS, several conclusions
+can be made. Analyses with STRUCTURE support the
+distinctness of ssp. texanus, as well as its clear inclusion
+within H. annuus, as initially postulated by Heiser (1951).
+Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
+analyses all confirm previous report of introgression
+between H. annuus and H. debilis (Rieseberg et al.
+1990, 2007) and suggest that the introgression is genomewide
+when it occurs. Contrary to the Heiser’s scenario of
+a recent Holocene origin of ssp. texanus, however, our
+results are more consistent with a much longer history of
+contact between H. annuus and H. debilis. Otherwise, it is
+difficult to account for the presence of significant longterm
+migration between currently allopatric populations
+of ssp. annuus and H. debilis cucumerifolius. This revised
+scenario makes sense in the light of numerous glacialinterglacial
+cycles over the past million years (Hewitt
+2000), which probably resulted in intermittent contact
+between H. annuus and H. debilis, with the last contact
+likely occurring during the Wisconsin glaciation,
+18 000 BP. The current contact between the species differs
+from the previous ones in that it appears to have been
+human-aided. Possibly, some of the molecular evidence
+of introgression, particularly into allopatric populations,
+stems from past periods of contact.
+
+One of the four outlier loci, HT0414, is associated
+with a heat-shock protein and was previously shown to
+have been the target of selection in H. annuus ssp. annuus
+salt marsh populations from the state of Utah
+(Kane and Rieseberg 2007), as well as in weedy sunflower populations
+from several different locations across the USA
+(Kane and Rieseberg 2008). In this study, populations of
+ssp. annuus from Kansas, Nebraska and Oklahoma
+monomorphic for a single allele, whereas populations
+of H. debilis and other populations of ssp. annuus are
+considerably more polymorphic. These results imply
+that HT0414 may be involved in adaptation to a range
+of different habitats or to conditions shared by several
+different habitats (Kane and Rieseberg 2008).
 

### 2017-06-16 paper.tex
Update paper.tex

@@
 Notes for review:
+
+http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
+
+http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
+
+file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
+
+http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
+
+https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
 

### 2017-06-28 paper.tex
Update paper.tex
minor edits to intro

@@
 
-\title{Review: Redefining Domestication: Adaptive Introgression during Crop Expansion}
+\title{Review: Adaptive Introgression has Expanded the Genetic Base of Crops during post-Domestication Spread}
 
@@
 
-The process of domestication was once thought to be rapid and geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation.
+The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation.
 \mbh{will need citation for this}
-However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017}.
-This new conception of domestication emphasizes the role of beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion after initial domestication.
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
+An important aspect of this emerging complexity is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
 
 Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
-In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and removal of undesirable wild background.
+In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
 To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Emerging results suggest a need for reevaluating the existing domestication paradigm.
+Preliminary results suggest a need to expand our conception of domestication to include the broadened genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
 
@@
 
+%note from Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice:
+%Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.
+
 \end{enumerate}

### 2017-06-28 paper.tex
edits to introduction

@@
 The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation.
-\mbh{will need citation for this}
+\mbh{will need citation for this, possibly Gepts review or a Harlan paper}
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
-An important aspect of this emerging complexity is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
+An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
 
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Preliminary results suggest a need to expand our conception of domestication to include the broadened genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
+Preliminary results suggest a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
 
-In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of their application for detecting crop-wild introgression, 2) review evidence suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 

### 2017-06-28 paper.tex
rearranging domestication concepts section

@@
 
-A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification.
-An history of introgression during diffusion appears to be the rule for crops rather than the exception.
-Theory suggests that colonizing species will overwhelmingly be recipients of introgression from locally-adapted native species \cite{Currat2008}.
-Crops, given their frequent history of diffusion from defined centers of origin, are therefore potential recipients of adaptive introgression.
+A framework in which crops were domesticated from a single population or even a single species is, in many instances, an oversimplification.
+An history of introgression during geographic expansion increasingly appears to be quite common.
+The addition of gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
 
-With this in mind, certain aspects of crop evolution must be re-evaluated:
+\subsection*{What is the progenitor of a crop?}
+When  highly-introgressed or hybrid populations are selected for domestication (as in the case of potato and tomato), identification of original progenitor(s) and domestication centers is difficult.
+Determining whether a crop 
+Crops arising from highly-interfertile compleces of wild relatives
 
-* Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
+%\gmj{It may not fit here perfectly, but we might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
+The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
+
+\subsection*{When was a crop domesticated?}
+Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
+The directionality of this effect is likewise dependent on $N_e$ of the donor population.
+
+\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
+Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
 Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
@@
 
-* Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
-The directionality of this effect is likewise dependent on $N_e$ of the donor population.
-
-* Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
+\subsection*{What candidate genes were targeted by selection during domestication?}
+Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
 Introgression may alter these signatures and confound detection of domestication loci.
@@
 But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
-
-* When  highly-introgressed or hybrid populations are selected for domestication (as in the case of potato and tomato), identification of original progenitor(s) and domestication centers is difficult.
-Determining whether a crop 
-Crops arising from highly-interfertile compleces of wild relatives
 

### 2017-06-29 paper.tex
added new reference to potato line in table
edits to domestication section

@@
 
-\title{Review: Adaptive Introgression has Expanded the Genetic Base of Crops during post-Domestication Spread}
+\title{Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread}
 
@@
 \mbh{will need citation for this, possibly Gepts review or a Harlan paper}
-However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
 An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
@@
 \emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
-Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+Maize was domesticated from \emph{Zea mays} ssp.
+\emph{parviglumis} (hereafter referred to as \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
@@
 A framework in which crops were domesticated from a single population or even a single species is, in many instances, an oversimplification.
-An history of introgression during geographic expansion increasingly appears to be quite common.
-The addition of gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
+A history of introgression during the geographic expansion of crops increasingly appears common.
+The addition of ongoing gene flow and adaptive introgression to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
 
 \subsection*{What is the progenitor of a crop?}
-When  highly-introgressed or hybrid populations are selected for domestication (as in the case of potato and tomato), identification of original progenitor(s) and domestication centers is difficult.
-Determining whether a crop 
-Crops arising from highly-interfertile compleces of wild relatives
+Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
+The level of divergence of a crop from these newly encountered populations will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
+For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{Matsuoka2002} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
+Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
+Similarly, extensive post-domestication introgression from wild relatives in both potato \cite{Spooner2014} and tomato
 

### 2017-07-03 paper.tex
citations in intro.

@@
 
-The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation.
-\mbh{will need citation for this, possibly Gepts review or a Harlan paper}
+The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation \cite{Harlan1992}.
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
@@
 In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
-To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}).
-\mbh{cite one paper here...the best, most recent and comprehensive review}
+To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
 The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
@@
 \hline
-Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing} \\
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing, gavrilenko2013genetic} \\
 \hline

### 2017-07-03 paper.tex
What is the progenitor?

@@
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
-Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
+Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002single, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
 However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
@@
 
-A framework in which crops were domesticated from a single population or even a single species is, in many instances, an oversimplification.
-A history of introgression during the geographic expansion of crops increasingly appears common.
-The addition of ongoing gene flow and adaptive introgression to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
+A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression during the geographic expansion of crops is extensive.
+The addition of ongoing gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
 
@@
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
-The level of divergence of a crop from these newly encountered populations will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
-For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{Matsuoka2002} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
+The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
+For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
 Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
-Similarly, extensive post-domestication introgression from wild relatives in both potato \cite{Spooner2014} and tomato
-
-%\gmj{It may not fit here perfectly, but we might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
-The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
+Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
+Recent work has shown that, following the original domestication event in the central Andes from \emph{Solanum tuberosum}, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
+Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
+In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to the genetic base of modern crops.
+Broad recognition of the role these wild relatives have played in crop adaptation could further their use in breeding and elevate their conservation status.
 

### 2017-07-05 paper.tex
edits to domestication section

@@
 
-\section*{Re-evaluating concepts of domestication}
+\section*{Re-evaluating domestication}
 
@@
 For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
-Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
+Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
 Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
-Recent work has shown that, following the original domestication event in the central Andes from \emph{Solanum tuberosum}, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
+Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
 Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
@@
 \subsection*{When was a crop domesticated?}
-Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
-The directionality of this effect is likewise dependent on $N_e$ of the donor population.
+Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
+In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
+In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and introgressing species.
+This phenomenon, in combination with divergence of samples from true ancestral populations and ongoing evolution of crop progenitors, may explain discrepancies between domestication dates based on genetic and archaeological data.
+\mbh{perhaps add citations here}
+More accurate genetic estimates of the timing of domestication may be obtained by excluding loci that show signatures of introgression.
 
 \subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
-Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
-Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
-For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.
-Conversely, if the donor population has a relatively small effective population size, the opposite bias may be imposed upon bottleneck estimations.
-In most cases, the effective population size of the domesticated crop will be lower than that of the wild progenitor and wild relative populations.
+Estimates of the strength of the initial domestication bottleneck may also be skewed when introgression is not considered.
+Introgressed chromosomal regions may have a distinct effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
+The influence of introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of introgression, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
+For example, substantial introgression at neutral loci from wild taxa with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
 

### 2017-07-05 paper.tex
Update paper.tex

@@
 The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
-\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
-On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-Adaptive introgression may in fact be the signal interpreted as independent domestications.
+Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
+On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups, \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+Others support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
+A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of indica and japonica \cite{molina2011molecular}.
+Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+%Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
+
+%The former first paragraph, replaced from above:
+%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
+%A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
+%\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
+%On the other hand, genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+%An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+%Adaptive introgression may in fact be the signal interpreted as independent domestications.
 
 The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
-The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high phenotypic diversity and exhibit locally-adaptive traits.
-\emph{Rufipogon} reproduction can be either primarily sexual or vegetative, and whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
-Likewise, domesticated rice varieties display patterns of local adaptation.
-For examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
-Gene flow between these wild and cultivated rices, though asymmetrical, is exceedingly common, producing nuisance weedy hybrids when the two are grown in sympatry.
+The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
+%From Huang2012, "Based on the SNP data, the sequence diversity (p) of O.
+rufipogon was estimated at 0.003, which is higher than that in O.
+sativa—the sequence diversity is 0.0024 for O.
+sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
+Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
 

### 2017-07-05 paper.tex
edits to domestication section

--- 2017-07-05 paper.tex
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Preliminary results suggest a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
+Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
 
-In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
@@
 In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
-In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and introgressing species.
-This phenomenon, in combination with divergence of samples from true ancestral populations and ongoing evolution of crop progenitors, may explain discrepancies between domestication dates based on genetic and archaeological data.
-\mbh{perhaps add citations here}
-More accurate genetic estimates of the timing of domestication may be obtained by excluding loci that show signatures of introgression.
+In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and the introgressing species.
+This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
+More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.
 
 \subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
-Estimates of the strength of the initial domestication bottleneck may also be skewed when introgression is not considered.
-Introgressed chromosomal regions may have a distinct effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
-The influence of introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of introgression, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
-For example, substantial introgression at neutral loci from wild taxa with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
+Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
+Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$). .
+The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
+For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
 
 \subsection*{What candidate genes were targeted by selection during domestication?}
-Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
-Introgression may alter these signatures and confound detection of domestication loci.
+Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \cite{Ross-Ibarra2007}.
+Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
+In this approach, diversity data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
+Due to the difficulty of this approach and the uncertainty associated with any given demographic model, many studies identify domestication loci using a strict outlier approach in which loci showing the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
+For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \cite{vanHeerwaarden2011}.
+Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
+Failure to account for introgression in maize would therefore compromise candidate detection, particularly if a study contained maize samples from the Mexican highlands.
+Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
+However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
 

### 2017-07-05 bib_gj.tex

@article{Ross-Ibarra2007,
author = {Ross-Ibarra, Jeffrey and Morrell, Peter L. and Gaut, Brandon S.}, 
title = {Plant domestication, a unique opportunity to identify the genetic basis of adaptation},
volume = {104}, 
number = {suppl 1}, 
pages = {8641-8648}, 
year = {2007}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@article{Zeder2006,
	Author = {Zeder, Melinda A. and Emshwiller, Eve and Smith, Bruce D. and Bradley, Daniel G.},
	Journal = {Trends in Genetics},
	Number = {3},
	Pages = {139--155},
	Title = {Documenting domestication: the intersection of genetics and archaeology},
	Volume = {22},
	Year = {2006}}

@article{Gavrilenko2013,
	Author = {Gavrilenko, Tatjana and Antonova, Olga and Shuvalova, Anna and Krylova, Ekaterina and Alpatyeva, Natalia and Spooner, David M. and Novikova, Lubov},
	Journal = {Genetic Resources and Crop Evolution},
	Number = {7},
	Pages = {1997--2015},
	Title = {Genetic diversity and origin of cultivated potatoes based on plastid microsatellite polymorphism},
	Volume = {60},
	Year = {2013}
}

@article{Dempewolf2017,
	Author = {Dempewolf, Hannes and Baute, Gregory and Anderson, Justin and Kilian, Benjamin and Smith, Chelsea and Guarino, Luigi},
	Pages = {1070--1082},
	Title = {Past and Future Use of Wild Relatives in Crop Breeding},
	Volume = {57},
	Year = {2017}
}

@article{wang2017,
  title={The interplay of demography and selection during maize domestication and expansion},
  author={Wang, Li and Beissinger, Timothy Mathes and Lorant, Anne and Ross-Ibarra, Claudia and Ross-Ibarra, Jeffrey and Hufford, Matthew},
  journal={bioRxiv},
  pages={114579},
  year={2017},
  publisher={Cold Spring Harbor Labs Journals}
}

@book{Harlan1992,
	Address = {Madison, WI},
	Author = {Harlan, Jack R. },
	Date-Added = {2017-07-03 15:15:48 +0000},
	Date-Modified = {2017-07-03 15:15:48 +0000},
	Doi = {10.2135/1992.cropsandman},
	Isbn = {978-0-89118-565-9},
	La = {English},
	Publisher = {American Society of Agronomy, Crop Science Society of America},
	Title = {Crops \& Man},
	Ty = {BOOK},
	Url = {http://dx.doi.org/10.2135/1992.cropsandman},
	Year = {1992},
	Bdsk-Url-1 = {http://dx.doi.org/10.2135/1992.cropsandman}
}

@article{li2011,
  title={Inference of human population history from individual whole-genome sequences},
  author={Li, Heng and Durbin, Richard},
  journal={Nature},
  volume={475},
  number={7357},
  pages={493--496},
  year={2011},
  publisher={Nature Publishing Group}
}

@article{zhou2017,
  title={The Evolutionary Genomics of Grape (Vitis vinifera ssp. vinifera) Domestication},
  author={Zhou, Yongfeng and Massonnet, Melanie and Sanjak, Jaleal and Cantu, Dario and Gaut, Brandon S},
  journal={bioRxiv},
  pages={146373},
  year={2017},
  publisher={Cold Spring Harbor Labs Journals}
}

@article{Fuller2014,
author = {Fuller, Dorian Q and Denham, Tim and Arroyo-Kalin, Manuel and Lucas, Leilani and Stevens, Chris J. and Qin, Ling and Allaby, Robin G. and Purugganan, Michael D.}, 
title = {Convergent evolution and parallelism in plant domestication revealed by an expanding archaeological record},
volume = {111}, 
number = {17}, 
pages = {6147-6152}, 
year = {2014}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@article {PBI:PBI207,
author = {Messeguer, Joaquima and Peñas, Gisela and Ballester, Jordi and Bas, Marta and Serra, Joan and Salvia, Jordi and Palaudelmàs, Montserrat and Melé, Enric},
title = {Pollen-mediated gene flow in maize in real situations of coexistence},
journal = {Plant Biotechnology Journal},
volume = {4},
number = {6},
publisher = {Blackwell Publishing Ltd},
issn = {1467-7652},
url = {http://dx.doi.org/10.1111/j.1467-7652.2006.00207.x},
doi = {10.1111/j.1467-7652.2006.00207.x},
pages = {633--645},
keywords = {coexistence, maize, pollen gene flow},
year = {2006},
}

@article{wang2017asian,
  title={Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice},
  author={Wang, Hongru and Vieira, Filipe G and Crawford, Jacob E and Chu, Chengcai and Nielsen, Rasmus},
  journal={Genome Research},
  volume={27},
  number={6},
  pages={1029--1038},
  year={2017},
  publisher={Cold Spring Harbor Lab}
}

@article{morrell2007genetic,
  title={Genetic evidence for a second domestication of barley (Hordeum vulgare) east of the Fertile Crescent},
  author={Morrell, Peter L and Clegg, Michael T},
  journal={Proceedings of the national academy of sciences},
  volume={104},
  number={9},
  pages={3289--3294},
  year={2007},
  publisher={National Acad Sciences}
}

@article{dai2012tibet,
  title={Tibet is one of the centers of domestication of cultivated barley},
  author={Dai, Fei and Nevo, Eviatar and Wu, Dezhi and Comadran, Jordi and Zhou, Meixue and Qiu, Long and Chen, Zhonghua and Beiles, Avigdor and Chen, Guoxiong and Zhang, Guoping},
  journal={Proceedings of the National Academy of Sciences},
  volume={109},
  number={42},
  pages={16969--16973},
  year={2012},
  publisher={National Acad Sciences}
}

@article{vaughan2008evolving,
  title={The evolving story of rice evolution},
  author={Vaughan, Duncan A and Lu, Bao-Rong and Tomooka, Norihiko},
  journal={Plant science},
  volume={174},
  number={4},
  pages={394--408},
  year={2008},
  publisher={Elsevier}
}

@article{broich1981evolutionary,
  title={Evolutionary studies of the soybean: the frequency and distribution of alleles among collections of Glycine max and G. soja of various origin},
  author={Broich, Steven L and Palmer, Reid G},
  journal={Euphytica},
  volume={30},
  number={1},
  pages={55--64},
  year={1981},
  publisher={Springer}
}

@article{wendel1992genetic,
  title={Genetic diversity in Gossypium hirsutum and the origin of upland cotton},
  author={Wendel, Jonathan F and Brubaker, Curt L and Percival, A Edward},
  journal={American Journal of Botany},
  pages={1291--1310},
  year={1992},
  publisher={JSTOR}
}

@article{leotard2009phylogeography,
  title={Phylogeography and the origin of cassava: new insights from the northern rim of the Amazonian basin},
  author={L{\'e}otard, Guillaume and Duputi{\'e}, Anne and Kjellberg, Finn and Douzery, Emmanuel JP and Debain, Chantal and De Granville, Jean-Jacques and McKey, Doyle},
  journal={Molecular Phylogenetics and Evolution},
  volume={53},
  number={1},
  pages={329--334},
  year={2009},
  publisher={Academic Press}
}

@article{gebrekidan1982utilization,
  title={Utilization of germplasm in sorghum improvement},
  author={Gebrekidan, Brhane},
  journal={Sorghum in the eighties. Patancheru, India: International Crops Research Institute for the Semi-Arid Tropics},
  pages={335--345},
  year={1982}
}

@article{robertson2007genetic,
  title={Genetic resources of tomato (Lycopersicon esculentum Mill.) and wild relatives},
  author={Robertson, LARRY D and Labate, JOANNE A and Razdan, MK and Mattoo, AK},
  journal={Genetic Improvement of Solanaceous Crops. Tomato},
  volume={2},
  pages={25--75},
  year={2007}
}

@article{rick1958role,
  title={The role of natural hybridization in the derivation of cultivated tomatoes of western South America},
  author={Rick, Charles M},
  journal={Economic Botany},
  volume={12},
  number={4},
  pages={346--367},
  year={1958},
  publisher={Springer}
}

@article{fujita1997extent,
  title={The extent of natural cross-pollination in wild soybean (Glycine soja)},
  author={Fujita, R and Ohara, M and Okazaki, K and Shimamoto, Y},
  journal={Journal of Heredity},
  volume={88},
  number={2},
  pages={124--128},
  year={1997},
  publisher={Am Genetic Assoc}
}

@article{concibido2003introgression,
  title={Introgression of a quantitative trait locus for yield from Glycine soja into commercial soybean cultivars},
  author={Concibido, V and La Vallee, B and Mclaird, P and Pineda, N and Meyer, J and Hummel, L and Yang, J and Wu, K and Delannay, X},
  journal={Theoretical and Applied Genetics},
  volume={106},
  number={4},
  pages={575--582},
  year={2003},
  publisher={Springer}
}

@article{Meyer2016,
	Author = {Meyer, Rachel S and Choi, Jae Young and Sanches, Michelle and Plessis, Anne and Flowers, Jonathan M and Amas, Junrey and Dorph, Katherine and Barretto, Annie and Gross, Briana and Fuller, Dorian Q and Bimpong, Isaac Kofi and Ndjiondjop, Marie-Noelle and Hazzouri, Khaled M and Gregorio, Glenn B and Purugganan, Michael D},
	Journal = {Nat Genet},
	Month = {09},
	Number = {9},
	Pages = {1083--1088},
	Title = {Domestication history and geographical adaptation inferred from a SNP map of African rice},
	Volume = {48},
	Year = {2016}}

@inproceedings{besnard2013complex,
  title={The complex history of the olive tree: from Late Quaternary diversification of Mediterranean lineages to primary domestication in the northern Levant},
  author={Besnard, G and Khadari, B and Navascu{\'e}s, M and Fern{\'a}ndez-Mazuecos, M and El Bakkali, Ahmed and Arrigo, N and Baali-Cherif, D and de Caraffa, V Brunini-Bronzini and Santoni, S and Vargas, P and others},
  booktitle={Proc. R. Soc. B},
  volume={280},
  number={1756},
  pages={20122833},
  year={2013},
  organization={The Royal Society}
}

@inproceedings{besnard2013complex,
  title={The complex history of the olive tree: from Late Quaternary diversification of Mediterranean lineages to primary domestication in the northern Levant},
  author={Besnard, G and Khadari, B and Navascu{\'e}s, M and Fern{\'a}ndez-Mazuecos, M and El Bakkali, Ahmed and Arrigo, N and Baali-Cherif, D and de Caraffa, V Brunini-Bronzini and Santoni, S and Vargas, P and others},
  booktitle={Proc. R. Soc. B},
  volume={280},
  number={1756},
  pages={20122833},
  year={2013},
  organization={The Royal Society}
}

@article{percy1990allozyme,
  title={Allozyme evidence for the origin and diversification of Gossypium barbadense L.},
  author={Percy, RG and Wendel, Jonathan F},
  journal={Theoretical and Applied Genetics},
  volume={79},
  number={4},
  pages={529--542},
  year={1990},
  publisher={Springer}
}

@article{wendel1990allozyme,
  title={Allozyme diversity and introgression in the Galapagos Islands endemic Gossypium darwinii and its relationship to continental G. barbadense},
  author={Wendel, Jonathan F and Percy, Richard G},
  journal={Biochemical Systematics and Ecology},
  volume={18},
  number={7},
  pages={517--528},
  year={1990},
  publisher={Elsevier}
}

@article{fryxell1979natural,
  title={The natural history ofthe cotton tribe},
  author={Fryxell, PA},
  journal={Texas A},
  year={1979}
}

@article{wendel1989genetic,
  title={Genetic diversity, introgression, and independent domestication of old world cultivated cottons},
  author={Wendel, Jonathan F and Olson, Paul D and McD, James},
  journal={American Journal of Botany},
  pages={1795--1806},
  year={1989},
  publisher={JSTOR}
}

@article{abdalla2001genetic,
  title={Genetic diversity and relationships of diploid and tetraploid cottons revealed using AFLP},
  author={Abdalla, AM and Reddy, OUK and El-Zik, KM and Pepper, AE},
  journal={Theoretical and Applied Genetics},
  volume={102},
  number={2-3},
  pages={222--229},
  year={2001},
  publisher={Springer}
}

@article{gunn2011independent,
  title={Independent origins of cultivated coconut (Cocos nucifera L.) in the old world tropics},
  author={Gunn, Bee F and Baudouin, Luc and Olsen, Kenneth M},
  journal={Plos one},
  volume={6},
  number={6},
  pages={e21143},
  year={2011},
  publisher={Public Library of Science}
}

@article{broich1981evolutionary,
  title={Evolutionary studies of the soybean: the frequency and distribution of alleles among collections of Glycine max and G. soja of various origin},
  author={Broich, Steven L and Palmer, Reid G},
  journal={Euphytica},
  volume={30},
  number={1},
  pages={55--64},
  year={1981},
  publisher={Springer}
}

@article{kwon1972studies,
  title={Studies on the diversity of seed weight in the Korean soybean land races and wild soybean},
  author={Kwon, SH and Im, KH and Kim, JR},
  journal={Korean J Breed},
  volume={4},
  number={1},
  pages={70--74},
  year={1972}
}

@article{singh1988genomic,
  title={The genomic relationship between Glycine max (L.) Merr. and G. soja Sieb. and Zucc. as revealed by pachytene chromosome analysis},
  author={Singh, RJ and Hymowitz, T},
  journal={Theoretical and Applied Genetics},
  volume={76},
  number={5},
  pages={705--711},
  year={1988},
  publisher={Springer}
}

@article{ray2003soybean,
  title={Soybean natural cross-pollination rates under field conditions},
  author={Ray, Jeffery D and Kilen, Thomas C and Abel, Craig A and Paris, Robert L},
  journal={Environmental Biosafety Research},
  volume={2},
  number={2},
  pages={133--138},
  year={2003},
  publisher={EDP Sciences}
}

@article{ellstrand1999gene,
  title={Gene flow and introgression from domesticated plants into their wild relatives},
  author={Ellstrand, Norman C and Prentice, Honor C and Hancock, James F},
  journal={Annual review of Ecology and Systematics},
  pages={539--563},
  year={1999},
  publisher={JSTOR}
}

@article{quist2001transgenic,
  title={Transgenic DNA introgressed into traditional maize landraces in Oaxaca, Mexico},
  author={Quist, David and Chapela, Ignacio H},
  journal={Nature},
  volume={414},
  number={6863},
  pages={541--543},
  year={2001},
  publisher={Nature Publishing Group}
}

@article{doebley1990molecular,
  title={Molecular evidence for gene flow among Zea species},
  author={Doebley, John},
  journal={BioScience},
  pages={443--448},
  year={1990},
  publisher={JSTOR}
}

@article{viaud2008spatial,
  title={Spatial sensitivity of maize gene-flow to landscape pattern: a simulation approach},
  author={Viaud, Val{\'e}rie and Monod, Herv{\'e} and Lavigne, Claire and Angevin, Fr{\'e}d{\'e}rique and Adamczyk, Katarzyna},
  journal={Landscape Ecology},
  volume={23},
  number={9},
  pages={1067--1079},
  year={2008},
  publisher={Springer}
}

@article{baltazar2005pollination,
  title={Pollination between maize and teosinte: an important determinant of gene flow in Mexico},
  author={Baltazar, Baltazar M and de Jesus Sanchez-Gonzalez, Jose and de la Cruz-Larios, Lino and Schoper, John B},
  journal={Theoretical and Applied Genetics},
  volume={110},
  number={3},
  pages={519--526},
  year={2005},
  publisher={Springer}
}

@article{galinat1973intergenomic,
  title={Intergenomic mapping of maize, teosinte and Tripsacum},
  author={Galinat, Walton C},
  journal={Evolution},
  pages={644--655},
  year={1973},
  publisher={JSTOR}
}

@article{ellstrand2007spontaneous,
  title={Spontaneous hybridization between maize and teosinte},
  author={Ellstrand, Norman C and Garner, Lauren C and Hegde, Subray and Guadagnuolo, Roberto and Blancas, Lesley},
  journal={Journal of Heredity},
  volume={98},
  number={2},
  pages={183--187},
  year={2007},
  publisher={Am Genetic Assoc}
}

@article{gore2009first,
  title={A first-generation haplotype map of maize},
  author={Gore, Michael A and Chia, Jer-Ming and Elshire, Robert J and Sun, Qi and Ersoz, Elhan S and Hurwitz, Bonnie L and Peiffer, Jason A and McMullen, Michael D and Grills, George S and Ross-Ibarra, Jeffrey and others},
  journal={Science},
  volume={326},
  number={5956},
  pages={1115--1117},
  year={2009},
  publisher={American Association for the Advancement of Science}
}

@article{ross2009historical,
  title={Historical divergence and gene flow in the genus Zea},
  author={Ross-Ibarra, Jeffrey and Tenaillon, Maud and Gaut, Brandon S},
  journal={Genetics},
  volume={181},
  number={4},
  pages={1399--1413},
  year={2009},
  publisher={Genetics Soc America}
}

@article{van2011genetic,
  title={Genetic signals of origin, spread, and introgression in a large sample of maize landraces},
  author={van Heerwaarden, Joost and Doebley, John and Briggs, William H and Glaubitz, Jeffrey C and Goodman, Major M and Gonzalez, Jose de Jesus Sanchez and Ross-Ibarra, Jeffrey},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={3},
  pages={1088--1092},
  year={2011},
  publisher={National Acad Sciences}
}

@article{zhang2009selection,
  title={Selection on grain shattering genes and rates of rice domestication},
  author={Zhang, Lin-Bin and Zhu, Qihui and Wu, Zhi-Qiang and Ross-Ibarra, Jeffrey and Gaut, Brandon S and Ge, Song and Sang, Tao},
  journal={New Phytologist},
  volume={184},
  number={3},
  pages={708--720},
  year={2009},
  publisher={Wiley Online Library}
}

@article{matsuoka2002single,
  title={A single domestication for maize shown by multilocus microsatellite genotyping},
  author={Matsuoka, Yoshihiro and Vigouroux, Yves and Goodman, Major M and Sanchez, Jesus and Buckler, Edward and Doebley, John},
  journal={Proceedings of the National Academy of Sciences},
  volume={99},
  number={9},
  pages={6080--6084},
  year={2002},
  publisher={National Acad Sciences}
}

@article{doebley1990molecular,
  title={Molecular evidence and the evolution of maize},
  author={Doebley, John},
  journal={Economic Botany},
  volume={44},
  number={3},
  pages={6--27},
  year={1990},
  publisher={Springer}
}

@article{warburton2011gene,
  title={Gene flow among different teosinte taxa and into the domesticated maize gene pool},
  author={Warburton, Marilyn L and Wilkes, Garrison and Taba, S and Charcosset, Alain and Mir, Celine and Dumas, Fabrice and Madur, Delphine and Dreisigacker, Susanne and Bedoya, Claudia and Prasanna, BM and others},
  journal={Genetic resources and crop evolution},
  volume={58},
  number={8},
  pages={1243--1261},
  year={2011},
  publisher={Springer}
}

@article{mallet2005hybridization,
  title={Hybridization as an invasion of the genome},
  author={Mallet, James},
  journal={Trends in Ecology \& Evolution},
  volume={20},
  number={5},
  pages={229--237},
  year={2005},
  publisher={Elsevier}
}

@article{crossa1987introgression,
  title={Introgression of an exotic germplasm for improving an adapted maize population},
  author={Crossa, J and Gardner, CO},
  journal={Crop Science},
  volume={27},
  number={2},
  pages={187--190},
  year={1987},
  publisher={Crop Science Society of America}
}

@article{arnold2009adaptation,
  title={Adaptation by introgression},
  author={Arnold, Michael L and Martin, Noland H},
  journal={J Biol},
  volume={8},
  number={9},
  pages={82},
  year={2009}
}

@article{doebley1984maize,
  title={Maize introgression into teosinte-a reappraisal},
  author={Doebley, John F},
  journal={Annals of the Missouri Botanical Garden},
  pages={1100--1113},
  year={1984},
  publisher={JSTOR}
}

@article{ribaut2007marker,
  title={Marker-assisted selection to improve drought adaptation in maize: the backcross approach, perspectives, limitations, and alternatives},
  author={Ribaut, Jean-Marcel and Ragot, Michel},
  journal={Journal of Experimental Botany},
  volume={58},
  number={2},
  pages={351--360},
  year={2007},
  publisher={Soc Experiment Biol}
}

@article{hufford2013genomic,
  title={The genomic signature of crop-wild introgression in maize},
  author={Hufford, Matthew B and Lubinksy, Pesach and Pyh{\"a}j{\"a}rvi, Tanja and Devengenzo, Michael T and Ellstrand, Norman C and Ross-Ibarra, Jeffrey},
  year={2013}
}

@article{michelini1991evaluation,
  title={Evaluation of exotic and adapted maize (Zea mays L.) germplasm crosses},
  author={Michelini, Luiz Antonio},
  year={1991},
  publisher={Digital Repository@ Iowa State University, http://lib. dr. iastate. edu/}
}

@article{bernardo2009genomewide,
  title={Genomewide selection for rapid introgression of exotic germplasm in maize},
  author={Bernardo, Rex},
  journal={Crop Science},
  volume={49},
  number={2},
  pages={419--425},
  year={2009},
  publisher={Crop Science Society of America}
}

@article{jiang1999genetic,
  title={Genetic analysis of adaptation differences between highland and lowland tropical maize using molecular markers},
  author={Jiang, C and Edmeades, GO and Armstead, I and Lafitte, HR and Hayward, MD and Hoisington, D},
  journal={Theoretical and Applied Genetics},
  volume={99},
  number={7-8},
  pages={1106--1119},
  year={1999},
  publisher={Springer}
}

@article{wilkes1977hybridization,
  title={Hybridization of maize and teosinte, in Mexico and Guatemala and the improvement of maize},
  author={Wilkes, HG},
  journal={Economic Botany},
  pages={254--293},
  year={1977},
  publisher={JSTOR}
}

@article{salvi2011genetic,
  title={Genetic dissection of maize phenology using an intraspecific introgression library},
  author={Salvi, Silvio and Corneti, Simona and Bellotti, Massimo and Carraro, Nicola and Sanguineti, Maria C and Castelletti, Sara and Tuberosa, Roberto},
  journal={BMC plant biology},
  volume={11},
  number={1},
  pages={4},
  year={2011},
  publisher={BioMed Central Ltd}
}

@article{bitocchi2009introgression,
  title={Introgression from modern hybrid varieties into landrace populations of maize (Zea mays ssp. mays L.) in central Italy},
  author={Bitocchi, Elena and Nanni, Laura and Rossi, Monica and Rau, Domenico and Bellucci, Elisa and Giardini, Alessandro and Buonamici, Anna and Vendramin, Giovanni Giuseppe and Papa, Roberto},
  journal={Molecular Ecology},
  volume={18},
  number={4},
  pages={603--621},
  year={2009},
  publisher={Wiley Online Library}
}

@article{gouesnard1996evaluation,
  title={Evaluation of agronomic traits and analysis of exotic germ plasm polymorphism in adapted x exotic maize crosses},
  author={Gouesnard, B and Sanou, J and Panouille, A and Bourion, V and Boyat, A},
  journal={Theoretical and Applied Genetics},
  volume={92},
  number={3-4},
  pages={368--374},
  year={1996},
  publisher={Springer}
}

@article{placido2013introgression,
  title={Introgression of novel traits from a wild wheat relative improves drought adaptation in wheat},
  author={Placido, Dante F and Campbell, Malachy T and Folsom, Jing J and Cui, Xinping and Kruger, Greg R and Baenziger, P Stephen and Walia, Harkamal},
  journal={Plant physiology},
  volume={161},
  number={4},
  pages={1806--1819},
  year={2013},
  publisher={Am Soc Plant Biol}
}

@article{arnold2004transfer,
  title={Transfer and origin of adaptations through natural hybridization: were Anderson and Stebbins right?},
  author={Arnold, Michael L},
  journal={The Plant Cell},
  volume={16},
  number={3},
  pages={562--570},
  year={2004},
  publisher={Am Soc Plant Biol}
}

@article{wilkes1970teosinte,
  title={Teosinte introgression in the maize of the Nobogame Valley},
  author={Wilkes, H Garrison},
  journal={Botanical Museum Leaflets, Harvard University},
  pages={297--311},
  year={1970},
  publisher={JSTOR}
}

@article{lam2010resequencing,
  title={Resequencing of 31 wild and cultivated soybean genomes identifies patterns of genetic diversity and selection},
  author={Lam, Hon-Ming and Xu, Xun and Liu, Xin and Chen, Wenbin and Yang, Guohua and Wong, Fuk-Ling and Li, Man-Wah and He, Weiming and Qin, Nan and Wang, Bo and others},
  journal={Nature genetics},
  volume={42},
  number={12},
  pages={1053--1059},
  year={2010},
  publisher={Nature Publishing Group}
}

@article{kato2002introgression,
  title={Introgression of chromosome knobs from Zea diploperennis into maize},
  author={KATO Y, TA and SANCHEZ G, JJ},
  journal={Maydica},
  volume={47},
  number={1},
  pages={33--50},
  year={2002},
  publisher={Maydica}
}

@article{fukunaga2005genetic,
  title={Genetic diversity and population structure of teosinte},
  author={Fukunaga, Kenji and Hill, Jason and Vigouroux, Yves and Matsuoka, Yoshihiro and Sanchez, Jesus and Liu, Kejun and Buckler, Edward S and Doebley, John},
  journal={Genetics},
  volume={169},
  number={4},
  pages={2241--2254},
  year={2005},
  publisher={Genetics Soc America}
}

@article{reagon2006cultivated,
  title={Cultivated Helianthus annuus (Asteraceae) volunteers as a genetic “bridge” to weedy sunflower populations in North America},
  author={Reagon, Michael and Snow, Allison A},
  journal={American Journal of Botany},
  volume={93},
  number={1},
  pages={127--133},
  year={2006},
  publisher={Botanical Soc America}
}

@article{broyles2002hybrid,
  title={Hybrid bridges to gene flow: a case study in milkweeds (Asclepias)},
  author={Broyles, Steven B},
  journal={Evolution},
  volume={56},
  number={10},
  pages={1943--1953},
  year={2002},
  publisher={Wiley Online Library}
}

@article{chapman2006letting,
  title={Letting the gene out of the bottle: the population genetics of genetically modified crops},
  author={Chapman, Mark A and Burke, John M},
  journal={New Phytologist},
  volume={170},
  number={3},
  pages={429--443},
  year={2006},
  publisher={Wiley Online Library}
}

@article{collins1921dominance,
  title={Dominance and the vigor of first generation hybrids},
  author={Collins, GN},
  journal={American Naturalist},
  pages={116--133},
  year={1921},
  publisher={JSTOR}
}

@article{doebley1987patterns,
  title={Patterns of isozyme variation between maize and Mexican annual teosinte},
  author={Doebley, John and Goodman, Major M and Stuber, Charles W},
  journal={Economic Botany},
  volume={41},
  number={2},
  pages={234--246},
  year={1987},
  publisher={Springer}
}

@article{slatkin1987gene,
  title={Gene flow and the geographic structure of natural populations},
  author={Slatkin, Montgomery},
  journal={Science},
  volume={236},
  number={4803},
  pages={787--792},
  year={1987},
  publisher={American Association for the Advancement of Science}
}

@article{zohary1969wild,
  title={The wild diploid progenitors of wheat and their breeding value},
  author={Zohary, Daniel and Harlan, Jack R and Vardi, A},
  journal={Euphytica},
  volume={18},
  number={1},
  pages={58--65},
  year={1969},
  publisher={Springer}
}

@article{von2006ab,
  title={AB-QTL analysis in spring barley: II. Detection of favourable exotic alleles for agronomic traits introgressed from wild barley (H. vulgare ssp. spontaneum)},
  author={Von Korff, M and Wang, H and L{\'e}on, J and Pillen, K},
  journal={Theoretical and Applied Genetics},
  volume={112},
  number={7},
  pages={1221--1231},
  year={2006},
  publisher={Springer}
}

@article{nagy1997local,
  title={Local adaptation in two subspecies of an annual plant: implications for migration and gene flow},
  author={Nagy, Eric S and Rice, Kevin J},
  journal={Evolution},
  pages={1079--1089},
  year={1997},
  publisher={JSTOR}
}

@article{abbott2013hybridization,
  title={Hybridization and speciation},
  author={Abbott, R and Albach, D and Ansell, S and Arntzen, JW and Baird, SJE and Bierne, N and Boughman, J and Brelsford, A and Buerkle, CA and Buggs, R and others},
  journal={Journal of Evolutionary Biology},
  volume={26},
  number={2},
  pages={229--246},
  year={2013},
  publisher={Wiley Online Library}
}

@article{nevo2010drought,
  title={Drought and salt tolerances in wild relatives for wheat and barley improvement},
  author={Nevo, Eviatar and Chen, Guoxiong},
  journal={Plant, cell \& environment},
  volume={33},
  number={4},
  pages={670--685},
  year={2010},
  publisher={Wiley Online Library}
}

@article{hussain1997development,
  title={Development of a fertile genetic bridge between Trifolium ambiguum M. Bieb. and T. repens L.},
  author={Hussain, SW and Williams, WM},
  journal={Theoretical and applied genetics},
  volume={95},
  number={4},
  pages={678--690},
  year={1997},
  publisher={Springer}
}

@article{hegde2004hybridization,
  title={Hybridization and introgression between bread wheat and wild and weedy relatives in North America},
  author={Hegde, SG and Waines, JG},
  journal={Crop Science},
  volume={44},
  number={4},
  pages={1145--1155},
  year={2004},
  publisher={Crop Science Society of America}
}

@article{autrique1995molecular,
  title={Molecular markers for four leaf rust resistance genes introgressed into wheat from wild relatives},
  author={Autrique, Enrique and Tanksley, Steven D and Sorrells, Mark E and Singh, Ravi P},
  journal={Genome},
  volume={38},
  number={1},
  pages={75--83},
  year={1995},
  publisher={NRC Research Press}
}

@article{Feuillet200824,
title = "Cereal breeding takes a walk on the wild side ",
journal = "Trends in Genetics ",
volume = "24",
number = "1",
pages = "24 - 32",
year = "2008",
note = "",
issn = "0168-9525",
doi = "http://dx.doi.org/10.1016/j.tig.2007.11.001",
url = "http://www.sciencedirect.com/science/article/pii/S0168952507003484",
author = "Catherine Feuillet and Peter Langridge and Robbie Waugh"
}

@article{able2007capturing,
  title={Capturing diversity in the cereals: many options but little promiscuity},
  author={Able, Jason A and Langridge, Peter and Milligan, Andrew S},
  journal={Trends in plant science},
  volume={12},
  number={2},
  pages={71--79},
  year={2007},
  publisher={Elsevier}
}

@article{varshney2006advances,
  title={Advances in cereal genomics and applications in crop breeding},
  author={Varshney, Rajeev K and Hoisington, David A and Tyagi, Akhilesh K},
  journal={Trends in biotechnology},
  volume={24},
  number={11},
  pages={490--499},
  year={2006},
  publisher={Elsevier}
}

@article{brown2009complex,
  title={The complex origins of domesticated crops in the Fertile Crescent},
  author={Brown, Terence A and Jones, Martin K and Powell, Wayne and Allaby, Robin G},
  journal={Trends in Ecology \& Evolution},
  volume={24},
  number={2},
  pages={103--109},
  year={2009},
  publisher={Elsevier}
}

@article{suneson1969dynamic,
  title={A dynamic population of weedy rye},
  author={Suneson, CA and Rachie, KO and Khush, GS},
  journal={Crop Science},
  volume={9},
  number={2},
  pages={121--124},
  year={1969},
  publisher={Crop Science Society of America}
}

@article{stewart2003transgene,
  title={Transgene introgression from genetically modified crops to their wild relatives},
  author={Stewart, C Neal and Halfhill, Matthew D and Warwick, Suzanne I},
  journal={Nature Reviews Genetics},
  volume={4},
  number={10},
  pages={806--817},
  year={2003},
  publisher={Nature Publishing Group}
}

@article{zamir2001improving,
  title={Improving plant breeding with exotic genetic libraries},
  author={Zamir, Dani},
  journal={Nature reviews genetics},
  volume={2},
  number={12},
  pages={983--989},
  year={2001},
  publisher={Nature Publishing Group}
}

@article{tanksley1997seed,
  title={Seed banks and molecular maps: unlocking genetic potential from the wild},
  author={Tanksley, Steven D and McCouch, Susan R},
  journal={Science},
  volume={277},
  number={5329},
  pages={1063--1066},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{xie2008wild,
  title={Wild emmer: genetic resources, gene mapping and potential for wheat improvement},
  author={Xie, Weilong and Nevo, Eviatar},
  journal={Euphytica},
  volume={164},
  number={3},
  pages={603--614},
  year={2008},
  publisher={Springer}
}

@book{grant1950genetic,
  title={Genetic and taxonomic studies in Gilia},
  author={Grant, Verne and Grant, Karen A},
  year={1950},
  publisher={Rancho Santa Ana Botanic Garden}
}

@article{hajjar2007use,
  title={The use of wild relatives in crop improvement: a survey of developments over the last 20 years},
  author={Hajjar, Reem and Hodgkin, Toby},
  journal={Euphytica},
  volume={156},
  number={1-2},
  pages={1--13},
  year={2007},
  publisher={Springer}
}

@article{jorgensen1992discovery,
  title={Discovery, characterization and exploitation of Mlo powdery mildew resistance in barley},
  author={J{\o}rgensen, I Helms},
  journal={Euphytica},
  volume={63},
  number={1-2},
  pages={141--152},
  year={1992},
  publisher={Springer}
}

@article{gepts2001origins,
  title={Origins of plant agriculture and major crop plants},
  author={Gepts, Paul},
  journal={Our fragile world: Challenges and opportunities for sustainable development. Oxford: EOLSS Publishers},
  pages={629--637},
  year={2001}
}

@article{smith1997initial,
  title={The initial domestication of Cucurbita pepo in the Americas 10,000 years ago},
  author={Smith, Bruce D},
  journal={Science},
  volume={276},
  number={5314},
  pages={932--934},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{piperno2001earliest,
  title={The earliest archaeological maize (Zea mays L.) from highland Mexico: new accelerator mass spectrometry dates and their implications},
  author={Piperno, Dolores R and Flannery, Kent V},
  journal={Proceedings of the National Academy of Sciences},
  volume={98},
  number={4},
  pages={2101--2103},
  year={2001},
  publisher={National Acad Sciences}
}

@article{colledge2004archaeobotanical,
  title={Archaeobotanical Evidence for the Spread of Farming in the Eastern Mediterranean 1},
  author={Colledge, Sue and Conolly, James and Shennan, Stephen and Bellwood, Peter and Bouby, Laurent and Hansen, Julie and Harris, DavidR and Kotsakis, Konstantinos and zdoan, Mehmet and Peltenburg, Edgar and others},
  journal={Current anthropology},
  volume={45},
  number={S4},
  pages={S35--S58},
  year={2004},
  publisher={JSTOR}
}

@article{zhijun1998middle,
  title={The Middle Yangtze region in China is one place where rice was domesticated: phytolith evidence from the Diaotonghuan Cave, Northern Jiangxi},
  author={Zhijun, Zhao},
  journal={Antiquity},
  volume={72},
  number={278},
  pages={885--897},
  year={1998},
  publisher={Cambridge Univ Press}
}

@article{hart2002age,
  title={The age of the common bean (Phaseolus vulgaris L.) in the northern Eastern Woodlands of North America},
  author={Hart, John P and Asch, David L and Scarry, C Margaret and Crawford, Gary W},
  journal={Antiquity},
  volume={76},
  number={292},
  pages={377--385},
  year={2002},
  publisher={Cambridge Univ Press}
}

@article{crites1993domesticated,
  title={Domesticated sunflower in fifth millennium BP temporal context: new evidence from Middle Tennessee},
  author={Crites, Gary D},
  journal={American Antiquity},
  pages={146--148},
  year={1993},
  publisher={JSTOR}
}

@article{piperno2000starch,
  title={Starch grains reveal early root crop horticulture in the Panamanian tropical forest},
  author={Piperno, Dolores R and Ranere, Anthony J and Holst, Irene and Hansell, Patricia},
  journal={Nature},
  volume={407},
  number={6806},
  pages={894--897},
  year={2000},
  publisher={Nature Publishing Group}
}

@article{pope2001origin,
  title={Origin and environmental setting of ancient agriculture in the lowlands of Mesoamerica},
  author={Pope, Kevin O and Pohl, Mary ED and Jones, John G and Lentz, David L and Von Nagy, Christopher and Vega, Francisco J and Quitmyer, Irvy R},
  journal={Science},
  volume={292},
  number={5520},
  pages={1370--1373},
  year={2001},
  publisher={American Association for the Advancement of Science}
}

@inproceedings{smith1989origins,
  title={Origins of agriculture in eastern North America},
  author={Smith, Bruce David},
  year={1989},
  organization={American Association for the Advancement of Science}
}

@book{ammerman2014neolithic,
  title={The Neolithic transition and the genetics of populations in Europe},
  author={Ammerman, Albert J and Cavalli-Sforza, Luigi Luca},
  year={2014},
  publisher={Princeton University Press}
}

@article{hufford2012comparative,
  title={Comparative population genomics of maize domestication and improvement},
  author={Hufford, Matthew B and Xu, Xun and Van Heerwaarden, Joost and Pyh{\"a}j{\"a}rvi, Tanja and Chia, Jer-Ming and Cartwright, Reed A and Elshire, Robert J and Glaubitz, Jeffrey C and Guill, Kate E and Kaeppler, Shawn M and others},
  journal={Nature genetics},
  volume={44},
  number={7},
  pages={808--811},
  year={2012},
  publisher={Nature Publishing Group}
}

@book{lumholtz1902unknown,
  title={Unknown Mexico: a record of five years' exploration among the tribes of the western Sierra Madre; in the Tierra Caliente of Tepic and Jalisco; and among the Tarascos of Michoacan},
  author={Lumholtz, Carl},
  volume={2},
  year={1902},
  publisher={C. Scribner's sons}
}

@book{smith1995emergence,
  title={The emergence of agriculture},
  author={Smith, Bruce D},
  year={1995},
  publisher={Scientific American Library New York}
}

@article{heun1997site,
  title={Site of einkorn wheat domestication identified by DNA fingerprinting},
  author={Heun, Manfred and Sch{\"a}fer-Pregl, Ralf and Klawan, Dieter and Castagna, Renato and Accerbi, Monica and Borghi, Basilio and Salamini, Francesco},
  journal={Science},
  volume={278},
  number={5341},
  pages={1312--1314},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{nesbitt1998wheat,
  title={Wheat domestication: archaeobotanical evidence},
  author={Nesbitt, Mark and Samuel, Delwen},
  journal={Science},
  volume={279},
  number={5356},
  pages={1431--1431},
  year={1998}
}

@article{lev2000cradle,
  title={The cradle of agriculture},
  author={Lev-Yadun, Simcha and Gopher, Avi and Abbo, Shahal},
  journal={Science},
  volume={288},
  number={5471},
  pages={1602},
  year={2000},
  publisher={The American Association for the Advancement of Science}
}

@book{zohary2012domestication,
  title={Domestication of Plants in the Old World: The origin and spread of domesticated plants in Southwest Asia, Europe, and the Mediterranean Basin},
  author={Zohary, Daniel and Hopf, Maria and Weiss, Ehud},
  year={2012},
  publisher={Oxford University Press on Demand}
}

@article{ozkan2005reconsideration,
  title={A reconsideration of the domestication geography of tetraploid wheats},
  author={Ozkan, Hakan and Brandolini, Andrea and Pozzi, Carlo and Effgen, Sieglinde and Wunder, J and Salamini, Francesco},
  journal={Theoretical and Applied Genetics},
  volume={110},
  number={6},
  pages={1052--1060},
  year={2005},
  publisher={Springer}
}

@article {TPJ:TPJ2772,
author = {Yahiaoui, Nabila and Brunner, Susanne and Keller, Beat},
title = {Rapid generation of new powdery mildew resistance genes after wheat domestication},
journal = {The Plant Journal},
volume = {47},
number = {1},
publisher = {Blackwell Publishing Ltd},
issn = {1365-313X},
url = {http://dx.doi.org/10.1111/j.1365-313X.2006.02772.x},
doi = {10.1111/j.1365-313X.2006.02772.x},
pages = {85--98},
keywords = {disease resistance, evolution, wheat, powdery mildew, arms race, domestication},
year = {2006},
}

@article{tanno2006fast,
  title={How fast was wild wheat domesticated?},
  author={Tanno, Ken-ichi and Willcox, George},
  journal={Science},
  volume={311},
  number={5769},
  pages={1886--1886},
  year={2006},
  publisher={American Association for the Advancement of Science}
}

@article{kilian2006haplotype,
  title={Haplotype structure at seven barley genes: relevance to gene pool bottlenecks, phylogeny of ear type and site of barley domestication},
  author={Kilian, Benjamin and {\"O}zkan, Hakan and Kohl, Jochen and von Haeseler, Arndt and Barale, Francesca and Deusch, Oliver and Brandolini, Andrea and Yucel, Cemal and Martin, William and Salamini, Francesco},
  journal={Molecular Genetics and Genomics},
  volume={276},
  number={3},
  pages={230--241},
  year={2006},
  publisher={Springer}
}

@article{badr2000origin,
  title={On the origin and domestication history of barley (Hordeum vulgare)},
  author={Badr, Abdelfattah and Sch, R and El Rabey, H and Effgen, S and Ibrahim, HH and Pozzi, C and Rohde, W and Salamini, F and others},
  journal={Molecular Biology and Evolution},
  volume={17},
  number={4},
  pages={499--510},
  year={2000},
  publisher={SMBE}
}

@article{azhaguvel2007phylogenetic,
  title={A phylogenetic analysis based on nucleotide sequence of a marker linked to the brittle rachis locus indicates a diphyletic origin of barley},
  author={Azhaguvel, Perumal and Komatsuda, Takao},
  journal={Annals of botany},
  volume={100},
  number={5},
  pages={1009--1015},
  year={2007},
  publisher={Annals Botany Co}
}

@article{von1995ecogeographical,
  title={An ecogeographical study of the genus Hordeum. Systematic and ecogeographic studies on crop genepools 7},
  author={von Bothmer, R von and Jacobsen, N and Baden, C and J{\o}rgensen, RB and Linde-Laursen, I},
  journal={International Plant Genetic Resources Institute, Rome},
  year={1995}
}

@article{hanson2000mapping,
  title={Mapping a wild tomato introgression associated with tomato yellow leaf curl virus resistance in a cultivated tomato line},
  author={Hanson, Peter M and Bernacchi, Dario and Green, Sylvia and Tanksley, Steven D and Muniyappa, Venkataramappa and Padmaja, Attiganal S and Chen, Huei-mei and Kuo, George and Fang, Denise and Chen, Jen-tzu},
  journal={Journal of the American Society for Horticultural Science},
  volume={125},
  number={1},
  pages={15--20},
  year={2000},
  publisher={American Society for Horticultural Science}
}

@article{banerjee1990transfer,
  title={Transfer of tomato leaf curl virus resistance from Lycopersicon hirsutum f. glabratum to L. esculentum},
  author={Banerjee, MK and others},
  journal={Plant Breeding},
  volume={105},
  number={2},
  pages={156--159},
  year={1990},
  publisher={Wiley Online Library}
}

@inproceedings{sims1979history,
  title={History of tomato production for industry around the world},
  author={Sims, WL},
  booktitle={Symposium on Production of Tomatoes for Processing 100},
  pages={25--26},
  year={1979}
}

@article{rick1988tomato,
  title={Tomato-like nightshades: affinities, autoecology, and breeders’ opportunities},
  author={Rick, Charles M},
  journal={Economic Botany},
  volume={42},
  number={2},
  pages={145--154},
  year={1988},
  publisher={Springer}
}

@article{miller1990rflp,
  title={RFLP analysis of phylogenetic relationships and genetic variation in the genus Lycopersicon},
  author={Miller, JC and Tanksley, SD},
  journal={Theoretical and applied genetics},
  volume={80},
  number={4},
  pages={437--448},
  year={1990},
  publisher={Springer}
}

@article{rick1982potential,
  title={The potential of exotic germplasm for tomato improvement},
  author={Rick, CM},
  year={1982},
  publisher={Academic Press}
}

@article{li2006rice,
  title={Rice domestication by reducing shattering},
  author={Li, Changbao and Zhou, Ailing and Sang, Tao},
  journal={science},
  volume={311},
  number={5769},
  pages={1936--1939},
  year={2006},
  publisher={American Association for the Advancement of Science}
}

@book{oka2012origin,
  title={Origin of cultivated rice},
  author={Oka, H-I},
  year={2012},
  publisher={Elsevier}
}

@incollection{brar1997alien,
  title={Alien introgression in rice},
  author={Brar, DS and Khush, GS},
  booktitle={Oryza: From Molecule to Plant},
  pages={35--47},
  year={1997},
  publisher={Springer}
}

@article{khush1974inheritance,
  title={Inheritance of resistance to grassy stunt virus and its vector in rice},
  author={KHUSH, GURDEV S and Ling, KC},
  journal={Journal of Heredity},
  volume={65},
  number={3},
  pages={135--136},
  year={1974},
  publisher={Am Genetic Assoc}
}

@article{spooner2005single,
  title={A single domestication for potato based on multilocus amplified fragment length polymorphism genotyping},
  author={Spooner, David M and McLean, Karen and Ramsay, Gavin and Waugh, Robbie and Bryan, Glenn J},
  journal={Proceedings of the National Academy of Sciences of the United States of America},
  volume={102},
  number={41},
  pages={14694--14699},
  year={2005},
  publisher={National Acad Sciences}
}

@article{second1982origin,
  title={Origin of the genic diversity of cultivated rice (Oryza spp.): study of the polymorphism scored at 40 isozyme loci.},
  author={Second, Gerard},
  journal={遺伝学雑誌},
  volume={57},
  number={1},
  pages={25--57},
  year={1982},
  publisher={The Genetics Society of Japan}
}

@article{van1999tight,
  title={Tight physical linkage of the nematode resistance gene Gpa2 and the virus resistance gene Rx on a single segment introgressed from the wild species Solanum tuberosum subsp. andigena CPC 1673 into cultivated potato},
  author={van der Voort, Jeroen Rouppe and Kanyuka, Konstantin and van der Vossen, Edwin and Bendahmane, Abdelhafid and Mooijman, Paul and Klein-Lankhorst, Ren{\'e} and Stiekema, Willem and Baulcombe, David and Bakker, Jaap},
  journal={Molecular plant-microbe interactions},
  volume={12},
  number={3},
  pages={197--206},
  year={1999},
  publisher={Am Phytopath Society}
}

@article{van2003ancient,
  title={An ancient R gene from the wild potato species Solanum bulbocastanum confers broad-spectrum resistance to Phytophthora infestans in cultivated potato and tomato},
  author={Van Der Vossen, Edwin and Sikkema, Anne and Hekkert, Bas te Lintel and Gros, Jack and Stevens, Patricia and Muskens, Marielle and Wouters, Doret and Pereira, Andy and Stiekema, Willem and Allefs, Sjefke},
  journal={The Plant Journal},
  volume={36},
  number={6},
  pages={867--882},
  year={2003},
  publisher={Wiley Online Library}
}

@article{johnston1980significance,
  title={The significance of genic balance to endosperm development in interspecific crosses},
  author={Johnston, SA and Den Nijs, TPM and Peloquin, SJ and Hanneman Jr, RE},
  journal={Theoretical and applied genetics},
  volume={57},
  number={1},
  pages={5--9},
  year={1980},
  publisher={Springer}
}

@article{jarvis1999wild,
  title={Wild relatives and crop cultivars: detecting natural introgression and farmer selection of new genetic combinations in agroecosystems},
  author={Jarvis, Devra I and Hodgkin, Toby},
  journal={Molecular ecology},
  volume={8},
  number={s1},
  pages={S159--S173},
  year={1999},
  publisher={Wiley Online Library}
}

@article{rieseberg1999introgression,
  title={Introgression between the cultivated sunflower and a sympatric wild relative, Helianthus petiolaris (Asteraceae)},
  author={Rieseberg, Loren H and Kim, Min Ju and Seiler, Gerald J},
  journal={International Journal of Plant Sciences},
  volume={160},
  number={1},
  pages={102--108},
  year={1999},
  publisher={JSTOR}
}

@article{kim1999genetic,
  title={Genetic architecture of species differences in annual sunflowers: implications for adaptive trait introgression},
  author={Kim, Seung-Chul and Rieseberg, Loren H},
  journal={Genetics},
  volume={153},
  number={2},
  pages={965--977},
  year={1999},
  publisher={Genetics Soc America}
}

@article{papa2003asymmetry,
  title={Asymmetry of gene flow and differential geographical structure of molecular diversity in wild and domesticated common bean (Phaseolus vulgaris L.) from Mesoamerica},
  author={Papa, R and Gepts, P},
  journal={Theoretical and Applied Genetics},
  volume={106},
  number={2},
  pages={239--250},
  year={2003},
  publisher={Springer}
}

@article{blackman2011sunflower,
  title={Sunflower domestication alleles support single domestication center in eastern North America},
  author={Blackman, Benjamin K and Scascitelli, Moira and Kane, Nolan C and Luton, Harry H and Rasmussen, David A and Bye, Robert A and Lentz, David L and Rieseberg, Loren H},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={34},
  pages={14360--14365},
  year={2011},
  publisher={National Acad Sciences}
}

@article{wills2006chloroplast,
  title={Chloroplast DNA variation confirms a single origin of domesticated sunflower (Helianthus annuus L.)},
  author={Wills, David M and Burke, John M},
  journal={Journal of Heredity},
  volume={97},
  number={4},
  pages={403--408},
  year={2006},
  publisher={Am Genetic Assoc}
}

@article{vandenberg1998collapse,
  title={Collapse of morphological species in the wild potato Solanum brevicaule complex (Solanaceae: sect. Petota).},
  author={Vandenberg, R and Miller, J and Ugarte, M and Kardolus, J and Villand, Julie and Nienhuis, James and Spooner, D},
  journal={American Journal of Botany},
  volume={85},
  number={1},
  pages={92--92},
  year={1998},
  publisher={Botanical Soc America}
}

@book{hawkes1990potato,
  title={The potato: evolution, biodiversity and genetic resources.},
  author={Hawkes, John Gregory and others},
  year={1990},
  publisher={Belhaven Press}
}

@article{spooner2001potato,
  title={Potato systematics and germplasm collecting, 1989--2000},
  author={Spooner, David M and Hijmans, Robert J},
  journal={American Journal of Potato Research},
  volume={78},
  number={4},
  pages={237--268},
  year={2001},
  publisher={Springer}
}

@article{sharma2000origin,
  title={Origin of O. sativa and its ecotypes},
  author={Sharma, SD and Tripathy, Smita and Biswal, Jyostnamayee},
  journal={Rice breeding and genetics: research priorities and challenges. Enfield, NH, USA: Science Publications},
  pages={349--369},
  year={2000}
}

@article{oka1988origin,
  title={Origin of cultivated rice. Jpn},
  author={Oka, HI},
  journal={Sci. Soc. Press, Elsevier, Tokyo},
  volume={254},
  year={1988}
}

@article{robertson2007genetic,
  title={Genetic resources of tomato (Lycopersicon esculentum Mill.) and wild relatives},
  author={Robertson, LARRY D and Labate, JOANNE A and Razdan, MK and Mattoo, AK},
  journal={Genetic Improvement of Solanaceous Crops. Tomato},
  volume={2},
  pages={25--75},
  year={2007}
}

@article{giles1985progenitor,
  title={The progenitor of barley (Hordeum vulgare spp. spontaneum)-its importance as a gene resource [incl. hordeins, Erysiphe graminis].},
  author={Giles, BE and others},
  journal={Sveriges Utsaedesfoerenings Tidskrift},
  year={1985}
}

@article{bai2007domestication,
  title={Domestication and breeding of tomatoes: what have we gained and what can we gain in the future?},
  author={Bai, Yuling and Lindhout, Pim},
  journal={Annals of Botany},
  volume={100},
  number={5},
  pages={1085--1094},
  year={2007},
  publisher={Annals Botany Co}
}

@article{rick1953novel,
  title={Novel variation in tomato species hybrids},
  author={Rick, Charles M and Smith, Paul G},
  journal={American Naturalist},
  pages={359--373},
  year={1953},
  publisher={JSTOR}
}

http://www.landscapeimagery.com/tomato.html
Cox S. 2000. I Say Tomayto, You Say Tomahto. http://www.landscapeimagery.com/tomato.html

@article{nesbitt2002comparative,
  title={Comparative sequencing in the genus Lycopersicon: implications for the evolution of fruit size in the domestication of cultivated tomatoes},
  author={Nesbitt, T Clint and Tanksley, Steven D},
  journal={Genetics},
  volume={162},
  number={1},
  pages={365--379},
  year={2002},
  publisher={Genetics Soc America}
}

@article{heiser1951hybridization,
  title={Hybridization in the annual sunflowers: Helianthus annuus$\times$ H. debilis var. cucumerifolius},
  author={Heiser Jr, Charles B},
  journal={Evolution},
  pages={42--51},
  year={1951},
  publisher={JSTOR}
}

@article{rieseberg1999hybrid,
  title={Hybrid zones and the genetic architecture of a barrier to gene flow between two sunflower species},
  author={Rieseberg, Loren H and Whitton, Jeannette and Gardner, Keith},
  journal={Genetics},
  volume={152},
  number={2},
  pages={713--727},
  year={1999},
  publisher={Genetics Soc America}
}

@article{kwak2009putative,
  title={The Putative Mesoamerican Domestication Center of Is Located in the Lerma--Santiago Basin of Mexico},
  author={Kwak, Myounghai and Kami, James A and Gepts, Paul},
  journal={Crop Science},
  volume={49},
  number={2},
  pages={554--563},
  year={2009},
  publisher={Crop Science Society of America}
}

@article{kami1995identification,
  title={Identification of presumed ancestral DNA sequences of phaseolin in Phaseolus vulgaris},
  author={Kami, James and Vel{\'a}squez, V Becerra and Debouck, Daniel G and Gepts, Paul},
  journal={Proceedings of the National Academy of Sciences},
  volume={92},
  number={4},
  pages={1101--1104},
  year={1995},
  publisher={National Acad Sciences}
}

@article{park1987transfer,
  title={Transfer of common blight (Xanthomonas campestris pv phaseoli) resistance from Phaseolus coccineus Lam. to P. vulgaris L. through interspecific hybridization},
  author={Park, SJ and Dhanvantari, BN},
  journal={Canadian journal of plant science},
  volume={67},
  number={3},
  pages={685--695},
  year={1987},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{schwartz2006inheritance,
  title={Inheritance of white mold resistance in Phaseolus vulgaris$\times$ P. coccineus crosses},
  author={Schwartz, Howard F and Otto, Kristen and Ter{\'a}n, Henry and Lema, Margarita and Singh, Shree P},
  journal={Plant Disease},
  volume={90},
  number={9},
  pages={1167--1170},
  year={2006},
  publisher={Am Phytopath Society}
}

@article{voysest1991bean,
  title={Bean cultivars: classes and commercial seed types},
  author={Voysest, Oswaldo and Dessert, Michael},
  journal={Common beans: research for crop improvement},
  pages={119--162},
  year={1991},
  publisher={CAB Int Wallingford, UK and CIAT, Cali, Colombia}
}

@article{gutierrez1985heterosis,
  title={Heterosis and inbreeding depression in dry bush beans, Phaseolus vulgaris L.},
  author={GUTIERREZ, J ARIEL and SINGH, SHREE P},
  journal={Canadian journal of plant science},
  volume={65},
  number={2},
  pages={243--249},
  year={1985},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{singh1995potential,
  title={Potential of wild common bean for seed yield improvement of cultivars in the tropics},
  author={Singh, Shree P and Molina, Albeiro and Gepts, Paul},
  journal={Canadian Journal of Plant Science},
  volume={75},
  number={4},
  pages={807--813},
  year={1995},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{burle2010microsatellite,
  title={Microsatellite diversity and genetic structure among common bean (Phaseolus vulgaris L.) landraces in Brazil, a secondary center of diversity},
  author={Burle, Mar{\'\i}lia Lobo and Fonseca, Jaime Roberto and Kami, James A and Gepts, Paul},
  journal={Theoretical and Applied Genetics},
  volume={121},
  number={5},
  pages={801--813},
  year={2010},
  publisher={Springer}
}

@article{ramalho2006,
[[Ramalho MAP, Abreu AFB (2006) Cultivares. In: Vieira C et al (eds) Feijao, 2nd edn. UFV, Vicosa, pp415-436]]

@article{gepts1986phaseolin,
  title={Phaseolin variability among wild and cultivated common beans (Phaseolus vulgaris) from Colombia},
  author={Gepts, P and Bliss, FA},
  journal={Economic Botany},
  volume={40},
  number={4},
  pages={469--478},
  year={1986},
  publisher={Springer}
}

@article{paredes1995extensive,
  title={Extensive introgression of Middle American germplasm into Chilean common bean cultivars},
  author={Paredes, Oscar M and Gepts, Paul},
  journal={Genetic Resources and Crop Evolution},
  volume={42},
  number={1},
  pages={29--41},
  year={1995},
  publisher={Springer}
}

@article{sotelo1995comparative,
  title={Comparative study of the chemical composition of wild and cultivated beans (Phaseolus vulgaris)},
  author={Sotelo, Angela and Sousa, H and Sanchez, M},
  journal={Plant Foods for Human Nutrition},
  volume={47},
  number={2},
  pages={93--100},
  year={1995},
  publisher={Springer}
}

@inproceedings{debouck1988implications,
  title={Implications for bean breeders of studies on the origin of common beans, Phaseolus vulgaris L},
  author={Debouck, DG and THOME, S},
  booktitle={THE BEAN breeders. Cali, Workshop at CIAT},
  year={1988}
}

@article{debouck1993genetic,
  title={Genetic diversity and ecological distribution ofPhaseolus vulgaris (Fabaceae) in northwestern South America},
  author={Debouck, Daniel G and Toro, Orlando and Paredes, Oscar M and Johnson, William C and Gepts, Paul},
  journal={Economic botany},
  volume={47},
  number={4},
  pages={408--423},
  year={1993},
  publisher={Springer}
}

@article{kornegay1991inheritance,
  title={Inheritance of resistance to Acanthoscelides obtectus in a wild common bean accession crossed to commercial bean cultivars},
  author={Kornegay, Julia L and Cardona, Cesar},
  journal={Euphytica},
  volume={52},
  number={2},
  pages={103--111},
  year={1991},
  publisher={Springer}
}

@article{komatsuda2007six,
  title={Six-rowed barley originated from a mutation in a homeodomain-leucine zipper I-class homeobox gene},
  author={Komatsuda, Takao and Pourkheirandish, Mohammad and He, Congfen and Azhaguvel, Perumal and Kanamori, Hiroyuki and Perovic, Dragan and Stein, Nils and Graner, Andreas and Wicker, Thomas and Tagiri, Akemi and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={104},
  number={4},
  pages={1424--1429},
  year={2007},
  publisher={National Acad Sciences}
}

@article{harlan1968origin,
  title={On the origin of barley},
  author={Harlan, Jack R and others},
  journal={USDA Agriculture Handbook},
  volume={338},
  pages={9--31},
  year={1968}
}

@misc{von1995ecographical,
  title={An ecographical study of genus Hordeum},
  author={Von Bothmer, R and Jacobsen, N and Baden, C and J{\o}rgensen, RB and Linde-Laursen, I},
  year={1995},
  publisher={Systematic and ecographic studies on crop gene pools}
}

@article{ellis2000wild,
  title={Wild barley: a source of genes for crop improvement in the 21st century?},
  author={Ellis, RP and Forster, BP and Robinson, D and Handley, LL and Gordon, DC and Russell, JR and Powell, W},
  journal={Journal of experimental botany},
  volume={51},
  number={342},
  pages={9--17},
  year={2000},
  publisher={Soc Experiment Biol}
}

@article{handley1994chromosome,
  title={Chromosome 4 controls potential water use efficiency ($\delta$13C) in barley},
  author={Handley, Linda L and Nevo, Eviatar and Raven, John A and MartInez-Carrasco, Rafael and Scrimgeour, Charles M and Pakniyat, Hassan and Forster, Brian P},
  journal={Journal of Experimental Botany},
  volume={45},
  number={11},
  pages={1661--1663},
  year={1994},
  publisher={Soc Experiment Biol}
}

@article{brown1978outcrossing,
  title={Outcrossing rates and heterozygosity in natural populations of Hordeum spontaneum Koch in Israel},
  author={Brown, AHD and Zohary, D and Nevo, E},
  journal={Heredity},
  volume={41},
  number={1},
  pages={49--62},
  year={1978}
}

@article{takahashi1955origin,
  title={The origin and evolution of cultivated barley},
  author={Takahashi, Ryuhei},
  journal={Advances in genetics},
  volume={7},
  pages={227--266},
  year={1955},
  publisher={Elsevier}
}

@article{huaman2002reclassification,
  title={Reclassification of landrace populations of cultivated potatoes (Solanum sect. Petota)},
  author={Huam{\'a}n, Z{\'o}simo and Spooner, David M},
  journal={American Journal of Botany},
  volume={89},
  number={6},
  pages={947--965},
  year={2002},
  publisher={Botanical Soc America}
}

@article{harlan1966distribution,
  title={Distribution of wild wheats and barley},
  author={Harlan, Jack R and Zohary, Daniel},
  journal={Science},
  volume={153},
  number={3740},
  pages={1074--1080},
  year={1966},
  publisher={American Association for the Advancement of Science}
}

@article{correll1962potato,
  title={The potato and its wild relatives.},
  author={Correll, Donovan Stewart and others},
  journal={The potato and its wild relatives.},
  year={1962},
  publisher={Texas Research Foundation, Renner.}
}

@article{dussle2004pl,
  title={Pl Arg from Helianthus argophyllus is unlinked to other known downy mildew resistance genes in sunflower},
  author={Dussle, CM and Hahn, V and Knapp, SJ and Bauer, E},
  journal={Theoretical and applied genetics},
  volume={109},
  number={5},
  pages={1083--1086},
  year={2004},
  publisher={Springer}
}

@article{rahim2002inheritance,
  title={Inheritance of resistance to sunflower downy mildew races 1, 2 and 3 in cultivated sunflower},
  author={Rahim, M and Jan, CC and Gulya, TJ},
  journal={Plant Breeding},
  volume={121},
  number={1},
  pages={57--60},
  year={2002},
  publisher={Wiley Online Library}
}

@article{miller1991inheritance,
  title={Inheritance of resistance to race 4 of downy mildew derived from interspecific crosses in sunflower},
  author={Miller, JF and Gulya, TJ},
  journal={Crop science},
  volume={31},
  number={1},
  pages={40--43},
  year={1991},
  publisher={Crop Science Society of America}
}

@article{cohen1973factors,
  title={Factors affecting infection of sunflowers by Plasmopara halstedii},
  author={Cohen, Yigal and Sackston, WE},
  journal={Canadian Journal of Botany},
  volume={51},
  number={1},
  pages={15--22},
  year={1973},
  publisher={NRC Research Press}
}

@article{heiser1978taxonomy,
  title={Taxonomy of Helianthus and origin of domesticated sunflower},
  author={Heiser, Charles B},
  journal={Sunflower science and technology, Agronomy Series},
  volume={19},
  pages={31--35},
  year={1978}
}

@book{smith2000sorghum,
  title={Sorghum: Origin, history, technology, and production},
  author={Smith, C Wayne and Frederiksen, Richard A},
  volume={2},
  year={2000},
  publisher={John Wiley \& Sons}
}

@book{doggett1988sorghum,
  title={Sorghum.},
  author={Doggett, Hugh and others},
  number={2. ed.},
  year={1988},
  publisher={Longman Scientific and Technical}
}

@article{harlan1971toward,
  title={Toward a rational classification of cultivated plants},
  author={Harlan, Jack R and de Wet, Jan MJ},
  journal={Taxon},
  pages={509--517},
  year={1971},
  publisher={JSTOR}
}

@webpage{USDAARS2007,
National Genetic Resources Program. Germplasm Resources Information Network (GRIN) [Online Database]
}

@article{garber1950cytotaxonomic,
  title={Cytotaxonomic studies in the genus Sorghum.},
  author={Garber, Edward D and others},
  journal={Univiversity of California Publications in Botany},
  volume={23},
  pages={283--362},
  year={1950}
}

@inproceedings{kuhlman2006interspecific,
  title={Interspecific sorghum breeding using S. macrospermum},
  author={Kuhlman, LC and Burson, BL and Klein, PE and Stelly, DM and Rooney, WL},
  booktitle={Proceedings of the ASA-CSSA-SSA 2006 International Meetings, Indianapolis},
  pages={12--16},
  year={2006}
}

@article{de1978systematics,
  title={Systematics and evolution of Sorghum sect. Sorghum (Gramineae)},
  author={De Wet, JMJ},
  journal={American Journal of Botany},
  pages={477--484},
  year={1978},
  publisher={JSTOR}
}

@article{yatabe2007rampant,
  title={Rampant gene exchange across a strong reproductive barrier between the annual sunflowers, Helianthus annuus and H. petiolaris},
  author={Yatabe, Yoko and Kane, Nolan C and Scotti-Saintagne, Caroline and Rieseberg, Loren H},
  journal={Genetics},
  volume={175},
  number={4},
  pages={1883--1893},
  year={2007},
  publisher={Genetics Soc America}
}

@article{schwarzbach2002likely,
  title={Likely multiple origins of a diploid hybrid sunflower species},
  author={Schwarzbach, AE and Rieseberg, LH},
  journal={Molecular Ecology},
  volume={11},
  number={9},
  pages={1703--1715},
  year={2002},
  publisher={Wiley Online Library}
}

@article{welch2002patterns,
  title={Patterns of genetic variation suggest a single, ancient origin for the diploid hybrid species Helianthus paradoxus},
  author={Welch, Mark E and Rieseberg, Loren H},
  journal={Evolution},
  volume={56},
  number={11},
  pages={2126--2137},
  year={2002},
  publisher={Wiley Online Library}
}

@article{saucă2010introgression,
  title={Introgression of drought-resistance gene (s) from Helianthus argophyllus to Helianthus annuus specie, using embryo rescue techniques.},
  author={Saucă, FLORENTINA and others},
  journal={Romanian Agricultural Research},
  number={27},
  pages={47--51},
  year={2010},
  publisher={National Agricultural Research and Development Institute}
}

@article{besnard1997specifying,
  title={Specifying the introgressed regions from H. argophyllus in cultivated sunflower (Helianthus annuus L.) to mark Phomopsis resistance genes},
  author={Besnard, G and Griveau, Y and Quillet, MC and Serieys, H and Lambert, P and Vares, D and Bervill{\'e}, A},
  journal={Theoretical and applied genetics},
  volume={94},
  number={1},
  pages={131--138},
  year={1997},
  publisher={Springer}
}

@article{shechter1975comparative,
  title={Comparative electrophoresis and isozyme analysis of seed proteins from cultivated races of sorghum},
  author={Shechter, Yaakov and De Wet, JMJ},
  journal={American Journal of Botany},
  pages={254--261},
  year={1975},
  publisher={JSTOR}
}

@article{po1982sorghum,
  title={Sorghum in the Eighties},
  author={PO, ICRISAT Patancheru},
  year={1982}
}

@article{gebrekidan1979sorghum,
  title={Sorghum Genetic Resources in Africa.},
  author={Gebrekidan, Brhane},
  journal={Ethiopian Journal of Agricultural Sciences},
  year={1979}
}

@article{reddy2006current,
  title={Current status of sorghum genetic resources at ICRISAT: their sharing and impacts},
  author={Reddy, V Gopal and Upadhyaya, HD and Gowda, CLL},
  journal={International Sorghum and Millets Newsletter},
  volume={47},
  pages={9--13},
  year={2006}
}

@article{johnson1979breeding,
  title={Breeding for arthropod resistance in sorghum},
  author={Johnson, Jerry W and Teetes, George L},
  journal={Harris, MK Biology and breeding for resistance to arthropods and pathogens in agricultural plants. Texas, USA},
  pages={168--180},
  year={1979}
}

@article{de1976cytogenetics,
  title={Cytogenetics of introgression from Saccharum into Sorghum},
  author={De Wet, JMJ and Gupta, SC and Harlan, JR and Grassl, CO},
  journal={Crop Science},
  volume={16},
  number={4},
  pages={568--572},
  year={1976},
  publisher={Crop Science Society of America}
}

@article{pickersgill1977origins,
  title={Origins and distribution of plants domesticated in the New World tropics},
  author={Pickersgill, Barbara and Heiser Jr, Charles B},
  journal={Origins of agriculture},
  year={1977}
}

@article{brush1995potato,
  title={Potato diversity in the Andean center of crop domestication},
  author={Brush, Stephen and Kesseli, Rick and Ortega, Ramiro and Cisneros, Pedro and Zimmerer, Karl and Quiros, Carlos},
  journal={Conservation Biology},
  volume={9},
  number={5},
  pages={1189--1198},
  year={1995},
  publisher={Wiley Online Library}
}

@article{rabinowitz1990high,
  title={High levels of interspecific hybridization betweenSolanum sparsipilum andS. stenotomum in experimental plots in the Andes},
  author={Rabinowitz, D and Linder, CR and Ortega, R and Begazo, D and Murguia, H and Douches, DS and Quiros, CF},
  journal={American Potato Journal},
  volume={67},
  number={2},
  pages={73--81},
  year={1990},
  publisher={Springer}
}

@article{johns1987relationships,
  title={Relationships among wild, weed, and cultivated potatoes in the Solanum x ajanhuiri complex},
  author={Johns, Timothy and Huaman, Zosimo and Ochoa, Carlos and Schmiediche, Peter E},
  journal={Systematic Botany},
  pages={541--552},
  year={1987},
  publisher={JSTOR}
}

@book{linder1987diversity,
  title={Diversity within traditional potato agriculture in the Peruvian Andes and its relevance to crop evolution},
  author={Linder, Craig Randal},
  year={1987},
  publisher={Cornell University, Aug.}
}

@article{johns1986ongoing,
  title={Ongoing evolution of the potato on the altiplano of western Bolivia},
  author={Johns, Timothy and Keen, Susan L},
  journal={Economic Botany},
  volume={40},
  number={4},
  pages={409--424},
  year={1986},
  publisher={Springer}
}

@article{hawkes1962origin,
  title={ORIGIN OF SOLANUM JUZEPCZUKII BUK AND S CURTILOBUM JUZ ET BUK},
  author={Hawkes, John Gregory},
  journal={ZEITSCHRIFT FUR PFLANZENZUCHTUNG-JOURNAL OF PLANT BREEDING},
  volume={47},
  number={1},
  pages={1},
  year={1962},
  publisher={Verlag Paul Parey SPITALERST. 12, D-2000 HAMBURG 1, GERMANY}
}

@article{schmiediche1980breeding,
  title={Breeding of the cultivated potato species Solanum x juzepczukii Buk. and Solanum x curtilobum Juz. etBuk.},
  author={Schmiediche, PE and Hawkes, JG and Ochoa, CM},
  journal={Euphytica},
  volume={29},
  number={3},
  pages={685--704},
  year={1980},
  publisher={Springer}
}

@article{brush1981dynamics,
  title={Dynamics of Andean potato agriculture},
  author={Brush, Stephen B and Carney, Heath J and Hum{\'a}n, Z{\'o}simo},
  journal={Economic Botany},
  volume={35},
  number={1},
  pages={70--88},
  year={1981},
  publisher={Springer}
}

@article{hawkes1988evolution,
  title={The evolution of cultivated potatoes and their tuber-bearing wild relatives},
  author={Hawkes, John G},
  journal={Die Kulturpflanze},
  volume={36},
  number={1},
  pages={189--208},
  year={1988},
  publisher={Springer}
}

@article{kamala2012interactions,
  title={Interactions of spotted stem borer Chilo partellus with wild relatives of sorghum},
  author={Kamala, Venkateswaran and Sharma, Hari C and Manohar Rao, Daggu and Varaprasad, Kodeboyina S and Bramel, Paula J and Chandra, Subhash},
  journal={Plant breeding},
  volume={131},
  number={4},
  pages={511--521},
  year={2012},
  publisher={Wiley Online Library}
}

@book{ellstrand2003dangerous,
  title={Dangerous liaisons?: when cultivated plants mate with their wild relatives},
  author={Ellstrand, Norman Carl},
  year={2003},
  publisher={JHU Press}
}

@article{harlan1995living,
  title={The living fields},
  author={Harlan, Jack R},
  journal={Our agricultural heritage.-271 S},
  year={1995}
}

@article{doggett1968disruptive,
  title={Disruptive selection in crop development},
  author={Doggett, H and Majisu, BN},
  journal={Heredity},
  volume={23},
  number={1},
  pages={1--23},
  year={1968}
}

@article{baker1972human,
  title={Human influences on plant evolution},
  author={Baker, Herbert G},
  journal={Economic Botany},
  volume={26},
  number={1},
  pages={32--43},
  year={1972},
  publisher={Springer}
}

@article{aldrich1992restriction,
  title={Restriction fragment variation in the nuclear and chloroplast genomes of cultivated and wild Sorghum bicolor},
  author={Aldrich, PR and Doebley, J},
  journal={Theoretical and Applied Genetics},
  volume={85},
  number={2-3},
  pages={293--302},
  year={1992},
  publisher={Springer}
}

@article{aldrich1992patterns,
  title={Patterns of allozyme variation in cultivated and wild Sorghum bicolor},
  author={Aldrich, PR and Doebley, J and Schertz, KF and Stec, A},
  journal={Theoretical and Applied Genetics},
  volume={85},
  number={4},
  pages={451--460},
  year={1992},
  publisher={Springer}
}

@article{arriola1996crop,
  title={Crop-to-weed gene flow in the genus Sorghum (Poaceae): spontaneous interspecific hybridization between johnsongrass, Sorghum halepense, and crop sorghum, S. bicolor},
  author={Arriola, Paul E and Ellstrand, Norman C},
  journal={American Journal of Botany},
  pages={1153--1159},
  year={1996},
  publisher={JSTOR}
}

@article{kwak2009structure,
  title={Structure of genetic diversity in the two major gene pools of common bean (Phaseolus vulgaris L., Fabaceae)},
  author={Kwak, Myounghai and Gepts, Paul},
  journal={Theoretical and Applied Genetics},
  volume={118},
  number={5},
  pages={979--992},
  year={2009},
  publisher={Springer}
}

@article{molina2011molecular,
  title={Molecular evidence for a single evolutionary origin of domesticated rice},
  author={Molina, Jeanmaire and Sikora, Martin and Garud, Nandita and Flowers, Jonathan M and Rubinstein, Samara and Reynolds, Andy and Huang, Pu and Jackson, Scott and Schaal, Barbara A and Bustamante, Carlos D and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={20},
  pages={8351--8356},
  year={2011},
  publisher={National Acad Sciences}
}

@article{zhao2010genomic,
  title={Genomic diversity and introgression in O. sativa reveal the impact of domestication and breeding on the rice genome},
  author={Zhao, Keyan and Wright, Mark and Kimball, Jennifer and Eizenga, Georgia and McClung, Anna and Kovach, Michael and Tyagi, Wricha and Ali, M Likat and Tung, Chih-Wei and Reynolds, Andy and others},
  journal={PloS one},
  volume={5},
  number={5},
  pages={e10780},
  year={2010}
}

@article{chang2003origin,
  title={Origin, domestication, and diversification},
  author={Chang, Te-Tzu},
  journal={Rice: origin, history, technology, and production. New Jersey USA: J. Wiley and Sons, Inc},
  pages={3--25},
  year={2003}
}

@article{khush2003classifying,
  title={Classifying rice germplasm by isozyme polymorphism and origin of cultivated rice},
  author={Khush, GS and Brar, DS and Virk, PS and Tang, SX and Malik, SS and Busto, GA and Lee, YT and McNally, R and Trinh, LN and Jiang, Y and others},
  journal={IRRI Discuss Pap},
  volume={46},
  pages={279},
  year={2003}
}

@article{glaszmann1987isozymes,
  title={Isozymes and classification of Asian rice varieties},
  author={Glaszmann, JC},
  journal={Theoretical and Applied Genetics},
  volume={74},
  number={1},
  pages={21--30},
  year={1987},
  publisher={Springer}
}

@article{zohary1959hordeum,
  title={Is Hordeum a griocrithon the Ancestor of Six-Rowed Cultivated Barley?},
  author={Zohary, Daniel},
  journal={Evolution},
  volume={13},
  number={2},
  pages={279--280},
  year={1959},
  publisher={JSTOR}
}

@article{staudt1961origin,
  title={The origin of cultivated barleys: a discussion},
  author={Staudt, G},
  journal={Economic Botany},
  volume={15},
  number={3},
  pages={205--212},
  year={1961},
  publisher={Springer}
}

@article{salamini2002genetics,
  title={Genetics and geography of wild cereal domestication in the Near East},
  author={Salamini, Francesco and {\"O}zkan, Hakan and Brandolini, Andrea and Sch{\"a}fer-Pregl, Ralf and Martin, William},
  journal={Nature Reviews Genetics},
  volume={3},
  number={6},
  pages={429--441},
  year={2002},
  publisher={Nature Publishing Group}
}

@article{civavn2013reticulated,
  title={Reticulated origin of domesticated emmer wheat supports a dynamic model for the emergence of agriculture in the fertile crescent},
  author={Civ{\'a}{\v{n}}, Peter and Ivani{\v{c}}ov{\'a}, Zuzana and Brown, Terence A},
  year={2013}
}

@book{hancock2012plant,
  title={Plant evolution and the origin of crop species},
  author={Hancock, James F},
  year={2012},
  publisher={CABI}
}

@article{hillman2001new,
  title={New evidence of Lateglacial cereal cultivation at Abu Hureyra on the Euphrates},
  author={Hillman, Gordon and Hedges, Robert and Moore, Andrew and Colledge, Susan and Pettitt, Paul},
  journal={The Holocene},
  volume={11},
  number={4},
  pages={383--393},
  year={2001},
  publisher={Sage Publications}
}

@article{khush1961cytogenetic,
  title={Cytogenetic and evolutionary studies in Secale. I. Some new data on the ancestry of S. cereale},
  author={Khush, Gurdev S and Stebbins, G Ledyard},
  journal={American Journal of Botany},
  pages={723--730},
  year={1961},
  publisher={JSTOR}
}

@article{kane2007selective,
  title={Selective sweeps reveal candidate genes for adaptation to drought and salt tolerance in common sunflower, Helianthus annuus},
  author={Kane, Nolan C and Rieseberg, Loren H},
  journal={Genetics},
  volume={175},
  number={4},
  pages={1823--1834},
  year={2007},
  publisher={Genetics Soc America}
}

@article{rieseberg1990helianthus,
  title={Helianthus annuus ssp. texanus has chloroplast DNA and nuclear ribosomal RNA genes of Helianthus debilis ssp. cucumerifolius},
  author={Rieseberg, Loren H and Beckstrom-Sternberg, Stephen and Doan, Khiem},
  journal={Proceedings of the National Academy of Sciences},
  volume={87},
  number={2},
  pages={593--597},
  year={1990},
  publisher={National Acad Sciences}
}

@article{stutz1972origin,
  title={On the origin of cultivated rye},
  author={Stutz, Howard C},
  journal={American Journal of Botany},
  pages={59--70},
  year={1972},
  publisher={JSTOR}
}

@article{fuller2010consilience,
  title={Consilience of genetics and archaeobotany in the entangled history of rice},
  author={Fuller, Dorian Q and Sato, Yo-Ichiro and Castillo, Cristina and Qin, Ling and Weisskopf, Alison R and Kingwell-Banham, Eleanor J and Song, Jixiang and Ahn, Sung-Mo and Van Etten, Jacob},
  journal={Archaeological and Anthropological Sciences},
  volume={2},
  number={2},
  pages={115--131},
  year={2010},
  publisher={Springer}
}

@article{baker1972migrations,
  title={Migrations of weeds},
  author={Baker, Herbert G},
  journal={Valentine, D, H ed (s). Taxonomy phytogeography and evolution},
  pages={327--347},
  year={1972}
}

@inproceedings{vavilov1928geographical,
  title={Geographical centers of our cultivated plants},
  author={Vavilov, NI},
  booktitle={Proceedings of V international congress on genetics, New York, USA},
  pages={342--369},
  year={1928}
}

@article{quiros1992increase,
  title={Increase of potato genetic resources in their center of diversity: the role of natural outcrossing and selection by the Andean farmer},
  author={Quiros, CF and Ortega, R and Van Raamsdonk, L and Herrera-Montoya, M and Cisneros, P and Schmidt, E and Brush, SB},
  journal={Genetic Resources and Crop Evolution},
  volume={39},
  number={2},
  pages={107--113},
  year={1992},
  publisher={Springer}
}

@article{zimmerer1998ecogeography,
  title={The ecogeography of Andean potatoes},
  author={Zimmerer, Karl S},
  journal={BioScience},
  pages={445--454},
  year={1998},
  publisher={JSTOR}
}

@article{zimmerer1991geographical,
  title={Geographical approaches to crop conservation: the partitioning of genetic diversity in Andean potatoes},
  author={Zimmerer, Karl S and Douches, David S},
  journal={Economic Botany},
  volume={45},
  number={2},
  pages={176--189},
  year={1991},
  publisher={Springer}
}

@article{hamrick1983distribution,
  title={The distribution of genetic variation within and among natural plant populations.},
  author={Hamrick, James L},
  journal={BIOL. CONSERV. SER.},
  pages={335--348},
  year={1983}
}

@article{hammer1987collection,
  title={Collection of plant genetic resources in South Italy, 1986},
  author={Hammer, Karl and Perrino, Pietro and Pignone, Domenico},
  journal={Die Kulturpflanze},
  volume={35},
  number={2},
  pages={389--399},
  year={1987},
  publisher={Springer}
}

@article{murphy1982origin,
  title={The origin of six-rowed ‘wild’barley from the western Himalaya},
  author={Murphy, PJ and Witcombe, JR and Shewry, PR and Miflin, BJ},
  journal={Euphytica},
  volume={31},
  number={1},
  pages={183--192},
  year={1982},
  publisher={Springer}
}

@article{huaman1980solanum,
  title={Solanum ajanhuiri: an important diploid potato cultivated in the Andean altiplano},
  author={Huam{\'a}n, Z and Hawkes, JG and Rowe, PR},
  journal={Economic Botany},
  volume={34},
  number={4},
  pages={335--343},
  year={1980},
  publisher={Springer}
}

@article{carney2000evolutionary,
  title={Evolutionary changes over the fifty-year history of a hybrid population of sunflowers (Helianthus)},
  author={Carney, Shanna E and Gardner, Keith A and Rieseberg, Loren H},
  journal={Evolution},
  volume={54},
  number={2},
  pages={462--474},
  year={2000},
  publisher={BioOne}
}

@article{rieseberg1988molecular,
  title={A molecular reexamination of introgression between Helianthus annuus and H bolanderi (Compositae)},
  author={Rieseberg, Loren H and Soltis, Douglas E and Palmer, Jeffrey D},
  journal={Evolution},
  pages={227--238},
  year={1988},
  publisher={JSTOR}
}

@article{gerechter1971transfer,
  title={Transfer of stem rust seedling resistance from wild diploid einkorn to tetraploid durum wheat by means of a triploid hybrid bridge},
  author={Gerechter-Amitai, ZK and Wahl, I and Vardi, Aliza and Zohary, D},
  journal={Euphytica},
  volume={20},
  number={2},
  pages={281--285},
  year={1971},
  publisher={Springer}
}

@article{blanco2008molecular,
  title={Molecular mapping of the novel powdery mildew resistance gene Pm36 introgressed from Triticum turgidum var. dicoccoides in durum wheat},
  author={Blanco, Antonio and Gadaleta, A and Cenci, A and Carluccio, AV and Abdelbacki, AMM and Simeone, R},
  journal={Theoretical and Applied Genetics},
  volume={117},
  number={1},
  pages={135--142},
  year={2008},
  publisher={Springer}
}

@article{kane2008genetics,
  title={Genetics and evolution of weedy Helianthus annuus populations: adaptation of an agricultural weed},
  author={Kane, Nolan C and Rieseberg, Loren H},
  journal={Molecular Ecology},
  volume={17},
  number={1},
  year={2008},
  publisher={Wiley Online Library}
}

@article{maron2004rapid,
  title={Rapid evolution of an invasive plant},
  author={Maron, John L and Vil{\`a}, Montserrat and Bommarco, Riccardo and Elmendorf, Sarah and Beardsley, Paul},
  journal={Ecological Monographs},
  volume={74},
  number={2},
  pages={261--280},
  year={2004},
  publisher={Eco Soc America}
}

@article{ozkan2002aflp,
  title={AFLP analysis of a collection of tetraploid wheats indicates the origin of emmer and hard wheat domestication in southeast Turkey},
  author={{\"O}zkan, Hakan and Brandolini, Andrea and Sch{\"a}fer-Pregl, Ralf and Salamini, Francesco},
  journal={Molecular Biology and Evolution},
  volume={19},
  number={10},
  pages={1797--1801},
  year={2002},
  publisher={SMBE}
}

@article{luo2007structure,
  title={The structure of wild and domesticated emmer wheat populations, gene flow between them, and the site of emmer domestication},
  author={Luo, M-C and Yang, Z-L and You, Frank M and Kawahara, Taihachi and Waines, J Giles and Dvorak, J},
  journal={Theoretical and Applied Genetics},
  volume={114},
  number={6},
  pages={947--959},
  year={2007},
  publisher={Springer}
}

@article{dvorak2006molecular,
  title={Molecular characterization of a diagnostic DNA marker for domesticated tetraploid wheat provides evidence for gene flow from wild tetraploid wheat to hexaploid wheat},
  author={Dvorak, Jan and Akhunov, Eduard D and Akhunov, Alina R and Deal, Karin R and Luo, Ming-Cheng},
  journal={Molecular biology and evolution},
  volume={23},
  number={7},
  pages={1386--1396},
  year={2006},
  publisher={SMBE}
}

@article{barton2001role,
  title={The role of hybridization in evolution},
  author={Barton, NH},
  journal={Molecular Ecology},
  volume={10},
  number={3},
  pages={551--568},
  year={2001},
  publisher={Wiley Online Library}
}

@article{stebbins1969significance,
  title={The significance of hybridization for plant taxonomy and evolution},
  author={Stebbins, G Ledyard},
  journal={Taxon},
  pages={26--35},
  year={1969},
  publisher={JSTOR}
}

@article{grun1990evolution,
  title={The evolution of cultivated potatoes},
  author={Grun, Paul},
  journal={Economic Botany},
  volume={44},
  number={3},
  pages={39--55},
  year={1990},
  publisher={Springer}
}

@article{ricepedia,
    url={http://ricepedia.org/rice-as-a-plant/rice-species/cultivated-rice-species},
}

@article{lentz2008sunflower,
  title={Sunflower (Helianthus annuus L.) as a pre-Columbian domesticate in Mexico},
  author={Lentz, David L and Pohl, Mary DeLand and Alvarado, Jos{\'e} Luis and Tarighat, Somayeh and Bye, Robert},
  journal={Proceedings of the National Academy of Sciences},
  volume={105},
  number={17},
  pages={6232--6237},
  year={2008},
  publisher={National Acad Sciences}
}

@article{harter2004origin,
  title={Origin of extant domesticated sunflowers in eastern North America},
  author={Harter, Abigail V and Gardner, Keith A and Falush, Daniel and Lentz, David L and Bye, Robert A and Rieseberg, Loren H},
  journal={Nature},
  volume={430},
  number={6996},
  pages={201--205},
  year={2004},
  publisher={Nature Publishing Group}
}

@article{mcfadden1946origin,
  title={The origin of Triticum spelta and its free-threshing hexaploid relatives},
  author={McFadden, ES and Sears, ER},
  journal={Journal of Heredity},
  volume={37},
  number={4},
  pages={107--116},
  year={1946},
  publisher={Am Genetic Assoc}
}

@article{marcussen2014ancient,
  title={Ancient hybridizations among the ancestral genomes of bread wheat},
  author={Marcussen, Thomas and Sandve, Simen R and Heier, Lise and Spannagl, Manuel and Pfeifer, Matthias and Jakobsen, Kjetill S and Wulff, Brande BH and Steuernagel, Burkhard and Mayer, Klaus FX and Olsen, Odd-Arne and others},
  journal={Science},
  volume={345},
  number={6194},
  pages={1250092},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{petersen2006phylogenetic,
  title={Phylogenetic relationships of Triticum and Aegilops and evidence for the origin of the A, B, and D genomes of common wheat (Triticum aestivum)},
  author={Petersen, Gitte and Seberg, Ole and Yde, Merete and Berthelsen, Kasper},
  journal={Molecular phylogenetics and evolution},
  volume={39},
  number={1},
  pages={70--82},
  year={2006},
  publisher={Elsevier}
}

@article{haberer2015barley,
  title={Barley: From Brittle to Stable Harvest},
  author={Haberer, Georg and Mayer, Klaus FX},
  journal={Cell},
  volume={162},
  number={3},
  pages={469--471},
  year={2015},
  publisher={Elsevier}
}

@article{giles2006gludy,
  title={GluDy allele variations in Aegilops tauschii and Triticum aestivum: implications for the origins of hexaploid wheats},
  author={Giles, Rachel J and Brown, Terence A},
  journal={Theoretical and applied genetics},
  volume={112},
  number={8},
  pages={1563--1572},
  year={2006},
  publisher={Springer}
}

@article{ting1933wild,
  title={Wild rice of Kwangtung and new variety bred from the hybrids of wild rice with cultivated rice},
  author={Ting, YING},
  journal={Coll. Agric. Sun Yatsen Univ., Agron. Bull},
  volume={3},
  pages={1--22},
  year={1933}
}

@webpage{CWR,
    url={http://www.cwrdiversity.org/distribution-map/}
}

@article{londo2006phylogeography,
  title={Phylogeography of Asian wild rice, Oryza rufipogon, reveals multiple independent domestications of cultivated rice, Oryza sativa},
  author={Londo, Jason P and Chiang, Yu-Chung and Hung, Kuo-Hsiang and Chiang, Tzen-Yuh and Schaal, Barbara A},
  journal={Proceedings of the National Academy of Sciences},
  volume={103},
  number={25},
  pages={9578--9583},
  year={2006},
  publisher={National Acad Sciences}
}

@article{zohary1975beginnings,
  title={Beginnings of fruit growing in the Old World},
  author={Zohary, Daniel and Spiegel-Roy, Pinhas},
  journal={Science},
  volume={187},
  number={4174},
  pages={319--327},
  year={1975},
  publisher={American Association for the Advancement of Science}
}

@article{austin1988taxonomy,
  title={The taxonomy, evolution and genetic diversity of sweet potatoes and related wild species},
  author={Austin, Daniel F},
  journal={Exploration, maintenance, and utilization of sweetpotato genetic resources. International Potato Center, Lima, Peru},
  pages={27--60},
  year={1988}
}

@phdthesis{akinbo2008introgression,
  title={Introgression of high protein and pest resistance genes from inter-specific hybrids of Manihot esculenta ssp flabellifolia into cassava (Manihot esculenta Crantz)},
  author={Akinbo, Olalekan Abiodun},
  year={2008},
  school={University of the Free State}
}

@article{tovar2015diversity,
  title={Diversity and genetic structure of cassava landraces and their wild relatives (Manihot spp.) in Colombia revealed by simple sequence repeats},
  author={Tovar, E and Bocanegra, JL and Villafa{\~n}e, C and Fory, L and Vel{\'a}squez, A and Gallego, G and Moreno, R},
  journal={Plant Genetic Resources},
  pages={1--11},
  year={2015},
  publisher={Cambridge Univ Press}
}

@article{roullier2013disentangling,
  title={Disentangling the origins of cultivated sweet potato (Ipomoea batatas (L.) Lam.)},
  author={Roullier, Caroline and Duputi{\'e}, Anne and Wennekes, Paul and Benoit, Laure and Bringas, V{\'\i}ctor Manuel Fern{\'a}ndez and Rossel, Genoveva and Tay, David and McKey, Doyle and Lebot, Vincent},
  journal={PLoS One},
  volume={8},
  number={5},
  pages={e62707},
  year={2013},
  publisher={Public Library of Science}
}

@article{roa1997aflp,
  title={AFLP analysis of relationships among cassava and other Manihot species},
  author={Roa, Ana Carolina and Maya, Mar{\'\i}a Mercedes and Duque, MC and Tohme, J and Allem, Antonio Costa and Bonierbale, Merideth W},
  journal={Theoretical and applied Genetics},
  volume={95},
  number={5-6},
  pages={741--750},
  year={1997},
  publisher={Springer}
}

@article{olsen1999evidence,
  title={Evidence on the origin of cassava: phylogeography of Manihot esculenta},
  author={Olsen, Kenneth M and Schaal, Barbara A},
  journal={Proceedings of the National Academy of Sciences},
  volume={96},
  number={10},
  pages={5586--5591},
  year={1999},
  publisher={National Acad Sciences}
}

@article{daniels1987taxonomy,
  title={Taxonomy and evolution},
  author={Daniels, J and Roach, Brian T},
  journal={Sugarcane improvement through breeding. Elsevier, Amsterdam},
  volume={7},
  pages={7--84},
  year={1987}
}

@article{jackson2005breeding,
  title={Breeding for improved sugar content in sugarcane},
  author={Jackson, Phillip A},
  journal={Field Crops Research},
  volume={92},
  number={2},
  pages={277--290},
  year={2005},
  publisher={Elsevier}
}

@article{brandes1958origin,
  title={Origin, classification and characteristics},
  author={Brandes, EW},
  journal={Sugarcane},
  pages={1--35},
  year={1958}
}

@article{irvine1999saccharum,
  title={Saccharum species as horticultural classes},
  author={Irvine, JE},
  journal={Theoretical and Applied Genetics},
  volume={98},
  number={2},
  pages={186--194},
  year={1999},
  publisher={Springer}
}

@inproceedings{roach1989origin,
  title={Origin and improvement of the genetic base of sugarcane},
  author={Roach, BT},
  booktitle={Proc Aust Soc Sugar Cane Technol},
  volume={11},
  pages={34--47},
  year={1989}
}

@webpage{keyroyalbotanicalgardens},
  url={http://www.kew.org/science-conservation/plants-fungi/saccharum-officinarum-sugar-cane},
}

@article{artschwager1958sugarcane,
  title={Sugarcane (Saccharum officinarum L.).},
  author={Artschwager, Ernst and Brandes, Elmer Walker and others},
  journal={Sugarcane (Saccharum officinarum L.).},
  year={1958},
  publisher={USDA}
}

@article{daniels1987taxonomy,
  title={Taxonomy and evolution},
  author={Daniels, J and Roach, Brian T},
  journal={Sugarcane improvement through breeding. Elsevier, Amsterdam},
  volume={7},
  pages={7--84},
  year={1987}
}

@article{roach1995sugar,
  title={Sugar canes--Saccharum (Gramineae-Andropogonae)},
  author={Roach, BT},
  year={1995},
  publisher={in: in: J. Smartt and NW Simmonds (Eds.), 160--166}
}

@article{sobral1994phylogenetic,
  title={Phylogenetic analysis of chloroplast restriction enzyme site mutations in the Saccharinae Griseb. subtribe of the Andropogoneae Dumort. tribe},
  author={Sobral, BWS and Braga, DPV and LaHood, ES and Keim, P},
  journal={Theoretical and Applied Genetics},
  volume={87},
  number={7},
  pages={843--853},
  year={1994},
  publisher={Springer}
}

@techreport{stevenson1965genetics,
  title={Genetics and breeding of sugar cane},
  author={Stevenson, Griffith Cross and others},
  year={1965},
  institution={Cambridge Univ Press}
}

@article{ellstrand1999gene,
  title={Gene flow and introgression from domesticated plants into their wild relatives},
  author={Ellstrand, Norman C and Prentice, Honor C and Hancock, James F},
  journal={Annual review of Ecology and Systematics},
  pages={539--563},
  year={1999},
  publisher={JSTOR}
}

#Li had these in a separate reference.bib file on Github:

@article{andrew2012sunflower,
  title={Adaptation with gene flow across the landscape in a dune sunflower},
  author={Andrew, Rose L and Ostevik, Katherine L and Ebert, Daniel P and Rieseberg, Loren H},
  journal={Molecular ecology},
  volume={21},
  number={9},
  pages={2078--2091},
  year={2012},
  publisher={Wiley Online Library}
}

@article{feder2012,
  title={The genomics of speciation-with-gene-flow},
  author={Feder, Jeffrey L and Egan, Scott P and Nosil, Patrik},
  journal={Trends in Genetics},
  volume={28},
  number={7},
  pages={342--350},
  year={2012},
  publisher={Elsevier}
}

@article{racimo2015,
  title={Evidence for archaic adaptive introgression in humans},
  author={Racimo, Fernando and Sankararaman, Sriram and Nielsen, Rasmus and Huerta-S{\'a}nchez, Emilia},
  journal={Nature Reviews Genetics},
  volume={16},
  number={6},
  pages={359--371},
  year={2015},
  publisher={Nature Publishing Group}
}

@article{Takuno2015,
author = {Takuno, Shohei and Ralph, Peter and Swarts, Kelly and Elshire, Rob J. and Glaubitz, Jeffrey C. and Buckler, Edward S. and Hufford, Matthew B. and Ross-Ibarra, Jeffrey}, 
title = {Independent Molecular Basis of Convergent Highland Adaptation in Maize},
year = {2015}, 
doi = {10.1534/genetics.115.178327}, 
URL = {http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.abstract}, 
eprint = {http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.full.pdf+html}, 
journal = {Genetics} 
}

@article{hufford2013,
  title={The genomic signature of crop-wild introgression in maize},
  author={Hufford, Matthew B and Lubinksy, Pesach and Pyh{\"a}j{\"a}rvi, Tanja and Devengenzo, Michael T and Ellstrand, Norman C and Ross-Ibarra, Jeffrey},
  year={2013},
  doi={10.1371/journal.pgen.1003477},
  journal={Plos genetics} 
}

@article{vanHeerwaarden2011,
  title={Genetic signals of origin, spread, and introgression in a large sample of maize landraces},
  author={van Heerwaarden, Joost and Doebley, John and Briggs, William H and Glaubitz, Jeffrey C and Goodman, Major M and Gonzalez, Jose de Jesus Sanchez and Ross-Ibarra, Jeffrey},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={3},
  pages={1088--1092},
  year={2011},
  publisher={National Acad Sciences}
}

@article{matsuoka2002,
  title={A single domestication for maize shown by multilocus microsatellite genotyping},
  author={Matsuoka, Yoshihiro and Vigouroux, Yves and Goodman, Major M and Sanchez, Jesus and Buckler, Edward and Doebley, John},
  journal={Proceedings of the National Academy of Sciences},
  volume={99},
  number={9},
  pages={6080--6084},
  year={2002},
  publisher={National Acad Sciences}
}

@article{wilkes1977,
  title={Hybridization of maize and teosinte, in Mexico and Guatemala and the improvement of maize},
  author={Wilkes, HG},
  journal={Economic Botany},
  pages={254--293},
  year={1977},
  publisher={JSTOR}
}

@article{lauter2004,
  title={The inheritance and evolution of leaf pigmentation and pubescence in teosinte},
  author={Lauter, Nick and Gustus, Charles and Westerbergh, Anna and Doebley, John},
  journal={Genetics},
  volume={167},
  number={4},
  pages={1949--1959},
  year={2004},
  publisher={Genetics Soc America}
}

@article{doebley1984,
  title={Maize introgression into teosinte-a reappraisal},
  author={Doebley, John F},
  journal={Annals of the Missouri Botanical Garden},
  pages={1100--1113},
  year={1984},
  publisher={JSTOR}
}

@article{doebley1987,
  title={Patterns of isozyme variation between maize and Mexican annual teosinte},
  author={Doebley, John and Goodman, Major M and Stuber, Charles W},
  journal={Economic Botany},
  volume={41},
  number={2},
  pages={234--246},
  year={1987},
  publisher={Springer}
}

@article{warburton2011,
  title={Gene flow among different teosinte taxa and into the domesticated maize gene pool},
  author={Warburton, Marilyn L and Wilkes, Garrison and Taba, S and Charcosset, Alain and Mir, Celine and Dumas, Fabrice and Madur, Delphine and Dreisigacker, Susanne and Bedoya, Claudia and Prasanna, BM and others},
  journal={Genetic resources and crop evolution},
  volume={58},
  number={8},
  pages={1243--1261},
  year={2011},
  publisher={Springer}
}

@article{fukunaga2005,
  title={Genetic diversity and population structure of teosinte},
  author={Fukunaga, Kenji and Hill, Jason and Vigouroux, Yves and Matsuoka, Yoshihiro and Sanchez, Jesus and Liu, Kejun and Buckler, Edward S and Doebley, John},
  journal={Genetics},
  volume={169},
  number={4},
  pages={2241--2254},
  year={2005},
  publisher={Genetics Soc America}
}

@article{martin2015,
  title={Evaluating the use of ABBA--BABA statistics to locate introgressed loci},
  author={Martin, Simon H and Davey, John W and Jiggins, Chris D},
  journal={Molecular biology and evolution},
  volume={32},
  number={1},
  pages={244--257},
  year={2015},
  publisher={SMBE}
}

#The following were in the TiG prospectus bib file.

@Article{Hufford2013,
author = {Hufford, MB and Lubinksy, P and Pyh\"{a}j\"{a}rvi, T and Devengenzo, MT and Ellstrand, NC and Ross-Ibarra, J}, 
title = {The genomic signature of crop-wild introgression in maize}, 
journal = {PLoS Genetics}, 
volume = {9}, 
number = {5}, 
pages = {e1003477}, 
year = {2013}, 
abstract = {The evolutionary significance of hybridization and subsequent introgression has long been appreciated, but evaluation of the genome-wide effects of these phenomena has only recently become possible. Crop-wild study systems represent ideal opportunities to examine evolution through hybridization. For example, maize and the conspecific wild teosinte Zea mays ssp. mexicana (hereafter, mexicana) are known to hybridize in the fields of highland Mexico. Despite widespread evidence of gene flow, maize and mexicana maintain distinct morphologies and have done so in sympatry for thousands of years. Neither the genomic extent nor the evolutionary importance of introgression between these taxa is understood. In this study we assessed patterns of genome-wide introgression based on 39,029 single nucleotide polymorphisms genotyped in 189 individuals from nine sympatric maize-mexicana populations and reference allopatric populations. While portions of the maize and mexicana genomes appeared resistant to introgression (notably near known cross-incompatibility and domestication loci), we detected widespread evidence for introgression in both directions of gene flow. Through further characterization of these genomic regions and preliminary growth chamber experiments, we found evidence suggestive of the incorporation of adaptive mexicana alleles into maize during its expansion to the highlands of central Mexico. In contrast, very little evidence was found for adaptive introgression from maize to mexicana. The methods we have applied here can be replicated widely, and such analyses have the potential to greatly inform our understanding of evolution through introgressive hybridization. Crop species, due to their exceptional genomic resources and frequent histories of spread into sympatry with relatives, should be particularly influential in these studies.}, 
location = {Department of Plant Sciences, University of California Davis, Davis, California, United States of America.}, 
keywords = {}}

@article{Racimo2015,
	Author = {Racimo, Fernando and Sankararaman, Sriram and Nielsen, Rasmus and Huerta-Sanchez, Emilia},
	Journal = {Nat Rev Genet},
	M3 = {Review},
	Number = {6},
	Pages = {359--371},
	Title = {Evidence for archaic adaptive introgression in humans},
	Ty = {JOUR},
	Volume = {16},
	Year = {2015}
	}

@article {Baute2015,
author = {Baute, Gregory J. and Kane, Nolan C. and Grassa, Christopher J. and Lai, Zhao and Rieseberg, Loren H.},
title = {Genome scans reveal candidate domestication and improvement genes in cultivated sunflower, as well as post-domestication introgression with wild relatives},
journal = {New Phytologist},
volume = {206},
number = {2},
pages = {830--838},
keywords = {breeding, domestication, genomics, introgression, sunflower (Helianthus spp.)},
year = {2015},
}

@article{Bredeson2016,
	Author = {Bredeson, Jessen V and Lyons, Jessica B and Prochnik, Simon E and Wu, G Albert and Ha, Cindy M and Edsinger-Gonzales, Eric and Grimwood, Jane and Schmutz, Jeremy and Rabbi, Ismail Y and Egesi, Chiedozie and Nauluvula, Poasa and Lebot, Vincent and Ndunguru, Joseph and Mkamilo, Geoffrey and Bart, Rebecca S and Setter, Tim L and Gleadow, Roslyn M and Kulakow, Peter and Ferguson, Morag E and Rounsley, Steve and Rokhsar, Daniel S},
	Journal = {Nat Biotech},
	Number = {5},
	Pages = {562--570},
	Title = {Sequencing wild and cultivated cassava and related species reveals extensive interspecific hybridization and genetic diversity},
	Volume = {34},
	Year = {2016},
	}

@article{Heliconius2012,
	Author = {The Heliconius Genome Consortium},
	Journal = {Nature},
	Number = {7405},
	Pages = {94--98},
	Title = {Butterfly genome reveals promiscuous exchange of mimicry adaptations among species},
	Volume = {487},
	Year = {2012},
	}

@article{Norris2015,
author = {Norris, Laura C. and Main, Bradley J. and Lee, Yoosook and Collier, Travis C. and Fofana, Abdrahamane and Cornel, Anthony J. and Lanzaro, Gregory C.}, 
title = {Adaptive introgression in an African malaria mosquito coincident with the increased usage of insecticide-treated bed nets},
volume = {112}, 
number = {3}, 
pages = {815--820}, 
year = {2015}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@Article{Poets2015,
author="Poets, Ana M.
and Fang, Zhou
and Clegg, Michael T.
and Morrell, Peter L.",
title="Barley landraces are characterized by geographically heterogeneous genomic origins",
journal="Genome Biology",
year="2015",
volume="16",
number="1",
pages="1--11",
}

@article{Huang2012,
	Author = {Huang, Xuehui and Kurata, Nori and Wei, Xinghua and Wang, Zi-Xuan and Wang, Ahong and Zhao, Qiang and Zhao, Yan and Liu, Kunyan and Lu, Hengyun and Li, Wenjun and Guo, Yunli and Lu, Yiqi and Zhou, Congcong and Fan, Danlin and Weng, Qijun and Zhu, Chuanrang and Huang, Tao and Zhang, Lei and Wang, Yongchun and Feng, Lei and Furuumi, Hiroyasu and Kubo, Takahiko and Miyabayashi, Toshie and Yuan, Xiaoping and Xu, Qun and Dong, Guojun and Zhan, Qilin and Li, Canyang and Fujiyama, Asao and Toyoda, Atsushi and Lu, Tingting and Feng, Qi and Qian, Qian and Li, Jiayang and Han, Bin},
	Journal = {Nature},
	Pages = {497--501},
	Title = {A map of rice genome variation reveals the origin of cultivated rice},
	Volume = {490},
	Year = {2012},
	}

@article {Diez2015,
author = {Diez, Concepcion M. and Trujillo, Isabel and Martinez-Urdiroz, Nieves and Barranco, Diego and Rallo, Luis and Marfil, Pedro and Gaut, Brandon S.},
title = {Olive domestication and diversification in the Mediterranean Basin},
journal = {New Phytologist},
volume = {206},
number = {1},
pages = {436--447},
keywords = {center of origin, demographic history, fruit crops, genetic admixture, olive (Olea europaea ssp. europaea), simple-sequence repeats (SSR)},
year = {2015},
}

@article {Stankowski2015,
	author = {Stankowski, Sean and Streisfeld, Matthew A.},
	title = {Introgressive hybridization facilitates adaptive divergence in a recent radiation of monkeyflowers},
	volume = {282},
	number = {1814},
	year = {2015},
	journal = {Proceedings of the Royal Society of London B: Biological Sciences}
}

@article {Currat2008,
author = {Currat, Mathias and Ruedi, Manuel and Petit, Rémy J. and Excoffier, Laurent},
title = {The hidden side of invasions: massive introgression by local genes},
journal = {Evolution},
volume = {62},
number = {8},
pages = {1908--1920},
year = {2008}
}

@article{Martin2015,
author = {Martin, Simon H. and Davey, John W. and Jiggins, Chris D.}, 
title = {Evaluating the Use of ABBA–BABA Statistics to Locate Introgressed Loci},
volume = {32}, 
number = {1}, 
pages = {244-257}, 
year = {2015}, 
journal = {Molecular Biology and Evolution} 
}

@article{Price2009,
    author = {Price, Alkes L. AND Tandon, Arti AND Patterson, Nick AND Barnes, Kathleen C. AND Rafaels, Nicholas AND Ruczinski, Ingo AND Beaty, Terri H. AND Mathias, Rasika AND Reich, David AND Myers, Simon},
    journal = {PLoS Genet},
    title = {Sensitive Detection of Chromosomal Segments of Distinct Ancestry in Admixed Populations},
    year = {2009},
    volume = {5},
    pages = {1-18},
    number = {6}
}

@article{Lawson2012,
    author = {Lawson, Daniel John AND Hellenthal, Garrett AND Myers, Simon AND Falush, Daniel},
    journal = {PLoS Genet},
    title = {Inference of Population Structure using Dense Haplotype Data},
    year = {2012},
    volume = {8},
    pages = {1-16},
    number = {1}
}

@article{pease2015,
  title={Detection and Polarization of Introgression in a Five-taxon Phylogeny},
  author={Pease, James B and Hahn, Matthew W},
  journal={Systematic biology},
  volume={64},
  number={4},
  pages={651--662},
  year={2015},
  publisher={Oxford University Press}
}

@article{rosenzweig2016,
  title={Powerful methods for detecting introgressed regions from population genomic data},
  author={Rosenzweig, Benjamin K and Pease, James B and Besansky, Nora J and Hahn, Matthew W},
  journal={Molecular ecology},
  year={2016},
  publisher={Wiley Online Library}
}

@article{geneva2015,
  title={A new method to scan genomes for introgression in a secondary contact model},
  author={Geneva, Anthony J and Muirhead, Christina A and Kingan, Sarah B and Garrigan, Daniel},
  journal={PloS one},
  volume={10},
  number={4},
  pages={e0118621},
  year={2015},
  publisher={Public Library of Science}
}

%% Li added references
@article{schraiber2015,
  title={Methods and models for unravelling human evolutionary history},
  author={Schraiber, Joshua G and Akey, Joshua M},
  journal={Nature Reviews Genetics},
  year={2015},
  publisher={Nature Publishing Group}
}

%fd:

@article{zhang2016genome,
  title={Genome-wide introgression among distantly related Heliconius butterfly species},
  author={Zhang, Wei and Dasmahapatra, Kanchon K and Mallet, James and Moreira, Gilson RP and Kronforst, Marcus R},
  journal={Genome biology},
  volume={17},
  number={1},
  pages={1},
  year={2016},
  publisher={BioMed Central}
}

@article{malinsky2015,
  title={Genomic islands of speciation separate cichlid ecomorphs in an East African crater lake},
  author={Malinsky, Milan and Challis, Richard J and Tyers, Alexandra M and Schiffels, Stephan and Terai, Yohey and Ngatunga, Benjamin P and Miska, Eric A and Durbin, Richard and Genner, Martin J and Turner, George F},
  journal={Science},
  volume={350},
  number={6267},
  pages={1493--1498},
  year={2015},
  publisher={American Association for the Advancement of Science}
}

%Hapmix
@article{price2009,
  title={Sensitive detection of chromosomal segments of distinct ancestry in admixed populations},
  author={Price, Alkes L and Tandon, Arti and Patterson, Nick and Barnes, Kathleen C and Rafaels, Nicholas and Ruczinski, Ingo and Beaty, Terri H and Mathias, Rasika and Reich, David and Myers, Simon},
  journal={PLoS Genet},
  volume={5},
  number={6},
  pages={e1000519},
  year={2009},
  publisher={Public Library of Science}
}

@article{suarez2016,
  title={Genomic and functional approaches reveal a case of adaptive introgression from Populus balsamifera (balsam poplar) in P. trichocarpa (black cottonwood)},
  author={Suarez-Gonzalez, Adriana and Hefer, Charles A and Christe, Camille and Corea, Oliver and Lexer, Christian and Cronk, Quentin CB and Douglas, Carl J},
  journal={Molecular ecology},
  year={2016},
  publisher={Wiley Online Library}
}

%PCAdmix
@article{brisbin2012pcadmix,
  title={PCAdmix: principal components-based assignment of ancestry along each chromosome in individuals with admixed ancestry from two or more populations},
  author={Brisbin, Abra and Bryc, Katarzyna and Byrnes, Jake and Zakharia, Fouad and Omberg, Larsson and Degenhardt, Jeremiah and Reynolds, Andrew and Ostrer, Harry and Mezey, Jason G and Bustamante, Carlos D},
  journal={Human biology},
  volume={84},
  number={4},
  pages={343},
  year={2012},
  publisher={NIH Public Access}
}

@article{moreno2014genetics,
  title={The genetics of Mexico recapitulates Native American substructure and affects biomedical traits},
  author={Moreno-Estrada, Andr{\'e}s and Gignoux, Christopher R and Fern{\'a}ndez-L{\'o}pez, Juan Carlos and Zakharia, Fouad and Sikora, Martin and Contreras, Alejandra V and Acu{\~n}a-Alonzo, Victor and Sandoval, Karla and Eng, Celeste and Romero-Hidalgo, Sandra and others},
  journal={Science},
  volume={344},
  number={6189},
  pages={1280--1285},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{pugach2016complex,
  title={The complex admixture history and recent southern origins of Siberian populations},
  author={Pugach, Irina and Matveev, Rostislav and Spitsyn, Viktor and Makarov, Sergey and Novgorodov, Innokentiy and Osakovsky, Vladimir and Stoneking, Mark and Pakendorf, Brigitte},
  journal={Molecular biology and evolution},
  pages={msw055},
  year={2016},
  publisher={SMBE}
}

%MULTIMIX
@article{churchhouse2013,
  title={Multiway admixture deconvolution using phased or unphased ancestral panels},
  author={Churchhouse, Claire and Marchini, Jonathan},
  journal={Genetic epidemiology},
  volume={37},
  number={1},
  pages={1--12},
  year={2013},
  publisher={Wiley Online Library}

}

@article{eyheramendy2015,
  title={Genetic structure characterization of Chileans reflects historical immigration patterns},
  author={Eyheramendy, Susana and Martinez, Felipe I and Manevy, Federico and Vial, Cecilia and Repetto, Gabriela M},
  journal={Nature communications},
  volume={6},
  year={2015},
  publisher={Nature Publishing Group}
}

%%Five taxon D statistic Dfoil

@article{fontaine2015,
  title={Extensive introgression in a malaria vector species complex revealed by phylogenomics},
  author={Fontaine, Michael C and Pease, James B and Steele, Aaron and Waterhouse, Robert M and Neafsey, Daniel E and Sharakhov, Igor V and Jiang, Xiaofang and Hall, Andrew B and Catteruccia, Flaminia and Kakani, Evdoxia and others},
  journal={Science},
  volume={347},
  number={6217},
  pages={1258524},
  year={2015},
  publisher={American Association for the Advancement of Science}
}

@techreport{terhorst2017,
  title={Robust and scalable inference of population history from hundreds of unphased whole genomes},
  author={Terhorst, Jonathan and Kamm, John A and Song, Yun S},
  year={2017},
  journal={Nature Genetics},
  volume={49},
  pages={303--309}
  institution={Nature Research}
}

@article{pease2016,
  title={Phylogenomics reveals three sources of adaptive variation during a rapid radiation},
  author={Pease, James B and Haak, David C and Hahn, Matthew W and Moyle, Leonie C},
  journal={PLoS Biol},
  volume={14},
  number={2},
  pages={e1002379},
  year={2016},
  publisher={Public Library of Science}
}

%D statistic
@article{durand2011,
  title={Testing for ancient admixture between closely related populations},
  author={Durand, Eric Y and Patterson, Nick and Reich, David and Slatkin, Montgomery},
  journal={Molecular biology and evolution},
  volume={28},
  number={8},
  pages={2239--2252},
  year={2011},
  publisher={SMBE}
}

@article{heliconius2012,
  title={Butterfly genome reveals promiscuous exchange of mimicry adaptations among species},
  author={Heliconius Genome Consortium and others},
  journal={Nature},
  volume={487},
  number={7405},
  pages={94--98},
  year={2012},
  publisher={Nature Publishing Group}
}

%ABC
@article{beaumont2002,
  title={Approximate Bayesian computation in population genetics},
  author={Beaumont, Mark A and Zhang, Wenyang and Balding, David J},
  journal={Genetics},
  volume={162},
  number={4},
  pages={2025--2035},
  year={2002},
  publisher={Genetics Soc America}
}

%dadi
@article{gutenkunst2009,
  title={Inferring the joint demographic history of multiple populations from multidimensional SNP frequency data},
  author={Gutenkunst, Ryan N and Hernandez, Ryan D and Williamson, Scott H and Bustamante, Carlos D},
  journal={PLoS Genet},
  volume={5},
  number={10},
  pages={e1000695},
  year={2009},
  publisher={Public Library of Science}
}

%IM isolation with migration
@article{hey2004,
  title={Multilocus methods for estimating population sizes, migration rates and divergence time, with applications to the divergence of Drosophila pseudoobscura and D. persimilis},
  author={Hey, Jody and Nielsen, Rasmus},
  journal={Genetics},
  volume={167},
  number={2},
  pages={747--760},
  year={2004},
  publisher={Genetics Soc America}
}

%multiple sequentially Markovian coalescent (MSMC)
@article{schiffels2014,
  title={Inferring human population size and separation history from multiple genome sequences},
  author={Schiffels, Stephan and Durbin, Richard},
  journal={Nature genetics},
  volume={46},
  number={8},
  pages={919--925},
  year={2014},
  publisher={Nature Publishing Group}
}

%RASPberry
@article{wegmann2011,
  title={Recombination rates in admixed individuals identified by ancestry-based inference},
  author={Wegmann, Daniel and Kessner, Darren E and Veeramah, Krishna R and Mathias, Rasika A and Nicolae, Dan L and Yanek, Lisa R and Sun, Yan V and Torgerson, Dara G and Rafaels, Nicholas and Mosley, Thomas and others},
  journal={Nature genetics},
  volume={43},
  number={9},
  pages={847--853},
  year={2011},
  publisher={Nature Publishing Group}
}

@article{christe2016,
  title={Selection against recombinant hybrids maintains reproductive isolation in hybridizing Populus species despite F1 fertility and recurrent gene flow},
  author={Christe, Camille and St{\"o}lting, Kai N and Bresadola, Luisa and Fussi, Barbara and Heinze, Berthold and Wegmann, Daniel and Lexer, Christian},
  journal={Molecular ecology},
  volume={25},
  number={11},
  pages={2482--2498},
  year={2016},
  publisher={Wiley Online Library}
}

%LAMP
@article{sankararaman2008,
  title={Estimating local ancestry in admixed populations},
  author={Sankararaman, Sriram and Sridhar, Srinath and Kimmel, Gad and Halperin, Eran},
  journal={The American Journal of Human Genetics},
  volume={82},
  number={2},
  pages={290--303},
  year={2008},
  publisher={Elsevier}
}

@article{patterson2012,
  title={Ancient admixture in human history},
  author={Patterson, Nick and Moorjani, Priya and Luo, Yontao and Mallick, Swapan and Rohland, Nadin and Zhan, Yiping and Genschoreck, Teri and Webster, Teresa and Reich, David},
  journal={Genetics},
  volume={192},
  number={3},
  pages={1065--1093},
  year={2012},
  publisher={Genetics Soc America}
}

%Gmin
@article{kingan2015,
  title={Genome divergence and gene flow between Drosophila simulans and D. mauritiana},
  author={Kingan, Sarah B and Geneva, Anthony J and Vedanayagam, Jeffrey P and Garrigan, Daniel},
  journal={bioRxiv},
  pages={024711},
  year={2015},
  publisher={Cold Spring Harbor Labs Journals}
}

%fineSTRUCTURE
@article{lawson2012,
  title={Inference of population structure using dense haplotype data},
  author={Lawson, Daniel John and Hellenthal, Garrett and Myers, Simon and Falush, Daniel},
  journal={PLoS Genet},
  volume={8},
  number={1},
  pages={e1002453},
  year={2012},
  publisher={Public Library of Science}
}

@article{skoglund2015,
  title={Genetic evidence for two founding populations of the Americas},
  author={Skoglund, Pontus and Mallick, Swapan and Bortolini, Maria C{\'a}tira and Chennagiri, Niru and H{\"u}nemeier, T{\'a}bita and Petzl-Erler, Maria Luiza and Salzano, Francisco Mauro and Patterson, Nick and Reich, David},
  journal={Nature},
  volume={525},
  number={7567},
  pages={104--108},
  year={2015},
  publisher={Nature Publishing Group}
}

%globetrotter
@article{hellenthal2014,
  title={A genetic atlas of human admixture history},
  author={Hellenthal, Garrett and Busby, George BJ and Band, Gavin and Wilson, James F and Capelli, Cristian and Falush, Daniel and Myers, Simon},
  journal={Science},
  volume={343},
  number={6172},
  pages={747--751},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{scott2016,
  title={Characterization of Greater Middle Eastern genetic variation for enhanced disease gene discovery},
  author={Scott, Eric M and Halees, Anason and Itan, Yuval and Spencer, Emily G and He, Yupeng and Azab, Mostafa Abdellateef and Gabriel, Stacey B and Belkadi, Aziz and Boisson, Bertrand and Abel, Laurent and others},
  journal={Nature Genetics},
  year={2016},
  publisher={Nature Publishing Group}
}

@article{pritchard2000,
  title={Inference of population structure using multilocus genotype data},
  author={Pritchard, Jonathan K and Stephens, Matthew and Donnelly, Peter},
  journal={Genetics},
  volume={155},
  number={2},
  pages={945--959},
  year={2000},
  publisher={Genetics Soc America}
}

@article{racimo2016,
  title={Signatures of archaic adaptive introgression in present-day human populations},
  author={Racimo, Fernando and Marnetto, Davide and Huerta-S{\'a}nchez, Emilia},
  journal={Molecular Biology and Evolution},
  pages={msw216},
  year={2016},
  publisher={SMBE}
}

@article{sams2016,
  title={Adaptively introgressed Neandertal haplotype at the OAS locus functionally impacts innate immune responses in humans},
  author={Sams, Aaron J and Dumaine, Anne and N{\'e}d{\'e}lec, Yohann and Yotova, Vania and Alfieri, Carolina and Tanner, Jerome E and Messer, Philipp W and Barreiro, Luis B},
  journal={Genome Biology},
  volume={17},
  number={1},
  pages={246},
  year={2016},
  publisher={BioMed Central}
}

@article{lam2010resequencing,
  title={Resequencing of 31 wild and cultivated soybean genomes identifies patterns of genetic diversity and selection},
  author={Lam, Hon-Ming and Xu, Xun and Liu, Xin and Chen, Wenbin and Yang, Guohua and Wong, Fuk-Ling and Li, Man-Wah and He, Weiming and Qin, Nan and Wang, Bo and others},
  journal={Nature genetics},
  volume={42},
  number={12},
  pages={1053--1059},
  year={2010},
  publisher={Nature Publishing Group}
}

@article{myles2011genetic,
  title={Genetic structure and domestication history of the grape},
  author={Myles, Sean and Boyko, Adam R and Owens, Christopher L and Brown, Patrick J and Grassi, Fabrizio and Aradhya, Mallikarjuna K and Prins, Bernard and Reynolds, Andy and Chia, Jer-Ming and Ware, Doreen and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={9},
  pages={3530--3535},
  year={2011},
  publisher={National Acad Sciences}
}

@article{diez2015olive,
  title={Olive domestication and diversification in the Mediterranean Basin},
  author={Diez, Concepcion M and Trujillo, Isabel and Martinez-Urdiroz, Nieves and Barranco, Diego and Rallo, Luis and Marfil, Pedro and Gaut, Brandon S},
  journal={New Phytologist},
  volume={206},
  number={1},
  pages={436--447},
  year={2015},
  publisher={Wiley Online Library}
}

@article{whitney2006adaptive,
  title={Adaptive introgression of herbivore resistance traits in the weedy sunflower Helianthus annuus},
  author={Whitney, Kenneth D and Randell, Rebecca A and Rieseberg, Loren H},
  journal={The American Naturalist},
  volume={167},
  number={6},
  pages={794--807},
  year={2006},
  publisher={The University of Chicago Press}
}

@article{whitney2010adaptive,
  title={Adaptive introgression of abiotic tolerance traits in the sunflower Helianthus annuus},
  author={Whitney, Kenneth D and Randell, Rebecca A and Rieseberg, Loren H},
  journal={New Phytologist},
  volume={187},
  number={1},
  pages={230--239},
  year={2010},
  publisher={Wiley Online Library}
}

@article{scascitelli2010genome,
  title={Genome scan of hybridizing sunflowers from Texas (Helianthus annuus and H. debilis) reveals asymmetric patterns of introgression and small islands of genomic differentiation},
  author={Scascitelli, M and Whitney, KD and Randell, RA and King, Matthew and Buerkle, CA and Rieseberg, LH},
  journal={Molecular Ecology},
  volume={19},
  number={3},
  pages={521--541},
  year={2010},
  publisher={Wiley Online Library}
}

@article{rieseberg2007hybridization,
  title={Hybridization and the colonization of novel habitats by annual sunflowers},
  author={Rieseberg, Loren H and Kim, Seung-Chul and Randell, Rebecca A and Whitney, Kenneth D and Gross, Briana L and Lexer, Christian and Clay, Keith},
  journal={Genetica},
  volume={129},
  number={2},
  pages={149--165},
  year={2007},
  publisher={Springer}
}

@article{steele2006marker,
  title={Marker-assisted selection to introgress rice QTLs controlling root traits into an Indian upland rice variety},
  author={Steele, KA and Price, AH and Shashidhar, HE and Witcombe, JR},
  journal={Theoretical and Applied Genetics},
  volume={112},
  number={2},
  pages={208--221},
  year={2006},
  publisher={Springer}
}

@article{heiser1951hybridization,
  title={Hybridization in the annual sunflowers: Helianthus annuus$\times$ H. debilis var. cucumerifolius},
  author={Heiser Jr, Charles B},
  journal={Evolution},
  pages={42--51},
  year={1951},
  publisher={JSTOR}
}

@article{dreiseitl2017heterogeneity,
  title={Heterogeneity of Powdery Mildew Resistance Revealed in Accessions of the ICARDA Wild Barley Collection},
  author={Dreiseitl, Antonin},
  journal={Frontiers in Plant Science},
  volume={8},
  year={2017},
  publisher={Frontiers Media SA}
}

@article{bredeson2016sequencing,
  title={Sequencing wild and cultivated cassava and related species reveals extensive interspecific hybridization and genetic diversity},
  author={Bredeson, Jessen V and Lyons, Jessica B and Prochnik, Simon E and Wu, G Albert and Ha, Cindy M and Edsinger-Gonzales, Eric and Grimwood, Jane and Schmutz, Jeremy and Rabbi, Ismail Y and Egesi, Chiedozie and others},
  journal={Nature biotechnology},
  volume={34},
  number={5},
  pages={562--570},
  year={2016},
  publisher={Nature Publishing Group}
}

@article{cornille2012new,
  title={New insight into the history of domesticated apple: secondary contribution of the European wild apple to the genome of cultivated varieties},
  author={Cornille, Amandine and Gladieux, Pierre and Smulders, Marinus JM and Rold{\'a}n-Ruiz, Isabel and Laurens, Fran{\c{c}}ois and Le Cam, Bruno and Nersesyan, Anush and Clavel, Joanne and Olonova, Marina and Feugey, Laurence and others},
  journal={PLoS Genet},
  volume={8},
  number={5},
  pages={e1002703},
  year={2012},
  publisher={Public Library of Science}
}

@article{purugganan2011archaeological,
  title={Archaeological data reveal slow rates of evolution during plant domestication},
  author={Purugganan, Michael D and Fuller, Dorian Q},
  journal={Evolution},
  volume={65},
  number={1},
  pages={171--183},
  year={2011},
  publisher={Wiley Online Library}
}

### 2017-07-05 paper.tex
edits to future studies

@@
 
+Basic:
+We're beginning to see the introgression has occurred across a number of crops during their expansion.
+What is the genomic architecture of this introgression and does the architecture suggest it's been adaptive?
+At what geographic scale is introgression adaptive?  Is it very local?  Do you see different architectures of introgression across different regions?
+At what taxonomic scale does introgression occur?  When do species become so diverged that introgression is either not possible or maladaptive?
+
 Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication histories of many agronomically-important crops.
@@
 To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
-Can colonizing species and/or hybrid swarms serve as bridges for gene flow between previously allopatric taxa?
 At what geographic scale does adaptive introgression occur?
@@
 To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?
-How did the conscious, subconscious, and unconscious decisions of early farmers facilitate or hinder adaptive introgression into their crops during early domestication?
-How do the practices of contemporary farmers affect the process of adaptive introgression today?
 
+Applied:
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.

### 2017-07-06 paper.tex
Update paper.tex

@@
 Barley (\emph{Hordeum vulgare} subsp.
-\emph{vulgare}) was domesticated at least twice roughly 10,000 BP: once from the wild subsp.
+\emph{vulgare}) was domesticated at least twice roughly 8,000 to 10,000 BP: once from the wild subsp.
 \emph{spontaneum} in the Fertile Crescent and once from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}.
-\gmj{the source morrell2007genetic will be used to update this paragraph once we've worked out for sure that the second domestication of barley was in the Zargos mountains, rather than Tibet.
-Also, this paper puts the earliest archaeological samples of barley at closer to 8500 calibrated years ago.}
-%However, details of barley domestication are still somewhat unclear.
-%Variety \emph{agriocrithon} is genetically diverse, and is found throughout much of the range of barley.
-%Some have suggested that \emph{agriocrithon} may be the progenitor of six-rowed barley, the product of a hybridization between eastern and western cultivated barley, or a wild-domesticate hybrid \cite{staudt1961origin, zohary1959hordeum, murphy1982origin}\, but \cite{azhaguvel2007phylogenetic}\ dispute these theories, suggesting instead that it arose from ssp.
-\emph{spontaneum}, perhaps more than once (in Israel as well as in Tibet).
+\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
+\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent.
+It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.
+I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.
+}
 Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
-%While little genetic investigation into spontaneous barley/\emph{spontaneum} hybrids \cite{ellstrand2003dangerous}.
 Barley/\emph{spontaneum} hybrids are known to be fertile and are found spontaneously when wild and domesticated barleys co-occur.
-In some cases, wild-to-crop introgression has been shown to occur over distances of more than a kilometer \cite{hillman2001new}.
+Introgression between wild and domesticated barleys appears to be frequent \cite{dai2012tibet}at times occuring over distances of more than a kilometer \cite{hillman2001new}.
 
 %The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}.
-\mbh{I think here a more focused discussion of potential for adaptive introgression based on the Poets paper is needed}
-The authors of \cite{Poets2015} used STRUCTURE to look for patterns of introgression from wild relatives in a dataset of 803 landraces, and found a high amount variability in the amount of contribution from wild relatives, as well as its location in the genome, within barley populations.
-This is indicative of contribution from numerous wild populations.
-Furthermore, the authors found that wild introgression contribution is generally greatest from geographically-proximate populations, and that introgressed regions might be combined from geographically-separate wild populations.
+
+The authors of \cite{Poets2015} investigated the contribution of wild barleys to domesticated barley landraces from either center of domestication, looking specifically for patterns of contribution across regions of the genome and across geographical ranges.
+Broadly, these authors found several lines of evidence that indicate that introgression from many wild populations of barley may have been important in the dispersal and adaptation of domesticated barleys.
+Barley landraces show greater genetic contributions from proximate wild relatives than from distant ones, suggesting that these contributions may be locally-adaptive.
+Furthermore, genomic segments with greater contribution from a wild population than the average are more likley to contain regionally-unique, locally-adaptive alleles, and unequal genetic contributions at specific genomic regions to individuals within landraces may suggest that adaptive alleles from diverse geographical areas may have been combined.
+\gmj{I'll be honest, I don't actually follow the reasoning in the second half of the above sentence.
+It is a point made in the Poets paper.
+I think my confusion arises from their specific "ancestry" term they define.
+I'll have to return to this at another point.}
 Low linkage disequilibrium and small blocks of identity by state indicate that these introgressed regions are old, perhaps dating back to the beginning of barley domestication.
-As landraces and nearby wild relatives share similar genomic sequences, the introgressed regions that are exclusive to that landrace are more likely to contain adaptive alleles. 
-Such alleles were not identified specifically, though wild-domesticate breeding experiments have shown that wild barleys have alleles for several important agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}.
+
+Genomic evidence points to a history of wild-crop introgression in the dispersal of barleys beyond their centers of domestication, but for the most part, the identity of the alleles critical to these adaptations have yet to be uncovered.
+One exception comes from the authors of \cite{dai2012tibet}, who hypothesize that alleles for cold stress tolerance and spring habit \gmj{"spring habit" are the words from the paper, but they're not defined.
+Does this mean that it grows more quickly in the spring?} from wild Tibetan 2-rowed barley allowed the barley to spread to the cold northern regions of Europe.
+Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.
 
@@
 
-file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
+%file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
 
@@
 Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
-On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups, \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-Others support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
-A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of indica and japonica \cite{molina2011molecular}.
-Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event.
+A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
+Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+%Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
+
+%The former first paragraph, replaced from above:
+%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
+%Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
+%On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
+%A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
+%Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
 %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
@@
 
-The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
-The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
+Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
+According to this hypothesis, the non-shattering allele arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some japonica-like, some indica-like) in the region.
+This would have facilitated the subsequent spread of other domestication alleles among those populations.
+As these domesticated populations then spread into new environments, they would have recieved introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
+Contemporary populations of wild and domesticated rice show evidence of this history of local adaptation.
+The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
+Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+
+%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
+%The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
+%The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
 %From Huang2012, "Based on the SNP data, the sequence diversity (p) of O.
@@
 sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
-Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
-Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+%Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+%Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
 
-To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples.
-\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgrogression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
+To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples (though \cite{zhao2010genomic} emphasizes the necessity of this research).
+%\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
 There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
@@
 This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-However, during early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
+Crop-wild ybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+During early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
 Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.

### 2017-07-06 paper.tex
Update paper.tex

@@
 Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
-Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$). .
+Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
 The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
@@
 However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
+\gmj{This is true in most cases I think, although I remember reading that an opposite pattern was found in sunflowers.  I'll go find that paper again and see if it warrants a mention here.}
 
@@
 
-Basic:
-We're beginning to see the introgression has occurred across a number of crops during their expansion.
-What is the genomic architecture of this introgression and does the architecture suggest it's been adaptive?
+\subsection*{Basic:}
+We're beginning to see that introgression has occurred across a number of crops during their expansion.
+What is the genomic architecture of this introgression and does the architecture suggest that it has been adaptive?
 At what geographic scale is introgression adaptive?  Is it very local?  Do you see different architectures of introgression across different regions?
@@
 
-Applied:
+\subsection*{Applied:}
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.

### 2017-07-06 paper.tex
edits to barley case study

--- 2017-07-06 paper.tex
@@
 
-In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how adaptive introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
@@
 The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
-These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
@@
 \emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
-\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent.
+%\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent.
 It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.
@@
 Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
-Barley/\emph{spontaneum} hybrids are known to be fertile and are found spontaneously when wild and domesticated barleys co-occur.
-Introgression between wild and domesticated barleys appears to be frequent \cite{dai2012tibet}at times occuring over distances of more than a kilometer \cite{hillman2001new}.
+Barley/\emph{spontaneum} hybrids are fertile and found spontaneously when wild and domesticated barley co-occur.
+Introgression between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.
 
@@
 
-The authors of \cite{Poets2015} investigated the contribution of wild barleys to domesticated barley landraces from either center of domestication, looking specifically for patterns of contribution across regions of the genome and across geographical ranges.
-Broadly, these authors found several lines of evidence that indicate that introgression from many wild populations of barley may have been important in the dispersal and adaptation of domesticated barleys.
-Barley landraces show greater genetic contributions from proximate wild relatives than from distant ones, suggesting that these contributions may be locally-adaptive.
-Furthermore, genomic segments with greater contribution from a wild population than the average are more likley to contain regionally-unique, locally-adaptive alleles, and unequal genetic contributions at specific genomic regions to individuals within landraces may suggest that adaptive alleles from diverse geographical areas may have been combined.
-\gmj{I'll be honest, I don't actually follow the reasoning in the second half of the above sentence.
-It is a point made in the Poets paper.
-I think my confusion arises from their specific "ancestry" term they define.
-I'll have to return to this at another point.}
-Low linkage disequilibrium and small blocks of identity by state indicate that these introgressed regions are old, perhaps dating back to the beginning of barley domestication.
+Poets and co-authors \cite{Poets2015} investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
+This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
+Genomic regions of shared ancestry were detected linking individual landraces to a number of wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
+However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
+Low linkage disequilibrium and small blocks of identity by state indicated even these locally introgressed regions are old, perhaps dating back to the early expansion of barley following domestication.
+While these results are suggestive, wild barley haplotypes have yet to be definitely linked to specific local adaptations in landraces.
 
-Genomic evidence points to a history of wild-crop introgression in the dispersal of barleys beyond their centers of domestication, but for the most part, the identity of the alleles critical to these adaptations have yet to be uncovered.
-One exception comes from the authors of \cite{dai2012tibet}, who hypothesize that alleles for cold stress tolerance and spring habit \gmj{"spring habit" are the words from the paper, but they're not defined.
+%Genomic evidence points to a history of wild-to-crop introgression during dispersal of barley beyond their centers of domestication, but for the most part, the identity of the alleles critical to these adaptations have yet to be uncovered.
+%One exception comes from the authors of \cite{dai2012tibet}, who hypothesize that alleles for cold stress tolerance and spring habit \gmj{"spring habit" are the words from the paper, but they're not defined.
 Does this mean that it grows more quickly in the spring?} from wild Tibetan 2-rowed barley allowed the barley to spread to the cold northern regions of Europe.
-Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.
+%Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.
 

### 2017-07-06 paper.tex
Update paper.tex

@@
 However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
-\gmj{This is true in most cases I think, although I remember reading that an opposite pattern was found in sunflowers.  I'll go find that paper again and see if it warrants a mention here.}
+\gmj{These expectations do not bear out in sunflower.
+The reintroduction of branching in the 70's is likely the cause of this.
+Here's the link to the paper: http://onlinelibrary.wiley.com/doi/10.1111/nph.13255/full}
 

### 2017-07-07 paper.tex
Update paper.tex
edits to rice case study

--- 2017-07-06 paper.tex
@@
 
-\item{Sunflower: }
+%\item{Sunflower: }
 
-The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
-The pre-Columbian \emph{H.
+%The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
+%The pre-Columbian \emph{H.
 annuus} distribution of cultivated sunflower spanned much of the Great Plains, from what is now north-central Texas to Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
-Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
+%Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
 petiolaris} and \emph{H.
 bolanderi} and forms stable hybrid populations with these taxa \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
-Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
-However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
-\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
-Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
-Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
-These introgressed hybrids formed a new lineage of sunflower (\emph{H.
+%Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
+%However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
+%\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
+%Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
+%Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
+%These introgressed hybrids formed a new lineage of sunflower (\emph{H.
 annuus} ssp.
@@
 debilis}-like traits adaptive to south-central Texas climate and ecology.
-These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
+%These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
 annuus} (as measured by higher seed production \cite{whitney2006adaptive}).
-Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
+%Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
 
-The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
-\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
-Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
+%The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
+%\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
+%Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
 argophyllus}, \emph{H.
@@
 annuus} \cite{miller1991inheritance}.
-PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
+%PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
 argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}.
-Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
-\emph{H.
+%Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
+%\emph{H.
 annuus} shows signs of persistent introgressive hybridization with \emph{H.
@@
 
-Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
-\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
-Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
-Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
+%Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
+%\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
+%Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
+%Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
 a.
 texanus} in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
-RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
-These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
+%RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
+%These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
 
-Notes for review:
+%Notes for review:
 
-http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
+%http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
 
-http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
+%http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
 
@@
 
-http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
+%http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
 
-https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
+%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
 
-Genome scan of hybridizing sunflowers from Texas
-(Helianthus annuus and H. debilis) reveals asymmetric
-patterns of introgression and small islands of genomic
-differentiation:
+%Genome scan of hybridizing sunflowers from Texas
+%(Helianthus annuus and H. debilis) reveals asymmetric
+%patterns of introgression and small islands of genomic
+%differentiation:
 
-In contrast, coalescent analyses of long-term migration
-rates imply that interspecific migration has been
-more important than mutation in providing new genetic
-variation in ssp. texanus populations. Immigration from
-H. debilis cucumerifolius into ssp. texanus (M = 4.05;
-Nem = 1.74) is higher than median estimates of intraspecific
-migration rates in plants (Morjan and Rieseberg 2004) 
-and only slightly lower than immigration from
-ssp. annuus into ssp. texanus (M = 5.00). Immigration
-rates from ssp. texanus into H. debilis cucumerifolius and
-ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
-implying that the putative hybrid lineage serves
-as a bridge for migration between the parental species.
+%In contrast, coalescent analyses of long-term migration
+%rates imply that interspecific migration has been
+%more important than mutation in providing new genetic
+%variation in ssp. texanus populations. Immigration from
+%H. debilis cucumerifolius into ssp. texanus (M = 4.05;
+%Nem = 1.74) is higher than median estimates of intraspecific
+%migration rates in plants (Morjan and Rieseberg 2004) 
+%and only slightly lower than immigration from
+%ssp. annuus into ssp. texanus (M = 5.00). Immigration
+%rates from ssp. texanus into H. debilis cucumerifolius and
+%ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
+%implying that the putative hybrid lineage serves
+%as a bridge for migration between the parental species.
 
-One puzzle is
-that the estimated immigration rate between the allopatric
-taxa (H. debilis cucumerifolius and ssp. annuus) is
-approximately comparable with that between the sympatric
-taxa (H. debilis cucumerifolius and ssp. texanus ssp.
-texanus). This observation may point to the effectiveness
-of ssp. texanus as a bridge for the transfer of alleles
-between H. debilis and ssp. annuus. An alternative explanation
-is that the BAYESASS estimates are unreliable
-because homoplasy of microsatellite alleles is high in
-taxa with large-effective population sizes (Estoup et al.
-2002) such as the annual sunflowers studied here.
+%One puzzle is
+%that the estimated immigration rate between the allopatric
+%taxa (H. debilis cucumerifolius and ssp. annuus) is
+%approximately comparable with that between the sympatric
+%taxa (H. debilis cucumerifolius and ssp. texanus ssp.
+%texanus). This observation may point to the effectiveness
+%of ssp. texanus as a bridge for the transfer of alleles
+%between H. debilis and ssp. annuus. An alternative explanation
+%is that the BAYESASS estimates are unreliable
+%because homoplasy of microsatellite alleles is high in
+%taxa with large-effective population sizes (Estoup et al.
+%2002) such as the annual sunflowers studied here.
 
-Despite the apparent discrepancy between the results
-from STRUCTURE and BAYEASS, several conclusions
-can be made. Analyses with STRUCTURE support the
-distinctness of ssp. texanus, as well as its clear inclusion
-within H. annuus, as initially postulated by Heiser (1951).
-Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
-analyses all confirm previous report of introgression
-between H. annuus and H. debilis (Rieseberg et al.
-1990, 2007) and suggest that the introgression is genomewide
-when it occurs. Contrary to the Heiser’s scenario of
-a recent Holocene origin of ssp. texanus, however, our
-results are more consistent with a much longer history of
-contact between H. annuus and H. debilis. Otherwise, it is
-difficult to account for the presence of significant longterm
-migration between currently allopatric populations
-of ssp. annuus and H. debilis cucumerifolius. This revised
-scenario makes sense in the light of numerous glacialinterglacial
-cycles over the past million years (Hewitt
-2000), which probably resulted in intermittent contact
-between H. annuus and H. debilis, with the last contact
-likely occurring during the Wisconsin glaciation,
-18 000 BP. The current contact between the species differs
-from the previous ones in that it appears to have been
-human-aided. Possibly, some of the molecular evidence
-of introgression, particularly into allopatric populations,
-stems from past periods of contact.
+%Despite the apparent discrepancy between the results
+%from STRUCTURE and BAYEASS, several conclusions
+%can be made. Analyses with STRUCTURE support the
+%distinctness of ssp. texanus, as well as its clear inclusion
+%within H. annuus, as initially postulated by Heiser (1951).
+%Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
+%analyses all confirm previous report of introgression
+%between H. annuus and H. debilis (Rieseberg et al.
+%1990, 2007) and suggest that the introgression is genomewide
+%when it occurs. Contrary to the Heiser’s scenario of
+%a recent Holocene origin of ssp. texanus, however, our
+%results are more consistent with a much longer history of
+%contact between H. annuus and H. debilis. Otherwise, it is
+%difficult to account for the presence of significant longterm
+%migration between currently allopatric populations
+%of ssp. annuus and H. debilis cucumerifolius. This revised
+%scenario makes sense in the light of numerous glacialinterglacial
+%cycles over the past million years (Hewitt
+%2000), which probably resulted in intermittent contact
+%between H. annuus and H. debilis, with the last contact
+%likely occurring during the Wisconsin glaciation,
+%18 000 BP. The current contact between the species differs
+%from the previous ones in that it appears to have been
+%human-aided. Possibly, some of the molecular evidence
+%of introgression, particularly into allopatric populations,
+%stems from past periods of contact.
 
-One of the four outlier loci, HT0414, is associated
-with a heat-shock protein and was previously shown to
-have been the target of selection in H. annuus ssp. annuus
-salt marsh populations from the state of Utah
-(Kane and Rieseberg 2007), as well as in weedy sunflower populations
-from several different locations across the USA
-(Kane and Rieseberg 2008). In this study, populations of
-ssp. annuus from Kansas, Nebraska and Oklahoma
-monomorphic for a single allele, whereas populations
-of H. debilis and other populations of ssp. annuus are
-considerably more polymorphic. These results imply
-that HT0414 may be involved in adaptation to a range
-of different habitats or to conditions shared by several
-different habitats (Kane and Rieseberg 2008).
+%One of the four outlier loci, HT0414, is associated
+%with a heat-shock protein and was previously shown to
+%have been the target of selection in H. annuus ssp. annuus
+%salt marsh populations from the state of Utah
+%(Kane and Rieseberg 2007), as well as in weedy sunflower populations
+%from several different locations across the USA
+%(Kane and Rieseberg 2008). In this study, populations of
+%ssp. annuus from Kansas, Nebraska and Oklahoma
+%monomorphic for a single allele, whereas populations
+%of H. debilis and other populations of ssp. annuus are
+%considerably more polymorphic. These results imply
+%that HT0414 may be involved in adaptation to a range
+%of different habitats or to conditions shared by several
+%different habitats (Kane and Rieseberg 2008).
 
@@
 
-The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
+The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
 On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event.
-A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
+Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from \emph{rufipogon}, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
 Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
@@
 In this approach, diversity data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
-Due to the difficulty of this approach and the uncertainty associated with any given demographic model, many studies identify domestication loci using a strict outlier approach in which loci showing the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
 Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
@@
 Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
-Failure to account for introgression in maize would therefore compromise candidate detection, particularly if a study contained maize samples from the Mexican highlands.
+Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands.
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.

### 2017-07-07 paper.tex

@@
 	\centering
-	\includegraphics[width=17.35cm]{par_figure_stars.png}
+	\includegraphics[width=17.35cm]{vertical.png}
 	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}
@@
 	\centering
-	\includegraphics[width=17.35cm]{boxplot.png}
+	\includegraphics[width=17.35cm]{boxplot_3.png}
 	\caption{The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}

### 2017-07-11 paper.tex
edit the table
Update paper.tex

@@
 Should we include these two methods, even in passing, in this portion of the paper?}
+\lwang{Garrett, I donot know these two methods. I am reading the paper you mentioned here.}
 
@@
 
-\rowcolor{white}
-{\emph{\bf population structure related}} &   &   &   \\
-\rowcolor{white}
-fineStructure &	phased haplotype &	Lawson et al. 2012	& Skoglund et al. 2015 \\
-\rowcolor{white}
-Globetrotter &	phased haplotype &	Hellenthal et al. 2014 & Scott et al. 2016 \\
 \end{tabular}
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=17.35cm]{vertical.png}
-	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}
-	\label{boxplot:map}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=17.35cm]{vertical.png}
+%	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}
+%	\label{boxplot:map}
+%\end{figure}
 
@@
 	\centering
-	\includegraphics[width=17.35cm]{boxplot_3.png}
-	\caption{The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
+	\includegraphics[width=17.35cm]{combination.png}
+	\caption{Map of the natural ranges of wild relatives of three domesticated crops, overlayed with average annual temperature.
+The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
 	\label{fig:map}

### 2017-07-12 paper.tex
Update paper.tex

@@
 
+\gmj{Might need to change the sunflower entry in this table.
+The best evidence I can now point to for adaptive introgression is the reintroduction of the branching trait, but that example is not natural.}
+
 \begin{table}
@@
 Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
-Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X & X & X & \cite{rieseberg2007hybridization}\\ 
+Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\ 
 \hline
@@
 However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
-\gmj{These expectations do not bear out in sunflower.
+%\gmj{These expectations do not bear out in sunflower.
 The reintroduction of branching in the 70's is likely the cause of this.
@@
 \subsection*{Basic:}
-We're beginning to see that introgression has occurred across a number of crops during their expansion.
-What is the genomic architecture of this introgression and does the architecture suggest that it has been adaptive?
-At what geographic scale is introgression adaptive?  Is it very local?  Do you see different architectures of introgression across different regions?
-At what taxonomic scale does introgression occur?  When do species become so diverged that introgression is either not possible or maladaptive?
+%We're beginning to see that introgression has occurred across a number of crops during their expansion.
+%What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
+%At what geographic scale is introgression adaptive?  Is it very local?  Does one see different architectures of introgression across different regions?
+%At what taxonomic scale does introgression occur?
+%When do species become so diverged that introgression is either maladaptive or impossible?
 
-Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication histories of many agronomically-important crops.
-However, the dynamics of the process in these cases are not yet fully understood.
+%Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication histories of many agronomically-important crops.
+%However, the dynamics of the process in particular cases are not yet fully understood.
+%To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
+\gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
+%At what geographic scale does adaptive introgression occur? Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
+%To what extent does this depend on the slope of environmental gradients, such as temperature, precipitation, and elevation, across these ranges?
+
+%\gmj{I am not seeing the reason behind splitting this up in into two paragraphs.  I'll try reformatting the same information below:}
+
+Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication and dispersal of many agronomically-important crops.
+However, the dynamics of this process are not yet fully understood in the context of individual cases, and many questions remain.
+At what taxonomic scale does introgression occur?
+What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
+When do species become so diverged that introgression is either maladaptive or impossible?
 To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
-At what geographic scale does adaptive introgression occur?
-Is introgression frequently restricted to very local populations, or is it often seen over broad geographic ranges?
-To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?
+\gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
+At what geographic scale does adaptive introgression occur? Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
+At what geographic scale is introgression adaptive?  Is it very local?  Does one see different architectures of introgression across different regions?
+To what extent does this depend on the slope of environmental gradients, such as temperature, precipitation, and elevation, across these ranges?
 
 \subsection*{Applied:}
+Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources (annotated genomic sequence assemblies and functional genomic data sets \cite{huang2012}).
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
@@
 Furthermore, adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
+\gmj{I could here include a handfull of examples of crops that farmers have bred adaptive alleles into.}
 

### 2017-07-14 paper.tex
Update paper.tex

@@
 
-%\gmj{I am not seeing the reason behind splitting this up in into two paragraphs.  I'll try reformatting the same information below:}
+%\gmj{I am not seeing the reason behind splitting this up in into these two paragraphs.  I'll try reformatting the same information below:}
 
 Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication and dispersal of many agronomically-important crops.
-However, the dynamics of this process are not yet fully understood in the context of individual cases, and many questions remain.
+However, the dynamics of this process are not yet fully understood, especially in the context of individual case examples, and many questions remain.
+
+What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
+After the hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
+
 At what taxonomic scale does introgression occur?
-What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-When do species become so diverged that introgression is either maladaptive or impossible?
+When do species become so diverged that introgression is either maladaptive or impossible (due to  Dobzhansky-Muller incompatiblities or other mechanisms)?
 To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
 \gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
-At what geographic scale does adaptive introgression occur? Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
-At what geographic scale is introgression adaptive?  Is it very local?  Does one see different architectures of introgression across different regions?
-To what extent does this depend on the slope of environmental gradients, such as temperature, precipitation, and elevation, across these ranges?
+
+Does one see different architectures of introgression across different regions?
+Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
+Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
+At what geographic scale is introgression adaptive?
+To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
 
 \subsection*{Applied:}
-Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources (annotated genomic sequence assemblies and functional genomic data sets \cite{huang2012}).
+Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources, such as annotated genomic sequence assemblies and functional genomic data sets \cite{huang2012}.
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
@@
 These loci are potentially beneficial targets for crop improvement.
-Furthermore, adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
+Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
 \gmj{I could here include a handfull of examples of crops that farmers have bred adaptive alleles into.}

### 2017-07-14 paper.tex
Update paper.tex

@@
 What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
+\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps, or by functional identification.}
+%\gmj{Kinds of functional identification?  Knockout experiments, sequence homology with annotated genes, proximity to QTLs?}
 After the hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
+\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
+\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
 
@@
 \gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
+\gmj{I could also talk about cross-incompatibility factors in maize/teosinte hybridization vs.
+highly compatible asian rice and its relatives.
+I probably wouldn't need a new source for the rice, but I might need to cite a paper about tcb1, ga1, and ga2.}
 

### 2017-07-17 paper.tex
Update paper.tex

@@
 
-\gmj{Might need to change the sunflower entry in this table.
-The best evidence I can now point to for adaptive introgression is the reintroduction of the branching trait, but that example is not natural.}
-
 \begin{table}
@@
 What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps, or by functional identification.}
+\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification.}
 %\gmj{Kinds of functional identification?  Knockout experiments, sequence homology with annotated genes, proximity to QTLs?}
@@
 At what geographic scale is introgression adaptive?
+\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
 To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
+\gmj{Refer to the map and bar graph here?}
 

### 2017-07-18 paper.tex
Update paper.tex

@@
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
+\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
 %\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent.
 It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.
-I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.
-}
+I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.}
 Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
@@
 %Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.
-
-%\item{Sunflower: }
-
-%The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
-%The pre-Columbian \emph{H.
-annuus} distribution of cultivated sunflower spanned much of the Great Plains, from what is now north-central Texas to Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
-%Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
-petiolaris} and \emph{H.
-bolanderi} and forms stable hybrid populations with these taxa \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
-%Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
-%However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
-%\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
-%Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
-%Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
-%These introgressed hybrids formed a new lineage of sunflower (\emph{H.
-annuus} ssp.
-\emph{texanus}, \emph{H.
-a.
-texanus} hereafter) which displays \emph{H.
-debilis}-like traits adaptive to south-central Texas climate and ecology.
-%These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
-annuus} (as measured by higher seed production \cite{whitney2006adaptive}).
-%Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
-
-%The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
-%\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
-%Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
-argophyllus}, \emph{H.
-tuberosus}, and \emph{H.
-praecox}) and have been successfully bred into modern \emph{H.
-annuus} \cite{miller1991inheritance}.
-%PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
-argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}.
-%Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
-%\emph{H.
-annuus} shows signs of persistent introgressive hybridization with \emph{H.
-petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}.
-
-%Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
-%\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
-%Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
-%Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
-a.
-texanus} in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
-%RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
-%These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
-
-%Notes for review:
-
-%http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
-
-%http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
-
-%file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
-
-%http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
-
-%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
-
-%Genome scan of hybridizing sunflowers from Texas
-%(Helianthus annuus and H. debilis) reveals asymmetric
-%patterns of introgression and small islands of genomic
-%differentiation:
-
-%In contrast, coalescent analyses of long-term migration
-%rates imply that interspecific migration has been
-%more important than mutation in providing new genetic
-%variation in ssp. texanus populations. Immigration from
-%H. debilis cucumerifolius into ssp. texanus (M = 4.05;
-%Nem = 1.74) is higher than median estimates of intraspecific
-%migration rates in plants (Morjan and Rieseberg 2004) 
-%and only slightly lower than immigration from
-%ssp. annuus into ssp. texanus (M = 5.00). Immigration
-%rates from ssp. texanus into H. debilis cucumerifolius and
-%ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
-%implying that the putative hybrid lineage serves
-%as a bridge for migration between the parental species.
-
-%One puzzle is
-%that the estimated immigration rate between the allopatric
-%taxa (H. debilis cucumerifolius and ssp. annuus) is
-%approximately comparable with that between the sympatric
-%taxa (H. debilis cucumerifolius and ssp. texanus ssp.
-%texanus). This observation may point to the effectiveness
-%of ssp. texanus as a bridge for the transfer of alleles
-%between H. debilis and ssp. annuus. An alternative explanation
-%is that the BAYESASS estimates are unreliable
-%because homoplasy of microsatellite alleles is high in
-%taxa with large-effective population sizes (Estoup et al.
-%2002) such as the annual sunflowers studied here.
-
-%Despite the apparent discrepancy between the results
-%from STRUCTURE and BAYEASS, several conclusions
-%can be made. Analyses with STRUCTURE support the
-%distinctness of ssp. texanus, as well as its clear inclusion
-%within H. annuus, as initially postulated by Heiser (1951).
-%Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
-%analyses all confirm previous report of introgression
-%between H. annuus and H. debilis (Rieseberg et al.
-%1990, 2007) and suggest that the introgression is genomewide
-%when it occurs. Contrary to the Heiser’s scenario of
-%a recent Holocene origin of ssp. texanus, however, our
-%results are more consistent with a much longer history of
-%contact between H. annuus and H. debilis. Otherwise, it is
-%difficult to account for the presence of significant longterm
-%migration between currently allopatric populations
-%of ssp. annuus and H. debilis cucumerifolius. This revised
-%scenario makes sense in the light of numerous glacialinterglacial
-%cycles over the past million years (Hewitt
-%2000), which probably resulted in intermittent contact
-%between H. annuus and H. debilis, with the last contact
-%likely occurring during the Wisconsin glaciation,
-%18 000 BP. The current contact between the species differs
-%from the previous ones in that it appears to have been
-%human-aided. Possibly, some of the molecular evidence
-%of introgression, particularly into allopatric populations,
-%stems from past periods of contact.
-
-%One of the four outlier loci, HT0414, is associated
-%with a heat-shock protein and was previously shown to
-%have been the target of selection in H. annuus ssp. annuus
-%salt marsh populations from the state of Utah
-%(Kane and Rieseberg 2007), as well as in weedy sunflower populations
-%from several different locations across the USA
-%(Kane and Rieseberg 2008). In this study, populations of
-%ssp. annuus from Kansas, Nebraska and Oklahoma
-%monomorphic for a single allele, whereas populations
-%of H. debilis and other populations of ssp. annuus are
-%considerably more polymorphic. These results imply
-%that HT0414 may be involved in adaptation to a range
-%of different habitats or to conditions shared by several
-%different habitats (Kane and Rieseberg 2008).
-
-%The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
-%This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}.
-
-%whitney2006adaptive
-%"these results suggest that introgression of biotic resistance traits was important in the adaptation of H. annuus to central and southern Texas."
-%fitness (seed production) high in texanus than in annuus
-%identified two pests to which resistance was imparted and important
-%Discussion points: 1. texanus has higher fitness than annuus in central texas.
-%2. texanus shared traits more in common with debilis than with annuus (herbivore resistance)
-%3. 2 of the three traits from point two above are important for adaptation to life in central south Texas.
-%The authors ask why these pest-resistance genes have not spread further north beyond this range.  They are uncertain, but conjecture that negative pleiotrophic effects are at play.
-
-%whitney2010adaptive
-%"We demonstrate that introgression has altered multiple aspects of the H.
-annuus phenotype in an adaptive manner, has affected traits relevant to both biotic and abiotic environments, and may have aided expansion of the H.
-annuus range into central Texas, USA."
-%This paper really does show that this case of natural introgression was adaptive.
-%The companion paper (2006 Whitney) will likely show the same thing, but I haven't read it yet.
-
-%scascitelli2010genome
-%"long-term migration rates were high, genome-wide and asymmetric, with higher migration rates from H. annuus texanus into the two parental taxa than vice versa."
-%"H. annuus texanus may serve as a bridge for the transfer of alleles between its parental taxa."
-%"contradict recent theory suggesting that introgression should predominantly be in the direction of the colonizing species."
-
-%rieseberg2007hybridization
-%this is a very good review.  i think i should use it for the table.
-%they also show that traits were imparted from debilis to texanus, and that these are adaptive.  i think they also show genetic evidence, but i'm not sure.
 
@@
 
+%\item{Sunflower: }
+
+%The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
+%The pre-Columbian \emph{H.
+annuus} distribution of cultivated sunflower spanned much of the Great Plains, from what is now north-central Texas to Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
+%Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
+petiolaris} and \emph{H.
+bolanderi} and forms stable hybrid populations with these taxa \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
+%Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
+%However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
+%\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
+%Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
+%Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
+%These introgressed hybrids formed a new lineage of sunflower (\emph{H.
+annuus} ssp.
+\emph{texanus}, \emph{H.
+a.
+texanus} hereafter) which displays \emph{H.
+debilis}-like traits adaptive to south-central Texas climate and ecology.
+%These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
+annuus} (as measured by higher seed production \cite{whitney2006adaptive}).
+%Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
+
+%The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
+%\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
+%Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
+argophyllus}, \emph{H.
+tuberosus}, and \emph{H.
+praecox}) and have been successfully bred into modern \emph{H.
+annuus} \cite{miller1991inheritance}.
+%PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
+argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}.
+%Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
+%\emph{H.
+annuus} shows signs of persistent introgressive hybridization with \emph{H.
+petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}.
+
+%Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
+%\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
+%Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
+%Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
+a.
+texanus} in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
+%RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
+%These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
+
+%Notes for review:
+
+%http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
+
+%http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
+
+%file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
+
+%http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
+
+%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
+
+%Genome scan of hybridizing sunflowers from Texas
+%(Helianthus annuus and H. debilis) reveals asymmetric
+%patterns of introgression and small islands of genomic
+%differentiation:
+
+%In contrast, coalescent analyses of long-term migration
+%rates imply that interspecific migration has been
+%more important than mutation in providing new genetic
+%variation in ssp. texanus populations. Immigration from
+%H. debilis cucumerifolius into ssp. texanus (M = 4.05;
+%Nem = 1.74) is higher than median estimates of intraspecific
+%migration rates in plants (Morjan and Rieseberg 2004) 
+%and only slightly lower than immigration from
+%ssp. annuus into ssp. texanus (M = 5.00). Immigration
+%rates from ssp. texanus into H. debilis cucumerifolius and
+%ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
+%implying that the putative hybrid lineage serves
+%as a bridge for migration between the parental species.
+
+%One puzzle is
+%that the estimated immigration rate between the allopatric
+%taxa (H. debilis cucumerifolius and ssp. annuus) is
+%approximately comparable with that between the sympatric
+%taxa (H. debilis cucumerifolius and ssp. texanus ssp.
+%texanus). This observation may point to the effectiveness
+%of ssp. texanus as a bridge for the transfer of alleles
+%between H. debilis and ssp. annuus. An alternative explanation
+%is that the BAYESASS estimates are unreliable
+%because homoplasy of microsatellite alleles is high in
+%taxa with large-effective population sizes (Estoup et al.
+%2002) such as the annual sunflowers studied here.
+
+%Despite the apparent discrepancy between the results
+%from STRUCTURE and BAYEASS, several conclusions
+%can be made. Analyses with STRUCTURE support the
+%distinctness of ssp. texanus, as well as its clear inclusion
+%within H. annuus, as initially postulated by Heiser (1951).
+%Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
+%analyses all confirm previous report of introgression
+%between H. annuus and H. debilis (Rieseberg et al.
+%1990, 2007) and suggest that the introgression is genomewide
+%when it occurs. Contrary to the Heiser’s scenario of
+%a recent Holocene origin of ssp. texanus, however, our
+%results are more consistent with a much longer history of
+%contact between H. annuus and H. debilis. Otherwise, it is
+%difficult to account for the presence of significant longterm
+%migration between currently allopatric populations
+%of ssp. annuus and H. debilis cucumerifolius. This revised
+%scenario makes sense in the light of numerous glacialinterglacial
+%cycles over the past million years (Hewitt
+%2000), which probably resulted in intermittent contact
+%between H. annuus and H. debilis, with the last contact
+%likely occurring during the Wisconsin glaciation,
+%18 000 BP. The current contact between the species differs
+%from the previous ones in that it appears to have been
+%human-aided. Possibly, some of the molecular evidence
+%of introgression, particularly into allopatric populations,
+%stems from past periods of contact.
+
+%One of the four outlier loci, HT0414, is associated
+%with a heat-shock protein and was previously shown to
+%have been the target of selection in H. annuus ssp. annuus
+%salt marsh populations from the state of Utah
+%(Kane and Rieseberg 2007), as well as in weedy sunflower populations
+%from several different locations across the USA
+%(Kane and Rieseberg 2008). In this study, populations of
+%ssp. annuus from Kansas, Nebraska and Oklahoma
+%monomorphic for a single allele, whereas populations
+%of H. debilis and other populations of ssp. annuus are
+%considerably more polymorphic. These results imply
+%that HT0414 may be involved in adaptation to a range
+%of different habitats or to conditions shared by several
+%different habitats (Kane and Rieseberg 2008).
+
+%The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
+%This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}.
+
+%whitney2006adaptive
+%"these results suggest that introgression of biotic resistance traits was important in the adaptation of H. annuus to central and southern Texas."
+%fitness (seed production) high in texanus than in annuus
+%identified two pests to which resistance was imparted and important
+%Discussion points: 1. texanus has higher fitness than annuus in central texas.
+%2. texanus shared traits more in common with debilis than with annuus (herbivore resistance)
+%3. 2 of the three traits from point two above are important for adaptation to life in central south Texas.
+%The authors ask why these pest-resistance genes have not spread further north beyond this range.  They are uncertain, but conjecture that negative pleiotrophic effects are at play.
+
+%whitney2010adaptive
+%"We demonstrate that introgression has altered multiple aspects of the H.
+annuus phenotype in an adaptive manner, has affected traits relevant to both biotic and abiotic environments, and may have aided expansion of the H.
+annuus range into central Texas, USA."
+%This paper really does show that this case of natural introgression was adaptive.
+%The companion paper (2006 Whitney) will likely show the same thing, but I haven't read it yet.
+
+%scascitelli2010genome
+%"long-term migration rates were high, genome-wide and asymmetric, with higher migration rates from H. annuus texanus into the two parental taxa than vice versa."
+%"H. annuus texanus may serve as a bridge for the transfer of alleles between its parental taxa."
+%"contradict recent theory suggesting that introgression should predominantly be in the direction of the colonizing species."
+
+%rieseberg2007hybridization
+%this is a very good review.  i think i should use it for the table.
+%they also show that traits were imparted from debilis to texanus, and that these are adaptive.  i think they also show genetic evidence, but i'm not sure.
+
 \bibliography{bib_gj}

### 2017-07-20 paper.tex
Update paper.tex

@@
 
-\subsection*{Basic:}
-%We're beginning to see that introgression has occurred across a number of crops during their expansion.
+\subsection*{Basic:}%We're beginning to see that introgression has occurred across a number of crops during their expansion.
 %What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
@@
 What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification.}
-%\gmj{Kinds of functional identification?  Knockout experiments, sequence homology with annotated genes, proximity to QTLs?}
-After the hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
-\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
-\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
+%Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
+At what geographic scale is introgression adaptive?
+%\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
+To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
+%\gmj{Refer to the map and bar graph here?}
+We can look at conservation of genomic architecture across landscapes and between populations, and make predictions about introgressions and their relations to local adaptation.
+If the genomic architecture of an introgressed region is conserved across a broad ecogeographical region, this suggests that the introgression imparts adaptation to general environmental or climatic variables.
+On the other hand, if the genomic architecture is conserved within populations but not between nearby populations in the region, this suggests that the introgressed regions offer adaptations to more local selective pressures.
+If the genomic architecture of an introgressed region is not conserved within a population, there is little evidence that the introgression is adaptive.
+%\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification.}
+
+After hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
+%\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
+%\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
+%Does one see different architectures of introgression across different regions?
+%Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
+Introgressed regions are easier to detect when there has been limited recombination to break them apart.
+Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
+Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).
 
@@
 When do species become so diverged that introgression is either maladaptive or impossible (due to  Dobzhansky-Muller incompatiblities or other mechanisms)?
-To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
-\gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
+Theory indicates that the most significant limiting factor to gene flow between progenitor and domesticate is divergence time.
+Over time, diverged populations drift and become increasingly incompatible.
+Small effective population size, and correspondingly high genetic load, of the introgressive population also limits gene flow.
+Although perhaps less applicable in crop systems, this effect is seen in other well-documented cases of introgression (for example, Neanderthal introgression into humans, \cite{harris2016genetic}).
+\gmj{I tracked down the sources that indicate that introgression is suppressed around genes.
+The case example was Neanderthal alleles into humans, the paper titled "The genetic history of Ice Age Europe" \cite{fu2016genetic}, cited by the Graham Coop paper.
+However, I'm unclear if this is a general rule, or simply a consequence of Neanderthal being high in deleterious genetic load.
+Wouldn't crop systems, with wild relatives harboring putatively adaptive alleles with low load, show a different pattern?}
+%To what extent does the level of introgression across taxa depend on divergence time and mutation load between donor and recipient taxa?
 \gmj{I could also talk about cross-incompatibility factors in maize/teosinte hybridization vs.
@@
 I probably wouldn't need a new source for the rice, but I might need to cite a paper about tcb1, ga1, and ga2.}
-
-Does one see different architectures of introgression across different regions?
-Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
-Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
-At what geographic scale is introgression adaptive?
-\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
-To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
-\gmj{Refer to the map and bar graph here?}
 

### 2017-07-21 paper.tex
edits throughout

--- 2017-07-18 paper.tex
@@
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
-An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
+An additional important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
 
@@
 
-Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and their wild relatives outside putative domestication centers.
-Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).
+Genome-wide data from extensive samples of crops and their wild relatives, in combination with the new methods described above, have recently allowed detailed analysis of wild-to-crop introgression in some of the world's most important crops (\textbf{Table 2}).
+Below we present a summary of findings from maize, barley, and rice, three promising systems in which introgression from wild relatives may have played an adaptive role.
 
@@
 \emph{mays}) and the teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
+\emph{mexicana} (hereafter, \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
 Maize was domesticated from \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter referred to as \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
-Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002single, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
+Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
 However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
@@
 
-Populations of \emph{mexicana} cannot be found outside of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
+Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
 %Several questions regarding \emph{mexicana} introgression yet remain.
@@
 %Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+Historically, \emph{mexicana} haplotypes appear to have played an important role in adaptation of maize to challenging high-elevation conditions.
+Breeding programs for highland maize may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
 

### 2017-07-28 paper.tex
minor edits

@@
 To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
-The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
+The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
@@
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
 These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
@@
 %More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
-Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
 %The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
@@
 %Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
-Historically, \emph{mexicana} haplotypes appear to have played an important role in adaptation of maize to challenging high-elevation conditions.
-Breeding programs for highland maize may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
+Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
 
@@
 Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
-Barley/\emph{spontaneum} hybrids are fertile and found spontaneously when wild and domesticated barley co-occur.
+Barley-\emph{spontaneum} hybrids are fertile and found spontaneously when these taxa co-occur.
 Introgression between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.
@@
 
-Poets and co-authors \cite{Poets2015} investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
+Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
 This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
-Genomic regions of shared ancestry were detected linking individual landraces to a number of wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
+Genomic regions of shared ancestry were detected linking particular landraces to several distinct wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
 However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
@@
 The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
-On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from \emph{rufipogon}, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
 Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.

### 2017-08-10 paper.tex
Update paper.tex

@@
 Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015}); introgression has conferred M\"{u}llerian mimicry \mbh{I would explain a bit more here...wing coloration loci, protects against predation...} across butterfly species \cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
+For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015});
+introgression of wing coloration loci has conferred M\"{u}llerian mimicry
+(a type of mimicry in which two unpalatable species resemble one another, compounding the efficacy of their repellant signal)
+%\mbh{I would explain a bit more here...wing coloration loci, protects against predation...}
+across butterfly species \cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
 

### 2018-02-07 paper.tex
minor edits, some notes (not finished yet)

@@
 Genome-wide data from extensive samples of crops and their wild relatives, in combination with the new methods described above, have recently allowed detailed analysis of wild-to-crop introgression in some of the world's most important crops (\textbf{Table 2}).
-Below we present a summary of findings from maize, barley, and rice, three promising systems in which introgression from wild relatives may have played an adaptive role.
+Below we present a summary of findings from maize, barley, and rice, three promising systems in which introgression from wild relatives appear to have played an adaptive role.
 
@@
 Maize was domesticated from \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
@@
 I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.}
-Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
+Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
 Barley-\emph{spontaneum} hybrids are fertile and found spontaneously when these taxa co-occur.
@@
 Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
+Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
 Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
@@
 This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-Crop-wild ybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
 During early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
@@
 \hline
-Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing, gavrilenko2013genetic} \\
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome, johns1986ongoing, gavrilenko2013genetic} \\
 \hline
@@
 \subsection*{What is the progenitor of a crop?}
-Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
-The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
+Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
+Introgression between a crop and newly-encountered populations and species decreases the level of divergence.
+This signal could be mistaken for origin rather than gene flow.
+%The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
 For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
@@
 Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
+\gmj{This sentence  needs to be updated with the new potato publication.}
+
 Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
@@
 For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
+\gmj{Does Li's research fit in well here? Should we add her citation?}
 
@@
 Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
+\gmj{or when the introgressed region is nearer the centromere, where recombination is less common?}
 Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).
@@
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
-Loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signal of introgression.
+Loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
 These loci are potentially beneficial targets for crop improvement.
 Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
-\gmj{I could here include a handfull of examples of crops that farmers have bred adaptive alleles into.}
+\gmj{I could here include a handfull of examples of crops that farmers have bred adaptive alleles into.
+Also, we could mention that identification of wild relatives responsible for valuable agronomic introgressed alleles could inform conservation efforts.}
 

### 2018-02-07 paper.tex
Update paper.tex

@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
+\gmj{Are there newer resequencing methods to discuss here?
+DART-Seq, etc?}
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
@@
 
-Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
+\gmj{This paragraph about application seems like it might not belong in this section.
+Maybe it should be a part of the introduction?
+Or, maybe each of these examples could be mentioned directly following the method(s) that was used in the study?}Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
 For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015});
@@
 \item{Maize:}
+
+\gmj{Li, can you update this section with your recent publication's findings?}
 
@@
 Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
-\gmj{or when the introgressed region is nearer the centromere, where recombination is less common?}
 Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).

### 2018-02-07 paper.tex
Update paper.tex

@@
 %Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.
+
+\item{Potato}
+
+\gmj{I had saved these notes from an earlier version of the paper.
+There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes.
+This new paper fills that gap.}
+
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}\.
+The northern members of the polyphyletic \emph{S.
+brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
+tuberosum} has a polyphyletic origin.
+Also, these northern members of the \emph{S.
+breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
+bukasovii}, \cite{spooner2005single}\).
+
+Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}\.
+Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}\.
+These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}\.
+However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}\.
+Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}\.
+
+Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}\.
+The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}\.
+Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}\.
+Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
+sparsipilum} \cite{rabinowitz1990high}\ and \emph{S.
+megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
+Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}\.
+Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
+tuberosum subsp.
+andigena}, \cite{van1999tight}, \emph{S.
+bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
 

### 2018-02-08 paper.tex
Update paper.tex

--- 2018-02-07 paper.tex
@@
 %Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.
+
+\item{Potato}
+
+\gmj{I had saved these notes from an earlier version of the paper.
+There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes.
+This new paper fills that gap.}
+
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
+The northern members of the polyphyletic \emph{S.
+brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
+tuberosum} has a polyphyletic origin.
+Also, these northern members of the \emph{S.
+breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
+bukasovii}, \cite{spooner2005single}).
+
+Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}.
+Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
+These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
+However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
+Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
+
+Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
+The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
+Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
+Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
+Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
+Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
+tuberosum subsp.
+andigena}, \cite{van1999tight}, \emph{S.
+bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
 

### 2018-02-08 paper.tex
Update paper.tex

@@
 Maybe it should be a part of the introduction?
-Or, maybe each of these examples could be mentioned directly following the method(s) that was used in the study?}Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
+Or, maybe each of these examples could be mentioned directly following the method(s) that was used in the study?}
+Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
 For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015});
@@
 There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes.
-This new paper fills that gap.}
+This new paper fills that gap, if we decide that the paper's merits are substantial enough that we can overlook its flaws.}
+\gmj{If we decide to keep and use the new potato introgression publication, that would be the natural choice for the potato reference, whether or not we include the write-up of the potato case study.}
 
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
 The northern members of the polyphyletic \emph{S.
@@
 However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
-%\gmj{These expectations do not bear out in sunflower.
-The reintroduction of branching in the 70's is likely the cause of this.
-Here's the link to the paper: http://onlinelibrary.wiley.com/doi/10.1111/nph.13255/full}
 
@@
 At what geographic scale is introgression adaptive?
-%\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
+\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
 To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
-%\gmj{Refer to the map and bar graph here?}
+\gmj{Refer to the map and bar graph here?}
 We can look at conservation of genomic architecture across landscapes and between populations, and make predictions about introgressions and their relations to local adaptation.
@@
 If the genomic architecture of an introgressed region is not conserved within a population, there is little evidence that the introgression is adaptive.
-%\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification.}
+\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification?}
 
@@
 
+%Omitted by Garrett
 %\item{Sunflower: }

### 2018-02-08 paper.tex
comments

--- 2018-02-08 paper.tex
@@
 Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
+\lwang{I think these paragraphs read pretty smooth. I pondered at the last sentence of this paragraph. 
+Suggestions: Preliminary results suggest a need to expand our conception of genetic bases of domestication in crops that inhabited new territory through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.}
 
@@
 
-\item{Potato}
-
-\gmj{I had saved these notes from an earlier version of the paper.
-There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes.
-This new paper fills that gap.}
-
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
-The northern members of the polyphyletic \emph{S.
-brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
-tuberosum} has a polyphyletic origin.
-Also, these northern members of the \emph{S.
-breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
-bukasovii}, \cite{spooner2005single}).
-
-Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}.
-Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
-These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
-However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
-Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
-
-Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
-The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
-Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
-Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
-Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
-Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
-tuberosum subsp.
-andigena}, \cite{van1999tight}, \emph{S.
-bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
-
 \end{enumerate}
@@
 \begin{table}
+
 \rowcolors{2}{white}{gray!25}
@@
 \caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabular}{llll}
+\begin{tabular}{0.9\textwidth}{llll}
 \\\toprule  
@@
 This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
+\lwang{maybe some citations for a protracted domestication time: African rice NG paper, our paper, grape PNAS paper}
 More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.
+\lwang{excluding the introgression regions might cause some inference method not working properly, such as MSMC, as it would reduce heterozygosity.
+Suggestions: "More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression and / or developing new statistic tools which taking potential introgression into account."}
 
 \subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
+\lwang{I suggest to delete "impacted by a domestication bottleneck" because that might confuse readers.
+Adaptive introgression could affect genome-wide diversity, owing to that introgressing haplotype have a higher historical Ne and domesticates have a lower historical Ne due to bottleneck.
+Adding "impacted by a domestication bottleneck" seems making the message not direct and not complete.}
 Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
@@
 Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
+\lwang{involve --> involving}
 Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).

### 2018-02-08 paper.tex
edits to intro.

@@
 In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
-To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
+To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
 The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
-\lwang{I think these paragraphs read pretty smooth. I pondered at the last sentence of this paragraph. 
-Suggestions: Preliminary results suggest a need to expand our conception of genetic bases of domestication in crops that inhabited new territory through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.}
+Preliminary results suggest there is a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred through adaptive gene flow with newly encountered wild relatives during post-domestication expansion.
 

### 2018-02-08 paper.tex
comments

--- 2018-02-08 paper.tex
@@
 \usepackage{booktabs}
+\usepackage{tabularx}
 \usepackage[x11names,dvipsnames,table]{xcolor}
@@
 DART-Seq, etc?}
+\lwang{I think it is fine to mention two of these methods, it is not our focus.}
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
@@
 
-\gmj{Li, can you update this section with your recent publication's findings?}
-
 The relationship between maize (\emph{Zea mays} ssp.
@@
 Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
-However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
+However, Hufford et al. \cite{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 %The following is not fully clear.  Cut or revise.
-The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
-These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX(Figure \ref{fig:introgressionMaize}).
+These introgressed segments overlapped with QTLs that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
@@
 %More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
-Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects.
-%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
-%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
-%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
-%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
+Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to Meso-American high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+
 Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
@@
 \caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabular}{0.9\textwidth}{llll}
+\begin{tabularx}{.9\textwidth}{llll}
 \\\toprule  
@@
 \rowcolor{gray!25}
-Hapmix	& phased haplotype; reference panel		& Price et al. 2009	&  Hufford et al. 2013;  Suarez et al. 2016 \\ 
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
 \rowcolor{gray!25}
-RASPberry &	phased haplotype &	Wegmann et al. 2011	 & Christe et al. 2016 \\
+RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
 \rowcolor{gray!25}
-MultiMix & phased/unphased genotype; reference panel &	Churchhouse and Marchini 2013 &	Eyheramendy et al. 2015 \\
+MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
 \rowcolor{gray!25}
-PCAdmix	 & phased haplotype	 & Brisbin et al. 2012	 & Moreno et al. 2014; Pugach et al. 2016 \\
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
 \rowcolor{gray!25}
-LAMP  &	phased haplotypes; reference panel	 & Sankararaman et al. 2008	 & Patterson et al. 2012 \\
+LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
 
@@
 \rowcolor{white}
-ABBA-BABA/D-statistics	 & biallelic SNP  &	Durand et al. 2011	 &  Heliconius Genome Consortium 2012 \\
+ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
 \rowcolor{white}
-fd statistic &	biallelic SNP &	Martin et al. 2015  &	Malinsky et al. 2015; Zhang et al. 2016 \\ 
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
 \rowcolor{white}
-five taxon D statistics	& biallelic SNP	&  Pease and Hahn 2015	& Fontaine et al. 2015; Pease et al. 2016 \\
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
 
@@
 \rowcolor{gray!25}
-Gmin &	biallelic SNP	&  Geneva et al. 2015	&  Kingan et al. 2015 \\
+Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
 \rowcolor{gray!25}
-RNDmin	& phased haplotype	& Rosenzweig et al. 2016	&  NA \\
+RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  NA \\
 \rowcolor{gray!25}
-(see .tex file for comment) & biallelic SNP & Racimo et al. 2016 & Sams et al. 2016 \\
+$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
 
-%I was not able to add Li's entry to this cell without errors.  I'll put her entry in the line below:
-%U_{A,B,C(w,x,y)} and Q95_{A,B,C(w,y)}
-
-\end{tabular}
+\end{tabularx}
 \end{center}

### 2018-02-08 paper.tex
edits to questions

--- 2018-02-08 paper.tex
@@
 \caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabular}{0.9\textwidth}{llll}
+\begin{tabular}{0.8\textwidth}{}
 \\\toprule  
@@
 
-A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression during the geographic expansion of crops is extensive.
-The addition of ongoing gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
+A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression is extensive during crop geographic expansion.
+The addition of ongoing gene flow to our understanding of crop demography could therefore bear importantly on fundamental questions of crop domestication:
 
@@
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
-Introgression between a crop and newly-encountered populations and species decreases the level of divergence.
+Introgression between a crop and newly-encountered taxa decreases divergence from these donors.
 This signal could be mistaken for origin rather than gene flow.
-%The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
 For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
@@
 Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
-Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
-\gmj{This sentence  needs to be updated with the new potato publication.}
+Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013, Hardigan2017}.
 
@@
 Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
-In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
-In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and the introgressing species.
-This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
-\lwang{maybe some citations for a protracted domestication time: African rice NG paper, our paper, grape PNAS paper}
-More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.
-\lwang{excluding the introgression regions might cause some inference method not working properly, such as MSMC, as it would reduce heterozygosity.
-Suggestions: "More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression and / or developing new statistic tools which taking potential introgression into account."}
+In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
+In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxon.
+This phenomenon, in combination with divergence of crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
+More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.
 
-\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
-\lwang{I suggest to delete "impacted by a domestication bottleneck" because that might confuse readers.
-Adaptive introgression could affect genome-wide diversity, owing to that introgressing haplotype have a higher historical Ne and domesticates have a lower historical Ne due to bottleneck.
-Adding "impacted by a domestication bottleneck" seems making the message not direct and not complete.}
+\subsection*{How was genome-wide diversity shaped by domestication?}
+
 Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
@@
 For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
-\gmj{Does Li's research fit in well here? Should we add her citation?}
 
@@
 In this approach, diversity data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
-Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
 Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
@@
 
-\bibliography{bib_gj}
+\bibliography{bib_gj.bib}
 
 \end{document}
-
-\end{document}

### 2018-02-08 paper.tex
fixing merge conflict and table issue

--- 2018-02-08 paper.tex
@@
 \caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabularx}{.9\textwidth}{llll}
+\begin{tabular}{0.8\textwidth}{}
 \\\toprule  
@@
 
-\end{tabularx}
+\end{tabular}
 \end{center}

### 2018-02-08 paper.tex
update rice section of paper.tex

@@
 
-The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
+The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
+Certain genetic and archaeobotanical evidence points toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with possibly another domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
 Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of variety \emph{aromatic}.
+A third analysis \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which became the progenitors to the other Asian rice varieties.
  %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
@@
 
-Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
-According to this hypothesis, the non-shattering allele arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some japonica-like, some indica-like) in the region.
+The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues in 2008 \cite{vaughan2008evolving}.
+%Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
+According to this hypothesis, the non-shattering allele (\emph{sh1}) arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
 This would have facilitated the subsequent spread of other domestication alleles among those populations.
-As these domesticated populations then spread into new environments, they would have recieved introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
-Contemporary populations of wild and domesticated rice show evidence of this history of local adaptation.
-The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
-Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+As these domesticated populations then spread into new environments, they would have received introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
+Both Vaughan and Choi support the role of introgression between domesticated Asian rice and locally-adapted wild \emph{rufipogon} populations in the development and dispersal of Asian rice into new environments.
 
-%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
-%The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
-%The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
-%From Huang2012, "Based on the SNP data, the sequence diversity (p) of O.
-rufipogon was estimated at 0.003, which is higher than that in O.
-sativa—the sequence diversity is 0.0024 for O.
-sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
-%Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
-%Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
-
-To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples (though \cite{zhao2010genomic} emphasizes the necessity of this research).
-%\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
-There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
-As with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
+Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
+Contemporary populations of wild rices show evidence of this history of local adaptation.
+Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
+Domesticated varieties have likewise adapted to local environments.
+The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
+Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
 This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
-During early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
-Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
-
-%Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
-
-%note from Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice:
-%Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.
+Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
+Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
 

### 2018-02-09 paper.tex
Update paper.tex

@@
 
-The study of crop domestication has been revolutionized by the advent and application of genomic tools.
-The genomes of crops and their wild relatives tell a story of give-and-take that extends well beyond the initial stages of domestication.
-Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adapted alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
+\section*{Conclusions}
+
+In the same way that Darwin explored natural selection by studying domesticated species (going so far as to feature domestic pigeons as his "special group" at the beginning of his 
+\emph{On the Origin of Species} \cite{darwin2009origin}), we frequently find that modern crop systems, in tandem with their extant wild relatives, offer unparalleled opportunities for the study of several evolutionary questions, including the role of adaptive introgression in dispersal and range expansion.
+The advent and application of genomics tools has lead to mounting evidence which underscores the prevalence of genetic admixture between crops and wild relatives during and well beyond the initial stages of domestication.
+Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adpated alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
+Within this new conception, the process of crop range expansion is a concurrent element of domestication, rather than a distinct event subsequent to it.
+%This is a revolutionary departure from the simpler model of a temporally-constrained domestication event with subsequent dispersal, and as such, offers new questions and new opportunities for investigation.
+This is a revolutionary departure from the prevailing model, and as such, offers new questions and new opportunities for investigation.
+These questions take on practical importance as well as theoretical.
+%Understanding the histories of domestication and adaptation inform and empower efforts to develop and improve crop systems.
+Understanding a crop's history of domestication in terms of selection, adaptation, and admixture empowers efforts at crop development and improvement.
+As the historic role of wild relatives in the development of many crops is increasingly clarified, it is likely that we may reevaluate the importance of conservation of wild populations (in particular or in general) that harbor agronomically-valuable alleles that may be targeted by pre-breeding efforts.
+These efforts may be of particular value in the face of future climate volatility and change.
+
+As a species, domesticated crops are among our first and greatest inventions.
+Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agronomic Revolution.
+In this more nuanced and dynamic understanding of crop domestication, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.
+\gmj{I let these final few sentences get a little more poetic than I normally would.  I think maybe I should reel them back in, or cut them.}
+
+%The study of crop domestication has been revolutionized by the advent and application of genomic tools.
+%The genomes of crops and their wild relatives tell a story of give-and-take that extends well beyond the initial stages of domestication.
+%Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adapted alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
 

### 2018-02-09 paper.tex
Update paper.tex

@@
 	\centering
-	\includegraphics[width=17.35cm]{combination.png}
-	\caption{Map of the natural ranges of wild relatives of three domesticated crops, overlayed with average annual temperature.
+	\includegraphics[width=17.35cm]{temperature_plot.png}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.
 The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
+	\label{fig:map}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=17.35cm]{boxplot.png}
+	\caption{The distribution of average annual temperatures experienced in the geographic home ranges of wild relatives interfertile with four crops}
 	\label{fig:map}
@@
 
-As a species, domesticated crops are among our first and greatest inventions.
+Domesticated crops are among humanity's first and greatest inventions.
+%As a species, domesticated crops are among our first and greatest inventions.
 Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agronomic Revolution.

### 2018-02-09 paper.tex
more edits added

--- 2018-02-08 paper.tex
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
-\gmj{Are there newer resequencing methods to discuss here?
-DART-Seq, etc?}
-\lwang{I think it is fine to mention two of these methods, it is not our focus.}
+%\gmj{Are there newer resequencing methods to discuss here?  DART-Seq, etc?}
+%\lwang{I think it is fine to mention two of these methods, it is not our focus.}
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
@@
 The former two statistics are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
-Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup:
+Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup:  $RND_{min} = d_{min}/d_{out}$, where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
 
@@
 
- \begin{equation}
- 	RND_{min} = \frac{d_{min}}{d_{out}}
- \end{equation}
-where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
+ %\begin{equation}
+
+ %\end{equation}
 
@@
 
-\gmj{Li, are you familiar with the analytical tools MIGRATE-N and BAYESASS?  Rieseberg comapres these two to STRUCTURE at some length in the discussion of this paper:
-https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
-Should we include these two methods, even in passing, in this portion of the paper?}
-\lwang{Garrett, I donot know these two methods. I am reading the paper you mentioned here.}
+%\gmj{Li, are you familiar with the analytical tools MIGRATE-N and BAYESASS?  Rieseberg comapres these two to STRUCTURE at some length in the discussion of this paper:
+%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
+%Should we include these two methods, even in passing, in this portion of the paper?}
+%\lwang{These are more complex statistics to infer the overall demography, like Isolate with migration, which we have defined in the beginning that we will not touch it in the manuscript}
 
@@
 
-\gmj{This paragraph about application seems like it might not belong in this section.
+%\gmj{This paragraph about application seems like it might not belong in this section.
 Maybe it should be a part of the introduction?
 Or, maybe each of these examples could be mentioned directly following the method(s) that was used in the study?}
-Application of these approaches across a number of plant and animal species suggests introgression can play an adaptive role.
+Application of these approaches combined with population genetic statistics to detect selection across a number of plant and animal species suggests introgression can play an adaptive role.
 For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015});
@@
 (a type of mimicry in which two unpalatable species resemble one another, compounding the efficacy of their repellant signal)
-%\mbh{I would explain a bit more here...wing coloration loci, protects against predation...}
 across butterfly species \cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
@@
 	\centering
-	\includegraphics[width=17.35cm]{li_figure.png}
-	\caption{Li's caption here.}
+	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1}
+	\caption{Hapmix vectors indicating introgression level from \emph{mexicana} to Mexican highland maize on chromosome 4.
+Data adopted from Hufford et al.
+\cite{hufford2013genomic}.
+ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.}
 	\label{fig:introgressionMaize}
@@
 
-The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-Certain genetic and archaeobotanical evidence points toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with possibly another domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
-Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
+The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
+Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
 Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of variety \emph{aromatic}.
-A third analysis \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which became the progenitors to the other Asian rice varieties.
 %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
@@
 
-The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues in 2008 \cite{vaughan2008evolving}.
-%Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
-According to this hypothesis, the non-shattering allele (\emph{sh1}) arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
+Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
+According to this hypothesis, the non-shattering allele arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some japonica-like, some indica-like) in the region.
 This would have facilitated the subsequent spread of other domestication alleles among those populations.
-As these domesticated populations then spread into new environments, they would have received introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
-Both Vaughan and Choi support the role of introgression between domesticated Asian rice and locally-adapted wild \emph{rufipogon} populations in the development and dispersal of Asian rice into new environments.
+As these domesticated populations then spread into new environments, they would have recieved introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
+Contemporary populations of wild and domesticated rice show evidence of this history of local adaptation.
+The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
+Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
 
-Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
-Contemporary populations of wild rices show evidence of this history of local adaptation.
-Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
-Domesticated varieties have likewise adapted to local environments.
-The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
+%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
+%The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
+%The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
+%From Huang2012, "Based on the SNP data, the sequence diversity (p) of O.
+rufipogon was estimated at 0.003, which is higher than that in O.
+sativa—the sequence diversity is 0.0024 for O.
+sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
+%Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
+%Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+
+To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples (though \cite{zhao2010genomic} emphasizes the necessity of this research).
+%\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
+There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
+As with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
+This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
 Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
-However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
-This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
-Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
+During early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
+Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
+
+%Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
+
+%note from Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice:
+%Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.
 
@@
 \caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabular}{0.8\textwidth}{}
+\begin{tabularx}{\textwidth}{llll}
 \\\toprule  
 \rowcolor{white}
-{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule
+{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf examples } \\ \midrule
 
@@
 
-\end{tabular}
+\end{tabularx}
 \end{center}
@@
 
-A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression is extensive during crop geographic expansion.
-The addition of ongoing gene flow to our understanding of crop demography could therefore bear importantly on fundamental questions of crop domestication:
+A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression during the geographic expansion of crops is extensive.
+The addition of ongoing gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:
 
@@
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
-Introgression between a crop and newly-encountered taxa decreases divergence from these donors.
+Introgression between a crop and newly-encountered populations and species decreases the level of divergence.
 This signal could be mistaken for origin rather than gene flow.
+%The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
 For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
@@
 Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
-Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013, Hardigan2017}.
+Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
+\gmj{This sentence  needs to be updated with the new potato publication.}
 
@@
 Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
-In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
-In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxon.
-This phenomenon, in combination with divergence of crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
-More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.
+In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
+In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and the introgressing species.
+This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
+\lwang{maybe some citations for a protracted domestication time: African rice NG paper, our paper, grape PNAS paper}
+More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.
+\lwang{excluding the introgression regions might cause some inference method not working properly, such as MSMC, as it would reduce heterozygosity.
+Suggestions: "More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression and / or developing new statistic tools which taking potential introgression into account."}
 
-\subsection*{How was genome-wide diversity shaped by domestication?}
-
+\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
+\lwang{I suggest to delete "impacted by a domestication bottleneck" because that might confuse readers.
+Adaptive introgression could affect genome-wide diversity, owing to that introgressing haplotype have a higher historical Ne and domesticates have a lower historical Ne due to bottleneck.
+Adding "impacted by a domestication bottleneck" seems making the message not direct and not complete.}
 Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
@@
 In this approach, diversity data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
-Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
 Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
@@
 
-\bibliography{bib_gj.bib}
+\bibliography{bib_gj}
 
 \end{document}
+
+\end{document}

### 2018-02-09 FigsAndTables/figure1.R

@@ -12,12 +12,17 @@ chr4.x2z <- chr4 %>% select(1:13)
 names(chr4.x2z)
 names(chr4.x2z)[5:13] <- gsub("_x2z", "", names(chr4.x2z)[5:13])
 names(chr4.x2z)
-chr4.long <- gather(chr4.x2z, pop, introgression, ELP:XOC)
+chr4.long <- gather(chr4.x2z, populations, introgression, ELP:XOC)
 head(chr4.long)
-ggplot(data=chr4.long, aes(x=gen, y=introgression, color=pop, group=pop)) + geom_line() + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
-ggsave("z2mIntro.pdf")
+#ggplot(data=chr4.long, aes(x=gen, y=introgression, color=populations, group=populations)) + geom_line() + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
+#ggsave("z2mIntro.pdf")
 
 
-ggplot(data=chr4.long, aes(x=gen, y=introgression, color=pop, group=pop)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
-ggsave("z2mIntro.loessRegression.pdf")
+#ggplot(data=chr4.long, aes(x=pos/1000000, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("physical position (mb)") +
+#ylab("Hapmix vectors")
+#ggsave("Figure1.pdf")
+
+ggplot(data=chr4.long, aes(x=gen, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position (cM)") +
+ylab("introgression level from mexicana to maize")
+ggsave("Figure1.pdf")
 

### 2018-02-09 paper.tex
typeset error fixed

--- 2018-02-09 paper.tex
@@
 
-\gmj{I had saved these notes from an earlier version of the paper.
-There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes.
-This new paper fills that gap, if we decide that the paper's merits are substantial enough that we can overlook its flaws.}
-\gmj{If we decide to keep and use the new potato introgression publication, that would be the natural choice for the potato reference, whether or not we include the write-up of the potato case study.}
-
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
-The northern members of the polyphyletic \emph{S.
-brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s.
-tuberosum} has a polyphyletic origin.
-Also, these northern members of the \emph{S.
-breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S.
-bukasovii}, \cite{spooner2005single}).
+Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000-10000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
+A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel (which included wild, landrace, and cultivated species) formed three groups (North American, South American, and cultivated) and identified \emph{S.
+candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.
 
@@
 
-Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
+Nevertheless, introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
 The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
-Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
-Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
-Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
-Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
+%Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
+sparsipilum} \cite{rabinowitz1990high} and \emph{S.
+megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
+%Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
 tuberosum subsp.
@@
 bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
+%Hardigan and colleagues \cite{hardigan2017genome} find evidence of extensive adaptive introgression into cultivated potatoes.
+In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
+candolleanum}) into Andean tetraploid cultivated landraces and North American cultivars (20.2\% and 32.1\%, respectively).
+Genes located within these introgressed regions were more likely to be highly-expressed, stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting that these introgressed regions impart stress response mechanisms critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion.
+Drought- and frost-resistance genes introgressed from wild relatives are also reported to have expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
+Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
 

### 2018-02-09 paper.tex
more edits

@@
 %\lwang{I think it is fine to mention two of these methods, it is not our focus.}
-High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{Martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
+High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
@@
 The former two statistics are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
-Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup:  $RND_{min} = d_{min}/d_{out}$, where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
-
-% \begin{equation}
-%    G_{min} = \frac{d_{min}}{d_{XY}}
-% \end{equation}
-% where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y.
-
- %\begin{equation}
-
- %\end{equation}
-
+Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
+%  $RND_{min} = d_{min}/d_{out}$, where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
 These statistics have recently been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
-For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($C$).
+For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.

### 2018-02-10 paper.tex
edits throughout
edits to introgression methods, Table 1

@@
 
-\mbh{In this section, I think the overall content is good, but we need to edit to make it more accessible and more explicit about how methods are implemented to detect adaptive introgression}
-
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
-%\gmj{Are there newer resequencing methods to discuss here?  DART-Seq, etc?}
-%\lwang{I think it is fine to mention two of these methods, it is not our focus.}
 High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
@@
 
-%\gmj{Li, are you familiar with the analytical tools MIGRATE-N and BAYESASS?  Rieseberg comapres these two to STRUCTURE at some length in the discussion of this paper:
-%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
-%Should we include these two methods, even in passing, in this portion of the paper?}
-%\lwang{These are more complex statistics to infer the overall demography, like Isolate with migration, which we have defined in the beginning that we will not touch it in the manuscript}
-
-Third, the ABBA-BABA statistic (also known as the D-statistic) and its derivatives are widely applied to introgression detection.
+Third, the ABBA-BABA statistic (also known as the D-statistic) and related statistics are widely applied to introgression detection.
 These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
@@
 
-%\gmj{This paragraph about application seems like it might not belong in this section.
-Maybe it should be a part of the introduction?
-Or, maybe each of these examples could be mentioned directly following the method(s) that was used in the study?}
 Application of these approaches combined with population genetic statistics to detect selection across a number of plant and animal species suggests introgression can play an adaptive role.
@@
 Genome-wide data from extensive samples of crops and their wild relatives, in combination with the new methods described above, have recently allowed detailed analysis of wild-to-crop introgression in some of the world's most important crops (\textbf{Table 2}).
-Below we present a summary of findings from maize, barley, and rice, three promising systems in which introgression from wild relatives appear to have played an adaptive role.
+Below we present a summary of findings from maize, barley, rice, and potato, four promising systems in which introgression from wild relatives appears to have played an adaptive role.
 
@@
 \rowcolor{gray!25}
-{\emph{\bf chromosome paiting}} &   &   &   \\
+{\emph{\bf chromosome painting}} &   &   &   \\
 \rowcolor{gray!25}
@@
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
-Introgression between a crop and newly-encountered taxa decreases divergence from these donors.
+Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
 This signal could be mistaken for origin rather than gene flow.

### 2018-02-10 paper.tex
edits to maize and barley

@@
 However, Hufford et al. \cite{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
-%The following is not fully clear.  Cut or revise.
 The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX(Figure \ref{fig:introgressionMaize}).
@@
 Height differences were not detected under lowland conditions.
-%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
-%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.
-
-%Even though introgression from \emph{mexicana} to Mexican highland maize has been highly supported, it is yet unknown whether and to what extent such introgression could be found in other highland maize populations which are allopatric to \emph{mexicana}.
-%A recent study \cite{Takuno2015} found little empirical or theoretical support for parallel highland adaptation in the Mexican and South American highland regions, which was explained partly by the difference in potential for adaptive introgression from wild relatives.
-%Adaptive SNPs in Mexican highland population were more likely located in the introgressed regions than those in South American highland populations.
-%Furthermore, the adaptive SNPs in the Mexican highland population were more likely also showing signatures of local adaptation in \emph{mexicana} and \emph {pulviglumis} populations than those South American highland population.
-%For these reasons, adaptive introgression from wild relatives may play a significant role in patterning the genetic differences of maize highland populations.
 
 Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
-%Several questions regarding \emph{mexicana} introgression yet remain.
-%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
-%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
 A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
-%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al.
-(2013) \citep{hufford2013}.
-%The signal of introgression is absent for the other three populations.
-%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
-%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
 Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to Meso-American high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
@@
 
-%Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
 Barley (\emph{Hordeum vulgare} subsp.
@@
 \emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
-%\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent.
-It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.
-I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.}
-Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
-Barley-\emph{spontaneum} hybrids are fertile and found spontaneously when these taxa co-occur.
-Introgression between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.
-
-%The barley domestication process has reduced the number of alleles in the domesticate to only 40\% of that found in wild barley, though there remains a great deal of phenotypic diversity among the wild barleys \cite{ellis2000wild}.
+Presently, the distribution of subsp.
+\emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
+Barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
+Gene flow between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.
 
@@
 This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
-Genomic regions of shared ancestry were detected linking particular landraces to several distinct wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
+Genomic regions of shared ancestry were detected linking particular landraces to numerous wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
 However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
-Low linkage disequilibrium and small blocks of identity by state indicated even these locally introgressed regions are old, perhaps dating back to the early expansion of barley following domestication.
+Limited linkage disequilibrium and small tracts of identity by state indicated even locally introgressed chromosomal regions are old, perhaps dating to the early expansion of barley  post-domestication.
 While these results are suggestive, wild barley haplotypes have yet to be definitely linked to specific local adaptations in landraces.

### 2018-02-10 FigsAndTables/figure1.R
edits

@@ -17,12 +17,24 @@ head(chr4.long)
 #ggplot(data=chr4.long, aes(x=gen, y=introgression, color=populations, group=populations)) + geom_line() + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position")
 #ggsave("z2mIntro.pdf")
 
+seg <- data.frame(x1 = 162.885280, x2 = 192.761788, y1 = 2, y2 = 2)
 
-#ggplot(data=chr4.long, aes(x=pos/1000000, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("physical position (mb)") +
-#ylab("Hapmix vectors")
-#ggsave("Figure1.pdf")
 
-ggplot(data=chr4.long, aes(x=gen, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position (cM)") +
-ylab("introgression level from mexicana to maize")
+p <- ggplot(data=chr4.long, aes(x=pos/1000000, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.01) + theme(legend.position=c(0.9, 0.8)) + xlab("physical position (mb)") +
+ylab("introgression level from mexicana to maize")+
+scale_y_continuous(limits=c(0, 2))
+p + geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = seg, size=2)+
+scale_y_continous(limits=c(0, 2))
+
 ggsave("Figure1.pdf")
 
+seg <- data.frame(x1 = 325.8, x2 = 397.4, y1 = 0, y2 = 0.05)
+
+ggplot(data=chr4.long, aes(x=gen, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.04) + theme(legend.position=c(0.9, 0.8)) + xlab("genetic position (cM)") +
+ylab("introgression level from mexicana to maize") +
+geom_rect(xmin=325.8, xmax=397.4, ymin=0, ymax=0.05, color="black") 
+
+ggsave("Figure1.pdf")
+
+325.8	397.4
+geom_rect(xmin=325.8, xmax=397.4, ymin=0, ymax=0.05), color="black") 

### 2018-02-10 paper.tex

@@
 
-%Genomic evidence points to a history of wild-to-crop introgression during dispersal of barley beyond their centers of domestication, but for the most part, the identity of the alleles critical to these adaptations have yet to be uncovered.
-%One exception comes from the authors of \cite{dai2012tibet}, who hypothesize that alleles for cold stress tolerance and spring habit \gmj{"spring habit" are the words from the paper, but they're not defined.
-Does this mean that it grows more quickly in the spring?} from wild Tibetan 2-rowed barley allowed the barley to spread to the cold northern regions of Europe.
-%Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.
-
 \item{Asian Rice:}
 
-The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-Certain genetic and archaeobotanical evidence points toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with possibly another domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
 Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
-Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of variety \emph{aromatic}.
-A third analysis \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which became the progenitors to the other Asian rice varieties.
- %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
+Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of \emph{aromatic} rice.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
 
-%The former first paragraph, replaced from above:
-%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-%Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
-%On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
-%A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
-%Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-%Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
+The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
+According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
+As these fully domesticated populations spread into new environments, they potentially received introgression from locally-adapted wild relatives, retaining alleles that improved fitness.
 
-%The former first paragraph, replaced from above:
-%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
-%A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
-%\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
-%On the other hand, genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
-%An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-%Adaptive introgression may in fact be the signal interpreted as independent domestications.
-
-The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues in 2008 \cite{vaughan2008evolving}.
-%Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
-According to this hypothesis, the non-shattering allele (\emph{sh1}) arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
-This would have facilitated the subsequent spread of other domestication alleles among those populations.
-As these domesticated populations then spread into new environments, they would have received introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
-Both Vaughan and Choi support the role of introgression between domesticated Asian rice and locally-adapted wild \emph{rufipogon} populations in the development and dispersal of Asian rice into new environments.
-
-Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
-Contemporary populations of wild rices show evidence of this history of local adaptation.
-Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
-Domesticated varieties have likewise adapted to local environments.
-The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
-Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
-However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
-This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
-Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
+%Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
+%Contemporary populations of wild rices show evidence of this history of local adaptation.
+%Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
+%Domesticated varieties have likewise adapted to local environments.
+%The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
+%Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
+%However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
+%This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
+%Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
+%Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
 

### 2018-02-10 paper.tex
edits

@@
 Patterns of allele sharing are interpreted in a phylogenetic context and the method is best suited to detection of introgression at the genome level.
-Elaborations of the D-statistic capable of localizing introgression to specific genomic regions include $\hat{f_{d}}$ \cite{Martin2015} and the five-taxon D-statistic \cite{pease2015}.
+Elaborations of the D-statistic capable of localizing introgression to specific genomic regions include $\hat{f_{d}}$ \cite{martin2015} and the five-taxon D-statistic \cite{pease2015}.
 The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern and is capable of determining introgression directionality.
@@
 Application of these approaches combined with population genetic statistics to detect selection across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, introgression from ancient hominins (\emph{e.g.}, Neanderthals and Denisovans) to humans has been detected at loci controlling skin pigmentation, defense against pathogens, and toleration of high altitude (reviewed in \cite{Racimo2015});
-introgression of wing coloration loci has conferred M\"{u}llerian mimicry
+For example, based on the sequence divergence, introgression has been detected across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species resulting in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015};
+Jeong et al. \cite{jeong2014} utilized HAPMIX to disclose introgression from Sherpa to Tibetan humans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015});
+ABBA-BABA statistics were employed to reveal introgression of wing coloration loci conferring M\"{u}llerian mimicry
 (a type of mimicry in which two unpalatable species resemble one another, compounding the efficacy of their repellant signal)
-across butterfly species \cite{Heliconius2012}; introgression has spread insecticide resistance across mosquito species \cite{Norris2015}, and introgression across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species has resulted in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015}.
+across butterfly species \cite{heliconius2012}.
+.
 
@@
 Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
-However, Hufford et al. \cite{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
+However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX(Figure \ref{fig:introgressionMaize}).
@@
 	\centering
-	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1}
+	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1withInset}
 	\caption{Hapmix vectors indicating introgression level from \emph{mexicana} to Mexican highland maize on chromosome 4.
 Data adopted from Hufford et al.
-\cite{hufford2013genomic}.
-ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.}
+\cite{Hufford2013}.
+ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
+Insets showed the phenotypes of macrohairs on lowland (red) and highland (blue) maize stems.}
 	\label{fig:introgressionMaize}
@@
 \rowcolor{gray!25}
-{\emph{\bf chromosome painting}} &   &   &   \\
-\rowcolor{gray!25}
-Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
-\rowcolor{gray!25}
-RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
-\rowcolor{gray!25}
-MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
-\rowcolor{gray!25}
-PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
-\rowcolor{gray!25}
-LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
-
-\rowcolor{white}
-{\emph{\bf phylogenetic relationship}} &   &   &   \\
-\rowcolor{white}
-ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
-\rowcolor{white}
-fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
-\rowcolor{white}
-five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
-
-\rowcolor{gray!25}
 {\emph{\bf divergence}} &   &   &   \\
@@
 $U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
+
+\rowcolor{white}
+{\emph{\bf chromosome painting}} &   &   &   \\
+\rowcolor{white}
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
+\rowcolor{white}
+RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
+\rowcolor{white}
+MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
+\rowcolor{white}
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
+\rowcolor{white}
+LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
+
+\rowcolor{gray!25}
+{\emph{\bf phylogenetic relationship}} &   &   &   \\
+\rowcolor{gray!25}
+ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
+\rowcolor{gray!25}
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
+\rowcolor{gray!25}
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
 
@@
 Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
-Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{hufford2013genomic} \\ 
+Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\ 
 \hline 
@@
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
-However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{hufford2013genomic}, likely due to ongoing selection favoring the domesticated phenotype.
+However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
 

### 2018-02-10 paper.tex
edits

@@
 ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
-Insets showed the phenotypes of macrohairs on lowland (red) and highland (blue) maize stems.}
+Insets showed the phenotypes of macrohairs on lowland (red) and highland (blue) maize stems.
+The black bar indicated one of the QTLs for macrohairs and pigment density in Lauter et al.
+\cite{lauter2004}.}
 	\label{fig:introgressionMaize}
@@
 \rowcolor{white}
-{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf examples } \\ \midrule
+{\bf methods}	& {\bf data type } &	{\bf references} &  {\bf applications } \\ \midrule
 
@@
 \rowcolor{gray!25}
-RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  NA \\
+RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
 \rowcolor{gray!25}

### 2018-02-10 FigsAndTables/figure1.R

@@ -20,9 +20,11 @@ head(chr4.long)
 seg <- data.frame(x1 = 162.885280, x2 = 192.761788, y1 = 2, y2 = 2)
 
 
-p <- ggplot(data=chr4.long, aes(x=pos/1000000, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.01) + theme(legend.position=c(0.9, 0.8)) + xlab("physical position (mb)") +
+p <- ggplot(data=chr4.long, aes(x=pos/1000000, y=introgression, color=populations, group=populations)) + geom_smooth(method="loess", size=1, formula = y ~ x, span=0.05) + theme(legend.position=c(0.9, 0.8)) + xlab("physical position (mb)") +
 ylab("introgression level from mexicana to maize")+
-scale_y_continuous(limits=c(0, 2))
+scale_y_continuous(limits=c(0, 2)) +
+geom_rect(xmin=162.9, xmax=192.8, ymin=0, ymax=0.05)
+
 p + geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = seg, size=2)+
 scale_y_continous(limits=c(0, 2))
 

### 2018-02-11 paper.tex
edits
edits to applications

@@
 Rather, we focus on methods that explicitly identify introgressed genomic segments based on the extent of differentiation, on patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
-%We focus here on analytical prospects and limitations and introduce only a few representatives of each category.
 
@@
 Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
-%  $RND_{min} = d_{min}/d_{out}$, where $d_{min}$ is the minimum sequence distance between haplotypes in species X and Y and $d_{out}$ equals $(d_{XO} + d_{YO})/2$, the average sequence distance between each species and the outgroup ($O$).
 These statistics have recently been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
@@
 
-Application of these approaches combined with population genetic statistics to detect selection across a number of plant and animal species suggests introgression can play an adaptive role.
-For example, based on the sequence divergence, introgression has been detected across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species resulting in adaptation to pollinator preference and contributed to speciation \cite{Stankowski2015};
-Jeong et al. \cite{jeong2014} utilized HAPMIX to disclose introgression from Sherpa to Tibetan humans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015});
-ABBA-BABA statistics were employed to reveal introgression of wing coloration loci conferring M\"{u}llerian mimicry
-(a type of mimicry in which two unpalatable species resemble one another, compounding the efficacy of their repellant signal)
-across butterfly species \cite{heliconius2012}.
-.
+Application of these approaches, in combination with population genetic statistics to detect selection, suggest introgression can play an adaptive role.
+For example, based on sequence divergence methods, introgression has been detected across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
+Likewise, the HAPMIX chromosome painting method was applied by Jeong et al.
+\cite{jeong2014} to detect introgression from Nepalese Sherpa to Tibetan humans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015}).
+Finally, the phylogenetic ABBA-BABA statistic has revealed introgression of wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
 
@@
 	\centering
-	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1withInset}
+	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1new}
 	\caption{Hapmix vectors indicating introgression level from \emph{mexicana} to Mexican highland maize on chromosome 4.
@@
 
-\section*{Conclusions}
-
 In the same way that Darwin explored natural selection by studying domesticated species (going so far as to feature domestic pigeons as his "special group" at the beginning of his 

### 2018-02-11 paper.tex
monkeying with tables

@@
 
-Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
-In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
+Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
+In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild haplotypes in the crop and selection against undesirable wild background.
 To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
-High-density marker data can be used with haplotype-based and other methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
+High-density marker data can be used with recently developed methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based on the extent of differentiation, on patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+Rather, we focus on methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 
@@
 Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
-These statistics have recently been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
+These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
 For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
-A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequencies in the admixed population (B) \cite{racimo2016}.
+A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \cite{racimo2016}.
 Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
@@
 Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations based on patterns of allele/haplotype sharing \cite{schraiber2015}. 
-One form of chromosome painting utilizes hidden Markov models to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
+One form of chromosome painting utilizes a hidden Markov models to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
 Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
@@
 
-\begin{table}
+\begin{table}[b]
 
@@
 \rowcolor{white}
-{\bf methods}	& {\bf data type } &	{\bf references} &  {\bf applications } \\ \midrule
+{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
 
 \rowcolor{gray!25}
-{\emph{\bf divergence}} &   &   &   \\
+{\emph{\bf Divergence}} &   &   &   \\
 \rowcolor{gray!25}
@@
 \rowcolor{white}
-{\emph{\bf chromosome painting}} &   &   &   \\
+{\emph{\bf Chromosome Painting}} &   &   &   \\
 \rowcolor{white}
@@
 \rowcolor{gray!25}
-{\emph{\bf phylogenetic relationship}} &   &   &   \\
+{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
 \rowcolor{gray!25}

### 2018-02-11 paper.tex
further edits

@@
 However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
-The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX(Figure \ref{fig:introgressionMaize}).
-These introgressed segments overlapped with QTLs that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
+These introgressed segments overlapped with QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
@@
 A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
-Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to Meso-American high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to Mesoamerican high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
 
-Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
+%Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
 
-\begin{figure}[h]
+\begin{figure}[b]
 	\centering
@@
 Barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
-Gene flow between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.
+Signatures of contemporary gene flow between wild and domesticated barley have been detected across distances greater than a kilometer \cite{hillman2001new}.
 
@@
 Limited linkage disequilibrium and small tracts of identity by state indicated even locally introgressed chromosomal regions are old, perhaps dating to the early expansion of barley  post-domestication.
-While these results are suggestive, wild barley haplotypes have yet to be definitely linked to specific local adaptations in landraces.
+While these results are suggestive, wild barley haplotypes have yet to be definitively linked to specific local adaptations in landraces.
 
@@
 The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
-Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
-Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop exapnsion.
+For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
 In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of \emph{aromatic} rice.
-A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
 
-The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
-According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
-As these fully domesticated populations spread into new environments, they potentially received introgression from locally-adapted wild relatives, retaining alleles that improved fitness.
-
-%Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
-%Contemporary populations of wild rices show evidence of this history of local adaptation.
-%Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
-%Domesticated varieties have likewise adapted to local environments.
-%The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
-%Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
-%However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
-%This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
-%Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
-%Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
+The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
+According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
+As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally-adapted wild relatives, retaining alleles that improved fitness.
+While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
 

### 2018-02-16 paper.tex
edits to case studies
Update paper.tex

@@
 \usepackage[x11names,dvipsnames,table]{xcolor}
+\usepackage{array}
+\newcolumntype{G}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 \bibliographystyle{ieeetr}
@@
 Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000-10000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
-A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel (which included wild, landrace, and cultivated species) formed three groups (North American, South American, and cultivated) and identified \emph{S.
+%A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel (which included wild, landrace, and cultivated species) formed three groups (North American, South American, and cultivated) and identified \emph{S.
+candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.
+A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel identified \emph{S.
 candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.
 
-Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}.
-Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
-These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
-However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
-Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
+Certain barriers to crop-wild introgression persist.
+In general, potoates are propogated clonally, restricting gene flow between wild and domesticated populations (although farmers also promote sexual hybridization at times to improve disease resistance and to develop new cultivars \cite{quiros1992increase}).
+As farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
+Also, natural selection for locally-adaptive alleles may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
 
@@
 The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
+Potatoes grow in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
+These complexes, combined with a diverse range of biotic and environmental selective pressures and conducive local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, intentional maintenance of a variety of landraces, and occasional intentional outcrossing), have fostered the expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
+In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
+candolleanum}) into Andean tetraploid cultivated landraces and North American cultivars (20.2\% and 32.1\%, respectively).
+
+In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
+candolleanum}) into North American cultivars and Andean tetraploid cultivated landraces (32.1\% and 20.2\%, respectively).
+These Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives \cite{zimmerer1998ecogeography}.
+Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting that these introgressed regions impart stress response mechanisms critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
+Drought- and frost-resistance genes introgressed from wild relatives are also reported to have expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
+
 %Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
@@
 %Hardigan and colleagues \cite{hardigan2017genome} find evidence of extensive adaptive introgression into cultivated potatoes.
-In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
-candolleanum}) into Andean tetraploid cultivated landraces and North American cultivars (20.2\% and 32.1\%, respectively).
-Genes located within these introgressed regions were more likely to be highly-expressed, stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting that these introgressed regions impart stress response mechanisms critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion.
-Drought- and frost-resistance genes introgressed from wild relatives are also reported to have expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
-Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
 
@@
 What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-%Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
 At what geographic scale is introgression adaptive?
-\gmj{Barley and maize most strongly show geographic patterns that we can talk about here, but I'm not sure how to discuss them again.}
 To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
-\gmj{Refer to the map and bar graph here?}
-We can look at conservation of genomic architecture across landscapes and between populations, and make predictions about introgressions and their relations to local adaptation.
+We can look at conservation of genomic architecture across landscapes and between populations, and make predictions about introgressions and their relationship with local adaptation.
 If the genomic architecture of an introgressed region is conserved across a broad ecogeographical region, this suggests that the introgression imparts adaptation to general environmental or climatic variables.
@@
 If the genomic architecture of an introgressed region is not conserved within a population, there is little evidence that the introgression is adaptive.
-\gmj{Determination of adaptiveness could be accomplished genomically (looking for high LD regions that indicate selective sweeps) or by functional identification?}
+Also, identification of environmental and/or disease tolerance loci within introgressed regions suggest adaptive significance.
 
 After hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
-%\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
-%\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
-%Does one see different architectures of introgression across different regions?
-%Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
 Introgressed regions are easier to detect when there has been limited recombination to break them apart.
 Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
-\lwang{involve --> involving}
 Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).
+Mutation rate, effective population size, and strength of selection also may influence the rate of breakup of introgressed regions.
 
@@
 Although perhaps less applicable in crop systems, this effect is seen in other well-documented cases of introgression (for example, Neanderthal introgression into humans, \cite{harris2016genetic}).
-\gmj{I tracked down the sources that indicate that introgression is suppressed around genes.
-The case example was Neanderthal alleles into humans, the paper titled "The genetic history of Ice Age Europe" \cite{fu2016genetic}, cited by the Graham Coop paper.
-However, I'm unclear if this is a general rule, or simply a consequence of Neanderthal being high in deleterious genetic load.
-Wouldn't crop systems, with wild relatives harboring putatively adaptive alleles with low load, show a different pattern?}
-%To what extent does the level of introgression across taxa depend on divergence time and mutation load between donor and recipient taxa?
-\gmj{I could also talk about cross-incompatibility factors in maize/teosinte hybridization vs.
-highly compatible asian rice and its relatives.
-I probably wouldn't need a new source for the rice, but I might need to cite a paper about tcb1, ga1, and ga2.}
 
 \subsection*{Applied:}
-Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources, such as annotated genomic sequence assemblies and functional genomic data sets \cite{huang2012}.
+Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources, such as annotated genomic sequence assemblies and functional genomic data sets \cite{Huang2012}.
 Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
@@
 Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
-\gmj{I could here include a handfull of examples of crops that farmers have bred adaptive alleles into.
-Also, we could mention that identification of wild relatives responsible for valuable agronomic introgressed alleles could inform conservation efforts.}
+Identification of wild relative species or populations that harbor putatively adaptive genotypes may facilitate greater prioritization of conservation efforts.
 
@@
 Domesticated crops are among humanity's first and greatest inventions.
-%As a species, domesticated crops are among our first and greatest inventions.
 Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agronomic Revolution.
@@
 
+\gmj{I was able to make the crop summary table look better, but I was unable so far to 1.) move it to the end of the document, and 2.) produce a working caption and table number label.}
+
+\begin{adjustbox}{angle=90}
+    %\begin{tabular}{llllll}
+        \small
+    \begin{tabular}{G{3cm}G{3.5cm}G{3cm}G{3.5cm}G{3.5cm}G{2cm}}
+        \rowcolor{white}
+    \\\toprule  
+    {\bf Domesticated Crop}	& {\bf Compatible Wild Relatives } &	{\bf Hybrids and/or Hybridization} &  {\bf Evidence of Crop Introgression } &	{\bf Evidence of Adaptativeness} & {\bf Sources}\\ \midrule
+    \rowcolor{gray!25}
+Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
+\rowcolor{white}
+Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
+\rowcolor{gray!25}
+Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\
+\rowcolor{white}
+Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\
+\rowcolor{gray!25}
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
+\rowcolor{white}
+Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
+\rowcolor{gray!25}
+Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\
+\rowcolor{white}
+Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
+\rowcolor{gray!25}
+
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome}\\
+\rowcolor{white}
+Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\
+
+\rowcolor{gray!25}
+Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
+\rowcolor{white}
+
+Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\
+\rowcolor{gray!25}
+Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
+
+\rowcolor{white}
+ Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
+    \end{tabular}
+\end{adjustbox}
+
 \bibliography{bib_gj.bib}

### 2018-06-25 paper.tex
edits throughout

@@
 
-\title{Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread}
+\title{Review: Adaptive Introgression Expanded the Genetic Base of Crops during Their post-Domestication Spread}
 
@@
 Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
-In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild haplotypes in the crop and selection against undesirable wild background.
+In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
 To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Preliminary results suggest there is a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred through adaptive gene flow with newly encountered wild relatives during post-domestication expansion.
+Emerging results from these studies suggest there is a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred through adaptive gene flow with newly encountered wild relatives during post-domestication expansion.
 
-In this review, we will: 1) briefly describe recently developed methods for detecting adaptive introgression and provide a summary of how they can be applied to detect crop-wild introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
 
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
-High-density marker data can be used with recently developed methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
+%High-density marker data can be used with recently developed methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
@@
 Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations based on patterns of allele/haplotype sharing \cite{schraiber2015}. 
-One form of chromosome painting utilizes a hidden Markov models to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
+One form of chromosome painting utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
 Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
@@
 
-Third, the ABBA-BABA statistic (also known as the D-statistic) and related statistics are widely applied to introgression detection.
+Phylogenetic relationships are evaluated and applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics.
 These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
-Patterns of allele sharing are interpreted in a phylogenetic context and the method is best suited to detection of introgression at the genome level.
-Elaborations of the D-statistic capable of localizing introgression to specific genomic regions include $\hat{f_{d}}$ \cite{martin2015} and the five-taxon D-statistic \cite{pease2015}.
+While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and the five-taxon D-statistic \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
 The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern and is capable of determining introgression directionality.
@@
 Likewise, the HAPMIX chromosome painting method was applied by Jeong et al.
-\cite{jeong2014} to detect introgression from Nepalese Sherpa to Tibetan humans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015}).
-Finally, the phylogenetic ABBA-BABA statistic has revealed introgression of wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
+\cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015}).
+Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
 
@@
 
-Genome-wide data from extensive samples of crops and their wild relatives, in combination with the new methods described above, have recently allowed detailed analysis of wild-to-crop introgression in some of the world's most important crops (\textbf{Table 2}).
+As range-wide genetic analyses of crops and their wild relatives have become feasible, evidence for substantial wild-to-crop introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
 Below we present a summary of findings from maize, barley, rice, and potato, four promising systems in which introgression from wild relatives appears to have played an adaptive role.
@@
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
-Height differences were not detected under lowland conditions.
+Height differences were not detected under lowland conditions, providing further evidence of local adaptation.
 
-Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
-A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
-Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to Mesoamerican high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-
-%Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
+Populations of \emph{mexicana} cannot be found outside the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
+A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
+Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Maize from the lowlands of northern South America appears to have been derived from Mesoamerican lowland maize and high elevation adaptation in Andean maize likely occurred \emph{de novo} from this genetic base.
 
@@
 \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
-Barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
-Signatures of contemporary gene flow between wild and domesticated barley have been detected across distances greater than a kilometer \cite{hillman2001new}.
+Cultivated barley is found throughout much of the \emph{spontaneum} distribution and barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
+In fact, signatures of contemporary gene flow between wild and domesticated barley have been detected across distances greater than a kilometer \cite{hillman2001new}.
 
-Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
-This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
-Genomic regions of shared ancestry were detected linking particular landraces to numerous wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
-However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
-Limited linkage disequilibrium and small tracts of identity by state indicated even locally introgressed chromosomal regions are old, perhaps dating to the early expansion of barley  post-domestication.
-While these results are suggestive, wild barley haplotypes have yet to be definitively linked to specific local adaptations in landraces.
+Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns of introgression.
+This study identified several lines of evidence consistent with wild introgression aiding the expansion and adaptation of domesticated barley.
+Using chromosome painting methods, the authors identified genomic regions of shared ancestry linking particular landraces to numerous wild relative populations.
+These results suggested landraces may have received wild introgression on a continual basis during post-domestication expansion.
+However, barley landraces also showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
+Limited admixture linkage disequilibrium and small tracts of identity by state suggested even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley  post-domestication.
+While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces.
 
@@
 Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
-Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop exapnsion.
+Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
 For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
@@
 According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
-As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally-adapted wild relatives, retaining alleles that improved fitness.
+As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
 While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
@@
 
-Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000-10000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
-%A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel (which included wild, landrace, and cultivated species) formed three groups (North American, South American, and cultivated) and identified \emph{S.
-candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.
-A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel identified \emph{S.
-candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.
+Modern potato (\emph{Solanum tuberosum}) was likely domesticated in southern Peru in sympatry with several wild relatives approximately 6,000-10,000 BP.
+The exact progenitor has remained in question for some time \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
+candolleanum} as the most probable progenitor \cite{hardigan2015taxonomy}.
+The lack of clarity regarding a progenitor has been due, in part, to extensive post-domestication hybridization between potato and a number of related species.
 
-Certain barriers to crop-wild introgression persist.
-In general, potoates are propogated clonally, restricting gene flow between wild and domesticated populations (although farmers also promote sexual hybridization at times to improve disease resistance and to develop new cultivars \cite{quiros1992increase}).
-As farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
-Also, natural selection for locally-adaptive alleles may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.
-
-Nevertheless, introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
-The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
-Potatoes grow in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
-These complexes, combined with a diverse range of biotic and environmental selective pressures and conducive local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, intentional maintenance of a variety of landraces, and occasional intentional outcrossing), have fostered the expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
-In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
-candolleanum}) into Andean tetraploid cultivated landraces and North American cultivars (20.2\% and 32.1\%, respectively).
-
-In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S.
-candolleanum}) into North American cultivars and Andean tetraploid cultivated landraces (32.1\% and 20.2\%, respectively).
-These Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives \cite{zimmerer1998ecogeography}.
-Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting that these introgressed regions impart stress response mechanisms critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
-Drought- and frost-resistance genes introgressed from wild relatives are also reported to have expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
-
-%Cultivated diploid potatoes have been shown introgressed with genes from \emph{S.
-sparsipilum} \cite{rabinowitz1990high} and \emph{S.
-megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
-%Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S.
-tuberosum subsp.
-andigena}, \cite{van1999tight}, \emph{S.
-bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
-%Hardigan and colleagues \cite{hardigan2017genome} find evidence of extensive adaptive introgression into cultivated potatoes.
+While potatoes are primarily propagated clonally, farmers do at times promote sexual reproduction for improvement of the crop and development of new cultivars \cite{quiros1992increase}.
+Close proximity of domesticated potatoes and wild relatives, active hybridization, and local selection pressure favoring wild haplotypes across a diverse range of biotic and environmental conditions have likely fostered an expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
+The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \cite{hardigan2017genome}.
+These authors discovered that both diploid and tetraploid domesticates had received extensive introgression from more than ten species of wild \emph{Solanum}, with a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
+In certain cultivars, wild ancestry was estimated at upwards of 30\%.
+Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting introgression conferred adaptations critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[b]
 	\centering
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=17.35cm]{boxplot.png}
-	\caption{The distribution of average annual temperatures experienced in the geographic home ranges of wild relatives interfertile with four crops}
-	\label{fig:map}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=17.35cm]{boxplot.png}
+%	\caption{The distribution of average annual temperatures experienced in the geographic home ranges of wild relatives interfertile with four crops}
+%	\label{fig:map}
+%\end{figure}
 
@@
 The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
-For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
+For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.
 
@@
 Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
-In this approach, diversity data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
+In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
 Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
@@
 
-%\lwang{I donot quite understand this point.
-So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right?
-But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
+\section*{Evaluating crop adaptation through introgression:}
 
-%\gmj{It may not fit here perfectly, but we might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.
-The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
+Research has so far shown that adaptive crop-wild introgression has played a significant role during domestication and dispersal of multiple important crops.
+However, the dynamics of this process are not yet fully understood and many questions remain:
 
-\section*{Future studies in crop-wild introgression}
+\subsection*{Do geographic patterns of introgression inform our understanding of adaptation?}
+Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is, in fact, adaptive.
+For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to general environmental or climatic variables.
+On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests much more local selective pressures.
+Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.
+Finally, introgression at genomic loci showing environmental association and/or a role in disease/pest resistance can be particularly suggestive of adaptive significance.
 
-\subsection*{Basic:}%We're beginning to see that introgression has occurred across a number of crops during their expansion.
-%What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-%At what geographic scale is introgression adaptive?  Is it very local?  Does one see different architectures of introgression across different regions?
-%At what taxonomic scale does introgression occur?
-%When do species become so diverged that introgression is either maladaptive or impossible?
+\subsection*{Over what timescales and in what genomic regions can we reliably detect adaptive introgression?}
+Introgressed haplotypes are most easily detected with limited recombination post-hybridization.
+Therefore, recent introgressions (limited meioses) or those occurring in low recombination regions such as centromeres or inversions are preferentially detected.
+Moreover, since recombination will progressively degrade tracts of introgression, the genome-wide distribution of tract lengths based on a genetic map can be used to date introgression events (as in \cite{Poets2015}).
+Mutation rate, effective population size, and strength of selection may also influence the rate of breakup of introgressed regions.
 
-%Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication histories of many agronomically-important crops.
-%However, the dynamics of the process in particular cases are not yet fully understood.
-%To what extent does the level of introgression across taxa depend on divergence time and/or mutation load between donor and recipient taxa?
-\gmj{I think we could include (effective) population size in this list of factors that determine extent of introgression.}
-%At what geographic scale does adaptive introgression occur? Is introgression generally restricted to very local populations, or does it pervasive over broad geographic ranges?
-%To what extent does this depend on the slope of environmental gradients, such as temperature, precipitation, and elevation, across these ranges?
+\subsection*{At what taxonomic scale does introgression occur?}
+As species become substantially diverged, introgression can become maladaptive or impossible due to, for example, Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers to hybridization.
+Divergence time may therefore be a useful predictor of the possibility of gene flow between a particular wild relative and a domesticate.
+Hybridization may also be limited between a crop and a particular wild relative due to demography. 
+For instance, gene flow from a wild relative with a small long-term effective population size, and correspondingly high genetic load may not be favored by selection.
+This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited in extent and relegated largely to non-genic regions due to the high genetic load found within Neanderthal donor individuals \cite{harris2016genetic}.
 
-%\gmj{I am not seeing the reason behind splitting this up in into these two paragraphs.  I'll try reformatting the same information below:}
+\subsection*{Can adaptive introgression inform crop improvement?}
 
-Research has so far shown that adaptive crop-wild introgression has played a significant role in the domestication and dispersal of many agronomically-important crops.
-However, the dynamics of this process are not yet fully understood, especially in the context of individual case examples, and many questions remain.
-
-What is the genomic architecture of this introgression, and does the architecture suggest that it has been adaptive?
-At what geographic scale is introgression adaptive?
-To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation across these ranges?
-We can look at conservation of genomic architecture across landscapes and between populations, and make predictions about introgressions and their relationship with local adaptation.
-If the genomic architecture of an introgressed region is conserved across a broad ecogeographical region, this suggests that the introgression imparts adaptation to general environmental or climatic variables.
-On the other hand, if the genomic architecture is conserved within populations but not between nearby populations in the region, this suggests that the introgressed regions offer adaptations to more local selective pressures.
-If the genomic architecture of an introgressed region is not conserved within a population, there is little evidence that the introgression is adaptive.
-Also, identification of environmental and/or disease tolerance loci within introgressed regions suggest adaptive significance.
-
-After hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
-Introgressed regions are easier to detect when there has been limited recombination to break them apart.
-Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
-Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).
-Mutation rate, effective population size, and strength of selection also may influence the rate of breakup of introgressed regions.
-
-At what taxonomic scale does introgression occur?
-When do species become so diverged that introgression is either maladaptive or impossible (due to  Dobzhansky-Muller incompatiblities or other mechanisms)?
-Theory indicates that the most significant limiting factor to gene flow between progenitor and domesticate is divergence time.
-Over time, diverged populations drift and become increasingly incompatible.
-Small effective population size, and correspondingly high genetic load, of the introgressive population also limits gene flow.
-Although perhaps less applicable in crop systems, this effect is seen in other well-documented cases of introgression (for example, Neanderthal introgression into humans, \cite{harris2016genetic}).
-
-\subsection*{Applied:}
-Our identification and understanding of introgression in agroecosystems would be augmented by the development of wild relative genomic resources, such as annotated genomic sequence assemblies and functional genomic data sets \cite{Huang2012}.
-Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics, and specifically the continued improvement of modern crops.
-Loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
+Additional study of introgression in agroecosystems could lead to advances in crop improvement.
+As described above, loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
 These loci are potentially beneficial targets for crop improvement.
-Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
-Identification of wild relative species or populations that harbor putatively adaptive genotypes may facilitate greater prioritization of conservation efforts.
+Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be directly utilized in crop breeding.
+Identification of wild relative species or populations that harbor such adaptive genotypes may lead to their prioritization as part of conservation efforts.
 
@@
 
-In the same way that Darwin explored natural selection by studying domesticated species (going so far as to feature domestic pigeons as his "special group" at the beginning of his 
-\emph{On the Origin of Species} \cite{darwin2009origin}), we frequently find that modern crop systems, in tandem with their extant wild relatives, offer unparalleled opportunities for the study of several evolutionary questions, including the role of adaptive introgression in dispersal and range expansion.
-The advent and application of genomics tools has lead to mounting evidence which underscores the prevalence of genetic admixture between crops and wild relatives during and well beyond the initial stages of domestication.
-Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adpated alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
-Within this new conception, the process of crop range expansion is a concurrent element of domestication, rather than a distinct event subsequent to it.
-%This is a revolutionary departure from the simpler model of a temporally-constrained domestication event with subsequent dispersal, and as such, offers new questions and new opportunities for investigation.
-This is a revolutionary departure from the prevailing model, and as such, offers new questions and new opportunities for investigation.
-These questions take on practical importance as well as theoretical.
-%Understanding the histories of domestication and adaptation inform and empower efforts to develop and improve crop systems.
-Understanding a crop's history of domestication in terms of selection, adaptation, and admixture empowers efforts at crop development and improvement.
-As the historic role of wild relatives in the development of many crops is increasingly clarified, it is likely that we may reevaluate the importance of conservation of wild populations (in particular or in general) that harbor agronomically-valuable alleles that may be targeted by pre-breeding efforts.
-These efforts may be of particular value in the face of future climate volatility and change.
+As recognized early on by Darwin, crop species represent promising model systems for evolutionary study. 
+The application of genomic tools has generated mounting evidence of crop-wild gene flow during and well beyond the initial stages of domestication.
+In support of these findings, population genetic theory reinforces the proclivity of native species (in this instance, wild relatives) to provide advantageous, locally adapted alleles to colonizing species (crops) as they disperse beyond their sites of origin into new habitats.
+The process of crop range expansion and adaptation through gene flow is closely linked to domestication and, as such, offers new questions and opportunities for practical and theoretical investigation.
+As the historic role of wild relatives in the adaptation of crops is clarified, their conservation may be more prioritized, particularly in the face of future climate volatility and change.
 
@@
 Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agronomic Revolution.
-In this more nuanced and dynamic understanding of crop domestication, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.
-\gmj{I let these final few sentences get a little more poetic than I normally would.  I think maybe I should reel them back in, or cut them.}
-
-%The study of crop domestication has been revolutionized by the advent and application of genomic tools.
-%The genomes of crops and their wild relatives tell a story of give-and-take that extends well beyond the initial stages of domestication.
-%Likewise, population genetic theory reinforces the proclivity of wild relatives to provide advantageous, locally-adapted alleles to crops as they disperse beyond their domestication centers into new geographies with new ecological pressures and niches.
-
-%Omitted by Garrett
-%\item{Sunflower: }
-
-%The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in the eastern United States \cite{harter2004origin, wills2006chloroplast}\, with potential for a second domestication center in Mexico \cite{lentz2008sunflower}.
-%The pre-Columbian \emph{H.
-annuus} distribution of cultivated sunflower spanned much of the Great Plains, from what is now north-central Texas to Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
-%Domesticated sunflower has long lived in sympatry with wild relatives such as \emph{H.
-petiolaris} and \emph{H.
-bolanderi} and forms stable hybrid populations with these taxa \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}.
-%Wild sunflowers are known to be locally-adapted, and weedy hybrid populations often share these adaptations \cite{kane2008genetics}.
-%However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
-%\mbh{I believe \emph{H. annuus} ssp. \emph{texanus} is a wild sunflower, right?  If so, this is more an example of hybrid speciation rather then adaptive wild-to-crop gene flow}
-%Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
-%Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
-%These introgressed hybrids formed a new lineage of sunflower (\emph{H.
-annuus} ssp.
-\emph{texanus}, \emph{H.
-a.
-texanus} hereafter) which displays \emph{H.
-debilis}-like traits adaptive to south-central Texas climate and ecology.
-%These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H.
-annuus} (as measured by higher seed production \cite{whitney2006adaptive}).
-%Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.
-
-%The genome of the common sunflower has been greatly influenced by introgression from wild relatives, due to both natural outcrossing events and concerted breeding efforts in crop improvement.
-%\emph{Helianthus} has several genes for downy mildew resistance, and each imparts resistance to one or more races of \emph{Plasmopara halstedii}, one of the most agronomically important diseases in sunflower cultivation \cite{cohen1973factors}.
-%Some of these downy mildew resistence genes were found in wild relatives (including \emph{H.
-argophyllus}, \emph{H.
-tuberosus}, and \emph{H.
-praecox}) and have been successfully bred into modern \emph{H.
-annuus} \cite{miller1991inheritance}.
-%PlArg, an allele found in wild silverleaf sunflowers (\emph{H.
-argophyllus}, inbred line Arg1575-2), confers resistance to all known (20 or more) races of downey mildew \cite{dussle2004pl}\, while others (Pl1-Pl11) are effective for one or more types \cite{rahim2002inheritance}.
-%Silverleaf sunflower has also been the focus of drought resistance breeding efforts \cite{saucă2010introgression}\ and \emph{Phomopsis} resistance breeding efforts \cite{besnard1997specifying}.
-%\emph{H.
-annuus} shows signs of persistent introgressive hybridization with \emph{H.
-petiolaris} with evidence of positive selection driving some of the genetic differentiation between the two species \cite{yatabe2007rampant}.
-
-%Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
-%\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
-%Using STRUCTURE, these authors found that introgressions from wild relatives exist on every chromosome in at least one modern line, covering over 10\% of the genome.
-%Of particular note is the modern line RHA 274, a modern line which was bred with \emph{H.
-a.
-texanus} in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
-%RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
-%These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.
-
-%Notes for review:
-
-%http://biology.unm.edu/Whitney/Whitney%20Reprints/Whitney%20et%20al.%202006%20Am%20Nat.pdf
-
-%http://www.jstor.org/stable/pdf/2409227.pdf?refreqid=excelsior:a686b7f06f9daa2b89bd5a654dade3b4
-
-%file:///home/gjanzen/Downloads/Whitney_et_al-2010-New_Phytologist%20(2).pdf
-
-%http://onlinelibrary.wiley.com/doi/10.1111/j.1469-8137.2010.03234.x/epdf
-
-%https://biology.unm.edu/Whitney/Whitney%20Reprints/Scascitelli%20et%20al%202010.pdf
-
-%Genome scan of hybridizing sunflowers from Texas
-%(Helianthus annuus and H. debilis) reveals asymmetric
-%patterns of introgression and small islands of genomic
-%differentiation:
-
-%In contrast, coalescent analyses of long-term migration
-%rates imply that interspecific migration has been
-%more important than mutation in providing new genetic
-%variation in ssp. texanus populations. Immigration from
-%H. debilis cucumerifolius into ssp. texanus (M = 4.05;
-%Nem = 1.74) is higher than median estimates of intraspecific
-%migration rates in plants (Morjan and Rieseberg 2004) 
-%and only slightly lower than immigration from
-%ssp. annuus into ssp. texanus (M = 5.00). Immigration
-%rates from ssp. texanus into H. debilis cucumerifolius and
-%ssp. annuus are still higher (M = 6.30 and 7.29, respectively),
-%implying that the putative hybrid lineage serves
-%as a bridge for migration between the parental species.
-
-%One puzzle is
-%that the estimated immigration rate between the allopatric
-%taxa (H. debilis cucumerifolius and ssp. annuus) is
-%approximately comparable with that between the sympatric
-%taxa (H. debilis cucumerifolius and ssp. texanus ssp.
-%texanus). This observation may point to the effectiveness
-%of ssp. texanus as a bridge for the transfer of alleles
-%between H. debilis and ssp. annuus. An alternative explanation
-%is that the BAYESASS estimates are unreliable
-%because homoplasy of microsatellite alleles is high in
-%taxa with large-effective population sizes (Estoup et al.
-%2002) such as the annual sunflowers studied here.
-
-%Despite the apparent discrepancy between the results
-%from STRUCTURE and BAYEASS, several conclusions
-%can be made. Analyses with STRUCTURE support the
-%distinctness of ssp. texanus, as well as its clear inclusion
-%within H. annuus, as initially postulated by Heiser (1951).
-%Likewise, the STRUCTURE, MIGRATE-N and BAYESASS
-%analyses all confirm previous report of introgression
-%between H. annuus and H. debilis (Rieseberg et al.
-%1990, 2007) and suggest that the introgression is genomewide
-%when it occurs. Contrary to the Heiser’s scenario of
-%a recent Holocene origin of ssp. texanus, however, our
-%results are more consistent with a much longer history of
-%contact between H. annuus and H. debilis. Otherwise, it is
-%difficult to account for the presence of significant longterm
-%migration between currently allopatric populations
-%of ssp. annuus and H. debilis cucumerifolius. This revised
-%scenario makes sense in the light of numerous glacialinterglacial
-%cycles over the past million years (Hewitt
-%2000), which probably resulted in intermittent contact
-%between H. annuus and H. debilis, with the last contact
-%likely occurring during the Wisconsin glaciation,
-%18 000 BP. The current contact between the species differs
-%from the previous ones in that it appears to have been
-%human-aided. Possibly, some of the molecular evidence
-%of introgression, particularly into allopatric populations,
-%stems from past periods of contact.
-
-%One of the four outlier loci, HT0414, is associated
-%with a heat-shock protein and was previously shown to
-%have been the target of selection in H. annuus ssp. annuus
-%salt marsh populations from the state of Utah
-%(Kane and Rieseberg 2007), as well as in weedy sunflower populations
-%from several different locations across the USA
-%(Kane and Rieseberg 2008). In this study, populations of
-%ssp. annuus from Kansas, Nebraska and Oklahoma
-%monomorphic for a single allele, whereas populations
-%of H. debilis and other populations of ssp. annuus are
-%considerably more polymorphic. These results imply
-%that HT0414 may be involved in adaptation to a range
-%of different habitats or to conditions shared by several
-%different habitats (Kane and Rieseberg 2008).
-
-%The species \emph{H. annuus} is a versatile species, capable of adapting to a wide range of environments across the Americas, Europe, Asia, and Australia \cite{kane2008genetics}.
-%This versatility may be due in part either to phenotypic plasticity or genetic adaptability \cite{maron2004rapid}.
-
-%whitney2006adaptive
-%"these results suggest that introgression of biotic resistance traits was important in the adaptation of H. annuus to central and southern Texas."
-%fitness (seed production) high in texanus than in annuus
-%identified two pests to which resistance was imparted and important
-%Discussion points: 1. texanus has higher fitness than annuus in central texas.
-%2. texanus shared traits more in common with debilis than with annuus (herbivore resistance)
-%3. 2 of the three traits from point two above are important for adaptation to life in central south Texas.
-%The authors ask why these pest-resistance genes have not spread further north beyond this range.  They are uncertain, but conjecture that negative pleiotrophic effects are at play.
-
-%whitney2010adaptive
-%"We demonstrate that introgression has altered multiple aspects of the H.
-annuus phenotype in an adaptive manner, has affected traits relevant to both biotic and abiotic environments, and may have aided expansion of the H.
-annuus range into central Texas, USA."
-%This paper really does show that this case of natural introgression was adaptive.
-%The companion paper (2006 Whitney) will likely show the same thing, but I haven't read it yet.
-
-%scascitelli2010genome
-%"long-term migration rates were high, genome-wide and asymmetric, with higher migration rates from H. annuus texanus into the two parental taxa than vice versa."
-%"H. annuus texanus may serve as a bridge for the transfer of alleles between its parental taxa."
-%"contradict recent theory suggesting that introgression should predominantly be in the direction of the colonizing species."
-
-%rieseberg2007hybridization
-%this is a very good review.  i think i should use it for the table.
-%they also show that traits were imparted from debilis to texanus, and that these are adaptive.  i think they also show genetic evidence, but i'm not sure.
-
-\gmj{I was able to make the crop summary table look better, but I was unable so far to 1.) move it to the end of the document, and 2.) produce a working caption and table number label.}
+In a more nuanced and dynamic understanding of crop domestication that includes sustained gene flow with wild relatives, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.
 

### 2018-06-25 paper.tex
minor edits to titles

@@
 
-\title{Review: Adaptive Introgression Expanded the Genetic Base of Crops during Their post-Domestication Spread}
+\title{Review: The Prevalence of Adaptive Introgression in Crop Histories}
 
@@
 
-In this review, we will: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.
+In this review, we will: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe remaining questions regarding crop adaptation through gene flow with wild relatives.
 

### 2018-06-26 paper.tex
more context added to introgression methods

@@
 
-The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation \cite{Harlan1992}.
+Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
-An additional important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.
+An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following initial domestication.
+It is this adaptive introgression that is the subject of this review.
 
@@
 However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Emerging results from these studies suggest there is a need to expand our conception of domestication to include a broadening of the genetic base of crops that occurred through adaptive gene flow with newly encountered wild relatives during post-domestication expansion.
+Emerging results from these studies suggest there is a need to expand our conception of domestication to encompass the broadening of the genetic base of crops that occurred through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.
 
-In this review, we will: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe remaining questions regarding crop adaptation through gene flow with wild relatives.
+In this review, we: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe some key questions regarding crop adaptation that is mediated by gene flow from wild relatives.
 
@@
 
-The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated comprehensive study of introgression across a number of species (\textbf{Table 1}).
-%High-density marker data can be used with recently developed methods to assign specific genomic regions to a taxon of origin and identify introgression across taxa \cite{martin2015,Price2009,Lawson2012,pease2015,rosenzweig2016,geneva2015}.
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated the comprehensive study of introgression across a broad spectrum of domesticated species (\textbf{Table 1}).
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
-Rather, we focus on methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+Rather, we focus on three broad categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 
@@
 These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
-For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
+For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes the number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \cite{racimo2016}.
 Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
+Since differentiation-based methods can be calculated site-by-site, high-density, genome-wide data are not necessarily required.
+However, accuracy of introgression estimates is improved with more comprehensive data.
 
-Second, local ancestry deconvolution (also known as chromosome painting) assigns genomic regions to various source populations based on patterns of allele/haplotype sharing \cite{schraiber2015}. 
-One form of chromosome painting utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
+Second, local ancestry deconvolution assigns genomic regions to various source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}. 
+One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
 Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
-And finally, introgression can be detected through chromosome painting by using a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+And finally, introgression can be detected through ancestry deconvolution by using a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+Ancestry deconvolution methods are better suited to high-density marker or whole genome sequencing data given the intent to assign ancestry genome-wide.
 
@@
 These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
-While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and the five-taxon D-statistic \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
-The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter detects introgression based on the localized phylogenetic pattern and is capable of determining introgression directionality.
+While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and $D_{\textrm{FOIL}}$ tests \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
+The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter can identify donor and recipient lineages of introgression in a more complex, five-taxon phylogeny.
+Like differentiation methods, phylogeny-based detection of introgression can be employed using low-density data.
+However, because these methods require knowledge of whether an allele is ancestral or derived, data from a sufficiently diverged outgroup must also be included in analyses.
 
-Application of these approaches, in combination with population genetic statistics to detect selection, suggest introgression can play an adaptive role.
-For example, based on sequence divergence methods, introgression has been detected across \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
+Application of these approaches, in combination with population genetic statistics to detect selection, suggest that introgression can play an adaptive role.
+For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
 Likewise, the HAPMIX chromosome painting method was applied by Jeong et al.

### 2018-06-26 paper.tex
edits throughout

@@
 
-As range-wide genetic analyses of crops and their wild relatives have become feasible, evidence for substantial wild-to-crop introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
-Below we present a summary of findings from maize, barley, rice, and potato, four promising systems in which introgression from wild relatives appears to have played an adaptive role.
+As comprehensive genetic analyses of crops and their wild relatives have become feasible across their geographic ranges, evidence for substantial wild-to-crop introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
+Below we present a summary of findings from maize, barley, rice, and potato, four systems in which introgression from wild relatives appears to have played an adaptive role.
+All four of these crops were domesticated in defined centers and have subsequently expanded to much broader distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=17.35cm]{temperature_plot.png}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.
+The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
+	\label{fig:map}
+\end{figure}
 
@@
 \emph{mays}) and the teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter, \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
+\emph{mexicana} (hereafter, \emph{mexicana}) offers a prime example of adaptive wild-to-crop introgression.
 Maize was domesticated from \emph{Zea mays} ssp.
@@
 However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
-The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
 These introgressed segments overlapped with QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
@@
 Cultivated barley is found throughout much of the \emph{spontaneum} distribution and barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
-In fact, signatures of contemporary gene flow between wild and domesticated barley have been detected across distances greater than a kilometer \cite{hillman2001new}.
 
@@
 This study identified several lines of evidence consistent with wild introgression aiding the expansion and adaptation of domesticated barley.
-Using chromosome painting methods, the authors identified genomic regions of shared ancestry linking particular landraces to numerous wild relative populations.
+Using ancestry deconvolution methods, the authors identified genomic regions of shared ancestry linking particular landraces to numerous wild relative populations.
 These results suggested landraces may have received wild introgression on a continual basis during post-domestication expansion.
 However, barley landraces also showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
-Limited admixture linkage disequilibrium and small tracts of identity by state suggested even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley  post-domestication.
-While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces.
+Limited admixture linkage disequilibrium and small tracts of identity by state suggest substantial recombination has occurred since initial crop-wild hybridization and that even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley post-domestication.
+While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in domesticated landraces.
 
@@
 The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
 Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
-For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
+For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
 In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of \emph{aromatic} rice.
@@
 
-Modern potato (\emph{Solanum tuberosum}) was likely domesticated in southern Peru in sympatry with several wild relatives approximately 6,000-10,000 BP.
+Modern potato (\emph{Solanum tuberosum}) was likely domesticated approximately 6,000-10,000 BP in southern Peru in sympatry with several wild relatives.
 The exact progenitor has remained in question for some time \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
@@
 Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting introgression conferred adaptations critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
+\end{enumerate}
 
-\end{enumerate}
+While the four crop systems described in detail here represent particularly well documented examples of adaptive wild introgression, given their similar histories, many additional crops have likely benefited from beneficial wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
+Across these four cases studies, potential generalities can be observed.
+Data thus far indicate that wild introgression is often local or regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S.
+highlands), newly introgressed wild haplotypes can be disseminated more broadly.
+Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
+Beyond these observations associated with post-domestication adaptation, pervasive wild-to-crop gene flow bears meaningfully on the study of domestication.
 
@@
 \rowcolor{white}
-{\emph{\bf Chromosome Painting}} &   &   &   \\
+{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
 \rowcolor{white}
@@
 
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[width=17.35cm]{vertical.png}
-%	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.}
-%	\label{boxplot:map}
-%\end{figure}
-
-\begin{figure}[b]
-	\centering
-	\includegraphics[width=17.35cm]{temperature_plot.png}
-	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.
-The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
-	\label{fig:map}
-\end{figure}
-
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[width=17.35cm]{boxplot.png}
-%	\caption{The distribution of average annual temperatures experienced in the geographic home ranges of wild relatives interfertile with four crops}
-%	\label{fig:map}
-%\end{figure}
-
 \section*{Re-evaluating domestication}
 
-A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression is extensive during crop geographic expansion.
-The addition of ongoing gene flow to our understanding of crop demography could therefore bear importantly on fundamental questions of crop domestication:
+A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression has been extensive throughout a crop's history.
+The addition of ongoing gene flow to our understanding of crop demography could therefore bear on fundamental questions of crop domestication:
 

### 2018-06-26 paper.tex
edits to domestication questions

@@
 
-\subsection*{What is the progenitor of a crop?}
-Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
+\subsection*{Where and from what taxon did a crop originate?}
+Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's origin may be complicated or confounded entirely.
 Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
@@
 Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
-Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
-Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013, Hardigan2017}.
+Similarly, as described above, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
 
-Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
-In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to the genetic base of modern crops.
-Broad recognition of the role these wild relatives have played in crop adaptation could further their use in breeding and elevate their conservation status.
+Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a more nuanced view of crop origins.
+In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to modern crops.
+While these crops may have originated from a single species or subspecies, the crops as we know them today have a broader genetic base.
 
@@
 In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
-In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxon.
-This phenomenon, in combination with divergence of crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
+In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxa.
+This phenomenon, in combination with divergence of modern crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
 More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.
 
-\subsection*{How was genome-wide diversity shaped by domestication?}
+\subsection*{How has genome-wide diversity been shaped by domestication?}
 
@@
 The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
-For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.
+For example, substantial introgression from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.
 
@@
 In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
-Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Due to the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
 Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
@@
 However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
+\vspace{5mm}
 
-\section*{Evaluating crop adaptation through introgression:}
+In summary, should post-domestication gene flow with wild relatives be pervasive during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.
+
+\section*{Investigating crop adaptation through introgression:}
 

### 2018-06-26 paper.tex
edits to adaptation section

@@
 
-Research has so far shown that adaptive crop-wild introgression has played a significant role during domestication and dispersal of multiple important crops.
-However, the dynamics of this process are not yet fully understood and many questions remain:
+Research has so far shown that adaptive crop-wild introgression has played a significant role during domestication and dispersal of many important crops.
+However, the scope and dynamics of this process are not yet fully described and remain unexplored in many systems.
+In determining the extent and nature of adaptation due to introgression, several questions should be considered:
 
@@
 Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is, in fact, adaptive.
-For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to general environmental or climatic variables.
-On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests much more local selective pressures.
+For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).
+On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests much more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
 Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.
-Finally, introgression at genomic loci showing environmental association and/or a role in disease/pest resistance can be particularly suggestive of adaptive significance.
 
@@
 Moreover, since recombination will progressively degrade tracts of introgression, the genome-wide distribution of tract lengths based on a genetic map can be used to date introgression events (as in \cite{Poets2015}).
-Mutation rate, effective population size, and strength of selection may also influence the rate of breakup of introgressed regions.
+Mutation rate, effective population size, and strength of selection on introgressed alleles may also influence the rate of breakup of introgressed chromosomal regions.
+Detection of introgression will also be affected by the extent of divergence of donor taxa from a crop's wild progenitor; highly divergent introgressed haplotypes will be easier to identify.
 
 \subsection*{At what taxonomic scale does introgression occur?}
-As species become substantially diverged, introgression can become maladaptive or impossible due to, for example, Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers to hybridization.
+As species become substantially diverged, introgression can become maladaptive or impossible due to, for example, Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
 Divergence time may therefore be a useful predictor of the possibility of gene flow between a particular wild relative and a domesticate.

### 2018-06-27 paper.tex
edits throughout

@@
 For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
-Likewise, the HAPMIX chromosome painting method was applied by Jeong et al.
+Likewise, the HAPMIX ancestry deconvolution method was applied by Jeong et al.
 \cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015}).
@@
 Populations of \emph{mexicana} cannot be found outside the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
-A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
+Wang and co-authors \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
 Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
@@
 
-While the four crop systems described in detail here represent particularly well documented examples of adaptive wild introgression, given their similar histories, many additional crops have likely benefited from beneficial wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
+The four crop systems described in detail here represent particularly compelling examples of putatively adaptive wild introgression.
+However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
 Across these four cases studies, potential generalities can be observed.
@@
 Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
-Beyond these observations associated with post-domestication adaptation, pervasive wild-to-crop gene flow bears meaningfully on the study of domestication.
+Beyond these observations associated with post-domestication adaptation, pervasive wild-to-crop gene flow is also relevant to the study of domestication itself.
 

### 2018-06-27 paper.tex
edits to adaptation questions

@@
 Therefore, recent introgressions (limited meioses) or those occurring in low recombination regions such as centromeres or inversions are preferentially detected.
-Moreover, since recombination will progressively degrade tracts of introgression, the genome-wide distribution of tract lengths based on a genetic map can be used to date introgression events (as in \cite{Poets2015}).
-Mutation rate, effective population size, and strength of selection on introgressed alleles may also influence the rate of breakup of introgressed chromosomal regions.
-Detection of introgression will also be affected by the extent of divergence of donor taxa from a crop's wild progenitor; highly divergent introgressed haplotypes will be easier to identify.
+While this can be problematic for the detection of ancient introgression, the fact that recombination degrades tracts of introgression at a relatively constant and predictable rate allows use of the genome-wide distribution of introgression tract lengths to date initial hybridization (as in \cite{Poets2015}).
+Detection of introgression will also be affected by mutation rate, effective population size, the strength of selection on introgressed alleles, and the extent of divergence between donor taxa and a crop's wild progenitor (\emph{e.g.}, highly divergent introgressed haplotypes will be easier to identify).
 

### 2018-06-27 paper.tex
edits to conclusions

@@
 As described above, loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
-These loci are potentially beneficial targets for crop improvement.
-Furthermore, adaptive introgression that is demonstrably tied to a specific environment may include beneficial alleles that can be directly utilized in crop breeding.
-Identification of wild relative species or populations that harbor such adaptive genotypes may lead to their prioritization as part of conservation efforts.
+These loci are potentially beneficial targets for crop improvement and their accurate identification is crucial.
+Furthermore, adaptive introgression that is demonstrably tied to a specific environment represents a promising source of beneficial alleles that can be directly utilized in breeding to adapt crops to similar conditions.
+Finally, as the historic role of wild relatives in the adaptation of crops is clarified, their conservation may be more prioritized, particularly as a resource for breeding in the face of future climate volatility and change.
 
@@
 The application of genomic tools has generated mounting evidence of crop-wild gene flow during and well beyond the initial stages of domestication.
-In support of these findings, population genetic theory reinforces the proclivity of native species (in this instance, wild relatives) to provide advantageous, locally adapted alleles to colonizing species (crops) as they disperse beyond their sites of origin into new habitats.
-The process of crop range expansion and adaptation through gene flow is closely linked to domestication and, as such, offers new questions and opportunities for practical and theoretical investigation.
-As the historic role of wild relatives in the adaptation of crops is clarified, their conservation may be more prioritized, particularly in the face of future climate volatility and change.
-
+The process of crop range expansion and adaptation through gene flow is closely linked to domestication and offers new questions and opportunities for practical and theoretical investigation.
 Domesticated crops are among humanity's first and greatest inventions.

### 2018-06-27 paper.tex

@@
 
-%Short Research reviews should be in the range 3500–4000 words, with up to 40 references and 6 figures/tables.
-
 Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
@@
 
-The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods, has facilitated the comprehensive study of introgression across a broad spectrum of domesticated species (\textbf{Table 1}).
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of domesticated species.
 The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
@@
 
+\begin{table}[h]
+
+\rowcolors{2}{white}{gray!25}
+\begin{center}
+\caption{List of recently developed methods for detecting introgression and examples of their use in empirical studies.} \label{tab:tools}
+\begin{tabularx}{\textwidth}{llll}
+\\\toprule  
+\rowcolor{white}
+{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
+
+\rowcolor{gray!25}
+{\emph{\bf Divergence}} &   &   &   \\
+\rowcolor{gray!25}
+Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
+\rowcolor{gray!25}
+RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
+\rowcolor{gray!25}
+$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
+
+\rowcolor{white}
+{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
+\rowcolor{white}
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
+\rowcolor{white}
+RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
+\rowcolor{white}
+MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
+\rowcolor{white}
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
+\rowcolor{white}
+LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
+
+\rowcolor{gray!25}
+{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
+\rowcolor{gray!25}
+ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
+\rowcolor{gray!25}
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
+\rowcolor{gray!25}
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
+
+\end{tabularx}
+\end{center}
+\end{table}
+
 \section*{Crop adaptation through introgression}
@@
 
-\begin{figure}[b]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=17.35cm]{temperature_plot.png}
-	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature.
-The distribution of average annual temperature experienced in the geographic home ranges of wild relatives interfertile with four crops}
+	\includegraphics[width=15cm]{temperature_plot.png}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature. Approximate domestication center for each crop is denoted by a triangle}
 	\label{fig:map}
@@
 
-\begin{figure}[b]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=17.35cm]{./FigsAndTables/Figure1new}
+	\includegraphics[width=12cm]{./FigsAndTables/Figure1new}
 	\caption{Hapmix vectors indicating introgression level from \emph{mexicana} to Mexican highland maize on chromosome 4.
@@
 ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
-Insets showed the phenotypes of macrohairs on lowland (red) and highland (blue) maize stems.
-The black bar indicated one of the QTLs for macrohairs and pigment density in Lauter et al.
+Insets show the phenotypic differences between lowland (red) and highland (blue) maize stems.
+The black bar indicates one of the QTLs for macrohairs and pigment density in Lauter et al.
 \cite{lauter2004}.}
@@
 
-\begin{table}[b]
-
-\rowcolors{2}{white}{gray!25}
-\begin{center}
-\caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
-\begin{tabularx}{\textwidth}{llll}
-\\\toprule  
-\rowcolor{white}
-{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
-
-\rowcolor{gray!25}
-{\emph{\bf Divergence}} &   &   &   \\
-\rowcolor{gray!25}
-Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
-\rowcolor{gray!25}
-RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
-\rowcolor{gray!25}
-$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
-
-\rowcolor{white}
-{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
-\rowcolor{white}
-Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
-\rowcolor{white}
-RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
-\rowcolor{white}
-MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
-\rowcolor{white}
-PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
-\rowcolor{white}
-LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
-
-\rowcolor{gray!25}
-{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
-\rowcolor{gray!25}
-ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
-\rowcolor{gray!25}
-fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
-\rowcolor{gray!25}
-five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
-
-\end{tabularx}
-\end{center}
-\end{table}
-
-\begin{table}
-\centering
-\begin{adjustbox}{width=1\textwidth}
-\small
-\label{my-label}
-\begin{tabular}{|p{5cm}|p{5cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|l|}
-\hline
-Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
-Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\ 
-\hline 
-Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
-\hline
-Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
-Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\ 
-\hline
-Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\ 
-\hline
-Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome, johns1986ongoing, gavrilenko2013genetic} \\
-\hline
-Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
-\hline
-Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
-\hline
-Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\ 
-\hline
-Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
-\hline
-Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
-\hline
-Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
-\hline
-Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
-\hline
-Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
-\hline
-\end{tabular}
-\end{adjustbox}
-\end{table}
+%\centering
+%\begin{adjustbox}{width=1\textwidth}
+%\small
+%\label{my-label}
+%\begin{tabular}{|p{5cm}|p{5cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|l|}
+%\hline
+%Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
+%Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\ 
+%\hline 
+%Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
+%\hline
+%Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
+%Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\ 
+%\hline
+%Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\ 
+%\hline
+%Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome, johns1986ongoing, gavrilenko2013genetic} \\
+%\hline
+%Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
+%\hline
+%Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
+%\hline
+%Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\ 
+%\hline
+%Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
+%\hline
+%Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
+%\hline
+%Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
+%\hline
+%Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
+%\hline
+%Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
+%\hline
+%\end{tabular}
+%\end{adjustbox}
+%\end{table}
 

### 2018-06-27 paper.tex
edits to references

@@
 Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
-However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{Meyer2016, Wang2017, Fuller2014}.
 An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following initial domestication.
@@
 
-Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
+Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role}.
 In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of domesticated species.
-The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and a series of methods utilizing the sequentially Markovian coalescent (PSMC, MSMC and SMC++) \cite{li2011, schiffels2014, terhorst2017}).
+The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
 Rather, we focus on three broad categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
@@
 Populations of \emph{mexicana} cannot be found outside the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
-Wang and co-authors \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
+Wang and co-authors \cite{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
 Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
@@
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
+\emph{agriocrithon} in Tibet \cite{badr2000origin, Morrell2007, dai2012tibet}.
 Presently, the distribution of subsp.
-\emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
+\emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
 Cultivated barley is found throughout much of the \emph{spontaneum} distribution and barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
@@
 Modern potato (\emph{Solanum tuberosum}) was likely domesticated approximately 6,000-10,000 BP in southern Peru in sympatry with several wild relatives.
-The exact progenitor has remained in question for some time \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
+The exact progenitor has remained in question for some time \cite{spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
 candolleanum} as the most probable progenitor \cite{hardigan2015taxonomy}.

### 2018-06-27 paper.tex
Table1

--- 2018-06-27 paper.tex
@@
 \rowcolor{white}
-Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013, suarez2016} \\ 
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013} \\ 
 \rowcolor{white}
@@
 \rowcolor{white}
-PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{pugach2016complex, moreno2014genetics} \\
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{moreno2014genetics} \\
 \rowcolor{white}
@@
 \rowcolor{gray!25}
-fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{malinsky2015, zhang2016genome} \\ 
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{zhang2016genome} \\ 
 \rowcolor{gray!25}
-five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015, pease2016} \\
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015} \\
 

### 2018-06-29 paper.tex
Update paper.tex

@@
 	\centering
-	\includegraphics[width=15cm]{temperature_plot.png}
+	\includegraphics[width=15cm]{temperature_plot_degC.pdf}
 	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature. Approximate domestication center for each crop is denoted by a triangle}
@@
 
-Barley (\emph{Hordeum vulgare} subsp.
-\emph{vulgare}) was domesticated at least twice roughly 8,000 to 10,000 BP: once from the wild subsp.
-\emph{spontaneum} in the Fertile Crescent and once from subsp.
+Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was domesticated from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
+Modern barley is most likely the result of multiple domestication events across the Middle East and Asia \cite{azhaguvel2007phylogenetic}.
+There is clear evidence of one domestication event of barley from the wild subsp.
+\emph{spontaneum} in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern center of barley domestication from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} in Tibet \cite{badr2000origin, Morrell2007, dai2012tibet}.
+\emph{agriocrithon} \cite{Morrell2007}, possibly Tibet \cite{dai2012tibet}.
+However, recent research from Pourkheirandish and colleagues \cite{pourkheirandish2018elucidation} suggests that the putative wild Tibetan progenitor \emph{agriocrithon} is a misidentified intercultivar hybrid of domesticated 6-row landraces, which they deem \emph{pseudo-agriocrithon}, as opposed to \emph{eu-agriocrithon}, which are six-rowed wild relatives derived from two-rowed \emph{ssp.
+spontaneum}.
+The non-brittle rachis mutations (\emph{btr1} and \emph{btr2}) appear to have arisen in populations from the Levant, whereas the two most common of four alleles conferring 6-rowed spike morphology (\emph{vrs1.a1} and \emph{vrs1.a4}) originated in \emph{eu-agriocrithon} populations in Central Asia (Turkmenistan and Uzbekistan, respectively).
+These two key barley domestication traits were combined when \emph{vrs1.a1} and \emph{vrs1.a4} introgressed from \emph{eu-agriocrithon} into cultivated barleys in the Levant.
+
 Presently, the distribution of subsp.
@@
 Limited admixture linkage disequilibrium and small tracts of identity by state suggest substantial recombination has occurred since initial crop-wild hybridization and that even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley post-domestication.
-While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in domesticated landraces.
+While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
 
-\item{Asian Rice:}
+%"The best current hypothesis holds that both vrs1.a1 and vrs1.a4 were introgressed from eu-agriocrithon into other wild barley genotypes and then outcrossed to cultivated barley (Figure 1)"
 
-The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
-Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
-For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
-In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of \emph{aromatic} rice.
-A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
+%"The proposal here is that vrs1.a1 and vrs1.a4 emerged in agriocrithon before they were introgressed into ssp. vulgare (Figure 1)."
 
-The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
-According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
-As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
-While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
+%"The origin of two alleles (vrs1.a2 and a3) was previously shown (Komatsuda et al., 2007) to involve single mutations at the vrs1 locus that changed the domesticated two-rowed to six-rowed phenotype (Figure 1)."
+
+%"Both vrs1.a1 and vrs1.a4 arose in Central Asia, some 2500 km distant from the site of emergence of non-brittle rachis.
+The occurrence of six-rowed types among wild barley implies that the six-rowed allele pre-dates domestication, when early farmers still relied on ssp.
+spontaneum.
+This time line agrees, in part at least, with the notion that six-rowed cultivated barley was derived from agriocrithon (Aberg, 1938).
+The non-brittle rachis trait represents a strong genetic bottleneck, and its geographical origin lies far from that of eu-agriocrithon, clearly implying that eu-agriocrithon could not have been the immediate ancestor of the non-brittle barleys that were donors of vrs1.a1 and vrs1.a4 to cultivated barley."
 

### 2018-06-29 paper.tex
starting edits to barley

@@
 
-Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was domesticated from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
-Modern barley is most likely the result of multiple domestication events across the Middle East and Asia \cite{azhaguvel2007phylogenetic}.
-There is clear evidence of one domestication event of barley from the wild subsp.
-\emph{spontaneum} in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern center of barley domestication from subsp.
+Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
+There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern domestication event from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} \cite{Morrell2007}, possibly Tibet \cite{dai2012tibet}.
-However, recent research from Pourkheirandish and colleagues \cite{pourkheirandish2018elucidation} suggests that the putative wild Tibetan progenitor \emph{agriocrithon} is a misidentified intercultivar hybrid of domesticated 6-row landraces, which they deem \emph{pseudo-agriocrithon}, as opposed to \emph{eu-agriocrithon}, which are six-rowed wild relatives derived from two-rowed \emph{ssp.
+\emph{agriocrithon} \cite{Morrell2007}, potentially in modern-day Tibet \cite{dai2012tibet}.
+However, recent research \cite{pourkheirandish2018elucidation} has suggested that var.
+\emph{agriocrithon} is a misidentified intercultivar hybrid of domesticated 6-row landraces, which they deem \emph{pseudo-agriocrithon}, as opposed to \emph{eu-agriocrithon}, which are six-rowed wild relatives derived from two-rowed \emph{ssp.
 spontaneum}.

### 2018-06-29 paper.tex
barley section update

@@
 Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
-There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern domestication event from subsp.
+There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern domestication event, potentially from \emph{subsp.
+spontaneum} east of the Zargos mountains \cite{Morrell2007} or from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} \cite{Morrell2007}, potentially in modern-day Tibet \cite{dai2012tibet}.
-However, recent research \cite{pourkheirandish2018elucidation} has suggested that var.
+\emph{agriocrithon} in modern-day Tibet \cite{dai2012tibet}.
+However, recent research \cite{pourkheirandish2018elucidation} suggests that Tibetan \emph{agriocrithon} is not a wild relative, but rather an intercultivar hybrid of domesticated landraces, casting doubt on the Tibetan domestication event from \emph{agriocrithon}.
+%However, recent research \cite{pourkheirandish2018elucidation} has suggested that var.
 \emph{agriocrithon} is a misidentified intercultivar hybrid of domesticated 6-row landraces, which they deem \emph{pseudo-agriocrithon}, as opposed to \emph{eu-agriocrithon}, which are six-rowed wild relatives derived from two-rowed \emph{ssp.
 spontaneum}.
-The non-brittle rachis mutations (\emph{btr1} and \emph{btr2}) appear to have arisen in populations from the Levant, whereas the two most common of four alleles conferring 6-rowed spike morphology (\emph{vrs1.a1} and \emph{vrs1.a4}) originated in \emph{eu-agriocrithon} populations in Central Asia (Turkmenistan and Uzbekistan, respectively).
-These two key barley domestication traits were combined when \emph{vrs1.a1} and \emph{vrs1.a4} introgressed from \emph{eu-agriocrithon} into cultivated barleys in the Levant.
+%The non-brittle rachis mutations (\emph{btr1} and \emph{btr2}) appear to have arisen in populations from the Levant, whereas the two most common of four alleles conferring 6-rowed spike morphology (\emph{vrs1.a1} and \emph{vrs1.a4}) originated in \emph{eu-agriocrithon} populations in Central Asia (Turkmenistan and Uzbekistan, respectively).
+%These two key barley domestication traits were combined when \emph{vrs1.a1} and \emph{vrs1.a4} introgressed from \emph{eu-agriocrithon} into cultivated barleys in the Levant.
 
@@
 Cultivated barley is found throughout much of the \emph{spontaneum} distribution and barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
+
+Two key barley domestication traits are non-brittle rachis and 6-row spike morphology.
+Haplotype analysis by Pourkheirandish and colleagues \cite{pourkheirandish2018elucidation} indicates that these two traits arose independently and were combined into domesticated barley when 6-row alleles (\emph{vrs1.a1} and \emph{vrs1.a4}) from \emph{spontaneum} introgressed into cultivated barleys from the Levant which harbored alleles for tough rachis (\emph{btr1} and \emph{btr2}).
 

### 2018-07-03 paper.tex
edits throughout

--- 2018-06-29 paper.tex
@@
 
-\title{Review: The Prevalence of Adaptive Introgression in Crop Histories}
+\title{Review: The Extent of Adaptive Wild Introgression in Crops}
 
@@
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{Meyer2016, Wang2017, Fuller2014}.
-An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following initial domestication.
+An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following domestication.
 It is this adaptive introgression that is the subject of this review.
@@
 The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
-However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
-Emerging results from these studies suggest there is a need to expand our conception of domestication to encompass the broadening of the genetic base of crops that occurred through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.
+Recently developed tools and methods have been applied to high-density marker data to detect genome-wide patterns of introgression, granting novel insight into the prevalence of adaptive introgression in crop histories.
+Results from these studies suggest there is a need to expand our conception of domestication to encompass the broadening of the genetic base of crops that occurred through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.
 
-In this review, we: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe some key questions regarding crop adaptation that is mediated by gene flow from wild relatives.
+In this review, we: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe key questions regarding crop adaptation through gene flow from wild relatives.
 
@@
 
-The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of domesticated species.
-The methods reviewed here do not include those marginally estimating introgression\slash migration rate as a component of demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
-Rather, we focus on three broad categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of species.
+The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
+Rather, we focus on three categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 
 First, introgressed segments are expected to show low differentiation from their source population.
-The $F_{st}$ and $d_{XY}$ statistics and their derivates including $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
-The former two statistics are insensitive to rare migrants and therefore lack power to detect recent introgression, while the latter two overcome this limitation.
+The $F_{st}$ and $d_{XY}$ statistics and derivates of $d_{XY}$ including $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
+The former two statistics are insensitive to rare migrants in a population and therefore lack power to detect very recent introgression, while the latter two overcome this limitation.
 Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
@@
 However, accuracy of introgression estimates is improved with more comprehensive data.
+Phased data are also not a prerequisite for differentiation-based methods.
 
-Second, local ancestry deconvolution assigns genomic regions to various source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}. 
+Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to various source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}.
 One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
 Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
-And finally, introgression can be detected through ancestry deconvolution by using a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+And, finally, a third version uses a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
 Ancestry deconvolution methods are better suited to high-density marker or whole genome sequencing data given the intent to assign ancestry genome-wide.
+Many such methods also require accurate phasing of haplotypes.
 
@@
 
-Application of these approaches, in combination with population genetic statistics to detect selection, suggest that introgression can play an adaptive role.
+Collectively, introgression detection methods have been applied across a wide range of systems, identifying several instances of adaptive introgression.
 For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
 Likewise, the HAPMIX ancestry deconvolution method was applied by Jeong et al.
-\cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation (reviewed in \cite{Racimo2015}).
+\cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
 Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
+Below we describe the nascent application of these methods to crop systems as well as implications for the study of domestication and adaptation.
 
@@
 
-As comprehensive genetic analyses of crops and their wild relatives have become feasible across their geographic ranges, evidence for substantial wild-to-crop introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
-Below we present a summary of findings from maize, barley, rice, and potato, four systems in which introgression from wild relatives appears to have played an adaptive role.
+As comprehensive genetic analyses of crops and their wild relatives have become feasible across their geographic ranges, evidence for substantial crop-wild introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
+Below we present a summary of findings from maize, barley, rice, and potato, four systems in which crop-wild gene flow appears to have played an adaptive role.
 All four of these crops were domesticated in defined centers and have subsequently expanded to much broader distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
+Below we describe for each case study what is known about a crop's domestication history and the prevalence of adaptive introgression during post-domestication  expansion.
 
@@
 Wang and co-authors \cite{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
-Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-Maize from the lowlands of northern South America appears to have been derived from Mesoamerican lowland maize and high elevation adaptation in Andean maize likely occurred \emph{de novo} from this genetic base.
+Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but that more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Maize from the lowlands of northern South America appears to have been derived from Mesoamerican lowland maize and high elevation adaptation in Andean maize likely occurred \emph{de novo} from this genetic base \cite{Takuno2015}.
 
@@
 Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
-There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and many have supported an additional eastern domestication event from subsp.
+There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from subsp.
+\emph{spontaneum} east of the Zagros Mountains \cite{Morrell2007} or from subsp.
 \emph{spontaneum} var.
-\emph{agriocrithon} \cite{Morrell2007}, potentially in modern-day Tibet \cite{dai2012tibet}.
-However, recent research \cite{pourkheirandish2018elucidation} has suggested that var.
-\emph{agriocrithon} is a misidentified intercultivar hybrid of domesticated 6-row landraces, which they deem \emph{pseudo-agriocrithon}, as opposed to \emph{eu-agriocrithon}, which are six-rowed wild relatives derived from two-rowed \emph{ssp.
-spontaneum}.
-The non-brittle rachis mutations (\emph{btr1} and \emph{btr2}) appear to have arisen in populations from the Levant, whereas the two most common of four alleles conferring 6-rowed spike morphology (\emph{vrs1.a1} and \emph{vrs1.a4}) originated in \emph{eu-agriocrithon} populations in Central Asia (Turkmenistan and Uzbekistan, respectively).
-These two key barley domestication traits were combined when \emph{vrs1.a1} and \emph{vrs1.a4} introgressed from \emph{eu-agriocrithon} into cultivated barleys in the Levant.
-
-Presently, the distribution of subsp.
-\emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
-Cultivated barley is found throughout much of the \emph{spontaneum} distribution and barley-\emph{spontaneum} hybrids are fertile and commonly found when these taxa co-occur.
+\emph{agriocrithon} in modern-day Tibet \cite{dai2012tibet}.
+However, recent research casts doubt on Tibetan domestication and suggests that var.
+\emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \cite{pourkheirandish2018elucidation}.
+Presently, the distribution of wild barley stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
+Cultivated barley is found throughout much of wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
 
@@
 This study identified several lines of evidence consistent with wild introgression aiding the expansion and adaptation of domesticated barley.
-Using ancestry deconvolution methods, the authors identified genomic regions of shared ancestry linking particular landraces to numerous wild relative populations.
+The authors utilized ancestry deconvolution methods to identify genomic regions of shared ancestry, which linked particular landraces to numerous wild relative populations.
 These results suggested landraces may have received wild introgression on a continual basis during post-domestication expansion.
@@
 While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
-
-%"The best current hypothesis holds that both vrs1.a1 and vrs1.a4 were introgressed from eu-agriocrithon into other wild barley genotypes and then outcrossed to cultivated barley (Figure 1)"
-
-%"The proposal here is that vrs1.a1 and vrs1.a4 emerged in agriocrithon before they were introgressed into ssp. vulgare (Figure 1)."
-
-%"The origin of two alleles (vrs1.a2 and a3) was previously shown (Komatsuda et al., 2007) to involve single mutations at the vrs1 locus that changed the domesticated two-rowed to six-rowed phenotype (Figure 1)."
-
-%"Both vrs1.a1 and vrs1.a4 arose in Central Asia, some 2500 km distant from the site of emergence of non-brittle rachis.
-The occurrence of six-rowed types among wild barley implies that the six-rowed allele pre-dates domestication, when early farmers still relied on ssp.
-spontaneum.
-This time line agrees, in part at least, with the notion that six-rowed cultivated barley was derived from agriocrithon (Aberg, 1938).
-The non-brittle rachis trait represents a strong genetic bottleneck, and its geographical origin lies far from that of eu-agriocrithon, clearly implying that eu-agriocrithon could not have been the immediate ancestor of the non-brittle barleys that were donors of vrs1.a1 and vrs1.a4 to cultivated barley."
 
@@
 
-\subsection*{Where and from what taxon did a crop originate?}
+\subsection*{Where and from what taxa did a crop originate?}
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's origin may be complicated or confounded entirely.

### 2018-07-03 paper.tex
figure2

--- 2018-06-29 paper.tex
@@
 	\centering
-	\includegraphics[width=12cm]{./FigsAndTables/Figure1new}
-	\caption{Hapmix vectors indicating introgression level from \emph{mexicana} to Mexican highland maize on chromosome 4.
+	\includegraphics[width=12cm]{./FigsAndTables/Figure1new2}
+	\caption{Evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize on chromosome 4.
+A.
+Stacked bar plots of the HAPMIX introgression scan.
 Data adopted from Hufford et al.
@@
 ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
-Insets show the phenotypic differences between lowland (red) and highland (blue) maize stems.
-The black bar indicates one of the QTLs for macrohairs and pigment density in Lauter et al.
-\cite{lauter2004}.}
+The dashed vertical lines indicate one of the QTLs for macrohairs and pigment density in Lauter et al.
+\cite{lauter2004}.}.
+B.
+Phenotypic differences between lowland (red) and highland (blue) maize stems.
+
 	\label{fig:introgressionMaize}

### 2018-07-05 paper.tex
figure2
continued edits

--- 2018-07-03 paper.tex
@@
 Limited admixture linkage disequilibrium and small tracts of identity by state suggest substantial recombination has occurred since initial crop-wild hybridization and that even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley post-domestication.
-While these results are consistent with adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
+While these patterns suggest the possibility of adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
 

### 2018-07-05 paper.tex

--- 2018-07-03 paper.tex
@@
 The dashed vertical lines indicate one of the QTLs for macrohairs and pigment density in Lauter et al.
-\cite{lauter2004}.}.
+\cite{lauter2004}.
 B.
 Phenotypic differences between lowland (red) and highland (blue) maize stems.
-
-	\label{fig:introgressionMaize}
+\label{fig:introgressionMaize}}.
 \end{figure}

### 2018-07-05 paper.tex
adding back rice section

@@
 
-\item{Potato}
+\item{Asian Rice:}
+
+The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
+For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
+In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
+
+The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
+According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
+As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
+While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
+
+\item{Potato:}
 

### 2018-07-05 paper.tex
edits throughout

@@
 Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
-For example, Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
+For example, Huang and colleagues \cite{Huang2012} measured genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was likely domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} potentially resulted from hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
 In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
-A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varieties.
 
-The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
+The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving} that also stresses the potential adaptive significance of crop-wild gene flow in rice.
 According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
@@
 The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \cite{hardigan2017genome}.
-These authors discovered that both diploid and tetraploid domesticates had received extensive introgression from more than ten species of wild \emph{Solanum}, with a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
+These authors discovered that, in particular, tetraploid domesticates had received extensive introgression from wild \emph{Solanum}, with a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
 In certain cultivars, wild ancestry was estimated at upwards of 30\%.
@@
 
-The four crop systems described in detail here represent particularly compelling examples of putatively adaptive wild introgression.
+The four crop systems described in detail here represent particularly compelling examples of wild introgression that was potentially adaptive.
 However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
-Across these four cases studies, potential generalities can be observed.
-Data thus far indicate that wild introgression is often local or regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S.
+Across these four cases studies, some generalities can be observed.
+Data thus far indicate that wild introgression is often regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S.
 highlands), newly introgressed wild haplotypes can be disseminated more broadly.
 Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
-Beyond these observations associated with post-domestication adaptation, pervasive wild-to-crop gene flow is also relevant to the study of domestication itself.
-
-%\centering
-%\begin{adjustbox}{width=1\textwidth}
-%\small
-%\label{my-label}
-%\begin{tabular}{|p{5cm}|p{5cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|l|}
-%\hline
-%Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
-%Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\ 
-%\hline 
-%Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
-%\hline
-%Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
-%Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\ 
-%\hline
-%Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\ 
-%\hline
-%Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome, johns1986ongoing, gavrilenko2013genetic} \\
-%\hline
-%Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
-%\hline
-%Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
-%\hline
-%Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\ 
-%\hline
-%Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
-%\hline
-%Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
-%\hline
-%Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
-%\hline
-%Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
-%\hline
-%Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
-%\hline
-%\end{tabular}
-%\end{adjustbox}
-%\end{table}
+The possibility of both substantial and common wild gene flow raises a number of questions regarding both domestication and subsequent adaptation of crops.
 
@@
 In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxa.
-This phenomenon, in combination with divergence of modern crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
+This phenomenon, in combination with divergence of modern crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and genetic structure among wild relative populations, may help explain discrepancies between domestication dates based on genetic and archaeological data.
 More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.
@@
 Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
-Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands.
+Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
@@
 
-In summary, should post-domestication gene flow with wild relatives be pervasive during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.
+In summary, since post-domestication gene flow with wild relatives appears frequent during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.
 
@@
 
-Research has so far shown that adaptive crop-wild introgression has played a significant role during domestication and dispersal of many important crops.
-However, the scope and dynamics of this process are not yet fully described and remain unexplored in many systems.
+While research in a subset of crops suggests adaptive wild introgression likely occurred, the scope and dynamics of this process are not yet fully described and remain unexplored in many systems.
 In determining the extent and nature of adaptation due to introgression, several questions should be considered:
@@
 
-\subsection*{At what taxonomic scale does introgression occur?}
-As species become substantially diverged, introgression can become maladaptive or impossible due to, for example, Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
+\subsection*{From which wild taxa will introgression occur?}
+As species become substantially diverged, introgression can become maladaptive or impossible due to Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
 Divergence time may therefore be a useful predictor of the possibility of gene flow between a particular wild relative and a domesticate.
-Hybridization may also be limited between a crop and a particular wild relative due to demography. 
+Hybridization may also be limited between a crop and a particular wild relative due to genetic load. 
 For instance, gene flow from a wild relative with a small long-term effective population size, and correspondingly high genetic load may not be favored by selection.

### 2018-07-05 paper.tex
edits to conclusions

@@
 
-As recognized early on by Darwin, crop species represent promising model systems for evolutionary study. 
-The application of genomic tools has generated mounting evidence of crop-wild gene flow during and well beyond the initial stages of domestication.
-The process of crop range expansion and adaptation through gene flow is closely linked to domestication and offers new questions and opportunities for practical and theoretical investigation.
-Domesticated crops are among humanity's first and greatest inventions.
-Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agronomic Revolution.
+Recent innovations in both high-density marker data and methods for characterizing genome-wide patterns of introgression have helped reveal the extent and timing of gene flow in a number of species.
+Application of these data and techniques has led to mounting evidence of crop-wild gene flow following initial domestication in several crops.
+Substantial post-domestication gene flow with wild relatives can affect inferences made regarding domestication and may be an important mechanism through which crops were able to adapt to novel conditions during global expansions.
+An accurate understanding of the extent of gene flow is therefore important to both the basic study of crop evolution and to the identification of adaptive alleles for continued crop improvement.
+
+Domesticated crops are among humanity's greatest inventions.
+Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agricultural Revolution.
 In a more nuanced and dynamic understanding of crop domestication that includes sustained gene flow with wild relatives, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.

### 2018-07-06 paper.tex
edits to Conclusions

@@
 
-Collectively, introgression detection methods have been applied across a wide range of systems, identifying several instances of adaptive introgression.
+Collectively, introgression detection methods have been applied across a wide range of systems, identifying several instances of adaptive introgression (see applications in \textbf{Table 1}).
 For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
@@
 All four of these crops were domesticated in defined centers and have subsequently expanded to much broader distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
-Below we describe for each case study what is known about a crop's domestication history and the prevalence of adaptive introgression during post-domestication  expansion.
+For each case study below we describe what is known about both a crop's domestication history and the prevalence of adaptive introgression during post-domestication  expansion.
 
@@
 An accurate understanding of the extent of gene flow is therefore important to both the basic study of crop evolution and to the identification of adaptive alleles for continued crop improvement.
+While multiple studies in crop systems have identified wild introgression, fewer have effectively linked introgressed alleles to adaptation.
+More comprehensive functional analyses and field evaluation will be critical for understanding the evolutionary and adaptive significance of introgression in crops.
 

### 2018-07-06 paper.tex
updated table
edits to formatting
edits for brevity

--- 2018-07-05 paper.tex
@@
 \usepackage[x11names,dvipsnames,table]{xcolor}
+\usepackage{lineno}
+\linenumbers
 \usepackage{array}
@@
 \bibliographystyle{ieeetr}
+\renewcommand{\baselinestretch}{1.5}
 \newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
@@
 \affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
-\affil[*]{Correspondence: mhufford@iastate.edu (M.B. Hufford)}
+\affil[*]{Correspondence: email: mhufford@iastate.edu, telephone: 1-515-294-8511}
 \date{}
@@
 
-Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
+Number of Figures: 2, both in color
+
+Number of Tables: 2
+
+Word Count:
+
+\clearpage
+
+\section*{Summary}
+The study of crop evolution has focused primarily on the process of initial domestication.
+Post-domestication adaptation during the expansion of crops from their centers of origin has received considerably less attention.
+Recent research has revealed that, in at least some instances, crops have continued a close relationship with their wild relatives that has facilitated adaptation to novel conditions encountered during expansion.
+Such adaptive introgression could bear importantly on the basic study of domestication, affecting estimates of several evolutionary processes of interest (\emph{e.g.}, the strength of the domestication bottleneck, the timing of domestication, the targets of selection during domestication).
+Identification of haplotypes introgressed from the wild may also aid the identification of alleles that are beneficial under particular environmental conditions.
+Here we review mounting evidence for substantial adaptive wild introgression in several crops and consider the implications of such gene flow to our understanding of crop histories.
+
+\textbf{Key Words:} adaptation, domestication, gene flow, introgression, wild relatives
+
+\section*{Introduction}
+
+Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
 However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{Meyer2016, Wang2017, Fuller2014}.
-An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following domestication.
+An additional aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following domestication.
 It is this adaptive introgression that is the subject of this review.
@@
 In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
-To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
-The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
-Recently developed tools and methods have been applied to high-density marker data to detect genome-wide patterns of introgression, granting novel insight into the prevalence of adaptive introgression in crop histories.
+To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desirable traits from wild relatives (for a review, \cite{Dempewolf2017}).
+The potential for adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
+Recently developed methods have been applied to high-density marker data to detect genome-wide patterns of introgression, granting novel insight into the prevalence of adaptive introgression in crop histories.
 Results from these studies suggest there is a need to expand our conception of domestication to encompass the broadening of the genetic base of crops that occurred through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.
 
-In this review, we: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe key questions regarding crop adaptation through gene flow from wild relatives.
+In this review, we: 1) briefly describe recent methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe key questions regarding crop adaptation through gene flow from wild relatives.
 
@@
 
-The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of species.
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated comprehensive study of introgression across a broad spectrum of species.
 The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
@@
 The former two statistics are insensitive to rare migrants in a population and therefore lack power to detect very recent introgression, while the latter two overcome this limitation.
-Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
 These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
@@
 
-Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to various source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}.
+Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}.
 One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
 Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
-And, finally, a third version uses a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
-Ancestry deconvolution methods are better suited to high-density marker or whole genome sequencing data given the intent to assign ancestry genome-wide.
+A third version uses a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+Ancestry deconvolution methods are better suited to high-density marker data given the intent to assign ancestry genome-wide.
 Many such methods also require accurate phasing of haplotypes.
 
-Phylogenetic relationships are evaluated and applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics.
+Phylogenetic relationships are applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics.
 These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
 While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and $D_{\textrm{FOIL}}$ tests \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
-The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter can identify donor and recipient lineages of introgression in a more complex, five-taxon phylogeny.
+The former is quite similar to the D-statistic but is based on allele frequencies, and the latter can identify donor and recipient lineages of introgression in a more complex, five-taxon phylogeny.
 Like differentiation methods, phylogeny-based detection of introgression can be employed using low-density data.
-However, because these methods require knowledge of whether an allele is ancestral or derived, data from a sufficiently diverged outgroup must also be included in analyses.
+However, because these methods require knowledge of whether an allele is ancestral or derived, data from a sufficiently diverged outgroup must be available.
 
-Collectively, introgression detection methods have been applied across a wide range of systems, identifying several instances of adaptive introgression.
-For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
+Collectively, introgression detection methods have been applied to several systems, frequently identifying instances of adaptive introgression (see applications in \textbf{Table 1}).
+For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in adaptation to pollinator preference and speciation \cite{Stankowski2015}.
 Likewise, the HAPMIX ancestry deconvolution method was applied by Jeong et al.
-\cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
+\cite{jeong2014} to detect introgression from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
 Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
@@
 
-As comprehensive genetic analyses of crops and their wild relatives have become feasible across their geographic ranges, evidence for substantial crop-wild introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
+As range-wide genetic analyses of crops and their wild relatives have become feasible, evidence for substantial crop-wild introgression has been discovered in many important crops (\textbf{Table 2}).
 Below we present a summary of findings from maize, barley, rice, and potato, four systems in which crop-wild gene flow appears to have played an adaptive role.
-All four of these crops were domesticated in defined centers and have subsequently expanded to much broader distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
-Below we describe for each case study what is known about a crop's domestication history and the prevalence of adaptive introgression during post-domestication  expansion.
+All four of these crops were domesticated in defined centers and have subsequently expanded to global distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
+For each case study we describe both what is known about a crop's domestication history and the prevalence of adaptive introgression during expansion.
 
@@
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
-Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
+Introgression from \emph{mexicana} to maize has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
 However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
-These introgressed segments overlapped with QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
-In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
+These introgressed segments overlapped with QTL that had previously been shown to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression had increased height (a proxy for fitness) under highland environmental conditions.
 Height differences were not detected under lowland conditions, providing further evidence of local adaptation.
@@
 Wang and co-authors \cite{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
-Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but that more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-Maize from the lowlands of northern South America appears to have been derived from Mesoamerican lowland maize and high elevation adaptation in Andean maize likely occurred \emph{de novo} from this genetic base \cite{Takuno2015}.
+Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high-elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but that more distant high-elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+These findings are consistent with previous work suggesting high elevation adaptation in Andean maize likely occurred \emph{de novo} \cite{Takuno2015}.
 
@@
 Stacked bar plots of the HAPMIX introgression scan.
-Data adopted from Hufford et al.
+Data obtained from Hufford et al.
 \cite{Hufford2013}.
 ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
-The dashed vertical lines indicate one of the QTLs for macrohairs and pigment density in Lauter et al.
+The dashed vertical lines indicate a previously identified QTL for macrohairs and pigment density in Lauter et al.
 \cite{lauter2004}.
@@
 
-Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
-There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from subsp.
-\emph{spontaneum} east of the Zagros Mountains \cite{Morrell2007} or from subsp.
+Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild ssp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
+There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from ssp.
+\emph{spontaneum} east of the Zagros Mountains \cite{Morrell2007} or from ssp.
 \emph{spontaneum} var.
@@
 \emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \cite{pourkheirandish2018elucidation}.
-Presently, the distribution of wild barley stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
-Cultivated barley is found throughout much of wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
+Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
+Cultivated barley is found throughout wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
 
@@
 The authors utilized ancestry deconvolution methods to identify genomic regions of shared ancestry, which linked particular landraces to numerous wild relative populations.
-These results suggested landraces may have received wild introgression on a continual basis during post-domestication expansion.
+These results suggest landraces may have received wild introgression on a continual basis during post-domestication expansion.
 However, barley landraces also showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
 Limited admixture linkage disequilibrium and small tracts of identity by state suggest substantial recombination has occurred since initial crop-wild hybridization and that even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley post-domestication.
-While these patterns suggest the possibility of adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
+While these patterns suggest the possibility of adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces.
 
@@
 The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon}, hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
 Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
@@
 In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
-A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varieties.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and concluded that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and were introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varietals.
 
-The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving} that also stresses the potential adaptive significance of crop-wild gene flow in rice.
+The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving} that stresses the potential adaptive significance of crop-wild gene flow in rice.
 According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
-As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
+As these domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
 While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
@@
 The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \cite{hardigan2017genome}.
-These authors discovered that, in particular, tetraploid domesticates had received extensive introgression from wild \emph{Solanum}, with a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
+These authors discovered that tetraploid domesticates in particular had received extensive introgression from wild \emph{Solanum}, documenting a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
 In certain cultivars, wild ancestry was estimated at upwards of 30\%.
@@
 Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
-The possibility of both substantial and common wild gene flow raises a number of questions regarding both domestication and subsequent adaptation of crops.
+The possibility of both substantial and common wild gene flow raises a number of questions regarding both domestication and adaptation of crops.
 
@@
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
-However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
+However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
 \vspace{5mm}
@@
 
-While research in a subset of crops suggests adaptive wild introgression likely occurred, the scope and dynamics of this process are not yet fully described and remain unexplored in many systems.
+While research in a subset of crops suggests adaptive wild introgression likely occurred, the scope and dynamics of this process remain poorly described or unexplored in many systems.
 In determining the extent and nature of adaptation due to introgression, several questions should be considered:
@@
 For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).
-On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests much more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
+On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
 Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.
@@
 As species become substantially diverged, introgression can become maladaptive or impossible due to Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
-Divergence time may therefore be a useful predictor of the possibility of gene flow between a particular wild relative and a domesticate.
+Divergence time may therefore be a useful predictor of the possibility of gene flow between taxa.
 Hybridization may also be limited between a crop and a particular wild relative due to genetic load. 
 For instance, gene flow from a wild relative with a small long-term effective population size, and correspondingly high genetic load may not be favored by selection.
-This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited in extent and relegated largely to non-genic regions due to the high genetic load found within Neanderthal donor individuals \cite{harris2016genetic}.
+This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited and relegated primarily to non-genic regions due to the high genetic load found within Neanderthal donor individuals \cite{harris2016genetic}.
 
@@
 As described above, loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
-These loci are potentially beneficial targets for crop improvement and their accurate identification is crucial.
 Furthermore, adaptive introgression that is demonstrably tied to a specific environment represents a promising source of beneficial alleles that can be directly utilized in breeding to adapt crops to similar conditions.
@@
 Recent innovations in both high-density marker data and methods for characterizing genome-wide patterns of introgression have helped reveal the extent and timing of gene flow in a number of species.
-Application of these data and techniques has led to mounting evidence of crop-wild gene flow following initial domestication in several crops.
-Substantial post-domestication gene flow with wild relatives can affect inferences made regarding domestication and may be an important mechanism through which crops were able to adapt to novel conditions during global expansions.
+Application of these data and techniques has led to mounting evidence of crop-wild gene flow following initial domestication in several species.
+Substantial post-domestication gene flow with wild relatives can affect inferences made regarding domestication and may be an important mechanism through which crops adapted to novel conditions during global expansion.
 An accurate understanding of the extent of gene flow is therefore important to both the basic study of crop evolution and to the identification of adaptive alleles for continued crop improvement.
+While some studies in crop systems have identified wild introgression, even fewer have effectively linked introgressed alleles to adaptation.
+More comprehensive functional analyses and field evaluation will be critical for understanding the evolutionary and adaptive significance of introgression in crops.
 
-Domesticated crops are among humanity's greatest inventions.
-Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agricultural Revolution.
-In a more nuanced and dynamic understanding of crop domestication that includes sustained gene flow with wild relatives, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.
+\section*{Acknowledgements}
+The authors thank Daniel Gates, Peter Morrell, Jeffrey Ross-Ibarra and Jonathan Wendel for comments on a previous version of this manuscript.
+This work was supported by funding from the National Science Foundation Plant Genome Research Program (IOS-1546719).
+
+\bibliography{bib_gj.bib}
 
@@
 
-\bibliography{bib_gj.bib}
-
 \end{document}

### 2018-07-06 paper.tex

@@
 \end{document}
+=======
+\documentclass[11pt]{article}
+\usepackage{color}
+\usepackage{authblk}%allows footnote format for authors
+\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
+\usepackage[numbers,sort]{natbib}
+\usepackage{amsmath}
+\usepackage{rotating}
+\usepackage{adjustbox}
+\usepackage[english]{babel}
+\usepackage{colortbl}
+\usepackage{booktabs}
+\usepackage{tabularx}
+\usepackage[x11names,dvipsnames,table]{xcolor}
+\usepackage{array}
+\newcolumntype{G}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\bibliographystyle{ieeetr}
+\newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
+\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
+\newcommand{\gmj}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments
+
+\title{Review: The Extent of Adaptive Wild Introgression in Crops}
+
+\author[1]{Authors: Garrett M. Janzen}%author information
+\author[1]{Li Wang}
+\author[1,*]{Matthew B. Hufford}
+\affil[1]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
+\affil[*]{Correspondence: mhufford@iastate.edu (M.B. Hufford)}
+\date{}
+
+\begin{document}
+
+\maketitle
+
+Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{Meyer2016, Wang2017, Fuller2014}.
+An additional important aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following domestication.
+It is this adaptive introgression that is the subject of this review.
+
+Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role}.
+In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
+To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
+The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
+Recently developed tools and methods have been applied to high-density marker data to detect genome-wide patterns of introgression, granting novel insight into the prevalence of adaptive introgression in crop histories.
+Results from these studies suggest there is a need to expand our conception of domestication to encompass the broadening of the genetic base of crops that occurred through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.
+
+In this review, we: 1) briefly describe recently developed methods for detecting introgression, 2) present case studies suggesting wild-to-crop introgression has conferred local adaptation, 3) consider how introgression bears upon fundamental questions of domestication, and 4) describe key questions regarding crop adaptation through gene flow from wild relatives.
+
+\section*{Introgression methods and their application}
+
+The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated the comprehensive study of introgression across a broad spectrum of species.
+The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
+Rather, we focus on three categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+
+First, introgressed segments are expected to show low differentiation from their source population.
+The $F_{st}$ and $d_{XY}$ statistics and derivates of $d_{XY}$ including $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
+The former two statistics are insensitive to rare migrants in a population and therefore lack power to detect very recent introgression, while the latter two overcome this limitation.
+Additionally, $RND_{min}$ accounts for variable mutation rate, which is detected based on branch length to an outgroup \cite{rosenzweig2016}.
+These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
+For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes the number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
+A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \cite{racimo2016}.
+Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
+Since differentiation-based methods can be calculated site-by-site, high-density, genome-wide data are not necessarily required.
+However, accuracy of introgression estimates is improved with more comprehensive data.
+Phased data are also not a prerequisite for differentiation-based methods.
+
+Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to various source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}.
+One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
+Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
+And, finally, a third version uses a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+Ancestry deconvolution methods are better suited to high-density marker or whole genome sequencing data given the intent to assign ancestry genome-wide.
+Many such methods also require accurate phasing of haplotypes.
+
+Phylogenetic relationships are evaluated and applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics.
+These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
+While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and $D_{\textrm{FOIL}}$ tests \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
+The former is quite similar to the D-statistic but uses allele frequencies from each population/species, and the latter can identify donor and recipient lineages of introgression in a more complex, five-taxon phylogeny.
+Like differentiation methods, phylogeny-based detection of introgression can be employed using low-density data.
+However, because these methods require knowledge of whether an allele is ancestral or derived, data from a sufficiently diverged outgroup must also be included in analyses.
+
+Collectively, introgression detection methods have been applied across a wide range of systems, identifying several instances of adaptive introgression (see applications in \textbf{Table 1}).
+For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in both adaptation to pollinator preference and speciation \cite{Stankowski2015}.
+Likewise, the HAPMIX ancestry deconvolution method was applied by Jeong et al.
+\cite{jeong2014} to detect introgression in human populations from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
+Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
+Below we describe the nascent application of these methods to crop systems as well as implications for the study of domestication and adaptation.
+
+\begin{table}[h]
+
+\rowcolors{2}{white}{gray!25}
+\begin{center}
+\caption{List of recently developed methods for detecting introgression and examples of their use in empirical studies.} \label{tab:tools}
+\begin{tabularx}{\textwidth}{llll}
+\\\toprule  
+\rowcolor{white}
+{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
+
+\rowcolor{gray!25}
+{\emph{\bf Divergence}} &   &   &   \\
+\rowcolor{gray!25}
+Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
+\rowcolor{gray!25}
+RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
+\rowcolor{gray!25}
+$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
+
+\rowcolor{white}
+{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
+\rowcolor{white}
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013} \\ 
+\rowcolor{white}
+RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
+\rowcolor{white}
+MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
+\rowcolor{white}
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{moreno2014genetics} \\
+\rowcolor{white}
+LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
+
+\rowcolor{gray!25}
+{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
+\rowcolor{gray!25}
+ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
+\rowcolor{gray!25}
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{zhang2016genome} \\ 
+\rowcolor{gray!25}
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015} \\
+
+\end{tabularx}
+\end{center}
+\end{table}
+
+\section*{Crop adaptation through introgression}
+
+As comprehensive genetic analyses of crops and their wild relatives have become feasible across their geographic ranges, evidence for substantial crop-wild introgression has been discovered in some of the world's most important crops (\textbf{Table 2}).
+Below we present a summary of findings from maize, barley, rice, and potato, four systems in which crop-wild gene flow appears to have played an adaptive role.
+All four of these crops were domesticated in defined centers and have subsequently expanded to much broader distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
+For each case study below we describe what is known about both a crop's domestication history and the prevalence of adaptive introgression during post-domestication  expansion.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=15cm]{temperature_plot_degC.pdf}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature. Approximate domestication center for each crop is denoted by a triangle}
+	\label{fig:map}
+\end{figure}
+
+\begin{enumerate}
+\item{Maize:}
+
+The relationship between maize (\emph{Zea mays} ssp.
+\emph{mays}) and the teosinte \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter, \emph{mexicana}) offers a prime example of adaptive wild-to-crop introgression.
+Maize was domesticated from \emph{Zea mays} ssp.
+\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
+Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
+However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
+These introgressed segments overlapped with QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
+Height differences were not detected under lowland conditions, providing further evidence of local adaptation.
+
+Populations of \emph{mexicana} cannot be found outside the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
+Wang and co-authors \cite{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
+Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but that more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
+Maize from the lowlands of northern South America appears to have been derived from Mesoamerican lowland maize and high elevation adaptation in Andean maize likely occurred \emph{de novo} from this genetic base \cite{Takuno2015}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=12cm]{./FigsAndTables/Figure1new2}
+	\caption{Evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize on chromosome 4.
+A.
+Stacked bar plots of the HAPMIX introgression scan.
+Data adopted from Hufford et al.
+\cite{Hufford2013}.
+ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
+The dashed vertical lines indicate one of the QTLs for macrohairs and pigment density in Lauter et al.
+\cite{lauter2004}.
+B.
+Phenotypic differences between lowland (red) and highland (blue) maize stems.
+\label{fig:introgressionMaize}}.
+\end{figure}
+
+\item{Barley:}
+
+Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild subsp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
+There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from subsp.
+\emph{spontaneum} east of the Zagros Mountains \cite{Morrell2007} or from subsp.
+\emph{spontaneum} var.
+\emph{agriocrithon} in modern-day Tibet \cite{dai2012tibet}.
+However, recent research casts doubt on Tibetan domestication and suggests that var.
+\emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \cite{pourkheirandish2018elucidation}.
+Presently, the distribution of wild barley stretches from the eastern Mediterranean through the Middle-East to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
+Cultivated barley is found throughout much of wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
+
+Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns of introgression.
+This study identified several lines of evidence consistent with wild introgression aiding the expansion and adaptation of domesticated barley.
+The authors utilized ancestry deconvolution methods to identify genomic regions of shared ancestry, which linked particular landraces to numerous wild relative populations.
+These results suggested landraces may have received wild introgression on a continual basis during post-domestication expansion.
+However, barley landraces also showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
+Limited admixture linkage disequilibrium and small tracts of identity by state suggest substantial recombination has occurred since initial crop-wild hybridization and that even locally introgressed chromosomal regions are ancient, perhaps dating to the early expansion of barley post-domestication.
+While these patterns suggest the possibility of adaptive introgression, wild barley haplotypes have yet to be definitively linked to specific adaptations in landraces post-domestication.
+
+\item{Asian Rice:}
+
+The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
+Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
+For example, Huang and colleagues \cite{Huang2012} measured genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was likely domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} potentially resulted from hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
+In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
+A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varieties.
+
+The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving} that also stresses the potential adaptive significance of crop-wild gene flow in rice.
+According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
+As these now fully domesticated populations spread further into new environments, they potentially received introgression from locally adapted wild relatives, retaining alleles that improved fitness.
+While the precise history of domesticated rice remains in question, multiple lines of evidence indicate diverse wild populations have contributed to domesticated germplasm and suggest adaptive introgression may have played a role during the expansion of this important crop.
+
+\item{Potato:}
+
+Modern potato (\emph{Solanum tuberosum}) was likely domesticated approximately 6,000-10,000 BP in southern Peru in sympatry with several wild relatives.
+The exact progenitor has remained in question for some time \cite{spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
+candolleanum} as the most probable progenitor \cite{hardigan2015taxonomy}.
+The lack of clarity regarding a progenitor has been due, in part, to extensive post-domestication hybridization between potato and a number of related species.
+
+While potatoes are primarily propagated clonally, farmers do at times promote sexual reproduction for improvement of the crop and development of new cultivars \cite{quiros1992increase}.
+Close proximity of domesticated potatoes and wild relatives, active hybridization, and local selection pressure favoring wild haplotypes across a diverse range of biotic and environmental conditions have likely fostered an expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
+The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \cite{hardigan2017genome}.
+These authors discovered that, in particular, tetraploid domesticates had received extensive introgression from wild \emph{Solanum}, with a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
+In certain cultivars, wild ancestry was estimated at upwards of 30\%.
+Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting introgression conferred adaptations critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
+\end{enumerate}
+
+The four crop systems described in detail here represent particularly compelling examples of wild introgression that was potentially adaptive.
+However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
+Across these four cases studies, some generalities can be observed.
+Data thus far indicate that wild introgression is often regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S.
+highlands), newly introgressed wild haplotypes can be disseminated more broadly.
+Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
+The possibility of both substantial and common wild gene flow raises a number of questions regarding both domestication and subsequent adaptation of crops.
+
+\section*{Re-evaluating domestication}
+
+A framework in which crops are domesticated from a single wild population or even a single species is an oversimplification when introgression has been extensive throughout a crop's history.
+The addition of ongoing gene flow to our understanding of crop demography could therefore bear on fundamental questions of crop domestication:
+
+\subsection*{Where and from what taxa did a crop originate?}
+Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's origin may be complicated or confounded entirely.
+Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
+This signal could be mistaken for origin rather than gene flow.
+For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
+Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
+Similarly, as described above, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
+
+Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a more nuanced view of crop origins.
+In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to modern crops.
+While these crops may have originated from a single species or subspecies, the crops as we know them today have a broader genetic base.
+
+\subsection*{When was a crop domesticated?}
+Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
+In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
+In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxa.
+This phenomenon, in combination with divergence of modern crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and genetic structure among wild relative populations, may help explain discrepancies between domestication dates based on genetic and archaeological data.
+More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.
+
+\subsection*{How has genome-wide diversity been shaped by domestication?}
+
+Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
+Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
+The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
+For example, substantial introgression from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.
+
+\subsection*{What candidate genes were targeted by selection during domestication?}
+Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \cite{Ross-Ibarra2007}.
+Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
+In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
+Due to the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
+Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
+For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \cite{vanHeerwaarden2011}.
+Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
+Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
+Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
+However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
+\vspace{5mm}
+
+In summary, since post-domestication gene flow with wild relatives appears frequent during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.
+
+\section*{Investigating crop adaptation through introgression:}
+
+While research in a subset of crops suggests adaptive wild introgression likely occurred, the scope and dynamics of this process are not yet fully described and remain unexplored in many systems.
+In determining the extent and nature of adaptation due to introgression, several questions should be considered:
+
+\subsection*{Do geographic patterns of introgression inform our understanding of adaptation?}
+Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is, in fact, adaptive.
+For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).
+On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests much more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
+Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.
+
+\subsection*{Over what timescales and in what genomic regions can we reliably detect adaptive introgression?}
+Introgressed haplotypes are most easily detected with limited recombination post-hybridization.
+Therefore, recent introgressions (limited meioses) or those occurring in low recombination regions such as centromeres or inversions are preferentially detected.
+While this can be problematic for the detection of ancient introgression, the fact that recombination degrades tracts of introgression at a relatively constant and predictable rate allows use of the genome-wide distribution of introgression tract lengths to date initial hybridization (as in \cite{Poets2015}).
+Detection of introgression will also be affected by mutation rate, effective population size, the strength of selection on introgressed alleles, and the extent of divergence between donor taxa and a crop's wild progenitor (\emph{e.g.}, highly divergent introgressed haplotypes will be easier to identify).
+
+\subsection*{From which wild taxa will introgression occur?}
+As species become substantially diverged, introgression can become maladaptive or impossible due to Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
+Divergence time may therefore be a useful predictor of the possibility of gene flow between a particular wild relative and a domesticate.
+Hybridization may also be limited between a crop and a particular wild relative due to genetic load. 
+For instance, gene flow from a wild relative with a small long-term effective population size, and correspondingly high genetic load may not be favored by selection.
+This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited in extent and relegated largely to non-genic regions due to the high genetic load found within Neanderthal donor individuals \cite{harris2016genetic}.
+
+\subsection*{Can adaptive introgression inform crop improvement?}
+
+Additional study of introgression in agroecosystems could lead to advances in crop improvement.
+As described above, loci underlying the domesticated phenotype can be more clearly identified by removing the confounding population genetic signals of introgression.
+These loci are potentially beneficial targets for crop improvement and their accurate identification is crucial.
+Furthermore, adaptive introgression that is demonstrably tied to a specific environment represents a promising source of beneficial alleles that can be directly utilized in breeding to adapt crops to similar conditions.
+Finally, as the historic role of wild relatives in the adaptation of crops is clarified, their conservation may be more prioritized, particularly as a resource for breeding in the face of future climate volatility and change.
+
+\section*{Conclusions}
+
+Recent innovations in both high-density marker data and methods for characterizing genome-wide patterns of introgression have helped reveal the extent and timing of gene flow in a number of species.
+Application of these data and techniques has led to mounting evidence of crop-wild gene flow following initial domestication in several crops.
+Substantial post-domestication gene flow with wild relatives can affect inferences made regarding domestication and may be an important mechanism through which crops were able to adapt to novel conditions during global expansions.
+An accurate understanding of the extent of gene flow is therefore important to both the basic study of crop evolution and to the identification of adaptive alleles for continued crop improvement.
+While multiple studies in crop systems have identified wild introgression, fewer have effectively linked introgressed alleles to adaptation.
+More comprehensive functional analyses and field evaluation will be critical for understanding the evolutionary and adaptive significance of introgression in crops.
+
+Domesticated crops are among humanity's greatest inventions.
+Interwoven in the genetic history of these crops is the story of our ancestors as they transitioned through the Agricultural Revolution.
+In a more nuanced and dynamic understanding of crop domestication that includes sustained gene flow with wild relatives, we gain insight into the evolutionary journey that crops and humanity took, and continue to take, together.
+
+\begin{adjustbox}{angle=90}
+    %\begin{tabular}{llllll}
+        \small
+    \begin{tabular}{G{3cm}G{3.5cm}G{3cm}G{3.5cm}G{3.5cm}G{2cm}}
+        \rowcolor{white}
+    \\\toprule  
+    {\bf Domesticated Crop}	& {\bf Compatible Wild Relatives } &	{\bf Hybrids and/or Hybridization} &  {\bf Evidence of Crop Introgression } &	{\bf Evidence of Adaptativeness} & {\bf Sources}\\ \midrule
+    \rowcolor{gray!25}
+Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{ma2017reduced} \\
+\rowcolor{white}
+Asian Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
+\rowcolor{gray!25}
+Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\
+\rowcolor{white}
+Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\
+\rowcolor{gray!25}
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{rendon2017genomic} \\
+\rowcolor{white}
+Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
+\rowcolor{gray!25}
+Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{Hufford2013} \\
+\rowcolor{white}
+Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
+\rowcolor{gray!25}
+
+Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{hardigan2017genome}\\
+\rowcolor{white}
+Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{han2016domestication} \\
+
+\rowcolor{gray!25}
+Sorghum (\emph{Sorghum bicolor} subsp. \emph{bicolor}) & \emph{S. b.} subsp. \emph{arundinaceum, S. b.} subsp. {drummondii} & X & X &  & \cite{aldrich1992patterns} \\
+\rowcolor{white}
+
+Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X &   &   & \cite{rieseberg2007hybridization}\\
+\rowcolor{gray!25}
+Tomato (\emph{Solanum lycopersicum}) & \emph{S. pimpinellifolium} & X & X & X & \cite{rick1958role} \\
+
+\rowcolor{white}
+ Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{dvorak2006molecular} \\
+%luo2007structure would be another appropriate citation for wheat (it is cited by dvorak2006molecular.  Strange that a paper from 2006 can cite a paper from 2007...
+    \end{tabular}
+\end{adjustbox}
+
+\bibliography{bib_gj.bib}
+
+\end{document}
+>>>>>>> 5130319157684b7bcffe4fe841969d5bc77b5a8d

### 2018-07-06 paper.tex
merge issues
edits to references
edits to reference

--- 2018-07-06 paper.tex
@@
 \usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
-\usepackage[numbers,sort]{natbib}
+\usepackage[sort]{natbib}
 \usepackage{amsmath}
@@
 \newcolumntype{G}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\bibliographystyle{ieeetr}
+\bibliographystyle{NewPhyt.bst}
 \renewcommand{\baselinestretch}{1.5}
@@
 
-Word Count:
+Word Count: 3,993
 
@@
 
-Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within defined centers followed by expansion to the modern-day range of cultivation \cite{Harlan1992}.
-However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{Meyer2016, Wang2017, Fuller2014}.
+Plant domestication is often conceptualized as a geographically constrained process, with crops originating from a wild progenitor within defined centers followed by expansion to the modern-day range of cultivation \citep{Harlan1992}.
+However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \citep{Meyer2016, Wang2017, Fuller2014}.
 An additional aspect of the emerging complexity of domestication is the occurrence of beneficial gene flow from locally adapted wild relatives to crops during their expansion following domestication.
@@
 
-Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \cite{barton2001role}.
+Adaptive introgression has three components: hybridization between differentiated taxa, backcrossing to one of the parents, and selection on recombinant genotypes with progressively diminished linkage drag \citep{barton2001role}.
 In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
-To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desirable traits from wild relatives (for a review, \cite{Dempewolf2017}).
+To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \citep{stewart2003transgene}) and on modern plant breeding efforts to introgress desirable traits from wild relatives (for a review, \citep{Dempewolf2017}).
 The potential for adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated comprehensive study of introgression across a broad spectrum of species.
-The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \cite{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$) \cite{gutenkunst2009}, isolation with migration models \cite{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC) \cite{li2011}).
-Rather, we focus on three categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \citealt{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$; \citealt{gutenkunst2009}), isolation with migration models \citep{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC; \citealt{li2011}).
+Rather, we focus on three categories of methods that explicitly identify introgressepd genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 
 First, introgressed segments are expected to show low differentiation from their source population.
-The $F_{st}$ and $d_{XY}$ statistics and derivates of $d_{XY}$ including $G_{min}$ \cite{geneva2015} and $RND_{min}$\cite{rosenzweig2016} gauge differentiation. 
+The $F_{st}$ and $d_{XY}$ statistics and derivates of $d_{XY}$ including $G_{min}$ \citep{geneva2015} and $RND_{min}$ \citep{rosenzweig2016} gauge differentiation. 
 The former two statistics are insensitive to rare migrants in a population and therefore lack power to detect very recent introgression, while the latter two overcome this limitation.
-These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \cite{racimo2016}. 
+These statistics have been further developed by adding differentiation between both non-admixed ($A$) and admixed populations ($B$) and a source population ($C$) \citep{racimo2016}. 
 For example, the $U_{A,B,C(w,x,y)}$ statistic summarizes the number of sites where an allele at frequency $y$ in the source population ($C$) has a frequency higher than $x$ in the admixed population ($B$) and lower than $w$ in the non-admixed population ($A$).
-A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \cite{racimo2016}.
+A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \citep{racimo2016}.
 Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
@@
 Below we describe the nascent application of these methods to crop systems as well as implications for the study of domestication and adaptation.
-
-\begin{table}[h]
-
-\rowcolors{2}{white}{gray!25}
-\begin{center}
-\caption{List of recently developed methods for detecting introgression and examples of their use in empirical studies.} \label{tab:tools}
-\begin{tabularx}{\textwidth}{llll}
-\\\toprule  
-\rowcolor{white}
-{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
-
-\rowcolor{gray!25}
-{\emph{\bf Divergence}} &   &   &   \\
-\rowcolor{gray!25}
-Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
-\rowcolor{gray!25}
-RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
-\rowcolor{gray!25}
-$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
-
-\rowcolor{white}
-{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
-\rowcolor{white}
-Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013} \\ 
-\rowcolor{white}
-RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
-\rowcolor{white}
-MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
-\rowcolor{white}
-PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{moreno2014genetics} \\
-\rowcolor{white}
-LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
-
-\rowcolor{gray!25}
-{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
-\rowcolor{gray!25}
-ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
-\rowcolor{gray!25}
-fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{zhang2016genome} \\ 
-\rowcolor{gray!25}
-five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015} \\
-
-\end{tabularx}
-\end{center}
-\end{table}
 
@@
 
+\begin{table}[h]
+
+\rowcolors{2}{white}{gray!25}
+\begin{center}
+\caption{List of recently developed methods for detecting introgression and examples of their use in empirical studies.} \label{tab:tools}
+\begin{tabularx}{\textwidth}{llll}
+\\\toprule  
+\rowcolor{white}
+{\bf Methods}	& {\bf Data Type } &	{\bf References} &  {\bf Applications } \\ \midrule
+
+\rowcolor{gray!25}
+{\emph{\bf Divergence}} &   &   &   \\
+\rowcolor{gray!25}
+Gmin &	biallelic SNP	&  \cite{geneva2015}	 &  \cite{kingan2015}\\
+\rowcolor{gray!25}
+RNDmin	& phased haplotype	& \cite{rosenzweig2016} &  \cite{roda2017} \\
+\rowcolor{gray!25}
+$U_{A,B,C(w,x,y)}$ and $Q95_{A,B,C(w,y)}$ & biallelic SNP & \cite{Racimo2015} & \cite{sams2016} \\
+
+\rowcolor{white}
+{\emph{\bf Ancestry Deconvolution}} &   &   &   \\
+\rowcolor{white}
+Hapmix	& phased haplotype; reference panel		& \cite{Price2009}	&  \cite{Hufford2013} \\ 
+\rowcolor{white}
+RASPberry &	phased haplotype &	\cite{wegmann2011}	 & \cite{christe2016} \\
+\rowcolor{white}
+MultiMix & phased/unphased genotype; reference panel &	\cite{churchhouse2013} &	\cite{eyheramendy2015} \\
+\rowcolor{white}
+PCAdmix	 & phased haplotype	 & \cite{brisbin2012pcadmix}	 &  \cite{moreno2014genetics} \\
+\rowcolor{white}
+LAMP  &	phased haplotypes; reference panel	 & \cite{sankararaman2008}	 & \cite{patterson2012} \\
+
+\rowcolor{gray!25}
+{\emph{\bf Phylogenetic Relationship}} &   &   &   \\
+\rowcolor{gray!25}
+ABBA-BABA/D-statistics	 & biallelic SNP  &	\cite{durand2011}	 &  \cite{heliconius2012} \\
+\rowcolor{gray!25}
+fd statistic &	biallelic SNP &	\cite{martin2015}  &	\cite{zhang2016genome} \\ 
+\rowcolor{gray!25}
+five taxon D statistics	& biallelic SNP	&  \cite{pease2015}	& \cite{fontaine2015} \\
+
+\end{tabularx}
+\end{center}
+\end{table}
+
 \begin{adjustbox}{angle=90}
@@
     \rowcolor{gray!25}
-Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{cornille2012new} \\
+Apple (\emph{Malus domesticus}) & \emph{M. sylvestris}, \emph{M. orientalis}, \emph{M. baccata}, \emph{M. sieversii}  & X & X & & \cite{ma2017reduced} \\
 \rowcolor{white}
@@
 \rowcolor{gray!25}
-Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{papa2003asymmetry} \\
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{rendon2017genomic} \\
 \rowcolor{white}
@@
 \rowcolor{white}
-Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\
+Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{han2016domestication} \\
 
@@
 \rowcolor{white}
- Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{zohary1969wild} \\
+ Wheat (\emph{Tritium monococcum, T. dicoccum, T. aestivum}) & \emph{T. m. boeoticum, T. dioccoides, T. urartu, Aegilops speltoides, A. tauschii} & X & X &  & \cite{dvorak2006molecular} \\
+
     \end{tabular}

### 2018-07-08 paper.tex
formatting edits

@@
 \newcolumntype{G}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 \bibliographystyle{NewPhyt.bst}
@@
 
-\title{Review: The Extent of Adaptive Wild Introgression in Crops}
+\title{Research Review: \\The Extent of Adaptive Wild Introgression in Crops}
 
@@
 Post-domestication adaptation during the expansion of crops from their centers of origin has received considerably less attention.
-Recent research has revealed that, in at least some instances, crops have continued a close relationship with their wild relatives that has facilitated adaptation to novel conditions encountered during expansion.
+Recent research has revealed that, in at least some instances, crops have received introgression from their wild relatives that has facilitated adaptation to novel conditions encountered during expansion.
 Such adaptive introgression could bear importantly on the basic study of domestication, affecting estimates of several evolutionary processes of interest (\emph{e.g.}, the strength of the domestication bottleneck, the timing of domestication, the targets of selection during domestication).
@@
 In domesticated species, adaptive introgression would consist of crop-wild hybrids backcrossing to a crop followed by increase in frequency of adaptive wild alleles in the crop and selection against undesirable wild background.
-To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \citep{stewart2003transgene}) and on modern plant breeding efforts to introgress desirable traits from wild relatives (for a review, \citep{Dempewolf2017}).
+To date, the literature on crop-wild gene flow has largely focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \citealt{stewart2003transgene}) and on modern plant breeding efforts to introgress desirable traits from wild relatives (for a review, \citealt{Dempewolf2017}).
 The potential for adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
@@
 The decreasing cost of genome-wide resequencing and availability of reduced-representation genotyping (\emph{e.g.}, GBS and RAD-Seq), combined with new analytical methods (\textbf{Table 1}), has facilitated comprehensive study of introgression across a broad spectrum of species.
-The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC) \citealt{beaumont2002}, diffusion approximations for demographic inference ($\delta a\delta i$; \citealt{gutenkunst2009}), isolation with migration models \citep{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC; \citealt{li2011}).
-Rather, we focus on three categories of methods that explicitly identify introgressepd genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
+The methods reviewed here do not include those estimating introgression\slash migration rate as a component of broader demographic history (\emph{e.g.}, Approximate Bayesian Computation (ABC; \citealt{beaumont2002}), diffusion approximations for demographic inference ($\delta a\delta i$; \citealt{gutenkunst2009}), isolation with migration models \citep{hey2004}, and methods utilizing the sequentially Markovian coalescent (\emph{e.g.}, PSMC; \citealt{li2011}).
+Rather, we focus on three categories of methods that explicitly identify introgressed genomic segments based on the extent of differentiation, patterns of nucleotide/haplotype sharing, and phylogenetic relationships.
 
@@
 A similar statistic, $Q95_{A,B,C(w,y)}$, sets a hard cutoff at the $95^{th}$ percentile of allele frequency in the admixed population (B) \citep{racimo2016}.
-Further modifications have allowed specification of more than one source population (see details in \cite{racimo2016}).
+Further modifications have allowed specification of more than one source population (see details in \citealt{racimo2016}).
 Since differentiation-based methods can be calculated site-by-site, high-density, genome-wide data are not necessarily required.
@@
 
-Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to source populations based on patterns of allele or haplotype sharing \cite{schraiber2015}.
-One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \cite{Price2009}).
-Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix \cite{brisbin2012pcadmix} and LAMP \cite{sankararaman2008}).
-A third version uses a Bayesian model \cite{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE \cite{Lawson2012}).
+Second, ancestry deconvolution (also known as local-ancestry inference and chromosome painting) assigns genomic regions to source populations based on patterns of allele or haplotype sharing \citep{schraiber2015}.
+One form of ancestry deconvolution utilizes a hidden Markov model to evaluate ancestry across admixed genomes through comparison to reference, non-admixed individuals (\emph{e.g.}, HAPMIX \citealt{Price2009}).
+Another clusters admixed populations with reference samples using a sliding-window approach (\emph{e.g.}, PCAdmix, \citealt{brisbin2012pcadmix} and LAMP, \citealt{sankararaman2008}).
+A third version uses a Bayesian model \citep{pritchard2000} in which deviations from Hardy-Weinberg equilibrium are minimized through creation of genetic groups (\emph{e.g.}, fineSTRUCTURE, \citealt{Lawson2012}).
 Ancestry deconvolution methods are better suited to high-density marker data given the intent to assign ancestry genome-wide.
@@
 
-Phylogenetic relationships are applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics.
+Phylogenetic relationships are applied to introgression detection using the ABBA-BABA statistic (also known as the D-statistic) and related metrics \citep{durand2011}.
 These statistics make inferences regarding introgression based on genomic patterns of derived variants that are shared between populations or species.
-While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \cite{martin2015} and $D_{\textrm{FOIL}}$ tests \cite{pease2015} are capable of localizing introgression to specific chromosomal regions.
+While the D-statistic is best suited to detection of introgression at the genome level, elaborations of the D-statistic including $\hat{f_{d}}$ \citep{martin2015} and $D_{\textrm{FOIL}}$ tests \citep{pease2015} are capable of localizing introgression to specific chromosomal regions.
 The former is quite similar to the D-statistic but is based on allele frequencies, and the latter can identify donor and recipient lineages of introgression in a more complex, five-taxon phylogeny.
@@
 Collectively, introgression detection methods have been applied to several systems, frequently identifying instances of adaptive introgression (see applications in \textbf{Table 1}).
-For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in adaptation to pollinator preference and speciation \cite{Stankowski2015}.
-Likewise, the HAPMIX ancestry deconvolution method was applied by Jeong et al.
-\cite{jeong2014} to detect introgression from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
-Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \cite{heliconius2012}.
+For example, based on sequence divergence methods, introgression has been detected in \emph{Mimulus} (\emph{i.e.}, monkeyflower) species and appears to play a role in adaptation to pollinator preference and speciation \citep{Stankowski2015}.
+Likewise, the HAPMIX ancestry deconvolution method was applied by \citet{jeong2014} to detect introgression from the Nepalese Sherpa to Tibetans at loci controlling high altitude adaptation.
+Finally, the ABBA-BABA statistic has revealed introgression at wing coloration loci conferring M\"{u}llerian mimicry across butterfly species \citep{heliconius2012}.
 Below we describe the nascent application of these methods to crop systems as well as implications for the study of domestication and adaptation.
@@
 Below we present a summary of findings from maize, barley, rice, and potato, four systems in which crop-wild gene flow appears to have played an adaptive role.
-All four of these crops were domesticated in defined centers and have subsequently expanded to global distributions, a migration that brought them into contact with new populations of wild relatives (Figure \ref{fig:map}).
+All four of these crops were domesticated in defined centers and have subsequently expanded to global distributions, a migration that brought them into contact with new populations of wild relatives that are distributed across broad environmental gradients (\textbf{Figure 1}).
 For each case study we describe both what is known about a crop's domestication history and the prevalence of adaptive introgression during expansion.
@@
 	\includegraphics[width=15cm]{temperature_plot_degC.pdf}
-	\caption{Map of the natural ranges of wild relatives of four domesticated crops, overlayed with average annual temperature. Approximate domestication center for each crop is denoted by a triangle}
+	\caption{Map of the natural ranges of wild relatives of four domesticated crops overlayed on average annual temperature. Approximate domestication center for each crop is denoted by a triangle}
 	\label{fig:map}
@@
 Maize was domesticated from \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
+\emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \citep{matsuoka2002single}.
 From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
-Introgression from \emph{mexicana} to maize has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
-However, Hufford et al. \cite{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
-The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (Figure \ref{fig:introgressionMaize}).
-These introgressed segments overlapped with QTL that had previously been shown to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
+Introgression from \emph{mexicana} to maize has been reported based on both morphological \citep{wilkes1977} and molecular \citep{vanHeerwaarden2011, doebley1987} data.
+However, \citet{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
+The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (\textbf{Figure 2}).
+These introgressed segments overlapped with QTL that had previously been shown to control anthocyanin content and leaf macrohairs \citep{lauter2004}, traits known to be adaptive at high elevation.
 In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression had increased height (a proxy for fitness) under highland environmental conditions.
@@
 Populations of \emph{mexicana} cannot be found outside the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
-Wang and co-authors \cite{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
+\citet{Wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate if maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
 Overall, analyses revealed that  \emph{mexicana} introgression was pervasive in maize from Mesoamerican high-elevation regions (the highlands of Mexico, Guatemala, and the southwestern United States), but that more distant high-elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry.
-These findings are consistent with previous work suggesting high elevation adaptation in Andean maize likely occurred \emph{de novo} \cite{Takuno2015}.
+These findings are consistent with previous work suggesting high elevation adaptation in Andean maize likely occurred \emph{de novo} \citep{Takuno2015}.
 
@@
 	\caption{Evidence of adaptive introgression from \emph{mexicana} to Mexican highland maize on chromosome 4.
-A.
-Stacked bar plots of the HAPMIX introgression scan.
-Data obtained from Hufford et al.
-\cite{Hufford2013}.
+(a) Stacked bar plots of a HAPMIX introgression scan.
+For each population and chromosomal position, red indicates maize ancestry and blue indicates \emph{mexicana} ancestry.
+Data obtained from \citet{Hufford2013}.
 ELP: EL Porvenir; IXT: Ixtlan ; NAB: Nabogame; OPO: Opopeo; PUR: Puruandiro; SAC: Santa Clara; SAP: San Pedro; TEN: Tenango del Aire; XOC: Xochimilco.
-The dashed vertical lines indicate a previously identified QTL for macrohairs and pigment density in Lauter et al.
-\cite{lauter2004}.
-B.
-Phenotypic differences between lowland (red) and highland (blue) maize stems.
+The dashed vertical lines indicate a previously identified QTL for macrohairs and pigment density in \citet{lauter2004}.
+(b) Phenotypic differences between maize stems with (blue) and without (red) \emph{mexicana} introgression in QTL affecting presence of pigment and macrohairs.
 \label{fig:introgressionMaize}}.
@@
 
-Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was likely domesticated multiple times from wild ssp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
-There is clear evidence of one domestication center in the Fertile Crescent \cite{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from ssp.
-\emph{spontaneum} east of the Zagros Mountains \cite{Morrell2007} or from ssp.
+Barley (\emph{Hordeum vulgare} ssp. \emph{vulgare}) was likely domesticated multiple times from wild ssp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
+There is clear evidence of one domestication center in the Fertile Crescent \citep{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from ssp.
+\emph{spontaneum} east of the Zagros Mountains \citep{Morrell2007} or from ssp.
 \emph{spontaneum} var.
-\emph{agriocrithon} in modern-day Tibet \cite{dai2012tibet}.
+\emph{agriocrithon} in modern-day Tibet \citep{dai2012tibet}.
 However, recent research casts doubt on Tibetan domestication and suggests that var.
-\emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \cite{pourkheirandish2018elucidation}.
-Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \cite{Morrell2007}.
+\emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \citep{pourkheirandish2018elucidation}.
+Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \citep{Morrell2007}.
 Cultivated barley is found throughout wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
 
-Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns of introgression.
+\citet{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns of introgression.
 This study identified several lines of evidence consistent with wild introgression aiding the expansion and adaptation of domesticated barley.
@@
 The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
-Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon}, hereafter) in China and the Indian Ganges plain, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
-Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012} during crop expansion.
-For example, Huang and colleagues \cite{Huang2012} measured genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was likely domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} potentially resulted from hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
-In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
-A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and concluded that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and were introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varietals.
+Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon}, hereafter) in China and the Indian Ganges plain, respectively \citep{fuller2010consilience}, with a potential third domestication event giving rise to the varietal group \emph{aus} in Bangladesh or central India \citep{civavn2015three}.
+Other studies support a single domestication in China, with later divergence of \emph{japonica} and \emph{indica} \citep{molina2011molecular, Huang2012} during crop expansion.
+For example, \citet{Huang2012} measured genetic distance between a range-wide sample of wild and domesticated rice, finding that \emph{japonica} was likely domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} potentially resulted from hybridization between \emph{japonica} and local \emph{rufipogon} populations in southern and south-eastern Asia.
+In a re-examination of these same data, Civ\'{a}\v{n} and colleagues \citeyearpar{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of aromatic rice.
+A recent third analysis by \citet{choi2018multiple} compared these two disparate results and concluded that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose during a single domestication event of \emph{japonica}, and were introgressed into several wild \emph{rufipogon} subpopulations (crop-to-wild gene flow), which thereby became the progenitors of other Asian rice varietals.
 
-The findings of \cite{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving} that stresses the potential adaptive significance of crop-wild gene flow in rice.
+The findings of \citet{choi2018multiple} bear similarities to a hypothesis posited by Vaughan and colleagues \citeyearpar{vaughan2008evolving} that stresses the potential adaptive significance of crop-wild gene flow in rice.
 According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like).
@@
 Modern potato (\emph{Solanum tuberosum}) was likely domesticated approximately 6,000-10,000 BP in southern Peru in sympatry with several wild relatives.
-The exact progenitor has remained in question for some time \cite{spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
-candolleanum} as the most probable progenitor \cite{hardigan2015taxonomy}.
+The exact progenitor has remained in question for some time \citep{spooner2005single, pickersgill1977origins, hawkes1988evolution}, but a distance-based phylogeny constructed using genotypic data from a \emph{Solanum} diversity panel recently identified \emph{S.
+candolleanum} as the most probable progenitor \citep{hardigan2015taxonomy}.
 The lack of clarity regarding a progenitor has been due, in part, to extensive post-domestication hybridization between potato and a number of related species.
 
-While potatoes are primarily propagated clonally, farmers do at times promote sexual reproduction for improvement of the crop and development of new cultivars \cite{quiros1992increase}.
-Close proximity of domesticated potatoes and wild relatives, active hybridization, and local selection pressure favoring wild haplotypes across a diverse range of biotic and environmental conditions have likely fostered an expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
-The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \cite{hardigan2017genome}.
+While potatoes are primarily propagated clonally, farmers do at times promote sexual reproduction for improvement of the crop and development of new cultivars \citep{quiros1992increase}.
+Close proximity of domesticated potatoes and wild relatives, active hybridization, and local selection pressure favoring wild haplotypes across a diverse range of biotic and environmental conditions have likely fostered an expansion of genetic diversity within potatoes subsequent to domestication \citep{brush1995potato}.
+The prevalence of wild introgression was recently clarified in a broad survey of potato diversity by Hardigan and colleagues \citeyearpar{hardigan2017genome}.
 These authors discovered that tetraploid domesticates in particular had received extensive introgression from wild \emph{Solanum}, documenting a continued broadening of the genetic base of potato as it spread away from its Peruvian origin.
 In certain cultivars, wild ancestry was estimated at upwards of 30\%.
-Genes located within these introgressed regions were more likely to be highly-expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting introgression conferred adaptations critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \cite{hardigan2017genome}.
+Genes located within these introgressed regions were more likely to be highly expressed and stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting introgression conferred adaptations critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion \citep{hardigan2017genome}.
 \end{enumerate}
 
-The four crop systems described in detail here represent particularly compelling examples of wild introgression that was potentially adaptive.
+The four crop systems described in detail here represent particularly compelling examples of wild and potentially adaptive introgression.
 However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
@@
 Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
-The possibility of both substantial and common wild gene flow raises a number of questions regarding both domestication and adaptation of crops.
+The wild gene flow identified in these case studies raises a number of questions regarding both domestication and adaptation of crops.
 
@@
 Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's origin may be complicated or confounded entirely.
-Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
+Introgression between a crop and newly encountered taxa decreases divergence of the crop from these donors.
 This signal could be mistaken for origin rather than gene flow.
-For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
-Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
+For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \citeyearpar{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
+Several years later, \citet{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
 Similarly, as described above, extensive post-domestication adaptive introgression from potato wild relatives long obscured this crop's origin.
@@
 \subsection*{When was a crop domesticated?}
-Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
+Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \citep{matsuoka2002single, molina2011molecular}).
 In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
@@
 \subsection*{What candidate genes were targeted by selection during domestication?}
-Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \cite{Ross-Ibarra2007}.
+Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \citep{Ross-Ibarra2007}.
 Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
@@
 Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
-For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \cite{vanHeerwaarden2011}.
-Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
+For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \citep{vanHeerwaarden2011}.
+Takuno and co-authors \citeyearpar{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
 Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
-However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
+However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \citep{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
 \vspace{5mm}
 
-In summary, since post-domestication gene flow with wild relatives appears frequent during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.
+In summary, since post-domestication gene flow with wild relatives appears frequent during crop histories, investigations seeking to unravel fundamental questions of initial domestication must take this into account in order to accurately estimate parameters of interest.
 
@@
 Therefore, recent introgressions (limited meioses) or those occurring in low recombination regions such as centromeres or inversions are preferentially detected.
-While this can be problematic for the detection of ancient introgression, the fact that recombination degrades tracts of introgression at a relatively constant and predictable rate allows use of the genome-wide distribution of introgression tract lengths to date initial hybridization (as in \cite{Poets2015}).
+While this can be problematic for the detection of ancient introgression, the fact that recombination degrades tracts of introgression at a relatively constant and predictable rate allows use of the genome-wide distribution of introgression tract lengths to date initial hybridization (as in \citealt{Poets2015}).
 Detection of introgression will also be affected by mutation rate, effective population size, the strength of selection on introgressed alleles, and the extent of divergence between donor taxa and a crop's wild progenitor (\emph{e.g.}, highly divergent introgressed haplotypes will be easier to identify).
@@
 \subsection*{From which wild taxa will introgression occur?}
-As species become substantially diverged, introgression can become maladaptive or impossible due to Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
+As species become substantially diverged gene flow is limited due to Dobzhansky-Muller incompatibilities and other pre- and post-zygotic barriers.
 Divergence time may therefore be a useful predictor of the possibility of gene flow between taxa.
@@
 For instance, gene flow from a wild relative with a small long-term effective population size, and correspondingly high genetic load may not be favored by selection.
-This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited and relegated primarily to non-genic regions due to the high genetic load found within Neanderthal donor individuals \cite{harris2016genetic}.
+This effect has been observed in the case of Neanderthal introgression into humans, which was likely limited and relegated primarily to non-genic regions due to the high genetic load found within Neanderthal donor individuals \citep{harris2016genetic}.
 
@@
 Application of these data and techniques has led to mounting evidence of crop-wild gene flow following initial domestication in several species.
-Substantial post-domestication gene flow with wild relatives can affect inferences made regarding domestication and may be an important mechanism through which crops adapted to novel conditions during global expansion.
+Substantial post-domestication gene flow with wild relatives can affect inferences regarding domestication and may be an important mechanism through which crops adapted to novel conditions during global expansion.
 An accurate understanding of the extent of gene flow is therefore important to both the basic study of crop evolution and to the identification of adaptive alleles for continued crop improvement.
@@
 
+\nolinenumbers
+
 \begin{table}[h]
-
+\small
 \rowcolors{2}{white}{gray!25}
@@
 \caption{List of recently developed methods for detecting introgression and examples of their use in empirical studies.} \label{tab:tools}
-\begin{tabularx}{\textwidth}{llll}
+\begin{tabular}{G{2.8cm}G{3.5cm}G{4cm}G{4cm}}
 \\\toprule  
@@
 
-\begin{adjustbox}{angle=90}
-    %\begin{tabular}{llllll}
+\hspace*{-1.5cm}
+\begin{sidewaystable}
+\caption{\small Extent of evidence for adaptive introgression for major crops including whether hybrids are observed, introgression is detected, and introgression has been shown to be adaptive.} \label{tab:intro}
         \small
-    \begin{tabular}{G{3cm}G{3.5cm}G{3cm}G{3.5cm}G{3.5cm}G{2cm}}
+    \begin{tabular}{G{4.5cm}G{5.5cm}C{1.75cm}C{2.5cm}C{2.25cm}G{3.5cm}}
         \rowcolor{white}
     \\\toprule  
-    {\bf Domesticated Crop}	& {\bf Compatible Wild Relatives } &	{\bf Hybrids and/or Hybridization} &  {\bf Evidence of Crop Introgression } &	{\bf Evidence of Adaptativeness} & {\bf Sources}\\ \midrule
+    {\bf Domesticated Crop}	& {\bf Compatible Wild Relatives} &	{\bf Hybrids} &  {\bf Evidence of Introgression } &	{\bf Evidence of Adaptation} & {\bf Sources}\\ \midrule
     \rowcolor{gray!25}
@@
 \rowcolor{gray!25}
-Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} [[not in this source]]& X & X &  & \cite{rendon2017genomic} \\
+Common Bean (\emph{Phaseolus vulgaris}) & \emph{P. v.} var. \emph{aborigineus, P. v.} var. \emph{mexicanus} & X & X &  & \cite{rendon2017genomic} \\
 \rowcolor{white}
@@
     \end{tabular}
-\end{adjustbox}
+\end{sidewaystable}
 

### 2018-07-26 paper.tex
updated word count
updated (see slack for details)

@@
 
-Word Count: 3,993
+Word Count: 3,966
 
@@
 Introgression from \emph{mexicana} to maize has been reported based on both morphological \citep{wilkes1977} and molecular \citep{vanHeerwaarden2011, doebley1987} data.
-However, \citet{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
+\citet{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
 The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (\textbf{Figure 2}).
@@
 Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \citep{Morrell2007}.
-Cultivated barley is found throughout wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.
+Cultivated barley is found throughout wild barley's distribution \cite{harlan1966distribution} and crop-wild hybrids are fertile and common when these taxa co-occur.
 
@@
 
-The four crop systems described in detail here represent particularly compelling examples of wild and potentially adaptive introgression.
+These four crop system case studies represent particularly compelling examples of wild and potentially adaptive introgression.
+%The four crop systems described in detail here represent particularly compelling examples of wild and potentially adaptive introgression.
 However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
@@
 \subsection*{What candidate genes were targeted by selection during domestication?}
+
 Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \citep{Ross-Ibarra2007}.
@@
 \subsection*{Do geographic patterns of introgression inform our understanding of adaptation?}
-Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is, in fact, adaptive.
+Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is adaptive.
 For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).

### 2018-07-30 paper.tex
Update paper.tex
revised the Allaby 2015 sentence.

@@
 While these crops may have originated from a single species or subspecies, the crops as we know them today have a broader genetic base.
+%In addtion, the domestications of barley and rice might best be described as geographically noncentric due to the significant genetic contributions from wide-ranging locations \cite{allaby2015barley}.
+The recognized contributions of multiple wild relatives and proto-domesticated crop varieties from across distant geographies into modern crops (which appears to be the case in barley and rice, for example) highlight the importance of polycentric and concentric conceptions of domestication \cite{allaby2015barley}.
 

### 2018-07-30 paper.tex
revised Allaby 2015 sentence again

--- 2018-07-26 paper.tex
@@
 While these crops may have originated from a single species or subspecies, the crops as we know them today have a broader genetic base.
+In other cases (such as barley and rice), crops appear to have arisen from the combined contributions of multiple wild or proto-domesticated crop varieties from across distant geographies.
+Such reticulated evolutionary histories defy traditional centric conceptions of domestication in favor of polycentric or noncentric origins \cite{allaby2015barley}.
 

### 2018-07-30 paper.tex
Cut words in section "what candidate genes were.."

@@
 Takuno and co-authors \citeyearpar{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
-Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
+Failure to account for introgression in maize would therefore compromise domestication candidate detection.
 Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
-However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \citep{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
-\vspace{5mm}
-
-In summary, since post-domestication gene flow with wild relatives appears frequent during crop histories, investigations seeking to unravel fundamental questions of initial domestication must take this into account in order to accurately estimate parameters of interest.
 

### 2018-07-31 paper.tex
explained phasing
recency of introgressions

@@
 However, accuracy of introgression estimates is improved with more comprehensive data.
-Phased data are also not a prerequisite for differentiation-based methods.
+Phased data (i.e., data aligned to maternal or paternal chromosomes) are also not a prerequisite for differentiation-based methods.
 
@@
 However, because these methods require knowledge of whether an allele is ancestral or derived, data from a sufficiently diverged outgroup must be available.
+%Using such methods as the D-statistic or HAPMIX to identify introgressant haplotypes and measure their length, for a given rate of linkage disequilibrium breakdown due to outcrossing frequency, one can estimate the recency of an introgression event.
+
+The recency of an introgression event can be assessed by measuring patterns of linkage disequilibrium (LD) and identity by state (IBS) between the domesticated and wild populations.
+Recent introgression will result in high LD \cite{chakraborty1988recombination} and genomic regions of high IBS \cite{gusev2011architecture}, both of which break down over time.
+LD will break down faster in crops with high outcrossing frequencies due to greater recombination rates.
+LD breakdown can be assessed with D-statistics or HAPMIX, and decrease in IBS can be assessed with with several tools, including the PLINK toolset \cite{purcell2007plink}.
+Determining that a crop genome has low levels of LD or IBS with wild relatives can be cited as evidence that any introgressions are not recent (as in Poets et al. 2015 \cite{Poets2015).
 

### 2018-08-02 paper.tex

@@
 Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
+Due to the domestication bottleneck, domesticated organisms typically have lower $N_e$, less effective genome-wide selection, and therefore higher genetic load, compared to their wild relatives, a pattern known as the "cost of domestication" \cite{moyers2017genetic}.
 The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
 For example, substantial introgression from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.
+
+Depending on the $N_e$ of wild species relative to the domesticate, introgression into the domesticate can either ameliorate or exascerbate the prevalence of deleterious alleles.
+Many crops show evidence of genetic load, including maize \cite{wang2017interplay}, cassava \cite{ramu2017cassava}, and grape \cite{zhou2017evolutionary}.
+For this reason, introgression from high-$N_e$ wild relatives is more likely to reduce load.
+In accord with these predictions, Li et al.
+\cite{wang2017interplay} found that $N_e$ of maize dropped due to the domestication bottleneck, but that highland maize from Mexico, Guatemala, and the southwestern United States which maintained mexicana introgressions subsequently had fewer deleterious alleles.
 
